### 2016-11-29 paper/paper.tex
added some notes on method

#generate a paper tex file

#Method:
magSN = parallaxSN = 16 (including systematic parallax error in tgas)
minimum galactic latitude = 5 degrees
no NaNs in apass photometry, apass errors > 0
matched in Apass, wise, 2mass, rave
check that matching looks good with extreme colors
calculate distance modulus from sampling of gaia distances
correct for dust using Bayestar (Schlafly+Finkbeiner); minimal dust to matched stars
use absolute V band magnitude, B-V, g-r r-i colors

build KDTree of matched stars in Mv/color space
grab 2048 nearest neighbors
calculate chisq like distance metric between each chosen star and its neighbors
generate posterior belief of a chosen stars spectroscopic characteristics (ie Teff, logg, FeH)
  given the spectroscopic characteristics (and uncertainties) of its nearest neighbors in Mv/color space
.
.
.
.

#Discussion:
Apass HR Diagram looks very different from Rave HR Diagram, MS stars are incorrectly being labeled as giants

### 2016-12-28 paper/paper.tex
make paper to make

@@
-#generate a paper tex file
+% This file is part of the stellarTwins project.
+% Copyright 2016 the authors.
 
-#Method:
+\documentclass[11pt,letterpaper]{article}
+
+\begin{document}
+
+\section*{Title here}
+
+\noindent
+Lauren Anderson (Flatiron) \& David W. Hogg (Flatiron)
+
+\paragraph{Abstract:}
+foo and bar
+
+\section{Introduction}
+
+Label transfer, The Cannon, etc.
+
+\section{Method}
+
 magSN = parallaxSN = 16 (including systematic parallax error in tgas)
+
 minimum galactic latitude = 5 degrees
+
 no NaNs in apass photometry, apass errors > 0
+
 matched in Apass, wise, 2mass, rave
+
 check that matching looks good with extreme colors
+
 calculate distance modulus from sampling of gaia distances
+
 correct for dust using Bayestar (Schlafly+Finkbeiner); minimal dust to matched stars
+
 use absolute V band magnitude, B-V, g-r r-i colors
@@
 build KDTree of matched stars in Mv/color space
+
 grab 2048 nearest neighbors
+
 calculate chisq like distance metric between each chosen star and its neighbors
+
 generate posterior belief of a chosen stars spectroscopic characteristics (ie Teff, logg, FeH)
+
 given the spectroscopic characteristics (and uncertainties) of its nearest neighbors in Mv/color space
-.
-.
-.
-.
 
-#Discussion:
+\section{Discussion}
+
 Apass HR Diagram looks very different from Rave HR Diagram, MS stars are incorrectly being labeled as giants
+
+\end{document}

### 2016-12-28 paper/paper.tex
author
more structure; ack

@@
 % Copyright 2016 the authors.
+\documentclass[11pt,letterpaper]{article}
 
-\documentclass[11pt,letterpaper]{article}
+\newcommand{\acronym}[1]{\small{#1}}
+\newcommand{\project}[1]{\textsl{#1}}
+\newcommand{\Gaia}{\project{Gaia}}
 
@@
 
-\section*{Title here}
+\section*{A photometric method for obtaining spectroscopic parameters of stars}
 
 \noindent
-Lauren Anderson (Flatiron) \& David W. Hogg (Flatiron)
+Lauren Anderson (Flatiron), David W. Hogg (Flatiron), others
 
@@
 
+\paragraph{Acknowledgements:}
+This project was developed in part at the 2016 NYC Gaia Sprint, hosted
+by the Center for Computational Astrophysics at the Simons Foundation
+in New York City.
+
+This work has made use of data from the European Space Agency (ESA)
+mission \Gaia\ (http://www.cosmos.esa.int/gaia), processed by the \Gaia
+Data Processing and Analysis Consortium (\acronym{DPAC},
+http://www.cosmos.esa.int/web/gaia/dpac/consortium). Funding for the
+\acronym{DPAC} has been provided by national institutions, in particular the
+institutions participating in the Gaia Multilateral Agreement.
+
 \end{document}

### 2016-12-28 paper/paper.tex
typesetting tweaks

@@
 
-\newcommand{\acronym}[1]{\small{#1}}
+\newcommand{\acronym}[1]{{\small{#1}}}
 \newcommand{\project}[1]{\textsl{#1}}
@@
 
+\linespread{1.09}
+\setlength{\parindent}{1.2\baselineskip}
+\sloppy\sloppypar\raggedbottom\frenchspacing
 \begin{document}
 
-\section*{A photometric method for obtaining spectroscopic parameters of stars}
+\section*{\raggedright%
+A photometric method for obtaining spectroscopic parameters of stars}
 
@@
 This work has made use of data from the European Space Agency (ESA)
-mission \Gaia\ (http://www.cosmos.esa.int/gaia), processed by the \Gaia
+mission \Gaia\ (http://www.cosmos.esa.int/gaia), processed by the \Gaia\ %
 Data Processing and Analysis Consortium (\acronym{DPAC},

### 2016-12-29 paper/paper.tex
abstract

@@
 \newcommand{\Gaia}{\project{Gaia}}
+\newcommand{\GALEX}{\project{\acronym{GALEX}}}
+\newcommand{\APASS}{\project{\acronym{APASS}}}
+\newcommand{\RAVEon}{\project{\acronym{RAVE}-on}}
 
@@
 \section*{\raggedright%
-A photometric method for obtaining spectroscopic parameters of stars}
+A photometric method for obtaining spectroscopic parameters of stars,
+with no use of physical models}
 
@@
 \paragraph{Abstract:}
-foo and bar
+% Context
+Stellar photospheres, to first order, have their spectral properties
+set by the local quantities (spectroscopic parameters) effective
+temperature, surface gravity, and metallicity (chemical composition);
+As their name suggests, spectroscopic parameters can be determined
+with stellar spectra and photosphere models.
+Global stellar structure, including radii, masses, and bolometric
+luminosities cannot be predicted theoretically without a physical
+model of stellar structure or stellar evolution.
+The \Gaia\ Mission has already delivered parallaxes---and therefore
+luminosity information---for millions of stars; \GALEX\ and
+\APASS\ deliver multi-band photometry; for a subset, \RAVEon\ delivers
+spectroscopic parameters.
+% Aims
+In this project, we infer spectroscopic parameters for stars for which
+only photometry and astrometry is available.
+We make no use of stellar structure or evolution models.
+% Methods
+After showing empirically that stars with similar luminosities and
+colors have similar spectroscopic parameters, we train a very flexible
+regression model of the relationship between luminosity, color, and
+the spectroscopic parameters.
+The model is a modification of a k-nearest-neighbors method.
+We use this model to paint spectroscopic parameters onto stars that have
+not been observed spectroscopically.
+% Results
+We find XXX and YYY. Things degrade in ZZZ way when WWW.
 

### 2016-12-29 paper/paper.tex
abstract tweak

@@
 As their name suggests, spectroscopic parameters can be determined
-with stellar spectra and photosphere models.
+with (medium to high resolution) spectral data and photosphere models.
 Global stellar structure, including radii, masses, and bolometric

### 2017-01-16 demo_plots.py
added demo_plots.py to the repository

@@ -1,0 +1,186 @@
+import numpy as np
+import matplotlib
+matplotlib.use('pdf')
+from matplotlib import pyplot as plt
+from astroML.plotting.tools import draw_ellipse
+from astroML.plotting import setup_text_plots
+from sklearn.mixture import GMM as skl_GMM
+
+def plot_bic(param_range,bics,lowest_comp):
+    plt.clf()
+    setup_text_plots(fontsize=16, usetex=True)
+    fig = plt.figure(figsize=(12, 6))
+    plt.bar(param_range-0.25,bics,color='blue',width=0.5)
+    plt.text(lowest_comp, bics.min() * 0.97 + .03 * bics.max(), '*',
+             fontsize=14, ha='center')
+
+    plt.xticks(param_range)
+    plt.ylim(bics.min() - 0.01 * (bics.max() - bics.min()),
+             bics.max() + 0.01 * (bics.max() - bics.min()))
+    plt.xlim(param_range.min() - 1, param_range.max() + 1)
+
+    plt.xticks(param_range,fontsize=14)
+    plt.yticks(fontsize=14)
+
+
+    plt.xlabel('Number of components',fontsize=18)
+    plt.ylabel('BIC score',fontsize=18)
+
+    plt.show()
+
+def plot_val_curve(param_range, train_mean, train_std, test_mean,
+                   test_std, log=False):
+    plt.clf()
+    setup_text_plots(fontsize=16, usetex=True)
+    fig=plt.figure(figsize=(12,8))
+    plt.plot(param_range, train_mean, label="Training score",
+             color="red")
+    plt.fill_between(param_range, train_mean - train_std,
+                     train_mean + train_std, alpha=0.2, color="red")
+    plt.plot(param_range, test_mean,label="Cross-validation score",
+             color="green")
+    plt.fill_between(param_range, test_mean - test_std,
+                     test_mean + test_std, alpha=0.2, color="green")
+
+    plt.legend(loc="best")
+    plt.xlabel("Number of Components", fontsize=18)
+    plt.ylabel("Score", fontsize=18)
+    plt.xlim(param_range.min(),param_range.max())
[…]

### 2017-01-17 demo_plots.py
added functionality for gaussians drawn in absolute magnitude space

@@ -5,6 +5,7 @@ from matplotlib import pyplot as plt
 from astroML.plotting.tools import draw_ellipse
 from astroML.plotting import setup_text_plots
 from sklearn.mixture import GMM as skl_GMM
+import drawEllipse
 
 def plot_bic(param_range,bics,lowest_comp):
     plt.clf()
@@ -50,7 +51,7 @@ def plot_val_curve(param_range, train_mean, train_std, test_mean,
     plt.savefig('val_curve.png')
 
 def fixAbsMag(x):
-    return 5.*np.log10(10.*x) 
+    return 5.*np.log10(10.*x)
 
 def plot_sample(x_true, y_true, x, y, samplex, sampley, xdgmm, xlabel='x', ylabel='y', xerr=None, yerr=None):
     setup_text_plots(fontsize=16, usetex=True)
@@ -76,8 +77,10 @@ def plot_sample(x_true, y_true, x, y, samplex, sampley, xdgmm, xlabel='x', ylabe
 
     ax4 = fig.add_subplot(224)
     for i in range(xdgmm.n_components):
-        draw_ellipse(xdgmm.mu[i], xdgmm.V[i], scales=[2], ax=ax4,
-                     ec='None', fc='gray', alpha=xdgmm.weights[i]/np.max(xdgmm.weights)*0.1)
+        points = drawEllipse(xdgmm.mu[i], xdgmmV[i])
+        ax4.plot(points[0, :], fixAbsMag(points[1,:]), alpha=xdgmm.weights[i]/np.max(xdgmm.weights), 'k-')
+        #draw_ellipse(xdgmm.mu[i], xdgmm.V[i], scales=[2], ax=ax4,
+        #         ec='None', fc='gray', alpha=xdgmm.weights[i]/np.max(xdgmm.weights)*0.1)
 
     titles = ["Observed Distribution", "Obs+Noise Distribution",
               "Extreme Deconvolution\n  resampling",
@@ -85,7 +88,7 @@ def plot_sample(x_true, y_true, x, y, samplex, sampley, xdgmm, xlabel='x', ylabe
 
     ax = [ax1, ax2, ax3, ax4]
 
-    for i in range(3):
+    for i in range(4):
         ax[i].set_xlim(-2, 3)
         ax[i].set_ylim(ylim[0], ylim[1]*1.1)
 
[…]

### 2017-01-17 stellarTwins.py

@@ -367,7 +367,6 @@ def observationsCutMatched(maxlogg=5., minlogg=4.2, mintemp=4500., SNthreshold=1
     matched = goodDistance & noNans & posErrors & lowPhotError & apassMatch
 
 
-
     tgasMatched = tgas[matched]
     magsMatched = tgasApass[matched]
     wiseMatched = tgasWise[matched]

### 2017-01-17 testXD.py

@@ -144,7 +144,7 @@ if __name__ == '__main__':
     for thresholdSN in [16, 8, 4, 2, 1]:
         fig, axes = plt.subplots(figsize=(7,7))
 
-        parallaxSNcut = tgasCutMatched['parallax']/tgasCutMatched['parallax_error'] < thresholdSN
+        parallaxSNcut = tgasCutMatched['parallax']/tgasCutMatched['parallax_error'] >= thresholdSN
         sigMax = 1.086/thresholdSN
         lowPhotErrorcut = (apassCutMatched['e_bmag'] < sigMax) & (apassCutMatched['e_vmag'] < sigMax) & (apassCutMatched['e_gmag'] < sigMax) & (apassCutMatched['e_rmag'] < sigMax) & (apassCutMatched['e_imag'] < sigMax)
 
@@ -176,8 +176,8 @@ if __name__ == '__main__':
             Vs = xdgmm.V
             amps = xdgmm.weights
             sample = xdgmm.sample(N)
-            dp.plot_sample(data1[j], fixAbsMag(data2[j]), data1[j], fixAbsMag(data2[j]), sample[:,0],fixAbsMag(sample[:,1]),
-                            xdgmm, xlabel=xlabel[j], ylabel='M$_\mathrm{G}$', xerr=err1[j], yerr=fixAbsMag(err2[j]))
+            dp.plot_sample(data1[j][indices], fixAbsMag(data2[j][indices]), data1[j][indices], fixAbsMag(data2[j][indices]),
+                   sample[:,0],fixAbsMag(sample[:,1]),xdgmm, xerr=err1[j][indices], yerr=fixAbsMag(err2[j][indices])) #xlabel=xlabel[j], ylabel=r'M$_\mathrm{G}$',
 
             os.rename('plot_sample.png', 'plot_sample_ngauss'+str(ngauss)+'.SN'+str(thresholdSN) + '.png')
 """

### 2017-01-18 demo_plots.py
no seeding 1028 gaussians

@@ -77,8 +77,8 @@ def plot_sample(x_true, y_true, x, y, samplex, sampley, xdgmm, xlabel='x', ylabe
 
     ax4 = fig.add_subplot(224)
     for i in range(xdgmm.n_components):
-        points = drawEllipse(xdgmm.mu[i], xdgmmV[i])
-        ax4.plot(points[0, :], fixAbsMag(points[1,:]), alpha=xdgmm.weights[i]/np.max(xdgmm.weights), 'k-')
+        points = drawEllipse.plotvector(xdgmm.mu[i], xdgmm.V[i])
+        ax4.plot(points[0, :], drawEllipse.fixAbsMag(points[1,:]), 'k-', alpha=xdgmm.weights[i]/np.max(xdgmm.weights))
         #draw_ellipse(xdgmm.mu[i], xdgmm.V[i], scales=[2], ax=ax4,
         #         ec='None', fc='gray', alpha=xdgmm.weights[i]/np.max(xdgmm.weights)*0.1)
 

### 2017-01-18 testXD.py

@@ -131,7 +131,7 @@ if __name__ == '__main__':
     err2 = [temp_err, absMagKinda_err]
     xlabel = ['B-V', 'B-V']
     ylabel = ['Teff [kK]', r'$\varpi 10^{0.2*m_G}$']
-    ngauss = 8
+    ngauss = 1028
     N = 120000
     optimize = False
     subset = False
@@ -141,7 +141,10 @@ if __name__ == '__main__':
     #[np.array([[0.5, 6.], [1., 4.]]), np.array([[0.5, 1.], [1., 2.]])]
 
 
+        
     for thresholdSN in [16, 8, 4, 2, 1]:
+    #for ngauss in [8, 128]:
+        #thresholdSN = 1
         fig, axes = plt.subplots(figsize=(7,7))
 
         parallaxSNcut = tgasCutMatched['parallax']/tgasCutMatched['parallax_error'] >= thresholdSN
@@ -149,7 +152,7 @@ if __name__ == '__main__':
         lowPhotErrorcut = (apassCutMatched['e_bmag'] < sigMax) & (apassCutMatched['e_vmag'] < sigMax) & (apassCutMatched['e_gmag'] < sigMax) & (apassCutMatched['e_rmag'] < sigMax) & (apassCutMatched['e_imag'] < sigMax)
 
         indices = parallaxSNcut & lowPhotErrorcut
-
+        #filename = 'xdgmm.1028gauss.1.2M.fit'
         filename = 'xdgmm.'+ str(ngauss) + 'gauss.'+nstar+ '.SN' + str(thresholdSN) + '.fit'
         for j, ax in zip([1],[axes]):
             try:
@@ -177,9 +180,9 @@ if __name__ == '__main__':
             amps = xdgmm.weights
             sample = xdgmm.sample(N)
             dp.plot_sample(data1[j][indices], fixAbsMag(data2[j][indices]), data1[j][indices], fixAbsMag(data2[j][indices]),
-                   sample[:,0],fixAbsMag(sample[:,1]),xdgmm, xerr=err1[j][indices], yerr=fixAbsMag(err2[j][indices])) #xlabel=xlabel[j], ylabel=r'M$_\mathrm{G}$',
+                   sample[:,0],fixAbsMag(sample[:,1]),xdgmm, xerr=err1[j][indices], yerr=fixAbsMag(err2[j][indices]), xlabel=xlabel[j], ylabel=r'M$_\mathrm{G}$')
 
[…]

### 2017-01-23 testXD.py
plotting posteriors

@@ -14,6 +14,7 @@ from xdgmm import XDGMM
 from sklearn.learning_curve import validation_curve
 from sklearn.cross_validation import ShuffleSplit
 import demo_plots as dp
+import drawEllipse
 
 def fixAbsMag(x):
     return 5.*np.log10(10.*x)
@@ -76,6 +77,18 @@ def optimize(param_range=np.array([256, 512, 1024, 2048, 4096, 8182])):
     dp.plot_val_curve(param_range, train_scores_mean, train_scores_std, test_scores_mean, test_scores_std)
     return train_scores, test_scores
 
+def multiplyGaussians(mean1, cov1, mean2, cov2):
+    a = mean1
+    b = mean2
+    A = cov1
+    B = cov2
+    d = len(a)
+    C = np.linalg.inv(np.linalg.inv(A) + np.linalg.inv(B))
+    c = C*np.linalg.inv(A)*a + C*np.linalg.inv(B)*b
+    exponent = -0.5*(np.transpose(a)*np.linalg.inv(A)*a + np.transpose(b)*np.linalg.inv(B)*b-np.transpose(c)*np.linalg.inv(C)*c)
+    z_c = (2*np.pi)**(-d/2.)*np.linalg.det(C)**0.5*np.linalg.det(A)**-0.5*np.linalg.det(B)**-0.5*np.exp(exponent)
+    return c, C, z_c
+
 if __name__ == '__main__':
     b_v_lim = [0.25, 1.5]
     g_r_lim = None #[0, 1.5]
@@ -113,10 +126,23 @@ if __name__ == '__main__':
     i_RedCoeff = 1.698
     bayesDust = st.dust(tgasCutMatched['l']*units.deg, tgasCutMatched['b']*units.deg, np.median(distCutMatched, axis=1)*units.pc)
     #M_V = apassCutMatched['vmag'] - V_RedCoeff*bayesDust - meanMuMatched
-    g_r = apassCutMatched['gmag'] - g_RedCoeff*bayesDust - (apassCutMatched['rmag'] - r_RedCoeff*bayesDust)
-    r_i = apassCutMatched['rmag'] - r_RedCoeff*bayesDust - (apassCutMatched['imag'] - i_RedCoeff*bayesDust)
-    B_V = apassCutMatched['bmag'] - B_RedCoeff*bayesDust - (apassCutMatched['vmag'] - V_RedCoeff*bayesDust)
[…]

### 2017-01-23 demo_plots.py
removed dust

@@ -89,7 +89,7 @@ def plot_sample(x_true, y_true, x, y, samplex, sampley, xdgmm, xlabel='x', ylabe
     ax = [ax1, ax2, ax3, ax4]
 
     for i in range(4):
-        ax[i].set_xlim(-2, 3)
+        ax[i].set_xlim(-1, 3)
         ax[i].set_ylim(ylim[0], ylim[1]*1.1)
 
         #ax[i].xaxis.set_major_locator(plt.MultipleLocator([-1, 0, 1]))

### 2017-01-23 testXD.py

@@ -118,6 +118,12 @@ if __name__ == '__main__':
     B_V = apassCutMatched['bmag'] - B_RedCoeff*bayesDust - (apassCutMatched['vmag'] - V_RedCoeff*bayesDust)
     B_V_err = np.sqrt(apassCutMatched['e_bmag']**2. + apassCutMatched['e_vmag']**2.)
 
+    g_r = apassCutMatched['gmag'] - apassCutMatched['rmag'] 
+    r_i = apassCutMatched['rmag'] - apassCutMatched['imag'] 
+    B_V = apassCutMatched['bmag'] - apassCutMatched['vmag'] 
+    B_V_err = np.sqrt(apassCutMatched['e_bmag']**2. + apassCutMatched['e_vmag']**2.)
+
+
     temp = raveCutMatched['TEFF']/1000.
     temp_err = raveCutMatched['E_TEFF']/1000.
 
@@ -142,7 +148,7 @@ if __name__ == '__main__':
 
 
         
-    for thresholdSN in [16, 8, 4, 2, 1]:
+    for thresholdSN in [1]: #[16, 8, 4, 2, 1]:
     #for ngauss in [8, 128]:
         #thresholdSN = 1
         fig, axes = plt.subplots(figsize=(7,7))

### 2017-01-23 paperDistance/paperApJ.tex

%% This is emulateapj reformatting of the AASTEX sample document
%%
\documentclass[iop]{emulateapj}

\newcommand{\vdag}{(v)^\dagger}
\newcommand{\myemail}{landerson@simonsfoundation.org}

%% You can insert a short comment on the title page using the command below.

\slugcomment{Stars are kinda cool again.}

%% If you wish, you may supply running head information, although
%% this information may be modified by the editorial offices.
%% The left head contains a list of authors,
%% usually a maximum of three (otherwise use et al.).  The right
%% head is a modified title of up to roughly 44 characters.
%% Running heads will not print in the manuscript style.

\shorttitle{Measuring Mo' Betta Distances}
\shortauthors{Anderson et al}

%% This is the end of the preamble.  Indicate the beginning of the
%% paper itself with \begin{document}.

\begin{document}

%% LaTeX will automatically break titles if they run longer than
%% one line. However, you may use \\ to force a line break if
%% you desire.

\title{Mo' Betta Gaia Distance Measurements, \\
    Using Gaia + APASS photometry}

%% Use \author, \affil, and the \and command to format
%% author and affiliation information.
%% Note that \email has replaced the old \authoremail command
%% from AASTeX v4.0. You can use \email to mark an email address
%% anywhere in the paper, not just in the front matter.
%% As in the title, use \\ to force line breaks.

\author{Lauren Anderson and David Hogg\altaffilmark{1}}
\affil{Center for Computational Astrophysics}

\author{Boris Leistedt}
\affil{NYU}

\author{Adrian Price-Whelan}
\affil{Princeton}

\and

\author{Andrew Casey}
\affil{Somewhere in the UK}

\altaffiltext{1}{NYU}

\begin{abstract}

\end{abstract}

\keywords{stars}

\section{Introduction}

Given the likelihood measurements of the parallax to local stars from Gaia, and using prior information for the distances to these stars built from the data themselves, we make more precise measurements of the distances to local stars in Gaia than the parallax measurements alone. The prior, which is very similar to a color-magnitude diagram of stars, is informative because stars fall on a locus within this space. By deconvolving the set of observed points with their uncertainties we tighten this locus which therefore informs the individual measurements that they are drawn from a tighter distribution. We build this prior using Extreme Deconvolution. Extreme Deconvolution is a gaussian mixture model of data, specifically data that have heteroscedastic uncertainties, a quality that gaussin mixture models alone cannot take into account. This prior does not include any assumptions about the physical distribution of stars within the galaxy, nor does it include any stellar models.

\section{Data}

We use stars cross matched in the TGAS and APASS set of observations. We require there are no NaNs in any APASS photometric band, as well as every band having positive errors. We check that the match between APASS and TGAS is good using extreme APASS-WISE colors. We correct for dust using 3D Bayestar (Schlafly+Finkbeiner) with an estimate of the distance from sampling of posterior Gaia distances from Adrain. There is minimal dust to the matched stars. To build our model here we use the Gaia G band magnitude, parallax and APASS B-V color.

\section{Methods}

We build a prior for the distance measurement of Gaia stars using the Gaia measured parallaxes plus APASS photometry to build a color-magnitude-like diagram that serves as the prior for distance measurements of Gaia stars. Instead of making assumptions about the physical distributions of stars in the galaxy, or using stellar models for the modeled HR diagram of stars, we build a prior from the data themselves.

magSN = parallaxSN = 16 (excluding systematic parallax error in tgas)
train gaussian means, mus, Vs on progressively lower SN dataset

\begin{equation}
\label{eq:bayes}
P(\theta|\textbf{E}, \textbf{I}) = P(\theta ) \frac{P(\textbf{E} |\theta,\textbf{I})}{P(\textbf{E})},
\end{equation}

where $\theta$ is the model parameters that generate the evidence $\textbf{E}$, and P($\theta$) is the prior knowledge of the model, and $P(\textbf{E} |\theta,\textbf{I})$ is the likelihood of the evidence, given the model and some other prior assumptions. Our evidence is the parallaxes and G band magnitudes from Gaia, and the B and V band magnitudes from APASS. We build the prior from this evidence using extreme deconvolution

Crossvalidation

\subsection{Formalism} \label{bozomath}

\section{Results}

\section{Discussion}
Apass HR Diagram looks very different from Rave HR Diagram, MS stars are incorrectly being labeled as giants

\acknowledgments

\appendix

\section{Appendix material}

\begin{thebibliography}{}
\end{thebibliography}

\clearpage

\end{document}

%%
%% End of file `sample.tex'.

### 2017-01-24 testXD.py
trying to plot posterior

@@ -169,13 +169,7 @@ if __name__ == '__main__':
     #[np.array([[0.5, 6.], [1., 4.]]), np.array([[0.5, 1.], [1., 2.]])]
 
 
-<<<<<<< HEAD
-        
     for thresholdSN in [1]: #[16, 8, 4, 2, 1]:
-=======
-
-    for thresholdSN in [1]:
->>>>>>> da75fc4efe5b66654e4ab6282d3ef578d4abe5db
     #for ngauss in [8, 128]:
         #thresholdSN = 1
         fig, axes = plt.subplots(figsize=(7,7))
@@ -217,15 +211,17 @@ if __name__ == '__main__':
 
             os.rename('plot_sample.png', 'plot_sample_ngauss'+str(ngauss)+'.SN'+str(thresholdSN) + '.noSEED.png')
             index = 0
+            dimension = 0
             mean2, cov2 = matrixize(data1[j][index], data2[j][index], err1[j][index], err2[j][index])
             figtest, axtest = plt.subplots()
-            points = drawEllipse.plotvector(mean2, cov2)
+            points = drawEllipse.plotvector(mean2[dimension], cov2[dimension])
             axtest.plot(points[0, :], drawEllipse.fixAbsMag(points[1,:]), 'b-', alpha=1.0)
-            for gg in arange(xdgmm.n_components):
-                newMean, newCov, newAmp = multiplyGaussians(xdgmm.mu[gg], xdgmm.V[gg], mean2, cov2)
-                points = drawEllipse.plotvector(newMean, newCov)
+            for gg in range(xdgmm.n_components):
+                newMean, newCov, newAmp = multiplyGaussians(xdgmm.mu[gg], xdgmm.V[gg], mean2[dimension], cov2[dimension])
+                print newMean, newCov, newAmp
+                points = drawEllipse.plotvector(newMean[0], newCov)
                 axtest.plot(points[0, :], drawEllipse.fixAbsMag(points[1,:]), 'k-', alpha=newAmp/np.max(xdgmm.weights))
-
+            print mean2, cov2, xdgmm.mu, xdgmm.V
             figtest.savefig('posterior.png')
 """
         mean_guess = np.random.rand(ngauss,2)*10.

### 2017-02-03 testXD.py
generalized bands

@@ -188,12 +188,13 @@ if __name__ == '__main__':
     projectedDimension = 1
 
     dataFilename = 'cutMatchedArrays.tgasApassSN0.npz'
-    xdgmmFilename = 'xdgmm.'+ str(ngauss) + 'gauss.'+nstar+ '.SN' + str(thresholdSN) + '.fit'
+    xdgmmFilename = 'xdgmm.'+ str(ngauss) + 'gauss.'+nstar+ '.SN' + str(thresholdSN) + '.2MASS.fit'
 
     useDust = False
     optimize = False
     subset = False
     timing = False
+
 
     try:
         cutMatchedArrays  = np.load(dataFilename)
@@ -213,42 +214,56 @@ if __name__ == '__main__':
 
     indicesM67 = m67indices(tgasCutMatched, plot=False, db=0.5, dl=0.5)
 
+    if survey == 'APASS':
+        mag1 = 'B'
+        mag2 = 'V'
+        absmag = 'G'
+        xlabel='B-V'
+        ylabel = r'M$_\mathrm{G}$'
+
+    if survey == '2MASS':
+        mag1 = 'J'
+        mag2 = 'K'
+        absmag = 'J'
+        xlabel = 'J-K$_s$'
+        ylabel = r'M$_\mathrm{J}$'
+
+    bandDictionary = {'B':{key:'bmag', errKey:'e_bmag', array:'apassCutMatched'},
+                      'V':{key:'vmag', errKey:'e_vmag', array:'apassCutMatched'},
+                      'J':{key:'j_mag', errKey:'j_cmsig', array:'twoMassCutMatched'}
+                      'K':{key:'k_mag', errKey:'k_cmsig', array:'twoMassCutMatched'}
+                      'G':{key:'phot_g_mean_mag', array:'tgasCutMatched'}}
+
+
     if useDust:
         bayesDust = st.dust(tgasCutMatched['l']*units.deg, tgasCutMatched['b']*units.deg, np.median(distCutMatched, axis=1)*units.pc)
-        B_dustcorrected = dustCorrection(apassCutMatched['bmag'], bayesDust, 'B')
-        V_dustcorrected = dustCorrection(apassCutMatched['vmag'], bayesDust, 'V')
-        #g_dustcorrected = dustCorrection(apassCutMatched['gmag'], bayesDust, 'g')
-        #r_dustcorrected = dustCorrection(apassCutMatched['rmag'], bayesDust, 'r')
[…]

### 2017-02-03 testXD.py
some updates

@@ -181,7 +181,7 @@ if __name__ == '__main__':
 
     np.random.seed(2)
     thresholdSN = 0.001
-    ngauss = 128
+    ngauss = 1024
     nstar = '1.2M'
     Nsamples = 120000
     nPosteriorPoints = 10000
@@ -241,6 +241,11 @@ if __name__ == '__main__':
     xlabel = 'B-V'
     ylabel = r'M$_\mathrm{G}$'
 
+    ylabel_posterior_logd = r'P(log d|$\varpi, \sigma_{\varpi}$)'
+    ylabel_posterior_d = r'P(d|$\varpi, \sigma_{\varpi}$)'
+    ylabel_posterior_parallax = r'P($\varpi_\mathrm{true}|\varpi, \sigma_{\varpi}$)'
+    ylabel_likelihood_d = r'P($\varpi$|d, $\sigma_{\varpi}$)'
+    ylabel_likelihood_logd = r'P($\varpi$|log d, $\sigma_{\varpi}$)'
     summedPosterior = np.zeros((np.sum(indicesM67), nPosteriorPoints))
 
     parallaxSNcut = tgasCutMatched['parallax']/tgasCutMatched['parallax_error'] >= thresholdSN
@@ -249,9 +254,10 @@ if __name__ == '__main__':
     indices = parallaxSNcut & lowPhotErrorcut
 
 
-    figDist, axDist = plt.subplots(2, 2, figsize=(15, 15))
+    figDist, axDist = plt.subplots(2, 3, figsize=(25, 15))
     axDist = axDist.flatten()
 
+    figDistLin, axDistLin = plt.subplots(2, figsize=(7.5, 10), sharex=True)
 
     try:
         xdgmm = XDGMM(filename=xdgmmFilename)
@@ -264,7 +270,7 @@ if __name__ == '__main__':
         xdgmm = XDGMM(method='Bovy')
         xdgmm.n_components = ngauss
         xdgmm = xdgmm.fit(X, Xerr)
-        xdgmm.save_model(filename)
+        xdgmm.save_model(xdgmmFilename)
         sample = xdgmm.sample(Nsamples)
         figPrior, axPrior = plt.subplots()
         for gg in range(xdgmm.n_components):
@@ -295,7 +301,6 @@ if __name__ == '__main__':
 
         dimension = 0
         mean2, cov2 = matrixize(data1[index], data2[index], err1[index], err2[index])
[…]

### 2017-02-03 demo_plots.py
checking 2Mass

@@ -50,8 +50,14 @@ def plot_val_curve(param_range, train_mean, train_std, test_mean,
     if log: plt.xscale('log', basex=2)
     plt.savefig('val_curve.png')
 
-def fixAbsMag(x):
-    return 5.*np.log10(10.*x)
+
+def absMagKinda2absMag(absMagKinda):
+    """                                                                                                                                                                                                                                     
+    convert my funny units of parallax[mas]*10**(0.2*apparent magnitude[mag]) to an absolute magnitude [mag]                                                                                                                                
+    """
+    absMagKinda_in_arcseconds = absMagKinda/1e3 #first convert parallax from mas ==> arcseconds                                                                                                                                             
+    return 5.*np.log10(10.*absMagKinda_in_arcseconds)
+
 
 def plot_sample(x_true, y_true, x, y, samplex, sampley, xdgmm, xlabel='x', ylabel='y', xerr=None, yerr=None):
     setup_text_plots(fontsize=16, usetex=True)
@@ -78,7 +84,7 @@ def plot_sample(x_true, y_true, x, y, samplex, sampley, xdgmm, xlabel='x', ylabe
     ax4 = fig.add_subplot(224)
     for i in range(xdgmm.n_components):
         points = drawEllipse.plotvector(xdgmm.mu[i], xdgmm.V[i])
-        ax4.plot(points[0, :], drawEllipse.fixAbsMag(points[1,:]), 'k-', alpha=xdgmm.weights[i]/np.max(xdgmm.weights))
+        ax4.plot(points[0, :], drawEllipse.absMagKinda2absMag(points[1,:]), 'k-', alpha=xdgmm.weights[i]/np.max(xdgmm.weights))
         #draw_ellipse(xdgmm.mu[i], xdgmm.V[i], scales=[2], ax=ax4,
         #         ec='None', fc='gray', alpha=xdgmm.weights[i]/np.max(xdgmm.weights)*0.1)
 

### 2017-02-03 testXD.py

@@ -179,9 +179,10 @@ def dustCorrection(magnitude, EBV, band):
 
 if __name__ == '__main__':
 
+    survey = '2MASS'
     np.random.seed(2)
     thresholdSN = 0.001
-    ngauss = 1024
+    ngauss = 128
     nstar = '1.2M'
     Nsamples = 120000
     nPosteriorPoints = 10000
@@ -200,10 +201,10 @@ if __name__ == '__main__':
         cutMatchedArrays  = np.load(dataFilename)
         tgasCutMatched    = cutMatchedArrays['tgasCutMatched']
         apassCutMatched   = cutMatchedArrays['apassCutMatched']
-        raveCutMatched    = cutMatchedArrays['raveCutMatched']
+        #raveCutMatched    = cutMatchedArrays['raveCutMatched']
         twoMassCutMatched = cutMatchedArrays['twoMassCutMatched']
-        wiseCutMatched    = cutMatchedArrays['wiseCutMatched']
-        distCutMatched    = cutMatchedArrays['distCutMatched']
+        #wiseCutMatched    = cutMatchedArrays['wiseCutMatched']
+        #distCutMatched    = cutMatchedArrays['distCutMatched']
     except IOError:
         maxlogg = 20
         minlogg = 1
@@ -213,6 +214,7 @@ if __name__ == '__main__':
 
 
     indicesM67 = m67indices(tgasCutMatched, plot=False, db=0.5, dl=0.5)
+
 
 
     if survey == 'APASS':
@@ -229,27 +231,27 @@ if __name__ == '__main__':
         xlabel = 'J-K$_s$'
         ylabel = r'M$_\mathrm{J}$'
 
-    bandDictionary = {'B':{key:'bmag', errKey:'e_bmag', array:'apassCutMatched'},
-                      'V':{key:'vmag', errKey:'e_vmag', array:'apassCutMatched'},
-                      'J':{key:'j_mag', errKey:'j_cmsig', array:'twoMassCutMatched'}
-                      'K':{key:'k_mag', errKey:'k_cmsig', array:'twoMassCutMatched'}
-                      'G':{key:'phot_g_mean_mag', array:'tgasCutMatched'}}
+    bandDictionary = {'B':{'key':'bmag', 'err_key':'e_bmag', 'array':apassCutMatched},
+                      'V':{'key':'vmag', 'err_key':'e_vmag', 'array':apassCutMatched},
+                      'J':{'key':'j_mag', 'err_key':'j_cmsig', 'array':twoMassCutMatched},
[…]

### 2017-02-03 generateMatchedArrays.py
generate arrays for just matched to 2MASS

@@ -1,0 +1,5 @@
+import stellarTwins as st
+
+thresholdSN = 0.001
+filename = 'cutMatchedArrays.SN' + str(thresholdSN) +'.npz'
+observationsCutMatched(SNthreshold=1., filename=filename)

### 2017-02-03 stellarTwins.py

@@ -322,7 +322,7 @@ def tgasDistance(ndist=1024):
                 dist = np.concatenate((dist, f['distance'][:,:ndist]))
     return dist
 
-def observationsCutMatched(maxlogg=5., minlogg=4.2, mintemp=4500., SNthreshold=16., filename='cutMatchedArrays.npz'):
+def observationsCutMatched(SNthreshold=1., filename='cutMatchedArrays.npz'):
 
     #read in TGAS data for and matched sample for magnitudes
     tgas = fits.getdata("stacked_tgas.fits", 1)
@@ -365,7 +365,7 @@ def observationsCutMatched(maxlogg=5., minlogg=4.2, mintemp=4500., SNthreshold=1
     hasDust = medianDist >= minDist
     raveMatch = np.in1d(tgas['source_id'], tgasRave['source_id'])
     if galacticLatMax: inDisk = np.abs(tgas['b'] <= galacticLatMax)
-    matched = goodDistance & noNans & posErrors & lowPhotError & apassMatch
+    matched = goodDistance & noNans & posErrors & lowPhotError #& apassMatch
 
 
     tgasMatched = tgas[matched]

### 2017-02-03 testXD.py

@@ -206,10 +206,7 @@ if __name__ == '__main__':
         #wiseCutMatched    = cutMatchedArrays['wiseCutMatched']
         #distCutMatched    = cutMatchedArrays['distCutMatched']
     except IOError:
-        maxlogg = 20
-        minlogg = 1
-        mintemp = 100
-        tgasCutMatched, apassCutMatched, raveCutMatched, twoMassCutMatched, wiseCutMatched, distCutMatched = st.observationsCutMatched(maxlogg=maxlogg, minlogg=minlogg, mintemp=mintemp, SNthreshold=thresholdSN, filename=dataFilename)
+        tgasCutMatched, apassCutMatched, raveCutMatched, twoMassCutMatched, wiseCutMatched, distCutMatched = st.observationsCutMatched(SNthreshold=thresholdSN, filename=dataFilename)
     print 'Number of Matched stars is: ', len(tgasCutMatched)
 
 
@@ -240,11 +237,11 @@ if __name__ == '__main__':
 
     if useDust:
         bayesDust = st.dust(tgasCutMatched['l']*units.deg, tgasCutMatched['b']*units.deg, np.median(distCutMatched, axis=1)*units.pc)
-        mag1DustCorrected   = dustCorrection(bandDictionary[mag1]['array']  [bandDictionary[mag1]['key']], bayesDust, mag1) 
+        mag1DustCorrected   = dustCorrection(bandDictionary[mag1]['array']  [bandDictionary[mag1]['key']], bayesDust, mag1)
         mag2DustCorrected   = dustCorrection(bandDictionary[mag2]['array']  [bandDictionary[mag2]['key']], bayesDust, mag2)
         absMagDustCorrected = dustCorrection(bandDictionary[absmag]['array'][bandDictionary[absmag]['key']], bayesDust, absmag)
         #B_dustcorrected = dustCorrection(apassCutMatched['bmag'], bayesDust, 'B')
-        #need to define color_err and absMagKinda_err when including dust correction 
+        #need to define color_err and absMagKinda_err when including dust correction
         color = mag1DustCorrected - mag2DustCorrected
     else:
         color = bandDictionary[mag1]['array'][bandDictionary[mag1]['key']] - \
@@ -252,7 +249,7 @@ if __name__ == '__main__':
[…]

### 2017-02-06 testXD.py
making 2mass work

@@ -220,6 +220,8 @@ if __name__ == '__main__':
         absmag = 'G'
         xlabel='B-V'
         ylabel = r'M$_\mathrm{G}$'
+        xlim = [-0.2, 2]
+        ylim = [9, -2]
 
     if survey == '2MASS':
         mag1 = 'J'
@@ -227,6 +229,8 @@ if __name__ == '__main__':
         absmag = 'J'
         xlabel = 'J-K$_s$'
         ylabel = r'M$_\mathrm{J}$'
+        xlim = [-0.25, 1.25]
+        ylim = [6, -4]
 
     bandDictionary = {'B':{'key':'bmag', 'err_key':'e_bmag', 'array':apassCutMatched},
                       'V':{'key':'vmag', 'err_key':'e_vmag', 'array':apassCutMatched},
@@ -246,20 +250,20 @@ if __name__ == '__main__':
     else:
         color = bandDictionary[mag1]['array'][bandDictionary[mag1]['key']] - \
                 bandDictionary[mag2]['array'][bandDictionary[mag2]['key']]
-        color_err = np.sqrt(bandDictionary[mag1]['array'][bandDictionary[mag1]['err_key']]**2. - bandDictionary[mag2]['array'][bandDictionary[mag2]['err_key']]**2.)
+        color_err = np.sqrt(bandDictionary[mag1]['array'][bandDictionary[mag1]['err_key']]**2. + bandDictionary[mag2]['array'][bandDictionary[mag2]['err_key']]**2.)
         absMagKinda = tgasCutMatched['parallax']*10.**(0.2*bandDictionary[absmag]['array'][bandDictionary[absmag]['key']])
-        absMagKinda_err = tgasCutMatched['parallax']*10.**(0.2*bandDictionary[absmag]['array'][bandDictionary[absmag]['err_key']])
+        absMagKinda_err = tgasCutMatched['parallax_error']*10.**(0.2*bandDictionary[absmag]['array'][bandDictionary[absmag]['key']])
 
     data1 = color
     data2 = absMagKinda
     err1 = color_err
     err2 = absMagKinda_err
 
-    ylabel_posterior_logd = r'P(log d|$\varpi, \sigma_{\varpi}$)'
-    ylabel_posterior_d = r'P(d|$\varpi, \sigma_{\varpi}$)'
-    ylabel_posterior_parallax = r'P($\varpi_\mathrm{true}|\varpi, \sigma_{\varpi}$)'
[…]

### 2017-02-06 generateMatchedArrays.py
taking away apass match necessity

@@ -2,4 +2,4 @@ import stellarTwins as st
 
 thresholdSN = 0.001
 filename = 'cutMatchedArrays.SN' + str(thresholdSN) +'.npz'
-observationsCutMatched(SNthreshold=1., filename=filename)
+st.observationsCutMatched(SNthreshold=1., filename=filename)

### 2017-02-06 stellarTwins.py

@@ -334,8 +334,8 @@ def observationsCutMatched(SNthreshold=1., filename='cutMatchedArrays.npz'):
     medianDist = np.median(distances, axis=1)
     #cut out low logg and temperatures outside well populated area
     nonNans = ~np.isnan(tgasRave['TEFF']) & ~np.isnan(tgasRave['LOGG']) & ~np.isnan(tgasRave['FE_H'])
-    dwarfs = (tgasRave['LOGG'] < maxlogg) & (tgasRave['LOGG'] > minlogg) & (tgasRave['TEFF'] > mintemp)
-    tgasRave = tgasRave[nonNans & dwarfs]
+    #dwarfs = (tgasRave['LOGG'] < maxlogg) & (tgasRave['LOGG'] > minlogg) & (tgasRave['TEFF'] > mintemp)
+    tgasRave = tgasRave[nonNans] # & dwarfs]
 
     #various cuts to select sample
     magSN = SNthreshold

### 2017-02-07 demo_plots.py
demo plot update

@@ -52,10 +52,10 @@ def plot_val_curve(param_range, train_mean, train_std, test_mean,
 
 
 def absMagKinda2absMag(absMagKinda):
-    """                                                                                                                                                                                                                                     
-    convert my funny units of parallax[mas]*10**(0.2*apparent magnitude[mag]) to an absolute magnitude [mag]                                                                                                                                
     """
-    absMagKinda_in_arcseconds = absMagKinda/1e3 #first convert parallax from mas ==> arcseconds                                                                                                                                             
+    convert my funny units of parallax[mas]*10**(0.2*apparent magnitude[mag]) to an absolute magnitude [mag]
+    """
+    absMagKinda_in_arcseconds = absMagKinda/1e3 #first convert parallax from mas ==> arcseconds
     return 5.*np.log10(10.*absMagKinda_in_arcseconds)
 
 
@@ -84,7 +84,7 @@ def plot_sample(x_true, y_true, x, y, samplex, sampley, xdgmm, xlabel='x', ylabe
     ax4 = fig.add_subplot(224)
     for i in range(xdgmm.n_components):
         points = drawEllipse.plotvector(xdgmm.mu[i], xdgmm.V[i])
-        ax4.plot(points[0, :], drawEllipse.absMagKinda2absMag(points[1,:]), 'k-', alpha=xdgmm.weights[i]/np.max(xdgmm.weights))
+        ax4.plot(points[0, :], absMagKinda2absMag(points[1,:]), 'k-', alpha=xdgmm.weights[i]/np.max(xdgmm.weights))
         #draw_ellipse(xdgmm.mu[i], xdgmm.V[i], scales=[2], ax=ax4,
         #         ec='None', fc='gray', alpha=xdgmm.weights[i]/np.max(xdgmm.weights)*0.1)
 
@@ -95,7 +95,7 @@ def plot_sample(x_true, y_true, x, y, samplex, sampley, xdgmm, xlabel='x', ylabe
     ax = [ax1, ax2, ax3, ax4]
 
     for i in range(4):
-        ax[i].set_xlim(-1, 3)
+        ax[i].set_xlim(-0.5, 1.5)
         ax[i].set_ylim(ylim[0], ylim[1]*1.1)
 
         #ax[i].xaxis.set_major_locator(plt.MultipleLocator([-1, 0, 1]))

### 2017-02-07 paperDistance/paperApJ.tex
added abstract

@@
 \begin{abstract}
+The Gaia TGAS Catalog contains more than 2 millions parallax measurements.
+Safe conversion of a noisy parallax into a posterior belief over distances requires inference with a prior.
+Usually this prior represents some kind of belief about the Milky Way.
+However, there is multi-band photometry for the TGAS stars from imaging surveys.
+This imaging is incredibly informative about stellar distances, as is clear from physical models of stars, and also just from looking at the data themselves.
+Here we use color information on TGAS stars from [2MASS or APASS] to build a noise-deconvolved empirical prior distribution of stars in color-magnitude space.
+This data-driven model contains no knowledge of the physics of stellar interiors or photospheres, nor of the Milky Way, but rather derives its precision from its generative model of noisy parallax measurements and an assumption of stationarity.
+We use the Extreme Deconvolution (XD) algorithm, which is an Empirical Bayes approximation to a full hierarchical model of the true parallax and photometry of every star.
+The algorithm is run not in absolute-magnitude space but in a transformed space where the measurement uncertainty is closer to Gaussian.
+The XD optimized prior is used to perform parallax and distance inferences for every star, yielding a precise stellar distance estimate and uncertainty (and full posterior) for every star.
+The method is capitalizing on statistical shrinkage.
+Our posterior parallax estimates are more precise than the Gaia catalog outputs by a factor of [XXX] for the median TGAS star; and more precise than previous examples of Bayesian distance estimates by a factor of [YYY].
+We independently validate our distances by looking at members of Milky Way star clusters; for example, M67 is not visible at all in the TGAS parallax estimates, but appears clearly in our posterior parallax estimates.
+All our results, including a posterior parallax and distance sampling for [ZZZ] TGAS stars, are available in companion electronic tables.
 
@@
 
+Long discussion of dust
+
+Number of gaussians in XD, and the w parameter
+
 \acknowledgments

### 2017-02-08 paperDistance/paperApJ.tex
makefile, etc

--- 2017-01-23 paperDistance/paperApJ.tex
@@
 \documentclass[iop]{emulateapj}
-
-\newcommand{\vdag}{(v)^\dagger}
 \newcommand{\myemail}{landerson@simonsfoundation.org}
 
-%% You can insert a short comment on the title page using the command below.
-
 \slugcomment{Stars are kinda cool again.}
-
-%% If you wish, you may supply running head information, although
-%% this information may be modified by the editorial offices.
-%% The left head contains a list of authors,
-%% usually a maximum of three (otherwise use et al.).  The right
-%% head is a modified title of up to roughly 44 characters.
-%% Running heads will not print in the manuscript style.
 
@@
 
-%% This is the end of the preamble.  Indicate the beginning of the
-%% paper itself with \begin{document}.
+\newcommand{\acronym}[1]{{\small{#1}}}
+\newcommand{\project}[1]{\textsl{#1}}
+\newcommand{\tgas}{\project{\acronym{TGAS}}}
+\newcommand{\tmass}{\project{\acronym{2MASS}}}
+\newcommand{\gaia}{\project{Gaia}}
 
 \begin{document}
-
-%% LaTeX will automatically break titles if they run longer than
-%% one line. However, you may use \\ to force a line break if
-%% you desire.
 
@@
 
-%% Use \author, \affil, and the \and command to format
-%% author and affiliation information.
-%% Note that \email has replaced the old \authoremail command
-%% from AASTeX v4.0. You can use \email to mark an email address
-%% anywhere in the paper, not just in the front matter.
-%% As in the title, use \\ to force line breaks.
-
-\author{Lauren Anderson and David Hogg\altaffilmark{1}}
-\affil{Center for Computational Astrophysics}
-
-\author{Boris Leistedt}
-\affil{NYU}
-
-\author{Adrian Price-Whelan}
-\affil{Princeton}
-
-\and
-
-\author{Andrew Casey}
-\affil{Somewhere in the UK}
-
-\altaffiltext{1}{NYU}
+\author{Lauren Anderson\altaffilmark{1},
+        David W. Hogg\altaffilmark{1, 2, 3, 4},
+        Boris Leistedt\altaffilmark{2},
+        Adrian Price-Whelan\altaffilmark{5}}
+\altaffiltext[1]{Center for Computational Astrophysics, Flatiron Institute}
+\altaffiltext[2]{Center for Cosmology and Particle Physics, Department of Physics, New York University}
+\altaffiltext[3]{Center for Data Science, New York University}
+\altaffiltext[4]{Max-Planck-Institut f\"ur Astronomie}
+\altaffiltext[5]{Princeton University Observatory, Princeton University}
 
 \begin{abstract}
-
+The \gaia\ \tgas\ \project{Catalog} contains more than 2 million parallax measurements.
+Conversion of a noisy parallax measurement into a posterior belief over distances requires inference with a prior.
+Usually this prior represents some kind of belief about the Milky Way.
+However, there is multi-band photometry for the \tgas\ stars from imaging surveys;
+this imaging is incredibly informative about stellar distances.
+Here we use color information on TGAS stars from \tmass\ to build a noise-deconvolved empirical prior distribution of stars in color-magnitude space.
+This data-driven model contains no knowledge of the physics of stellar interiors or photospheres, nor of the Milky Way, but rather derives its precision from its generative model of noisy parallax measurements and an assumption of stationarity.
+We use the Extreme Deconvolution (XD) algorithm, which is an Empirical Bayes approximation to a full hierarchical model of the true parallax and photometry of every star.
+The algorithm is run not in absolute-magnitude space but in a transformed space where the measurement uncertainty is closer to Gaussian.
+The XD optimized prior is used to perform parallax and distance inferences for every star, yielding a precise stellar distance estimate and uncertainty (and full posterior) for every star.
+The method is capitalizing on statistical shrinkage.
+Our posterior parallax estimates are more precise than the Gaia catalog outputs by a factor of [XXX] for the median \tgas\ star; and more precise than previous examples of Bayesian distance estimates by a factor of [YYY].
+We independently validate our distances by looking at members of Milky Way star clusters; for example, M67 is not visible at all in the \tgas\ parallax estimates, but appears clearly in our posterior parallax estimates.
+All our results, including a posterior parallax and distance sampling for [ZZZ] \tgas\ stars, are available in companion electronic tables.
 \end{abstract}
@@
 
-\acknowledgments
+Long discussion of dust
+
+Number of gaussians in XD, and the w parameter
+
+\acknowledgments It is a pleasure to thank Andy Casey (Monash) and the
+attendees at the Stars Group Meeting at the CCA for comments and
+input.
 

### 2017-02-08 paperDistance/paper.tex
tweaks

--- 2017-02-08 paperDistance/paperApJ.tex
@@
 \newcommand{\gaia}{\project{Gaia}}
+\newcommand{\xd}{\acronym{XD}}
 
-\begin{document}
+\begin{document}\sloppy\sloppypar\raggedbottom\frenchspacing
 
@@
         Adrian Price-Whelan\altaffilmark{5}}
-\altaffiltext[1]{Center for Computational Astrophysics, Flatiron Institute}
-\altaffiltext[2]{Center for Cosmology and Particle Physics, Department of Physics, New York University}
-\altaffiltext[3]{Center for Data Science, New York University}
-\altaffiltext[4]{Max-Planck-Institut f\"ur Astronomie}
-\altaffiltext[5]{Princeton University Observatory, Princeton University}
+\altaffiltext{1}{Center for Computational Astrophysics, Flatiron Institute}
+\altaffiltext{2}{Center for Cosmology and Particle Physics, Department of Physics, New York University}
+\altaffiltext{3}{Center for Data Science, New York University}
+\altaffiltext{4}{Max-Planck-Institut f\"ur Astronomie}
+\altaffiltext{5}{Princeton University Observatory, Princeton University}
 
@@
 this imaging is incredibly informative about stellar distances.
-Here we use color information on TGAS stars from \tmass\ to build a noise-deconvolved empirical prior distribution of stars in color-magnitude space.
+Here we use color information on \tgas\ stars from \tmass\ to build a noise-deconvolved empirical prior distribution of stars in color--magnitude space.
 This data-driven model contains no knowledge of the physics of stellar interiors or photospheres, nor of the Milky Way, but rather derives its precision from its generative model of noisy parallax measurements and an assumption of stationarity.
-We use the Extreme Deconvolution (XD) algorithm, which is an Empirical Bayes approximation to a full hierarchical model of the true parallax and photometry of every star.
+We use the Extreme Deconvolution (\xd) algorithm, which is an Empirical Bayes approximation to a full hierarchical model of the true parallax and photometry of every star.
 The algorithm is run not in absolute-magnitude space but in a transformed space where the measurement uncertainty is closer to Gaussian.
-The XD optimized prior is used to perform parallax and distance inferences for every star, yielding a precise stellar distance estimate and uncertainty (and full posterior) for every star.
+The \xd-optimized prior is used to perform parallax and distance inferences for every star, yielding a precise stellar distance estimate and uncertainty (and full posterior) for every star.
 The method is capitalizing on statistical shrinkage.
@@
 
-\keywords{stars}
+\keywords{
+  catalogs
+  ---
+  Hertzsprung--Russell and C--M diagrams
+  ---
+  methods: statistical
+  ---
+  parallaxes
+  ---
+  stars: distances
+  ---
+  stars: statistics
+}
 

### 2017-02-08 paperDistance/paper.tex
acks

@@
 
+This project was developed in part at the 2016 \acronym{NYC} Gaia Sprint, hosted
+by the Center for Computational Astrophysics at the Simons Foundation
+in New York City.
+
+This work has made use of data from the European Space Agency (\acronym{ESA})
+mission Gaia (http://www.cosmos.esa.int/gaia), processed by the Gaia
+Data Processing and Analysis Consortium (\acronym{DPAC},
+http://www.cosmos.esa.int/web/gaia/dpac/consortium). Funding for the
+\acronym{DPAC} has been provided by national institutions, in particular the
+institutions participating in the Gaia Multilateral Agreement.
+
+[IS THERE A \tmass\ ACKNOWLEDGEMENT?]
+
+This project was partially supported by [DWH GIVE GRANT NUMBERS]. It
+made use of the \acronym{NASA} Astrophysics Data System.
+
 \appendix

### 2017-02-09 paperDistance/paper.tex
abstract tweak
some intro text, other tweaks

@@
-%% This is emulateapj reformatting of the AASTEX sample document
-%%
+% This project is part of the stellarTwins project.
+% Copyright 2017 the authors.
+
+% # To-do list
+% - zeroth draft the introduction
+% - zeroth draft the method section
+% - zeroth draft the assumptions list
+% - reformat single-column
+
 \documentclass[iop]{emulateapj}
@@
         Adrian Price-Whelan\altaffilmark{5}}
-\altaffiltext{1}{Center for Computational Astrophysics, Flatiron Institute}
+\altaffiltext{1}{Center for Computational Astrophysics, Flatiron Institute, Simons Foundation, New York City}
 \altaffiltext{2}{Center for Cosmology and Particle Physics, Department of Physics, New York University}
 \altaffiltext{3}{Center for Data Science, New York University}
-\altaffiltext{4}{Max-Planck-Institut f\"ur Astronomie}
+\altaffiltext{4}{Max-Planck-Institut f\"ur Astronomie, Heidelberg}
 \altaffiltext{5}{Princeton University Observatory, Princeton University}
@@
 The \xd-optimized prior is used to perform parallax and distance inferences for every star, yielding a precise stellar distance estimate and uncertainty (and full posterior) for every star.
-The method is capitalizing on statistical shrinkage.
 Our posterior parallax estimates are more precise than the Gaia catalog outputs by a factor of [XXX] for the median \tgas\ star; and more precise than previous examples of Bayesian distance estimates by a factor of [YYY].
+The precision can be attributed to the statistics concept of shrinkage.
 We independently validate our distances by looking at members of Milky Way star clusters; for example, M67 is not visible at all in the \tgas\ parallax estimates, but appears clearly in our posterior parallax estimates.
@@
 
-Given the likelihood measurements of the parallax to local stars from Gaia, and using prior information for the distances to these stars built from the data themselves, we make more precise measurements of the distances to local stars in Gaia than the parallax measurements alone.
+The \gaia\ Mission will deliver more than a billion stellar distances.
+Only a small fraction (but large number) of these distance
+measurements will be purely astrometric:
+\gaia\ uses astrometric parallax to determine the distances of the closer
+stars, and calibrate spectrophotometric models.
+These spectrophotometric models, in turn, along with \gaia's on-board
+low-resolution $B_p\,R_p$ spectrophotometry, are used to provide
+distance estimates to more distant stars.
+The full stack required for these distance inferences is complex.
+It involves modeling not just stars, but also the dust in the Milky Way,
+and the response of the telescope itself.
+
+With projects like \project{The Cannon} (DWH CITE) and \project{Avast}
+(DWH CITE), we are exploring the extent to which our predictive models of
+stars could be purely data-driven or statistical.
+That is, under what circumstances could the data themselves deliver
+more precise or more accurate models than any theoretical or
+physics-based model?
+
+[DWH: NEEDS WORK] Given the likelihood measurements of the parallax to local stars from Gaia, and using prior information for the distances to these stars built from the data themselves, we make more precise measurements of the distances to local stars in Gaia than the parallax measurements alone.
 The prior, which is very similar to a color-magnitude diagram of stars, is informative because stars fall on a locus within this space.

### 2017-02-09 paperDistance/paper.tex
more content; rearrange order

@@
 
-The \gaia\ Mission will deliver more than a billion stellar distances.
+The \gaia\ Mission (DWH CITE) will deliver more than a billion stellar distances.
 Only a small fraction (but large number) of these distance
@@
 low-resolution $B_p\,R_p$ spectrophotometry, are used to provide
-distance estimates to more distant stars.
+distance estimates to more distant stars (DWH CITE).
 The full stack required for these distance inferences is complex.
 It involves modeling not just stars, but also the dust in the Milky Way,
-and the response of the telescope itself.
+and the response of the telescope itself (DWH CITE).
 
@@
 physics-based model?
+The answer to this question is extremely context dependent: It depends
+on what data are available, how much data are available, and what
+questions are being answered.
+In general, data-driven models contain fewer assumptions than
+physics-driven models, but they are also usually less interpretable.
+However, the \gaia\ use case is ideal for thinking about purely statistical
+models of stars:
+The nearby stars are measured at very high signal-to-noise and can be
+used to anchor empirical models that can be applied to the more
+distant stars.
+
+Here we deliver a demonstration-of-concept project that shows that a
+data-driven model of the \gaia\ data could in principle deliver very
+precise distances, without any input of the physics of stellar
+evolution, interiors, or photospheres...[DWH CONTENT]
 
@@
 
-\section{Data}
-
-We use stars cross matched in the TGAS and APASS set of observations.
-We require there are no NaNs in any APASS photometric band, as well as every band having positive errors.
-We check that the match between APASS and TGAS is good using extreme APASS-WISE colors.
-We correct for dust using 3D Bayestar (Schlafly+Finkbeiner) with an estimate of the distance from sampling of posterior Gaia distances from Adrain.
-There is minimal dust to the matched stars.
-To build our model here we use the Gaia G band magnitude, parallax and APASS B-V color.
-
-\section{Methods}
+\section{Assumptions and methods}
 
@@
 
-\subsection{Formalism} \label{bozomath}
+\section{Data and results}
 
-\section{Results}
+We use stars cross matched in the TGAS and APASS set of observations.
+We require there are no NaNs in any APASS photometric band, as well as every band having positive errors.
+We check that the match between APASS and TGAS is good using extreme APASS-WISE colors.
+We correct for dust using 3D Bayestar (Schlafly+Finkbeiner) with an estimate of the distance from sampling of posterior Gaia distances from Adrain.
+There is minimal dust to the matched stars.
+To build our model here we use the Gaia G band magnitude, parallax and APASS B-V color.
 

### 2017-02-16 paperDistance/paper.tex
superficial changes; title

@@
 
-\documentclass[iop]{emulateapj}
-\newcommand{\myemail}{landerson@simonsfoundation.org}
+\documentclass[12pt, preprint]{aastex6}
+\hypersetup{colorlinks = false}
 
-\slugcomment{Stars are kinda cool again.}
-
-\shorttitle{Measuring Mo' Betta Distances}
-\shortauthors{Anderson et al}
+%% Begin DWH mods
+\setlength{\voffset}{0in}
+\setlength{\hoffset}{0in}
+\setlength{\textwidth}{6in}
+\setlength{\textheight}{9in}
+\setlength{\headheight}{0ex}
+\setlength{\footnotesep}{0in}
+\setlength{\topmargin}{-\headsep}
+\setlength{\oddsidemargin}{0.25in}
+\setlength{\evensidemargin}{0.25in}
+\linespread{1.09}
+\setlength{\parindent}{1.2\baselineskip}
+\makeatletter % you know you are living your life wrong when you need to do this
+\long\def\frontmatter@title@above{
+\vspace*{-\headsep}\vspace*{\headheight}
+\noindent\footnotesize
+{\noindent\footnotesize\textsc{\@journalinfo}}\par
+{\noindent\scriptsize Preprint typeset using \LaTeX\ style AASTeX6
+with modifications by DWH.
+}\par\vspace*{-\baselineskip}\vspace*{0.625in}
+}%
+\long\def\frontmatter@abstractheading{%
+\makeaffils
+  \vspace*{-\baselineskip}\vspace*{1.5pt}
+  \vspace*{0.13189in}
+ \begingroup
+  \centering
+  \abstractname
+  \vskip 1mm
+  \par
+ \endgroup
+ \everypar{\rightskip=0.0in\leftskip=\rightskip}\par
+}%
+\def\frontmatter@keys@format{\vspace*{0.5mm}%
+  \settowidth{\keys@width}{\normalsize\@keys@name}%
+  \rightskip=0.0in\leftskip=\rightskip\parindent=0pt%
+    \hangindent=\keys@width\hangafter=1\normalsize\raggedright}%
+\def\twodigits#1{\ifnum#1<10 0\fi\the#1}
+\def\mydate{\leavevmode\hbox{\the\year-\twodigits\month-\twodigits\day}}
+\makeatother
+%% End DWH mods
 
@@
 
+\newcommand{\myemail}{landerson@simonsfoundation.org}
+\slugcomment{Draft manuscript not ready for distribution. \textbf{Copyright 2017 the authors.}}
+\shorttitle{Improving Gaia parallaxes}
+\shortauthors{Anderson, Hogg, Leistedt, Price-Whelan}
 \begin{document}\sloppy\sloppypar\raggedbottom\frenchspacing
 
-\title{Mo' Betta Gaia Distance Measurements, \\
-    Using Gaia + APASS photometry}
-
+\title{Improving \textsl{Gaia} parallaxes
+       with a data-driven model of the color--magnitude diagram}
 \author{Lauren Anderson\altaffilmark{1},
-        David W. Hogg\altaffilmark{1, 2, 3, 4},
+        David W. Hogg\altaffilmark{1,2,3,4},
         Boris Leistedt\altaffilmark{2},
@@
 That is, under what circumstances could the data themselves deliver
-more precise or more accurate models than any theoretical or
+more precise or more accurate information than any theoretical or
 physics-based model?
-The answer to this question is extremely context dependent: It depends
+The answer to this question is extremely context-dependent: It depends
 on what data are available, how much data are available, and what

### 2017-02-16 paperDistance/paper.tex
intro text

@@
 
-[DWH: NEEDS WORK] Given the likelihood measurements of the parallax to local stars from Gaia, and using prior information for the distances to these stars built from the data themselves, we make more precise measurements of the distances to local stars in Gaia than the parallax measurements alone.
-The prior, which is very similar to a color-magnitude diagram of stars, is informative because stars fall on a locus within this space.
-By deconvolving the set of observed points with their uncertainties we tighten this locus which therefore informs the individual measurements that they are drawn from a tighter distribution.
-We build this prior using Extreme Deconvolution.
-Extreme Deconvolution is a gaussian mixture model of data, specifically data that have heteroscedastic uncertainties, a quality that gaussin mixture models alone cannot take into account.
-This prior does not include any assumptions about the physical distribution of stars within the galaxy, nor does it include any stellar models.
+For our data-driven model of the color--magnitude diagram (CMD), we use
+an Empirical-Bayes approach known as Extreme Deconvolution (DWH CITE).
+This method deconvolves heteroskedastic data to derive an estimate
+of the noise-free or low-noise distribution that \emph{would have}
+been observed with far better data.
+This method has been used in astrophysics
+previously to model the velocity distribution of stars in the Solar
+Neighborhood (DWH CITE BOTH) and to perform quasar target selection
+in Sloan Digital Sky Survey data (DWH CITE BOTH).
+Its advantage over other deconvolution methods is that it takes as input,
+and handles in a principled way, heteroskedastic data.
+Its principle disadvantage for the present purposes is that it requires
+or implies a strictly Gaussian noise model.
+In the below, this requires us to transform the CMD to a space in which
+the noise is approximately Gaussian.
+
+The [use of the XD output as prior...DWH]
 

### 2017-02-17 demo_plots.py
stepping away from distances of just the cluster to all distances to correct for dust

@@ -84,7 +84,7 @@ def plot_sample(x_true, y_true, x, y, samplex, sampley, xdgmm, xlabel='x', ylabe
     ax4 = fig.add_subplot(224)
     for i in range(xdgmm.n_components):
         points = drawEllipse.plotvector(xdgmm.mu[i], xdgmm.V[i])
-        ax4.plot(points[0, :], drawEllipse.absMagKinda2absMag(points[1,:]), 'k-', alpha=xdgmm.weights[i]/np.max(xdgmm.weights))
+        ax4.plot(points[0, :], absMagKinda2absMag(points[1,:]), 'k-', alpha=xdgmm.weights[i]/np.max(xdgmm.weights))
         #draw_ellipse(xdgmm.mu[i], xdgmm.V[i], scales=[2], ax=ax4,
         #         ec='None', fc='gray', alpha=xdgmm.weights[i]/np.max(xdgmm.weights)*0.1)
 

### 2017-02-17 stellarTwins.py

@@ -354,7 +354,9 @@ def observationsCutMatched(SNthreshold=1., filename='cutMatchedArrays.npz'):
 
     lowPhotError = (tgasApass['e_bmag'] < sigMax) & (tgasApass['e_vmag'] < sigMax) & (tgasApass['e_gmag'] < sigMax) & (tgasApass['e_rmag'] < sigMax) & (tgasApass['e_imag'] < sigMax)
 
-    lowPhotError_IR = (tgas2mass['j_cmsig'] < sigMax) & (tgas2mass['h_cmsig'] < sigMax) & (tgas2mass['k_cmsig'] < sigMax)  & (tgasWise['w1sigmpro'] < sigMax) & (tgasWise['w2sigmpro'] < sigMax) & (tgasWise['w3sigmpro'] < sigMax)
+    #lowPhotError_IR = (tgas2mass['j_cmsig'] < sigMax) & (tgas2mass['h_cmsig'] < sigMax) & (tgas2mass['k_cmsig'] < sigMax)  & (tgasWise['w1sigmpro'] < sigMax) & (tgasWise['w2sigmpro'] < sigMax) & (tgasWise['w3sigmpro'] < sigMax)
+
+    lowPhotError_IR = (tgas2mass['j_cmsig'] < sigMax) & (tgas2mass['k_cmsig'] < sigMax)
 
     noDust = (medianDist < maxDist) & (np.abs(tgas['b']) > galacticLatMin)
     apassMatch = tgasApass['matched']
@@ -365,7 +367,7 @@ def observationsCutMatched(SNthreshold=1., filename='cutMatchedArrays.npz'):
     hasDust = medianDist >= minDist
     raveMatch = np.in1d(tgas['source_id'], tgasRave['source_id'])
     if galacticLatMax: inDisk = np.abs(tgas['b'] <= galacticLatMax)
-    matched = goodDistance & noNans & posErrors & lowPhotError #& apassMatch
+    matched = goodDistance & lowPhotError_IR #& apassMatch
 
 
     tgasMatched = tgas[matched]

### 2017-02-17 testXD.py

@@ -165,156 +165,72 @@ def plotXarrays(minParallaxMAS, maxParallaxMAS, apparentMagnitude, nPosteriorPoi
     return xparallaxMAS, xabsMagKinda
 
 
-def dustCorrection(magnitude, EBV, band):
-    """
-    using Finkbeiner's dust model, correct the magnitude for dust extinction
-    """
+def absMagKindaPosterior(xdgmm, ndim, mean, cov, x, projectedDimension=1):
+    allMeans = np.zeros((xdgmm.n_components, ndim))
+    allAmps = np.zeros(xdgmm.n_components)
+    allCovs = np.zeros((xdgmm.n_components, ndim, ndim))
+    summpedPosterior = np.zeros(len(x))
+    individualPosterior = np.zeros((xdgmm.n_components, nPosteriorPoints))
+    for gg in range(xdgmm.n_components):
+        #print mean2[dimension], cov2[dimension], xdgmm.mu[gg], xdgmm.V[gg]
+        newMean, newCov, newAmp = multiplyGaussians(xdgmm.mu[gg], xdgmm.V[gg], mean, cov)
+        newAmp *= xdgmm.weights[gg]
+        allMeans[gg] = newMean
+        allAmps[gg] = newAmp
+        allCovs[gg] = newCov
+        
+        summedPosterior += st.gaussian(newMean[projectedDimension], np.sqrt(newCov[projectedDimension, projectedDimension]), x, amplitude=newAmp)
+        individualPosterior[gg,:] = st.gaussian(newMean[projectedDimension], np.sqrt(newCov[projectedDimension, projectedDimension]), x, amplitude=newAmp)
+    summedPosterior = summedPosterior/np.sum(allAmps)
+    return allMeans, allAmps, allCovs, summedPosterior
 
-    dustCoeff = {'B': 3.626,
-                 'V': 2.742,
-                 'g': 3.303,
-                 'r': 2.285,
-                 'i': 1.698}
-    return mag - dustCoeff[band]*EBV
+def posterior2d(mean, amp, cov, xbins, ybins, nperGauss=100000., plot=False):
+    previousIndex = 0
+    magicNumber = -99999.
+    nsamples = np.int(np.rint(np.sum(nperGauss*allAmps/np.max(allAmps))))
+    samplesX = np.zeros(nsamples) + magicNumber
+    samplesY = np.zeros(nsamples) + magicNumber
+    
+    for gg in range(xdgmm.n_components):
[…]

### 2017-02-17 paperSpec/paperApJ.tex

--- 2017-01-23 paperDistance/paperApJ.tex
@@
 
-\shorttitle{Measuring Mo' Betta Distances}
+\shorttitle{Measuring Mo Betta Distances}
 \shortauthors{Anderson et al}
@@
 
-\title{Mo' Betta Gaia Distance Measurements, \\
+\title{Mo Betta Gaia Distance Measurements, \\
     Using Gaia + APASS photometry}
@@
 \label{eq:bayes}
-P(\theta|\textbf{E}, \textbf{I}) = P(\theta ) \frac{P(\textbf{E} |\theta,\textbf{I})}{P(\textbf{E})},
+P(\theta|\textbf{D}, \textbf{I}) = P(\theta ) \frac{P(\textbf{D} |\theta,\textbf{I})}{P(\textbf{D})},
 \end{equation}
-
-where $\theta$ is the model parameters that generate the evidence $\textbf{E}$, and P($\theta$) is the prior knowledge of the model, and $P(\textbf{E} |\theta,\textbf{I})$ is the likelihood of the evidence, given the model and some other prior assumptions.
-Our evidence is the parallaxes and G band magnitudes from Gaia, and the B and V band magnitudes from APASS.
-We build the prior from this evidence using extreme deconvolution
 

### 2017-02-20 testXD.py
sectioned up distance test

@@ -141,11 +141,7 @@ def multiplyGaussians(a, A, b, B):
     C = np.linalg.inv(Ainv + Binv)
     Cinv = np.linalg.inv(C)
     d = len(a)
-    #print 'C is: ', C, 'A is: ', A, 'Ainv is: ', Ainv, 'B is: ', B, 'Binv is: ', Binv
-    #print 'the first term is : mean - ', a, ' first part - ', np.dot(C, Ainv), ' whole thing -', np.dot(np.dot(C,Ainv), a)
-    #print 'the second term is: mean - ', b, ' first part - ', np.dot(C, Binv), ' whole thing -', np.dot(np.dot(C,Binv), b)
     c = np.dot(np.dot(C,Ainv),a) + np.dot(np.dot(C,Binv),b)
-    #print 'the new mean is :', c
 
     exponent = -0.5*(np.dot(np.dot(np.transpose(a),Ainv),a) + \
                      np.dot(np.dot(np.transpose(b),Binv),b) - \
@@ -166,10 +162,13 @@ def plotXarrays(minParallaxMAS, maxParallaxMAS, apparentMagnitude, nPosteriorPoi
 
 
 def absMagKindaPosterior(xdgmm, ndim, mean, cov, x, projectedDimension=1):
+    """
+    calculate the posterior of data likelihood mean, cov with prior xdgmm
+    """
     allMeans = np.zeros((xdgmm.n_components, ndim))
     allAmps = np.zeros(xdgmm.n_components)
     allCovs = np.zeros((xdgmm.n_components, ndim, ndim))
-    summpedPosterior = np.zeros(len(x))
+    summedPosterior = np.zeros(len(x))
     individualPosterior = np.zeros((xdgmm.n_components, nPosteriorPoints))
     for gg in range(xdgmm.n_components):
         #print mean2[dimension], cov2[dimension], xdgmm.mu[gg], xdgmm.V[gg]
@@ -178,48 +177,142 @@ def absMagKindaPosterior(xdgmm, ndim, mean, cov, x, projectedDimension=1):
         allMeans[gg] = newMean
         allAmps[gg] = newAmp
         allCovs[gg] = newCov
-        
+
         summedPosterior += st.gaussian(newMean[projectedDimension], np.sqrt(newCov[projectedDimension, projectedDimension]), x, amplitude=newAmp)
         individualPosterior[gg,:] = st.gaussian(newMean[projectedDimension], np.sqrt(newCov[projectedDimension, projectedDimension]), x, amplitude=newAmp)
[…]

### 2017-02-20 testXD.py
measuring all distances in 2mass

@@ -520,16 +520,18 @@ if __name__ == '__main__':
     sigMax = 1.086/thresholdSN
     lowPhotErrorcut = (bandDictionary[mag1]['array'][bandDictionary[mag1]['err_key']] < sigMax) & \
                       (bandDictionary[mag2]['array'][bandDictionary[mag2]['err_key']] < sigMax)
+    nonzeroError = (bandDictionary[mag1]['array'][bandDictionary[mag1]['err_key']] != 0.0) & \
+                   (bandDictionary[mag2]['array'][bandDictionary[mag2]['err_key']] != 0.0)
 
     if survey == '2MASS':
         nonZeroColor = (bandDictionary[mag1]['array'][bandDictionary[mag1]['key']] -
                         bandDictionary[mag2]['array'][bandDictionary[mag2]['key']] != 0.0) & \
                        (bandDictionary[mag1]['array'][bandDictionary[mag1]['key']] != 0.0)
 
-        indices = parallaxSNcut & lowPhotErrorcut & nonZeroColor
+        indices = parallaxSNcut & lowPhotErrorcut & nonZeroColor & nonzeroError
 
     else:
-        indices = parallaxSNcut & lowPhotErrorcut
+        indices = parallaxSNcut & lowPhotErrorcut & nonzeroError
 
 
     try:
@@ -566,9 +568,18 @@ if __name__ == '__main__':
     nstars = len(tgasCutMatched)
     summedPosterior = np.zeros((nstars, nPosteriorPoints))
 
-    for index in range(nstars):
+    for i, index in enumerate(np.where(indices)[0]):
+        if np.mod(i, 1000) == 0.0: print i
         meanData, covData = matrixize(color[index], absMagKinda[index], color_err[index], absMagKinda_err[index])
-        apparentMagnitude = tgasCutMatched['phot_g_mean_mag'][index]
+        dimension = 0
+        windowFactor = 5. #the number of sigma to sample in mas for plotting
+        minParallaxMAS = tgasCutMatched['parallax'][index] - windowFactor*tgasCutMatched['parallax_error'][index]
+        maxParallaxMAS = tgasCutMatched['parallax'][index] + windowFactor*tgasCutMatched['parallax_error'][index]
+        apparentMagnitude = bandDictionary[absmag]['array'][bandDictionary[absmag]['key']][index]
[…]

### 2017-02-20 testXD.py
added distance array to save

@@ -567,7 +567,7 @@ if __name__ == '__main__':
     #calculate parallax-ish posterior for each star
     nstars = len(tgasCutMatched)
     summedPosterior = np.zeros((nstars, nPosteriorPoints))
-
+    distance = np.zeros((nstars, nPosteriorPoints))
     for i, index in enumerate(np.where(indices)[0]):
         if np.mod(i, 1000) == 0.0: print i
         meanData, covData = matrixize(color[index], absMagKinda[index], color_err[index], absMagKinda_err[index])
@@ -581,5 +581,6 @@ if __name__ == '__main__':
         positive = xparallaxMAS > 0.
         allMeans, allAmps, allCovs, summedPosteriorAbsmagKinda = absMagKindaPosterior(xdgmm, ndim, meanData[dimension], covData[dimension], xabsMagKinda, projectedDimension=1)
 
-        summedPosterior[i, :] = summedPosteriorAbsmagKinda[positive]*xparallaxMAS[positive]**2.*10.**(0.2*apparentMagnitude)
-    np.save('posteriorDistanceTgas', summedPosterior)
+        summedPosterior[i, :] = summedPosteriorAbsmagKinda*xparallaxMAS**2.*10.**(0.2*apparentMagnitude)
+        distance[i, :] = 1./xparallaxMAS
+    np.savez('posteriorDistanceTgas', posterior=summedPosterior, distance=distance)

### 2017-02-20 paperDistance/paper.tex
intro content

@@
 
-The [use of the XD output as prior...DWH]
+The novel use of XD here is to treat its output---the deconvolved CMD---as
+a prior for use in inference of individual-star distances.
+These inferences, one per star, provide much more precise parallax, distance,
+or absolute magnitude estimates than we get from each star's primary 
+\tgas\ parallax measurement alone.
+That is, we are using the XD model to de-noise the \gaia\ parallax
+measurements.
+Technically, since XD produces a maximum-marginalized-likelihood estimator
+of the deconvolved distribution,
+its use as a prior is only approximately valid; it constitutes using the
+data twice.
+(It wouldn't be using the data twice---that is, it would be valid from a
+probabilistic perspective---if we performed a full hierarchical Bayesian
+inference.)
+However, since the data set is large, the XD approximation is not bad; it is
+sometimes known as the ``empirical Bayes'' method, and is well studied (DWH REFS).
+
+This project is fundamentally a demonstration of concept:
+We are only using the ``small'' (relative to the expected final \gaia\ data set)
+\tgas\ Catalog.
+We are performing only an approximation to full Bayesian hierarchical inference.
+We have to use photometry that is ground-based, rather than the full-precision,
+space-based photometry that the \gaia\ Mission will deliver.
+However, as we show below, we get very good results.
+It is promising that we will be able to obtain similar or even better results
+in later \gaia\ data releases, and eventually put distances on $>10^9$
+\gaia\ stars using photometric distance methodologies but without any
+commitment to---or even use of---physical models of stars.
 

### 2017-02-21 paperDistance/paper.tex
intro content

@@
 precise distances, without any input of the physics of stellar
-evolution, interiors, or photospheres...[DWH CONTENT]
+evolution, interiors, or photospheres.
+Most stellar models used to generate photometric distances or photometric
+parallaxes have been physical models.
+Because there are small issues with physical models of stars, leading to small
+color and luminosity differences with the data, these models produce distance
+estimates that are biased.
+Besides, they build a long list of physical assumptions---assumptions about
+how stars work and how to model them numerically---into the distance estimates.
+That is, it is impossible (when using physical models)
+to deliver distance estimates that involve minimal in assumptions.
 
-For our data-driven model of the color--magnitude diagram (CMD), we use
+However, the use of physical models for distance estimation is not necessary.
+Because there are stars at all luminosities with good parallax information,
+it is possible to build a stellar photometric model from the data themselves.
+Different stars are observed at different levels of parallax precision,
+so it requires relatively high technology to build this data-driven model
+using all the data available, fairly and responsibly.
+
+Here we build and use just such a data-driven model.
+In particular, we use all of the \gaia\ \tgas\ data, matched to the \tmass\ 
+photometry, to make a model of the noise-free (or low-noise)
+color--magnitude diagram (CMD) of stars.
+We use
 an Empirical-Bayes approach known as Extreme Deconvolution (DWH CITE).
@@
 
-\begin{thebibliography}{}
-\end{thebibliography}
+%\begin{thebibliography}{}
+%\end{thebibliography}
 

### 2017-02-21 paperDistance/paper.tex
discussion

@@
 \section{Discussion}
-Apass HR Diagram looks very different from Rave HR Diagram, MS stars are incorrectly being labeled as giants
+What have we achieved and why is it important?
 
-Long discussion of dust
+Come back to the assumptions. What happens if we change or violate them?
 
-Number of gaussians in XD, and the w parameter
+Long discussion of the noise model. We took serious short-cuts. Call out to Boris.
 
-\acknowledgments It is a pleasure to thank Andy Casey (Monash) and the
-attendees at the Stars Group Meeting at the CCA for comments and
-input.
+Long discussion of dust.
+
+Some discussion of possible failure modes. Including very rare kinds of stars.
+
+Number of gaussians in XD, and the w parameter. That we set them heuristically.
+
+\acknowledgments It is a pleasure to thank
+  Jo Bovy (Toronto),
+  Andy Casey (Monash),
+and the attendees at the Stars Group Meeting at the Flatiron Institute
+Center for Computational Astrophysics for comments and input.
 

### 2017-02-21 paperDistance/paper.tex
sdss typography

@@
 Neighborhood (DWH CITE BOTH) and to perform quasar target selection
-in Sloan Digital Sky Survey data (DWH CITE BOTH).
+in \project{Sloan Digital Sky Survey} data (DWH CITE BOTH).
 Its advantage over other deconvolution methods is that it takes as input,

### 2017-02-21 stellarTwins.py
dust corrected prior

@@ -412,9 +412,9 @@ def distanceModulus(distCutMatched):
     varMuMatched = np.mean((muMatched - meanMuMatched[:,None]) ** 2, axis=1)
     return meanMuMatched, varMuMatched
 
-def dust(l, b, distance, plot=False):
+def dust(l, b, distance, plot=False, max_samples=2, mode='median'):
     sfd = SFDQuery()
-    bayes = BayestarQuery(max_samples=2)
+    bayes = BayestarQuery(max_samples=max_samples)
     c = SkyCoord(l, b,
             distance = distance,
             frame='galactic')
@@ -422,11 +422,11 @@ def dust(l, b, distance, plot=False):
     cNoDist = SkyCoord(l, b,
             frame='galactic')
 
-    bayesDust = bayes(c, mode='median')
-    bayesDustNoDist = bayes(cNoDist, mode='median')
+    bayesDust = bayes(c, mode=mode)
+    bayesDustNoDist = bayes(cNoDist, mode=mode)
 
     #!!!!! Do something else than setting it equal to 0 !!!!!
-    bayesDust[np.isnan(bayesDust)] = 0.0
+    #if len(bayesDust) > 1: bayesDust[np.isnan(bayesDust)] = 0.0
 
     if plot:
         fig, ax = plt.subplots(3, figsize=(5, 7.5))

### 2017-02-21 testXD.py

@@ -424,8 +424,56 @@ def distanceTest(tgasCutMatched, nPosteriorPoints, data1, data2, err1, err2, xli
     plt.tight_layout()
     figDist.savefig('distancesM67.png')
 
+def dustCorrectionPrior(tgasCutMatched, dataFilename, quantile=0.05, nDistanceSamples=512):
+    dustFile = 'dustCorrection_' + dataFilename
 
-def dustCorrection(magnitude, EBV, band):
+    try:
+        data = np.load(dustFile + '.npz')
+        dustEBV = data['ebv']
+        sourceID = data['sourceID']
+    except IOError:
+        print 'calculating dust corrections, this may take awhile'
+
+        nstars = len(tgasCutMatched)
+        max_samples = 2
+        sourceID = np.zeros(nstars, dtype='>i8')
+        dustEBV = np.zeros(nstars)
+
+        for i, index in enumerate(np.where(indices)[0]):
+            if np.mod(i, 1000) == 0.0:
+                print i
+                np.savez('dustCorrection_' + dataFilename, ebv=dustEBV, sourceID=sourceID)
+
+            #calculate parallax-ish posterior for each star
+            meanData, covData = matrixize(color[index], absMagKinda[index], color_err[index], absMagKinda_err[index])
+            dimension = 0
+            windowFactor = 5. #the number of sigma to sample in mas for plotting
+            minParallaxMAS = tgasCutMatched['parallax'][index] - windowFactor*tgasCutMatched['parallax_error'][index]
+            maxParallaxMAS = tgasCutMatched['parallax'][index] + windowFactor*tgasCutMatched['parallax_error'][index]
+            apparentMagnitude = bandDictionary[absmag]['array'][bandDictionary[absmag]['key']][index]
+            xparallaxMAS, xabsMagKinda = plotXarrays(minParallaxMAS, maxParallaxMAS, apparentMagnitude, nPosteriorPoints=nPosteriorPoints)
+
+            positive = xparallaxMAS > 0.
+            allMeans, allAmps, allCovs, summedPosteriorAbsmagKinda = absMagKindaPosterior(xdgmm, ndim, meanData[dimension], covData[dimension], xabsMagKinda, projectedDimension=1)
+
+            #normalize prior pdf
[…]

### 2017-02-21 paperDistance/paper.tex
dust subsection

@@
 
+\subsection{Dust}
+
+Dust corrections are non trivial for stars with poor signal to noise.
+Stars with poor signal to noise have a likelihood that is consistent with infinite distance and can therefore have severe dust corrections.
+This is most obvious in the giants, which tend to have the lowest signal to noise, and can therefore have dust corrections $\sim 1-2$ magnitudes.
+To apply a dust correction to the 2MASS photometry, we first generate our prior using XD on the observed, attenuated photometry.
+Using this raw prior we infer the distances to all the stars in our prior more precisely, increasing our signal to noise for each star and therefore minimize severe distances and their associated severe dust corrections.
+With more precise distance posteriors, we take the $5\%$ quantile of each distance posterior, so the closest bit of the posterior, and sample the Green et al 3D dust map at that distance.
+For each distance we take the median of the samples of the dust model.
+We apply each dust correction to our 2MASS photometry given by Equation \ref{eq:dust} where $K_{\lambda}$ = [3.626, 2.742, 3.303, 2.285, 1.698] for bands [B, V, g, r, i] respectively, taken from table (REF TABLE SFD), and $E(B-V)$ is the output of the Green et al dust models.
+We then regenerate our prior using the dust corrected photometry.
+We do not update the errors, but keep them the same as before the dust correction, which is wrong and under representing our errors but this leads again to a lighter deconvolution.
+
+\begin{equation}
+\label{eq:dust}
+m_{\mathrm{band, corrected}} = m_{\mathrm{band}} - K_{\lambda} \times E(B-V)
+\end{equation}
+
 \section{Discussion}

### 2017-02-21 testXD.py
added IR dust coeff

@@ -482,7 +482,11 @@ def dustCorrection(mag, EBV, band):
                  'V': 2.742,
                  'g': 3.303,
                  'r': 2.285,
-                 'i': 1.698}
+                 'i': 1.698,
+                 'J': 0.709,
+                 'H': 0.449,
+                 'K': 0.302}
+                 
     return mag - dustCoeff[band]*EBV
 
 def cdf(x, y):

### 2017-02-22 testXD.py
speeding up dust

@@ -18,6 +18,7 @@ import drawEllipse
 from astropy.coordinates import SkyCoord
 import astropy.units as units
 import scipy.integrate
+import time
 
 def convert2gal(ra, dec):
     """
@@ -424,7 +425,7 @@ def distanceTest(tgasCutMatched, nPosteriorPoints, data1, data2, err1, err2, xli
     plt.tight_layout()
     figDist.savefig('distancesM67.png')
 
-def dustCorrectionPrior(tgasCutMatched, dataFilename, quantile=0.05, nDistanceSamples=512):
+def dustCorrectionPrior(tgasCutMatched, dataFilename, quantile=0.05, nDistanceSamples=512, max_samples = 2):
     dustFile = 'dustCorrection_' + dataFilename
 
     try:
@@ -435,14 +436,19 @@ def dustCorrectionPrior(tgasCutMatched, dataFilename, quantile=0.05, nDistanceSa
         print 'calculating dust corrections, this may take awhile'
 
         nstars = len(tgasCutMatched)
-        max_samples = 2
         sourceID = np.zeros(nstars, dtype='>i8')
-        dustEBV = np.zeros(nstars)
-
+        #dustEBV = np.zeros(nstars)
+        #dustEBV50 = np.zeros(nstars)
+        distanceQuantile = np.zeros(nstars)
+        distanceQuantile50 = np.zeros(nstars)
+        start = time.time()
         for i, index in enumerate(np.where(indices)[0]):
             if np.mod(i, 1000) == 0.0:
-                print i
-                np.savez('dustCorrection_' + dataFilename, ebv=dustEBV, sourceID=sourceID)
+                end = time.time()
+                print i, ' took ', str(end - start), 'seconds, projecting will be ', str((end-start)*(nstars/1000.))
+                start = time.time()
+
+                #np.savez('dustCorrection_' + dataFilename, ebv=dustEBV, sourceID=sourceID)
 
             #calculate parallax-ish posterior for each star
             meanData, covData = matrixize(color[index], absMagKinda[index], color_err[index], absMagKinda_err[index])
@@ -464,13 +470,18 @@ def dustCorrectionPrior(tgasCutMatched, dataFilename, quantile=0.05, nDistanceSa
             sampleDistance = samples(distance[::-1], posteriorDistance[::-1], nDistanceSamples, plot=False)
 
[…]

### 2017-02-22 paperDistance/paper.tex
dust subsection
shout out to and cf Boris

@@
 data twice.
-(It wouldn't be using the data twice---that is, it would be valid from a
-probabilistic perspective---if we performed a full hierarchical Bayesian
-inference.)
 However, since the data set is large, the XD approximation is not bad; it is
 sometimes known as the ``empirical Bayes'' method, and is well studied (DWH REFS).
+
+It wouldn't be using the data twice---that is, it would be valid from a
+probabilistic perspective---if we performed a full hierarchical Bayesian
+inference.
+Exploration of that possibility, and its computational tractability,
+is among our long-term goals and motivations.
+Along those lines, this paper can be seen as a companion to
+a related project (CITE BORIS), in which we use a
+somewhat less flexible model for the CMD, but in which we take a much more principled
+approach to the inference.
 
@@
 We do not update the errors, but keep them the same as before the dust correction, which is wrong and under representing our errors but this leads again to a lighter deconvolution.
+Should I quote some statistics about the dust corrections?
+mostly small?
+maybe difference before after correcting for distance?
 

### 2017-02-22 paperDistance/paper.tex
intro tweaks; notes to selves

@@
 on what data are available, how much data are available, and what
-questions are being answered.
+questions are being asked.
 In general, data-driven models contain fewer assumptions than
@@
 models of stars:
-The nearby stars are measured at very high signal-to-noise and can be
-used to anchor empirical models that can be applied to the more
-distant stars.
+The distance to a star is a simple latent variable, and the nearby
+stars constitute a high signal-to-noise ``training set''.
+These nearby stars can be used to anchor empirical models that can
+then be applied to the more distant stars.
 
@@
 data-driven model of the \gaia\ data could in principle deliver very
-precise distances, without any input of the physics of stellar
+precise distances, without any input of the physics or numerical models of stellar
 evolution, interiors, or photospheres.
 Most stellar models used to generate photometric distances or photometric
-parallaxes have been physical models.
-Because there are small issues with physical models of stars, leading to small
-color and luminosity differences with the data, these models produce distance
-estimates that are biased.
-Besides, they build a long list of physical assumptions---assumptions about
-how stars work and how to model them numerically---into the distance estimates.
-That is, it is impossible (when using physical models)
-to deliver distance estimates that involve minimal in assumptions.
+parallaxes have been physical models, based on gravity, fluid mechanics,
+radiative transfer, nuclear reactions, and atomic and molecular transitions.
+Because there are small issues with each of these components, there are small
+color and luminosity differences with the data---the models are wrong in detail---and
+the physical models therefore produce distance estimates that are biased.
+Besides, they build a long list of physical assumptions (about
+nuclear reactions, convection, and thermal timescales, for example)
+into the distance estimates.
+That is, it is impossible, when using physical models,
+to deliver photometric distance estimates that involve minimal assumptions.
 
@@
 using all the data available, fairly and responsibly.
+We have this technology!
 
@@
 Instead of making assumptions about the physical distributions of stars in the galaxy, or using stellar models for the modeled HR diagram of stars, we build a prior from the data themselves.
+
+LA: Insert itemized or enumerated list of assumptions here. DWH will
+typeset it to look okay! Be exhaustive and then we will edit down.
 

### 2017-02-22 paperDistance/paper.tex
start discussion; need to make dinner now

@@
 \section{Discussion}
+
+We have shown that it is possible to obtain photometric parallaxes for distant
+stars in the \gaia\ data without any use of physical stellar models.
+Although we have not performed principled Bayesian inference in this project,
+we have performed a statistically responsible deconvolution of the
+\gaia\ \tgas\ data that is justifiable under a clear set of assumptions.
+...DWH NOT DONE THIS PARAGRAPH
+
 What have we achieved and why is it important?

### 2017-02-22 paperDistance/paper.tex
discussion content

--- 2017-02-22 paperDistance/paper.tex
@@
 \section{Discussion}
-What have we achieved and why is it important?
+
+We have shown that it is possible to obtain photometric parallaxes for
+distant stars in the \gaia\ data without any use of physical stellar
+models.
+Although we have not performed principled Bayesian inference in this
+project, we have performed a statistically responsible deconvolution
+of the \gaia\ \tgas\ data that is justifiable under a clear set of
+assumptions.
+This deconvolution makes use of a reasonable noise model and an
+assumption of stationarity to shrink the parallax uncertainties, and
+do so dramatically for the stars measured at lower signal-to-noise.
+Because the model we use (Extreme Deconvolution or XD) accounts for
+heteroskedastic noise, we were able to build our photometric parallax
+model---which is a model of the color--magnitude diagram (CMD)---using
+all the data, not just the data with the highest signal-to-noise or
+largest parallaxes.
+That is, the model for the CMD we have built is representative for
+the \tgas\ Catalog.
+Inasmuch as that is true, we expect the distance estimates to be
+weakly biased.
+
+What would be different if we wanted a model not of the \tgas\ data,
+but a model appropriate for the full billion stars, or for a
+volume-limited sample of stars?
+How could we convert our existing results?
+How would we do it properly in general?
 
@@
 Number of gaussians in XD, and the w parameter. That we set them heuristically.
+
+How this compares to other Bayesian attempts at inferring distances,
+like, eg, the Tri--Coryn papers. Favorably! And with \emph{fewer}
+assumptions, in fact!
 

### 2017-02-22 paperDistance/paper.tex
intro text re assumptions

@@
 
+On a different tack, there has been some discussion in the literature
+about how to use a parallax responsibly to infer a distance (CITE TRI+CBJ).
+These papers use the \gaia\ measurement to construct a parallax likelihood,
+and combine this with a distance prior.
+While this is sensible and correct, all present methods proposed along
+these lines build in informative (and known-to-be-wrong) assumptions
+about the line-of-sight distance distribution to \gaia\ stars.
+That is, they build in strong and informative assumptions about the Milky
+Way.
+The methodology presented in this paper makes exceedingly weak assumptions
+about the Milky Way, and similarly weak assumptions about the properties of
+stars.
+The strongest new assumption made in this work is that for every star
+in \tgas, there are other, photometrically (and bolometrically)
+similar stars.
+That is a much weaker assumption than has been made in other Bayesian
+distance estimations.
+
 This project is fundamentally a demonstration of concept:
@@
 \gaia\ stars using photometric distance methodologies but without any
-commitment to---or even use of---physical models of stars.
+commitment to---or even use of---physical models of stars or the Milky Way.
 

### 2017-02-22 paperDistance/paper.tex
discussion content

@@
 We have shown that it is possible to obtain photometric parallaxes for
-distant stars in the \gaia\ data without any use of physical stellar
-models.
-Although we have not performed principled Bayesian inference in this
-project, we have performed a statistically responsible deconvolution
+distant stars in the \gaia\ \tgas\ data without any use of physical stellar
+models, nor models of the Milky Way.
+We obtained these by building a data-driven model of the
+color--magnitude diagram (CMD) of the stars in the \tgas\ data set,
+and using it as a prior pdf for Bayesian interences.
+The posterior pdfs for distance that we obtain are, in general, much
+narrower than the likelihood functions delivered by the
+\gaia\ Mission, and therefore the distance estimates (or,
+equivalently, parallax estimates) are much more precise.
+It is not surprising that a Bayesian inference provides more
+precise inferences than the likelihood function alone; Bayesian
+inferences bring in new information that decrease variance (but
+often introduce bias).
+We have shown thatm, in addition to pure precision improvements, at
+least some aspects of \emph{accuracy} have been improved as well, by
+showing that the posterior distance estimates to stars in stellar
+clusters are much more clustered than likelihood-based distance
+estimates.
+
+Although we have not performed principled hierarchical Bayesian inference in this
+project, we built our prior for each star's distance by
+performing a statistically responsible deconvolution
 of the \gaia\ \tgas\ data that is justifiable under a clear set of
@@
 the \tgas\ Catalog.
-Inasmuch as that is true, we expect the distance estimates to be
-weakly biased.
+Inasmuch as that is true, we expect any (sensible) distance estimates
+we generate from our distance posterior pdfs to be weakly biased.
 

### 2017-02-22 paperDistance/paper.tex
acronym trouble

@@
 \newcommand{\xd}{\acronym{XD}}
+\newcommand{\cmd}{\acronym{CMD}}
 
@@
 photometry, to make a model of the noise-free (or low-noise)
-color--magnitude diagram (CMD) of stars.
+color--magnitude diagram (\cmd) of stars.
 We use
@@
 or implies a strictly Gaussian noise model.
-In the below, this requires us to transform the CMD to a space in which
+In the below, this requires us to transform the \cmd\ to a space in which
 the noise is approximately Gaussian.
 
-The novel use of XD here is to treat its output---the deconvolved CMD---as
+The novel use of \xd\ here is to treat its output---the deconvolved \cmd---as
 a prior for use in inference of individual-star distances.
@@
 \tgas\ parallax measurement alone.
-That is, we are using the XD model to de-noise the \gaia\ parallax
+That is, we are using the \xd\ model to de-noise the \gaia\ parallax
 measurements.
-Technically, since XD produces a maximum-marginalized-likelihood estimator
+Technically, since \xd\ produces a maximum-marginalized-likelihood estimator
 of the deconvolved distribution,
@@
 data twice.
-However, since the data set is large, the XD approximation is not bad; it is
+However, since the data set is large, the \xd\ approximation is not bad; it is
 sometimes known as the ``empirical Bayes'' method, and is well studied (DWH REFS).
@@
 a related project (CITE BORIS), in which we use a
-somewhat less flexible model for the CMD, but in which we take a much more principled
+somewhat less flexible model for the \cmd, but in which we take a much more principled
 approach to the inference.
@@
 This is most obvious in the giants, which tend to have the lowest signal to noise, and can therefore have dust corrections $\sim 1-2$ magnitudes.
-To apply a dust correction to the 2MASS photometry, we first generate our prior using XD on the observed, attenuated photometry.
+To apply a dust correction to the 2MASS photometry, we first generate our prior using \xd\ on the observed, attenuated photometry.
 Using this raw prior we infer the distances to all the stars in our prior more precisely, increasing our signal to noise for each star and therefore minimize severe distances and their associated severe dust corrections.
@@
 We obtained these by building a data-driven model of the
-color--magnitude diagram (CMD) of the stars in the \tgas\ data set,
+color--magnitude diagram (\cmd) of the stars in the \tgas\ data set,
 and using it as a prior pdf for Bayesian interences.
@@
 do so dramatically for the stars measured at lower signal-to-noise.
-Because the model we use (Extreme Deconvolution or XD) accounts for
+Because the method we use (Extreme Deconvolution or \xd, which is
+an Empirical Bayesian maximum-marginalized-likelihood estimator) accounts for
 heteroskedastic noise, we were able to build our photometric parallax
-model---which is a model of the color--magnitude diagram (CMD)---using
+model---which is a model of the \cmd---using
 all the data, not just the data with the highest signal-to-noise or
 largest parallaxes.
-That is, the model for the CMD we have built is representative for
+That is, the model for the \cmd\ we have built is representative for
 the \tgas\ Catalog.
@@
 
-Number of gaussians in XD, and the w parameter. That we set them heuristically.
+Number of gaussians in \xd, and the w parameter. That we set them heuristically.
 

### 2017-02-22 paperDistance/paper.tex
note to self

@@
 often introduce bias).
-We have shown thatm, in addition to pure precision improvements, at
+We have shown that, in addition to pure precision improvements, at
 least some aspects of \emph{accuracy} have been improved as well, by
 showing that the posterior distance estimates to stars in stellar
-clusters are much more clustered than likelihood-based distance
+clusters [HOGG: Is this too general?] are much more clustered than likelihood-based distance
 estimates.

### 2017-02-24 testXD.py
bugs in dust

@@ -481,7 +481,7 @@ def dustCorrectionPrior(tgasCutMatched, dataFilename, quantile=0.05, nDistanceSa
         end = time.time()
         print 'dust sampling ', str(nDistanceSamples), ' took ',str(end-start), ' seconds for index ', str(i)
 
-        np.savez('dustCorrection_' + dataFilename+'_alldistance', ebv=dustEBV, sourceID=sourceID, ebv50=dustEBV50)
+        np.savez('dustCorrection_' + dataFilename, ebv=dustEBV, sourceID=sourceID, ebv50=dustEBV50)
     return dustEBV, sourceID
 
 def dustCorrection(mag, EBV, band):
@@ -678,7 +678,7 @@ if __name__ == '__main__':
 
     dustEBV, sourceID = dustCorrectionPrior(tgasCutMatched, dataFilename, quantile=0.05, nDistanceSamples=128, max_samples=None)
 
-    assert np.sum(tgasCutMatched['sourceID'][indices] - sourceID) == 0.0, 'dust and data arrays are sorted differently !!!'
+    assert np.sum(tgasCutMatched['source_id'][indices] - sourceID) == 0.0, 'dust and data arrays are sorted differently !!!'
 
     for i, index in enumerate(np.where(indices)[0]):
         mag1DustCorrected   = dustCorrection(bandDictionary[mag1]['array']  [bandDictionary[mag1]['key']][index], dustEBV[i], mag1)

### 2017-02-24 testXD.py
more dust bugs

@@ -680,19 +680,18 @@ if __name__ == '__main__':
 
     assert np.sum(tgasCutMatched['source_id'][indices] - sourceID) == 0.0, 'dust and data arrays are sorted differently !!!'
 
-    for i, index in enumerate(np.where(indices)[0]):
-        mag1DustCorrected   = dustCorrection(bandDictionary[mag1]['array']  [bandDictionary[mag1]['key']][index], dustEBV[i], mag1)
-        mag2DustCorrected   = dustCorrection(bandDictionary[mag2]['array']  [bandDictionary[mag2]['key']][index], dustEBV[i], mag2)
-        apparentMagDustCorrected = dustCorrection(bandDictionary[absmag]['array'][bandDictionary[absmag]['key']][index], dustEBV[i], absmag)
-        absMagDustCorrected = tgasCutMatched['parallax_error']*10.**(0.2*apparentMagDustCorrected)
-        #B_dustcorrected = dustCorrection(apassCutMatched['bmag'], bayesDust, 'B')
-        #need to define color_err and absMagKinda_err when including dust correction
-        colorDustCorrected = mag1DustCorrected - mag2DustCorrected
-        data1 = colorDustCorrected
-        data2 = absMagDustCorrected
-        X, Xerr = matrixize(data1[indices], data2[indices], err1[indices], err2[indices])
-        xdgmm.fit(X, Xerr)
-        xdgmm.save_model(xdgmmFilenameDust)
+    mag1DustCorrected   = dustCorrection(bandDictionary[mag1]['array']  [bandDictionary[mag1]['key']][indices], dustEBV, mag1)
+    mag2DustCorrected   = dustCorrection(bandDictionary[mag2]['array']  [bandDictionary[mag2]['key']][indices], dustEBV, mag2)
+    apparentMagDustCorrected = dustCorrection(bandDictionary[absmag]['array'][bandDictionary[absmag]['key']][indices], dustEBV, absmag)
+    absMagDustCorrected = tgasCutMatched['parallax_error'][indices]*10.**(0.2*apparentMagDustCorrected)
+    #B_dustcorrected = dustCorrection(apassCutMatched['bmag'], bayesDust, 'B')
+    #need to define color_err and absMagKinda_err when including dust correction
+    colorDustCorrected = mag1DustCorrected - mag2DustCorrected
+    data1 = colorDustCorrected
+    data2 = absMagDustCorrected
[…]

### 2017-02-24 stellarTwins.py
cleaning up dust

@@ -412,18 +412,23 @@ def distanceModulus(distCutMatched):
     varMuMatched = np.mean((muMatched - meanMuMatched[:,None]) ** 2, axis=1)
     return meanMuMatched, varMuMatched
 
-def dust(l, b, distance, plot=False, max_samples=2, mode='median'):
-    sfd = SFDQuery()
-    bayes = BayestarQuery(max_samples=max_samples)
-    c = SkyCoord(l, b,
-            distance = distance,
-            frame='galactic')
+def dust(l, b, distance, plot=False, max_samples=2, mode='median', model='bayes'):
+    if model == 'sfd':
+        c = SkyCoord(l, b,
+                frame='galactic')
+        sfd = SFDQuery()
+        dust = sfd(c)
 
-    cNoDist = SkyCoord(l, b,
-            frame='galactic')
+    if model == 'bayes':
+        c = SkyCoord(l, b,
+                distance = distance,
+                frame='galactic')
+        bayes = BayestarQuery(max_samples=max_samples)
+        dust = bayes(c, mode=mode)
 
-    bayesDust = bayes(c, mode=mode)
-    bayesDustNoDist = bayes(cNoDist, mode=mode)
+    #cNoDist = SkyCoord(l, b,
+    #        frame='galactic')
+    #bayesDustNoDist = bayes(cNoDist, mode=mode)
 
     #!!!!! Do something else than setting it equal to 0 !!!!!
     #if len(bayesDust) > 1: bayesDust[np.isnan(bayesDust)] = 0.0
@@ -443,7 +448,7 @@ def dust(l, b, distance, plot=False, max_samples=2, mode='median'):
         plt.hist(magsMatched['bmag'], bins=bins, histtype='step')
         plt.hist(magsMatched['bmag'] - B_RedCoeff*bayesDust, bins=bins, histtype='step')
         plt.tight_layout()
-    return bayesDust
+    return dust
 
 def dustTightenMS(B_V_dust, M_V_dust, B_V, M_V):
     #check dust tightens main sequence

### 2017-02-24 testXD.py

@@ -434,52 +434,60 @@ def dustCorrectionPrior(tgasCutMatched, dataFilename, quantile=0.05, nDistanceSa
         sourceID = data['sourceID']
     except IOError:
         print 'calculating dust corrections, this may take awhile'
-
-        nstars = len(tgasCutMatched)
-        sourceID = np.zeros(nstars, dtype='>i8')
-        #dustEBV = np.zeros(nstars)
-        #dustEBV50 = np.zeros(nstars)
-        distanceQuantile = np.zeros(nstars)
-        distanceQuantile50 = np.zeros(nstars)
-        start = time.time()
-        for i, index in enumerate(np.where(indices)[0]):
-            if np.mod(i, 1000) == 0.0:
-                end = time.time()
-                print i, ' took ', str(end - start), 'seconds, projecting will be ', str((end-start)*(nstars/1000.))
-                start = time.time()
+        try:
+            data = np.load('distanceQuantiles'+dataFilename)
+            distanceQuntile = data['distanceQuntile']
+            distanceQuantile50 = data['distanceQuantile50']
+        except IOError:
+            nstars = len(tgasCutMatched)
+            sourceID = np.zeros(nstars, dtype='>i8')
+            #dustEBV = np.zeros(nstars)
+            #dustEBV50 = np.zeros(nstars)
+            distanceQuantile = np.zeros(nstars)
+            distanceQuantile50 = np.zeros(nstars)
+            start = time.time()
+            for i, index in enumerate(np.where(indices)[0]):
+                if np.mod(i, 1000) == 0.0:
+                    end = time.time()
+                    print i, ' took ', str(end - start), 'seconds, projecting will be ', str((end-start)*(nstars/1000.))
+                    start = time.time()
 
                 #np.savez('dustCorrection_' + dataFilename, ebv=dustEBV, sourceID=sourceID)
 
-            #calculate parallax-ish posterior for each star
-            meanData, covData = matrixize(color[index], absMagKinda[index], color_err[index], absMagKinda_err[index])
-            dimension = 0
-            windowFactor = 5. #the number of sigma to sample in mas for plotting
-            minParallaxMAS = tgasCutMatched['parallax'][index] - windowFactor*tgasCutMatched['parallax_error'][index]
[…]

### 2017-02-24 testXD.py
cleaning up dust

@@ -483,7 +483,7 @@ def dustCorrectionPrior(tgasCutMatched, dataFilename, quantile=0.05, nDistanceSa
         start = time.time()
         dustEBV, dustEBV50 = st.dust([l,l], [b,b], [distanceQuantile*units.kpc, distanceQuantile50*units.kpc], mode='median')
         nan = np.isnan(dustEBV)
-        dustNan = st.dust(l, b, None, model='sfd')
+        dustNan = st.dust(l[nan], b[nan], None, model='sfd')
         dustEBV[nan] = dustNan
         dustEBV50[nan] = dustNan
         end = time.time()

### 2017-02-24 testXD.py
cleaning up dust 3

@@ -436,7 +436,7 @@ def dustCorrectionPrior(tgasCutMatched, dataFilename, quantile=0.05, nDistanceSa
         print 'calculating dust corrections, this may take awhile'
         try:
             data = np.load('distanceQuantiles'+dataFilename)
-            distanceQuntile = data['distanceQuntile']
+            distanceQuntile = data['distanceQuantile']
             distanceQuantile50 = data['distanceQuantile50']
         except IOError:
             nstars = len(tgasCutMatched)

### 2017-02-24 testXD.py
cleaning up dust 4

@@ -436,7 +436,7 @@ def dustCorrectionPrior(tgasCutMatched, dataFilename, quantile=0.05, nDistanceSa
         print 'calculating dust corrections, this may take awhile'
         try:
             data = np.load('distanceQuantiles'+dataFilename)
-            distanceQuntile = data['distanceQuantile']
+            distanceQuantile = data['distanceQuantile']
             distanceQuantile50 = data['distanceQuantile50']
         except IOError:
             nstars = len(tgasCutMatched)

### 2017-02-25 paperDistance/paper.tex
limitations of Empirical Bayes

@@
 
+The Empirical Bayesian methodology is a good approximation to full
+hierarchical inference in the limit of large numbers of data points.
+That is, when no individual star is carrying a lot of weight in the
+model of the \cmd.
+With its millions of stars, the \tgas\ Catalog may appear to safely
+live in this limit.
+However, because there are parts of the \cmd\ that are poorly populated,
+even in \tgas, the Empirical Bayes approximation may be bad for some
+color ranges.
+In particular, there appear to be essentially no white dwarf stars,
+and other classes (like hot main-sequence stars) have few to no exemplars
+in the data set at high signal-to-noise (in parallax).
+For all these reasons, we do think that this approximation is not perfectly
+safe, and it is a goal to explore full hierarchical modeling in the coming
+years (and see, for example, CITE BORIS).
+
 What would be different if we wanted a model not of the \tgas\ data,

### 2017-02-26 testXD.py
dust keywords

@@ -117,7 +117,7 @@ def matrixize(data1, data2, err1, err2):
     Xerr = np.zeros(X.shape + X.shape[-1:])
     diag = np.arange(X.shape[-1])
     Xerr[:, diag, diag] = np.vstack([err1**2., err2**2.]).T
-    return X, Xerr
+    return X[0], Xerr[0]
 
 def optimize(X, Xerr, param_name='n_components', param_range=np.array([256, 512, 1024, 2048, 4096, 8182])):
     """
@@ -352,19 +352,18 @@ def distanceTest(tgasCutMatched, nPosteriorPoints, data1, data2, err1, err2, xli
         #for distances, only want positive parallaxes
         positive = xparallaxMAS > 0.
 
-        dimension = 0
         meanData, covData = matrixize(data1[index], data2[index], err1[index], err2[index])
         ndim = 2
 
         #calculate the posterior, a gaussian for each xdgmm component
-        allMeans, allAmps, allCov, summedPosterior[k,:] = absMagKindaPosterior(xdgmm, ndim, meanData[dimension], covData[dimension], xabsMagKinda)
+        allMeans, allAmps, allCov, summedPosterior[k,:] = absMagKindaPosterior(xdgmm, ndim, meanData, covData, xabsMagKinda)
 
         #for 2D visual of posterior, sample all the xdgmm component gaussians into 2d historgram
         if plot2DPost: xedges, yedges, Z, samplesY = posterior2d(allMeans, allAmps, allCov, xbins, ybins, nperGauss=nperGauss)
         else:
             X = Y = Z = samplesY = None
         #plot the posterior of each star in example.png and feed it the ax for all the stars to accumualte into one plot
-        plotPosteriorEachStar(allMeans, allCov, allAmps, summedPosterior[k,:], meanData[dimension], covData[dimension], xdgmm, xparallaxMAS, xabsMagKinda, apparentMagnitude, X=X, Y=Y, Z=Z, samplesY=samplesY, axAll=axDist, plot2D=plot2DPost)
+        plotPosteriorEachStar(allMeans, allCov, allAmps, summedPosterior[k,:], meanData, covData, xdgmm, xparallaxMAS, xabsMagKinda, apparentMagnitude, X=X, Y=Y, Z=Z, samplesY=samplesY, axAll=axDist, plot2D=plot2DPost)
         os.rename('example.png', 'example.' + str(k) + '.png')
 
[…]

### 2017-02-26 testXD.py
dust keywords

@@ -654,50 +654,58 @@ if __name__ == '__main__':
     else:
         indices = parallaxSNcut & lowPhotErrorcut & nonzeroError
 
-
     try:
-        xdgmm = XDGMM(filename=xdgmmFilename)
+        xdgmm = XDGMM(filename=xdgmmFilenameDust)
+        print 'dust corrected XD read in'
     except IOError:
-        if subset:
-            X, Xerr = subset(data1[indices], data2[indices], err1[indices], err2[indices], nsamples=1024)
-        else:
-            X, Xerr = matrixize(data1[indices], data2[indices], err1[indices], err2[indices])
-        xdgmm = XDGMM(method='Bovy')
-        xdgmm.n_components = ngauss
-        xdgmm = xdgmm.fit(X, Xerr)
-        xdgmm.save_model(xdgmmFilename)
+        try:
+            xdgmm = XDGMM(filename=xdgmmFilename)
+            print 'non corrected XD read in'
+        except IOError:
+            if subset:
+                X, Xerr = subset(data1[indices], data2[indices], err1[indices], err2[indices], nsamples=1024)
+            else:
+                X, Xerr = matrixize(data1[indices], data2[indices], err1[indices], err2[indices])
+            xdgmm = XDGMM(method='Bovy')
+            xdgmm.n_components = ngauss
+            xdgmm = xdgmm.fit(X, Xerr)
+            xdgmm.save_model(xdgmmFilename)
 
 
-    #plot XD prior
-    figPrior, axPrior = plt.subplots()
-    for gg in range(xdgmm.n_components):
-        points = drawEllipse.plotvector(xdgmm.mu[gg], xdgmm.V[gg])
-        axPrior.plot(points[0,:],absMagKinda2absMag(points[1,:]), 'r', lw=0.5, alpha=xdgmm.weights[gg]/np.max(xdgmm.weights))
-        axPrior.invert_yaxis()
-    figPrior.savefig('prior.png')
+        #plot XD prior
+        figPrior, axPrior = plt.subplots()
+        for gg in range(xdgmm.n_components):
+            points = drawEllipse.plotvector(xdgmm.mu[gg], xdgmm.V[gg])
+            axPrior.plot(points[0,:],absMagKinda2absMag(points[1,:]), 'r', lw=0.5, alpha=xdgmm.weights[gg]/np.max(xdgmm.weights))
[…]

### 2017-02-26 testXD.py
dust keywords

@@ -578,7 +578,7 @@ if __name__ == '__main__':
     optimize = False
     subset = False
     timing = False
-
+    dustCorrectedArraysGenerated = False
 
     try:
         cutMatchedArrays  = np.load(dataFilename)
@@ -694,6 +694,7 @@ if __name__ == '__main__':
         mag2DustCorrected   = dustCorrection(bandDictionary[mag2]['array']  [bandDictionary[mag2]['key']][indices], dustEBV, mag2)
         apparentMagDustCorrected = dustCorrection(bandDictionary[absmag]['array'][bandDictionary[absmag]['key']][indices], dustEBV, absmag)
         absMagDustCorrected = tgasCutMatched['parallax_error'][indices]*10.**(0.2*apparentMagDustCorrected)
+        dustCorrectedArraysGenerated = True
         #B_dustcorrected = dustCorrection(apassCutMatched['bmag'], bayesDust, 'B')
         #need to define color_err and absMagKinda_err when including dust correction
         colorDustCorrected = mag1DustCorrected - mag2DustCorrected
@@ -712,4 +713,14 @@ if __name__ == '__main__':
     #distanceTest(tgasCutMatched, nPosteriorPoints, data1, data2, err1, err2, xlim, ylim, plot2DPost=False)
 
     #calculate parallax-ish posterior for each star
+    if not dustCorrectedArraysGenerated:
+        dustEBV, sourceID = dustCorrectionPrior(tgasCutMatched, dataFilename, quantile=0.05, nDistanceSamples=128, max_samples=None)
+        mag1DustCorrected   = dustCorrection(bandDictionary[mag1]['array']  [bandDictionary[mag1]['key']][indices], dustEBV, mag1)
+        mag2DustCorrected   = dustCorrection(bandDictionary[mag2]['array']  [bandDictionary[mag2]['key']][indices], dustEBV, mag2)
+        apparentMagDustCorrected = dustCorrection(bandDictionary[absmag]['array'][bandDictionary[absmag]['key']][indices], dustEBV, absmag)
+        absMagDustCorrected = tgasCutMatched['parallax_error'][indices]*10.**(0.2*apparentMagDustCorrected)
+        #B_dustcorrected = dustCorrection(apassCutMatched['bmag'], bayesDust, 'B')
+        #need to define color_err and absMagKinda_err when including dust correction
+        colorDustCorrected = mag1DustCorrected - mag2DustCorrected
+
[…]

### 2017-02-26 testXD.py
cleaning up dust 4

@@ -703,16 +703,6 @@ if __name__ == '__main__':
         X, Xerr = matrixize(data1, data2, err1, err2)
         xdgmm.fit(X, Xerr)
         xdgmm.save_model(xdgmmFilenameDust)
-    sample = xdgmm.sample(Nsamples)
-    dp.plot_sample(data1[indices], absMagKinda2absMag(data2[indices]), data1[indices], absMagKinda2absMag(data2[indices]),
-                sample[:,0],absMagKinda2absMag(sample[:,1]),xdgmm, xerr=err1[indices], yerr=absMagKinda2absMag(err2[indices]), xlabel=xlabel, ylabel=ylabel)
-    os.rename('plot_sample.png', 'plot_sample_ngauss'+str(ngauss)+'.SN'+str(thresholdSN) + '.2Mass.dustCorrected.png')
-
-
-    #check it's working by inferring distances to M67
-    #distanceTest(tgasCutMatched, nPosteriorPoints, data1, data2, err1, err2, xlim, ylim, plot2DPost=False)
-
-    #calculate parallax-ish posterior for each star
     if not dustCorrectedArraysGenerated:
         dustEBV, sourceID = dustCorrectionPrior(tgasCutMatched, dataFilename, quantile=0.05, nDistanceSamples=128, max_samples=None)
         mag1DustCorrected   = dustCorrection(bandDictionary[mag1]['array']  [bandDictionary[mag1]['key']][indices], dustEBV, mag1)
@@ -722,5 +712,16 @@ if __name__ == '__main__':
         #B_dustcorrected = dustCorrection(apassCutMatched['bmag'], bayesDust, 'B')
         #need to define color_err and absMagKinda_err when including dust correction
         colorDustCorrected = mag1DustCorrected - mag2DustCorrected
+
+    sample = xdgmm.sample(Nsamples)
+    dp.plot_sample(colorDustCorrected, absMagDustCorrected, colorDustCorrected, absMagDustCorrected,
+                sample[:,0],absMagKinda2absMag(sample[:,1]),xdgmm, xerr=err1[indices], yerr=absMagKinda2absMag(err2[indices]), xlabel=xlabel, ylabel=ylabel)
+    os.rename('plot_sample.png', 'plot_sample_ngauss'+str(ngauss)+'.SN'+str(thresholdSN) + '.2Mass.dustCorrected.png')
+
+
+    #check it's working by inferring distances to M67
[…]

### 2017-02-26 testXD.py
parallax not parallax_error

@@ -117,7 +117,7 @@ def matrixize(data1, data2, err1, err2):
     Xerr = np.zeros(X.shape + X.shape[-1:])
     diag = np.arange(X.shape[-1])
     Xerr[:, diag, diag] = np.vstack([err1**2., err2**2.]).T
-    return X[0], Xerr[0]
+    return X, Xerr
 
 def optimize(X, Xerr, param_name='n_components', param_range=np.array([256, 512, 1024, 2048, 4096, 8182])):
     """
@@ -353,6 +353,8 @@ def distanceTest(tgasCutMatched, nPosteriorPoints, data1, data2, err1, err2, xli
         positive = xparallaxMAS > 0.
 
         meanData, covData = matrixize(data1[index], data2[index], err1[index], err2[index])
+        meanData = meanData[0]
+        covData = covData[0]
         ndim = 2
 
         #calculate the posterior, a gaussian for each xdgmm component
@@ -455,6 +457,8 @@ def dustCorrectionPrior(tgasCutMatched, dataFilename, quantile=0.05, nDistanceSa
 
                 #calculate parallax-ish posterior for each star
                 meanData, covData = matrixize(color[index], absMagKinda[index], color_err[index], absMagKinda_err[index])
+                meanData = meanData[0]
+                covData = covData[0]
                 windowFactor = 5. #the number of sigma to sample in mas for plotting
                 minParallaxMAS = tgasCutMatched['parallax'][index] - windowFactor*tgasCutMatched['parallax_error'][index]
                 maxParallaxMAS = tgasCutMatched['parallax'][index] + windowFactor*tgasCutMatched['parallax_error'][index]
@@ -539,6 +543,8 @@ def posteriorDistanceAllStars(tgasCutMatched, nPosteriorPoints, color, absMagKin
 
 
         meanData, covData = matrixize(color[index], absMagKinda[index], color_err[index], absMagKinda_err[index])
+        meanData = meanData[0]
+        covData = covData[0]
         windowFactor = 5. #the number of sigma to sample in mas for plotting
         minParallaxMAS = tgasCutMatched['parallax'][index] - windowFactor*tgasCutMatched['parallax_error'][index]
         maxParallaxMAS = tgasCutMatched['parallax'][index] + windowFactor*tgasCutMatched['parallax_error'][index]
[…]

### 2017-02-27 testXD.py
push to 1.2M

@@ -436,7 +436,7 @@ def dustCorrectionPrior(tgasCutMatched, dataFilename, quantile=0.05, nDistanceSa
     except IOError:
         print 'calculating dust corrections, this may take awhile'
         try:
-            data = np.load('distanceQuantiles'+dataFilename)
+            data = np.load('distanceQuantiles_'+dataFilename)
             distanceQuantile = data['distanceQuantile']
             distanceQuantile50 = data['distanceQuantile50']
         except IOError:
@@ -479,7 +479,7 @@ def dustCorrectionPrior(tgasCutMatched, dataFilename, quantile=0.05, nDistanceSa
                 #find the distance at the 5% quantile
                 distanceQuantile[i] = np.percentile(sampleDistance, quantile*100.)
                 distanceQuantile50[i] = np.percentile(sampleDistance, 0.5*100.)
-            np.savez('distanceQuantiles'+dataFilename, distanceQuantile=distanceQuantile, distanceQuantile50=distanceQuantile50)
+            np.savez('distanceQuantiles_'+dataFilename, distanceQuantile=distanceQuantile, distanceQuantile50=distanceQuantile50)
         sourceID = tgasCutMatched['source_id'][indices]
         l = tgasCutMatched['l'][indices]*units.deg
         b = tgasCutMatched['b'][indices]*units.deg
@@ -572,7 +572,7 @@ if __name__ == '__main__':
     ngauss = 128
     nstar = '1.2M'
     Nsamples = 120000
-    nPosteriorPoints = 10000
+    nPosteriorPoints = 1000
     projectedDimension = 1
     ndim = 2
     #dataFilename = 'cutMatchedArrays.SN0.001.npz'

### 2017-02-27 testXD.py
push to 1.2M

@@ -577,9 +577,10 @@ if __name__ == '__main__':
     ndim = 2
     #dataFilename = 'cutMatchedArrays.SN0.001.npz'
     #dataFilename = 'cutMatchedArrays.tgasApassSN0.npz'
-    dataFilename = 'cutMatchedArrays.1_20_100_4.npz'
-    xdgmmFilename = 'xdgmm.'+ str(ngauss) + 'gauss.'+nstar+ '.SN' + str(thresholdSN) + '.2MASS.fit'
-    xdgmmFilenameDust = 'xdgmm.'+ str(ngauss) + 'gauss.'+nstar+ '.SN' + str(thresholdSN) + '.2MASS.dustCorrected.fit'
+    #dataFilename = 'cutMatchedArrays.1_20_100_4.npz'
+    dataFilename = 'cutMatchedArrays.SN0.001.npz'
+    xdgmmFilename = 'xdgmm.'+ str(ngauss) + 'gauss.'+dataFilename+'.' + survey + '.fit'
+    xdgmmFilenameDust = 'xdgmm.'+ str(ngauss) + 'gauss.'+dataFilename+'.' + survey + '.dustCorrected.fit'
     useDust = False
     optimize = False
     subset = False

### 2017-02-27 testXD.py
push to 1.2M

@@ -691,7 +691,7 @@ if __name__ == '__main__':
         sample = xdgmm.sample(Nsamples)
         dp.plot_sample(data1[indices], absMagKinda2absMag(data2[indices]), data1[indices], absMagKinda2absMag(data2[indices]),
                     sample[:,0],absMagKinda2absMag(sample[:,1]),xdgmm, xerr=err1[indices], yerr=absMagKinda2absMag(err2[indices]), xlabel=xlabel, ylabel=ylabel)
-        os.rename('plot_sample.png', 'plot_sample_ngauss'+str(ngauss)+'.SN'+str(thresholdSN) + '.2Mass.png')
+        os.rename('plot_sample.png', 'prior.ngauss'+str(ngauss)+'.' + dataFilename + '.' + survey + '.png')
 
         dustEBV, sourceID = dustCorrectionPrior(tgasCutMatched, dataFilename, quantile=0.05, nDistanceSamples=128, max_samples=None)
 
@@ -726,7 +726,7 @@ if __name__ == '__main__':
     sample = xdgmm.sample(Nsamples)
     dp.plot_sample(colorDustCorrected, absMagKinda2absMag(absMagKindaDustCorrected), colorDustCorrected, absMagKinda2absMag(absMagKindaDustCorrected),
                 sample[:,0],absMagKinda2absMag(sample[:,1]),xdgmm, xerr=err1[indices], yerr=absMagKinda2absMag(err2[indices]), xlabel=xlabel, ylabel=ylabel)
-    os.rename('plot_sample.png', 'plot_sample_ngauss'+str(ngauss)+'.SN'+str(thresholdSN) + '.2Mass.dustCorrected.png')
+    os.rename('plot_sample.png', 'prior.ngauss'+str(ngauss)+'.' + dataFilename + '.' + survey + '.dustCorrected.png')
 
 
     #check it's working by inferring distances to M67

### 2017-02-27 testXD.py
trying 16 gaussians

@@ -427,16 +427,16 @@ def distanceTest(tgasCutMatched, nPosteriorPoints, data1, data2, err1, err2, xli
     figDist.savefig('distancesM67.png')
 
 def dustCorrectionPrior(tgasCutMatched, dataFilename, quantile=0.05, nDistanceSamples=512, max_samples = 2):
-    dustFile = 'dustCorrection_' + dataFilename
-
+    dustFile = 'dustCorrection_' + ngauss + '_' + dataFilename
+    distanceFile = 'distanceQuantiles_'ngauss + '_' +dataFilename
     try:
-        data = np.load(dustFile + '.npz')
+        data = np.load(dustFile)
         dustEBV = data['ebv']
         sourceID = data['sourceID']
     except IOError:
         print 'calculating dust corrections, this may take awhile'
         try:
-            data = np.load('distanceQuantiles_'+dataFilename)
+            data = np.load(distanceFile)
             distanceQuantile = data['distanceQuantile']
             distanceQuantile50 = data['distanceQuantile50']
         except IOError:
@@ -479,7 +479,7 @@ def dustCorrectionPrior(tgasCutMatched, dataFilename, quantile=0.05, nDistanceSa
                 #find the distance at the 5% quantile
                 distanceQuantile[i] = np.percentile(sampleDistance, quantile*100.)
                 distanceQuantile50[i] = np.percentile(sampleDistance, 0.5*100.)
-            np.savez('distanceQuantiles_'+dataFilename, distanceQuantile=distanceQuantile, distanceQuantile50=distanceQuantile50)
+            np.savez(distanceFile, distanceQuantile=distanceQuantile, distanceQuantile50=distanceQuantile50)
         sourceID = tgasCutMatched['source_id'][indices]
         l = tgasCutMatched['l'][indices]*units.deg
         b = tgasCutMatched['b'][indices]*units.deg
@@ -492,7 +492,7 @@ def dustCorrectionPrior(tgasCutMatched, dataFilename, quantile=0.05, nDistanceSa
         end = time.time()
         #print 'dust sampling ', str(nDistanceSamples), ' took ',str(end-start), ' seconds for index ', str(i)
 
-        np.savez('dustCorrection_' + dataFilename, ebv=dustEBV, sourceID=sourceID, ebv50=dustEBV50)
[…]

### 2017-02-27 testXD.py
trying 16 gaussians

@@ -428,7 +428,7 @@ def distanceTest(tgasCutMatched, nPosteriorPoints, data1, data2, err1, err2, xli
 
 def dustCorrectionPrior(tgasCutMatched, dataFilename, quantile=0.05, nDistanceSamples=512, max_samples = 2):
     dustFile = 'dustCorrection_' + ngauss + '_' + dataFilename
-    distanceFile = 'distanceQuantiles_'ngauss + '_' +dataFilename
+    distanceFile = 'distanceQuantiles_' + ngauss + '_' +dataFilename
     try:
         data = np.load(dustFile)
         dustEBV = data['ebv']

### 2017-02-27 testXD.py
dust corrections

@@ -447,10 +447,10 @@ def dustCorrectionPrior(tgasCutMatched, dataFilename, quantile=0.05, nDistanceSa
             distanceQuantile = np.zeros(nstars)
             distanceQuantile50 = np.zeros(nstars)
             start = time.time()
-            for i, index in enumerate(np.where(indices)[0]):
-                if np.mod(i, 1000) == 0.0:
+            for index in range(nstars):
+                if np.mod(index, 1000) == 0.0:
                     end = time.time()
-                    print i, ' took ', str(end - start), 'seconds, projecting will be ', str((end-start)*(nstars/1000.))
+                    print index, ' took ', str(end - start), 'seconds, projecting will be ', str((end-start)*(nstars/1000.))
                     start = time.time()
 
                 #np.savez('dustCorrection_' + dataFilename, ebv=dustEBV, sourceID=sourceID)
@@ -477,12 +477,12 @@ def dustCorrectionPrior(tgasCutMatched, dataFilename, quantile=0.05, nDistanceSa
                 sampleDistance = samples(distance[::-1], posteriorDistance[::-1], nDistanceSamples, plot=False)
 
                 #find the distance at the 5% quantile
-                distanceQuantile[i] = np.percentile(sampleDistance, quantile*100.)
-                distanceQuantile50[i] = np.percentile(sampleDistance, 0.5*100.)
+                distanceQuantile[index] = np.percentile(sampleDistance, quantile*100.)
+                distanceQuantile50[index] = np.percentile(sampleDistance, 0.5*100.)
             np.savez(distanceFile, distanceQuantile=distanceQuantile, distanceQuantile50=distanceQuantile50)
-        sourceID = tgasCutMatched['source_id'][indices]
-        l = tgasCutMatched['l'][indices]*units.deg
-        b = tgasCutMatched['b'][indices]*units.deg
+        sourceID = tgasCutMatched['source_id']
+        l = tgasCutMatched['l']*units.deg
+        b = tgasCutMatched['b']*units.deg
         start = time.time()
         dustEBV, dustEBV50 = st.dust([l,l], [b,b], [distanceQuantile*units.kpc, distanceQuantile50*units.kpc], mode='median')
         nan = np.isnan(dustEBV)
[…]

### 2017-02-27 testXD.py
dust corrections

@@ -427,8 +427,8 @@ def distanceTest(tgasCutMatched, nPosteriorPoints, data1, data2, err1, err2, xli
     figDist.savefig('distancesM67.png')
 
 def dustCorrectionPrior(tgasCutMatched, dataFilename, quantile=0.05, nDistanceSamples=512, max_samples = 2):
-    dustFile = 'dustCorrection_' + ngauss + '_' + dataFilename
-    distanceFile = 'distanceQuantiles_' + ngauss + '_' +dataFilename
+    dustFile = 'dustCorrection_' + str(ngauss) + '_' + dataFilename
+    distanceFile = 'distanceQuantiles_' + str(ngauss) + '_' +dataFilename
     try:
         data = np.load(dustFile)
         dustEBV = data['ebv']
@@ -535,7 +535,7 @@ def posteriorDistanceAllStars(tgasCutMatched, nPosteriorPoints, color, absMagKin
     distancePosterior = np.zeros((nstars, nPosteriorPoints))
     colorDustCorrected = np.zeros(nstars)
     absMagDustCorrected = np.zeros(nstars)
-    posteriorFile = 'posteriorDistanceTgas_' + ngauss + '_' + dataFilename
+    posteriorFile = 'posteriorDistanceTgas_' + str(ngauss) + '_' + dataFilename
 
     for i, index in enumerate(np.where(indices)[0]):
         if np.mod(i, 1000) == 0.0:

### 2017-02-27 testXD.py
fixed indexing error

@@ -708,7 +708,7 @@ if __name__ == '__main__':
         colorDustCorrected = mag1DustCorrected - mag2DustCorrected
         data1 = colorDustCorrected
         data2 = absMagKindaDustCorrected
-        X, Xerr = matrixize(data1, data2, err1, err2)
+        X, Xerr = matrixize(data1, data2, err1[indices], err2[indices])
         xdgmm.fit(X, Xerr)
         xdgmm.save_model(xdgmmFilenameDust)
     if not dustCorrectedArraysGenerated:

### 2017-02-27 testXD.py
fixed indexing error

@@ -535,7 +535,7 @@ def posteriorDistanceAllStars(tgasCutMatched, nPosteriorPoints, color, absMagKin
     distancePosterior = np.zeros((nstars, nPosteriorPoints))
     colorDustCorrected = np.zeros(nstars)
     absMagDustCorrected = np.zeros(nstars)
-    posteriorFile = 'posteriorDistanceTgas_' + ngauss + '_' + dataFilename
+    posteriorFile = 'posteriorDistanceTgas_' + str(ngauss) + '_' + dataFilename
     nstars = len(tgasCutMatched)
     for index in range(nstars):
         if np.mod(index, 1000) == 0.0:

### 2017-02-27 testXD.py
random sample

@@ -693,7 +693,7 @@ if __name__ == '__main__':
                     sample[:,0],absMagKinda2absMag(sample[:,1]),xdgmm, xerr=err1[indices], yerr=absMagKinda2absMag(err2[indices]), xlabel=xlabel, ylabel=ylabel)
         os.rename('plot_sample.png', 'prior.ngauss'+str(ngauss)+'.' + dataFilename + '.' + survey + '.png')
 
-        dustEBV, sourceID = dustCorrectionPrior(tgasCutMatched[indices], dataFilename, quantile=0.05, nDistanceSamples=128, max_samples=None)
+        dustEBV, sourceID = dustCorrectionPrior(tgasCutMatched[indices], dataFilename, quantile=0.05, nDistanceSamples=128, max_samples=1, mode='random_sample')
 
         assert np.sum(tgasCutMatched['source_id'][indices] - sourceID) == 0.0, 'dust and data arrays are sorted differently !!!'
 
@@ -712,7 +712,7 @@ if __name__ == '__main__':
         xdgmm.fit(X, Xerr)
         xdgmm.save_model(xdgmmFilenameDust)
     if not dustCorrectedArraysGenerated:
-        dustEBV, sourceID  = dustCorrectionPrior(tgasCutMatched[indices], dataFilename, quantile=0.05, nDistanceSamples=128, max_samples=None)
+        dustEBV, sourceID  = dustCorrectionPrior(tgasCutMatched[indices], dataFilename, quantile=0.05, nDistanceSamples=128, max_samples=1, mode='random_sample')
         mag1DustCorrected   = dustCorrection(bandDictionary[mag1]['array']  [bandDictionary[mag1]['key']][indices], dustEBV, mag1)
         mag2DustCorrected   = dustCorrection(bandDictionary[mag2]['array']  [bandDictionary[mag2]['key']][indices], dustEBV, mag2)
         apparentMagnitude = bandDictionary[absmag]['array'][bandDictionary[absmag]['key']][indices]

### 2017-02-27 testXD.py
random sample

@@ -426,7 +426,7 @@ def distanceTest(tgasCutMatched, nPosteriorPoints, data1, data2, err1, err2, xli
     plt.tight_layout()
     figDist.savefig('distancesM67.png')
 
-def dustCorrectionPrior(tgasCutMatched, dataFilename, quantile=0.05, nDistanceSamples=512, max_samples = 2):
+def dustCorrectionPrior(tgasCutMatched, dataFilename, quantile=0.05, nDistanceSamples=512, max_samples = 2, plot=False, mode='median'):
     dustFile = 'dustCorrection_' + str(ngauss) + '_' + dataFilename
     distanceFile = 'distanceQuantiles_' + str(ngauss) + '_' +dataFilename
     try:
@@ -484,7 +484,7 @@ def dustCorrectionPrior(tgasCutMatched, dataFilename, quantile=0.05, nDistanceSa
         l = tgasCutMatched['l']*units.deg
         b = tgasCutMatched['b']*units.deg
         start = time.time()
-        dustEBV, dustEBV50 = st.dust([l,l], [b,b], [distanceQuantile*units.kpc, distanceQuantile50*units.kpc], mode='median')
+        dustEBV, dustEBV50 = st.dust([l,l], [b,b], [distanceQuantile*units.kpc, distanceQuantile50*units.kpc], mode=mode)
         nan = np.isnan(dustEBV)
         dustNan = st.dust(l[nan], b[nan], None, model='sfd')
         dustEBV[nan] = dustNan
@@ -493,6 +493,21 @@ def dustCorrectionPrior(tgasCutMatched, dataFilename, quantile=0.05, nDistanceSa
         #print 'dust sampling ', str(nDistanceSamples), ' took ',str(end-start), ' seconds for index ', str(i)
 
         np.savez(dustFile, ebv=dustEBV, sourceID=sourceID, ebv50=dustEBV50)
+    if plot:
+        figDust, axDust = plt.subplots(2)
+        from matplotlib.colors import LogNorm
+        ax[0].hist2d(np.log10(distanceQuantile), np.log10(dustEBV), bins=100, norm=LogNorm())
+        ax[1].hist2d( np.log10(distanceQuantile50), np.log10(dustEBV50), bins=100, norm=LogNorm())
+        ax[0].colorbar()
+        ax[1].colorbar()
+        ax[0].set_title('Dust for 0.05 quantile distance')
+        ax[1].set_title('Dust for 0.5 quantile distance')
[…]

### 2017-02-27 testXD.py
random sample

@@ -496,16 +496,16 @@ def dustCorrectionPrior(tgasCutMatched, dataFilename, quantile=0.05, nDistanceSa
     if plot:
         figDust, axDust = plt.subplots(2)
         from matplotlib.colors import LogNorm
-        ax[0].hist2d(np.log10(distanceQuantile), np.log10(dustEBV), bins=100, norm=LogNorm())
-        ax[1].hist2d( np.log10(distanceQuantile50), np.log10(dustEBV50), bins=100, norm=LogNorm())
-        ax[0].colorbar()
-        ax[1].colorbar()
-        ax[0].set_title('Dust for 0.05 quantile distance')
-        ax[1].set_title('Dust for 0.5 quantile distance')
-        ax[0].set_xlabel('log Distance [kpc]')
-        ax[1].set_xlabel('log Distance [kpc]')
-        ax[0].set_ylabel('E(B-V)')
-        ax[1].set_ylabel('E(B-V)')
+        axDust[0].hist2d(np.log10(distanceQuantile), np.log10(dustEBV), bins=100, norm=LogNorm())
+        axDust[1].hist2d( np.log10(distanceQuantile50), np.log10(dustEBV50), bins=100, norm=LogNorm())
+        axDust[0].colorbar()
+        axDust[1].colorbar()
+        #axDust[0].set_title('Dust for 0.05 quantile distance')
+        #axDust[1].set_title('Dust for 0.5 quantile distance')
+        axDust[0].set_xlabel('log Distance [kpc]')
+        axDust[1].set_xlabel('log Distance [kpc]')
+        axDust[0].set_ylabel('E(B-V) 5% quantile')
+        axDust[1].set_ylabel('E(B-V) 50% quantile')
         figDust.save('ebvDistribution.png')
 
     return dustEBV, sourceID

### 2017-02-27 testXD.py
random sample

@@ -494,6 +494,10 @@ def dustCorrectionPrior(tgasCutMatched, dataFilename, quantile=0.05, nDistanceSa
 
         np.savez(dustFile, ebv=dustEBV, sourceID=sourceID, ebv50=dustEBV50)
     if plot:
+        data = np.load(distanceFile)
+        distanceQuantile = data['distanceQuantile']
+        distanceQuantile50 = data['distanceQuantile50']
+
         figDust, axDust = plt.subplots(2)
         from matplotlib.colors import LogNorm
         axDust[0].hist2d(np.log10(distanceQuantile), np.log10(dustEBV), bins=100, norm=LogNorm())

### 2017-02-27 testXD.py
random sample

@@ -432,6 +432,7 @@ def dustCorrectionPrior(tgasCutMatched, dataFilename, quantile=0.05, nDistanceSa
     try:
         data = np.load(dustFile)
         dustEBV = data['ebv']
+        dustEBV50 = data['ebv50']
         sourceID = data['sourceID']
     except IOError:
         print 'calculating dust corrections, this may take awhile'

### 2017-02-27 testXD.py
random sample

@@ -503,8 +503,8 @@ def dustCorrectionPrior(tgasCutMatched, dataFilename, quantile=0.05, nDistanceSa
         from matplotlib.colors import LogNorm
         axDust[0].hist2d(np.log10(distanceQuantile), np.log10(dustEBV), bins=100, norm=LogNorm())
         axDust[1].hist2d( np.log10(distanceQuantile50), np.log10(dustEBV50), bins=100, norm=LogNorm())
-        axDust[0].colorbar()
-        axDust[1].colorbar()
+        figDust.colorbar()
+
         #axDust[0].set_title('Dust for 0.05 quantile distance')
         #axDust[1].set_title('Dust for 0.5 quantile distance')
         axDust[0].set_xlabel('log Distance [kpc]')

### 2017-02-27 testXD.py
random sample

@@ -503,7 +503,7 @@ def dustCorrectionPrior(tgasCutMatched, dataFilename, quantile=0.05, nDistanceSa
         from matplotlib.colors import LogNorm
         axDust[0].hist2d(np.log10(distanceQuantile), np.log10(dustEBV), bins=100, norm=LogNorm())
         axDust[1].hist2d( np.log10(distanceQuantile50), np.log10(dustEBV50), bins=100, norm=LogNorm())
-        figDust.colorbar()
+        #figDust.colorbar()
 
         #axDust[0].set_title('Dust for 0.05 quantile distance')
         #axDust[1].set_title('Dust for 0.5 quantile distance')

### 2017-02-27 testXD.py
random sample

@@ -511,7 +511,7 @@ def dustCorrectionPrior(tgasCutMatched, dataFilename, quantile=0.05, nDistanceSa
         axDust[1].set_xlabel('log Distance [kpc]')
         axDust[0].set_ylabel('E(B-V) 5% quantile')
         axDust[1].set_ylabel('E(B-V) 50% quantile')
-        figDust.save('ebvDistribution.png')
+        figDust.savefig('ebvDistribution.png')
 
     return dustEBV, sourceID
 

### 2017-02-28 paperDistance/paper.tex

@@
 % # To-do list
+% - We need a standard name for the \tgas--\tmass\ Intersection.
 % - zeroth draft the introduction
@@
 
+\newcommand{\foreign}[1]{\textsl{#1}}
 \newcommand{\acronym}[1]{{\small{#1}}}
@@
 models, nor models of the Milky Way.
-We obtained these by building a data-driven model of the
+That is, the geometric parallaxes can calibrate a set of photometric
+models that are purely statistical---a model of the data rather than
+a model of stars \foreign{per se}.
+This opens up the possibility of completing the goals of the \gaia\ Mission
+without building in unnecessary assumptions about the mechanical properties
+of stars or the Galaxy.
+
+We obtained the photometric parallaxes in this project
+by building a data-driven model of the
 color--magnitude diagram (\cmd) of the stars in the \tgas\ data set,
@@
 
-What would be different if we wanted a model not of the \tgas\ data,
-but a model appropriate for the full billion stars, or for a
-volume-limited sample of stars?
-How could we convert our existing results?
-How would we do it properly in general?
+In some sense, the biggest limitation of the data-driven \cmd\ model
+created in this project is that we made no use of any kind of \tgas\ selection
+function or completeness estimate or inverse-selection-volume corrections.
+That is, the model is a model of the \tgas\ Catalog
+(or really the \tgas--\tmass\ intersection), not of the stars in
+the Milky Way, nor any volume-limited subsample of the Milky Way.
+Importantly, because we used \emph{all} of the stars in the \tgas--\tmass\
+intersection, the model \emph{is} representive of the full catalog,
+even the distant parts, where all the parallaxes are measured at low
+signal-to-noise.
+This is in contrast to techniques that might build the model from only
+the high signal-to-noise sources; such a model would be biased towards
+stellar types and compositions found locally, lower-luminosity stars,
+and stars which happened to get good parallaxes.
+However, our use of the entire \tgas--\tmass\ intersection without
+any selection function restricts the use of our \cmd\ model as a prior
+to inferences within that same \tgas--\tmass\ intersection.
+For example, it would be biasing to use this same prior to infer
+distances for the full billion-star catalog released in \gaia\ DR1
+alongside \tgas, or for stars fainter or brighter or bluer than those
+included in the \tgas--\tmass\ intersection.
+
+Note the dangers of using the stars from our output in hierarchical inferences!
 

### 2017-03-01 paperDistance/paper.tex
fiddling with dust

@@
 \newcommand{\cmd}{\acronym{CMD}}
+
+\usepackage{amsmath}
 
@@
 
+The data are a transformation of the dust corrected absolute magnitude, and the dust corrected colors.
+
+\begin{equation}
+\label{eq:data}
+\begin{aligned}
+\mathrm{data vector} \;\;\;\; &\vec{y} = [10^{0.2M_{J,n}}, J_n - K_n] \;  n \in \mathrm{stars} \\
+\mathrm{with \; absolute \; magnitude} \;\;\;\; &M_{J,n} = J_n - DM_n - Q_JE(B-V)_n \\
+& DM_n = -5\mathrm{log_{10}}\frac{1}{\varpi} \\
+\mathrm{and \; color} \;\;\;\; &J_n - K_n = J_n - K_n - Q_{JK}E(B-V)_n
+
+\end{aligned}
+\end{equation}
+
+We fit the data using XD which maximizes the likelihood
+
+\begin{equation}
+\label{eq:xdlikelihood}
+P(\vec{y}|\alpha, I)
+\end{equation}
+
+with the output mixture of gaussians
+
+\begin{equation}
+\label{eq:xd}
+\begin{aligned}
+\alpha = &\left\{\mathrm{amp}_k, \mu_k, V_k\right\}_{k=1}^K \; k \in \mathrm{Gaussians} \\
+I = &\left\{\left\{\sigma^2_n\right\}_{n=1}^N, \mathrm{assumptions}\right\}
+\end{aligned}
+\end{equation}
+
 Crossvalidation

### 2017-03-01 stellarTwins.py

@@ -12,6 +12,9 @@ from matplotlib.patches import Ellipse
 import pdb
 from dustmaps.sfd import SFDQuery
 from dustmaps.bayestar import BayestarQuery
+from dustmaps.iphas import IPHASQuery
+from dustmaps.marshall import MarshallQuery
+from dustmaps.chen2014 import Chen2014Query
 from astropy.coordinates import SkyCoord
 import astropy.units as units
 from scipy.integrate import cumtrapz
@@ -425,7 +428,26 @@ def dust(l, b, distance, plot=False, max_samples=2, mode='median', model='bayes'
                 frame='galactic')
         bayes = BayestarQuery(max_samples=max_samples)
         dust = bayes(c, mode=mode)
+    if model == 'iphas':
+        c = SkyCoord(l, b,
+                distance = distance,
+                frame='galactic')
+        iphas = IPHASQuery()
+        dust = iphas(c, mode=mode)
 
+    if model == 'marshall':
+        c = SkyCoord(l, b,
+                distance = distance,
+                frame='galactic')
+        marshall = MarshallQuery()
+        dust = marshall(c)
+
+    if model == 'chen':
+        c = SkyCoord(l, b,
+                distance = distance,
+                frame='galactic')
+        chen = Chen2014Query()
+        dust = chen(c)
     #cNoDist = SkyCoord(l, b,
     #        frame='galactic')
     #bayesDustNoDist = bayes(cNoDist, mode=mode)

### 2017-03-01 testXD.py

@@ -19,6 +19,7 @@ from astropy.coordinates import SkyCoord
 import astropy.units as units
 import scipy.integrate
 import time
+from matplotlib.colors import LogNorm
 
 def convert2gal(ra, dec):
     """
@@ -486,8 +487,32 @@ def dustCorrectionPrior(tgasCutMatched, dataFilename, quantile=0.05, nDistanceSa
         b = tgasCutMatched['b']*units.deg
         start = time.time()
         dustEBV, dustEBV50 = st.dust([l,l], [b,b], [distanceQuantile*units.kpc, distanceQuantile50*units.kpc], mode=mode)
+        for model in ['marshall', 'iphas', 'chen']:
+            nan = np.isnan(dustEBV)
+            firstGo =  np.sum(nan)
+            dustEBVNan, dustEBV50Nan = st.dust([l[nan],l[nan]], [b[nan],b[nan]], [distanceQuantile[nan]*units.kpc, distanceQuantile50[nan]*units.kpc], model=model, mode=mode)
+            dustEBV[nan] = dustEBVNan
+            dustEBV50[nan] = dustEBV50Nan
+            nan = np.isnan(dustEBV)
+            secondGo = np.sum(nan)
+            print model, ' filled in ', str(firstGo - secondGo), ' still ', str(secondGo), ' Nans'
+
+        """
+        dustEBVNan, dustEBV50Nan = st.dust([l[nan],l[nan]], [b[nan],b[nan]], [distanceQuantile[nan]*units.kpc, distanceQuantile50[nan]*units.kpc], model='iphas', mode=mode)
+        dustEBV[nan] = dustEBVNan
+        dustEBV50[nan] = dustEBV50Nan
         nan = np.isnan(dustEBV)
+        thirdGo = np.sum(nan)
+        print 'iphas filled in ', str(secondGo - thirdGo), ' still ', str(thirdGo), ' Nans'
+        """
+
         dustNan = st.dust(l[nan], b[nan], None, model='sfd')
+        figSFD, axSFD = plt.subplots()
+        nonzero = dustNan > 0.
+        axSFD.hist2d(color[indices][nan][nonzero], np.log10(dustNan[nonzero]), bins=100, norm=LogNorm(), cmap='Greys')
+        axSFD.set_xlabel('J-K')
+        axSFD.set_ylabel('log E(B-V)')
+        figSFD.savefig('dustsfd.png')
         dustEBV[nan] = dustNan
         dustEBV50[nan] = dustNan
         end = time.time()
[…]

### 2017-03-01 testXD.py
only good dust

@@ -20,6 +20,7 @@ import astropy.units as units
 import scipy.integrate
 import time
 from matplotlib.colors import LogNorm
+from dustmaps.bayestar import BayestarQuery
 
 def convert2gal(ra, dec):
     """
@@ -487,6 +488,7 @@ def dustCorrectionPrior(tgasCutMatched, dataFilename, quantile=0.05, nDistanceSa
         b = tgasCutMatched['b']*units.deg
         start = time.time()
         dustEBV, dustEBV50 = st.dust([l,l], [b,b], [distanceQuantile*units.kpc, distanceQuantile50*units.kpc], mode=mode)
+        """
         for model in ['marshall', 'iphas', 'chen']:
             nan = np.isnan(dustEBV)
             firstGo =  np.sum(nan)
@@ -496,15 +498,13 @@ def dustCorrectionPrior(tgasCutMatched, dataFilename, quantile=0.05, nDistanceSa
             nan = np.isnan(dustEBV)
             secondGo = np.sum(nan)
             print model, ' filled in ', str(firstGo - secondGo), ' still ', str(secondGo), ' Nans'
-
-        """
         dustEBVNan, dustEBV50Nan = st.dust([l[nan],l[nan]], [b[nan],b[nan]], [distanceQuantile[nan]*units.kpc, distanceQuantile50[nan]*units.kpc], model='iphas', mode=mode)
         dustEBV[nan] = dustEBVNan
         dustEBV50[nan] = dustEBV50Nan
         nan = np.isnan(dustEBV)
         thirdGo = np.sum(nan)
         print 'iphas filled in ', str(secondGo - thirdGo), ' still ', str(thirdGo), ' Nans'
-        """
+
 
         dustNan = st.dust(l[nan], b[nan], None, model='sfd')
         figSFD, axSFD = plt.subplots()
@@ -515,10 +515,12 @@ def dustCorrectionPrior(tgasCutMatched, dataFilename, quantile=0.05, nDistanceSa
         figSFD.savefig('dustsfd.png')
         dustEBV[nan] = dustNan
         dustEBV50[nan] = dustNan
+        """
         end = time.time()
         #print 'dust sampling ', str(nDistanceSamples), ' took ',str(end-start), ' seconds for index ', str(i)
-
-        np.savez(dustFile, ebv=dustEBV, sourceID=sourceID, ebv50=dustEBV50)
+        print 'calculating dust took ', str(end - start), ' seconds'
[…]

### 2017-03-07 testXD.py
iterating dust

@@ -428,7 +428,7 @@ def distanceTest(tgasCutMatched, nPosteriorPoints, data1, data2, err1, err2, xli
     plt.tight_layout()
     figDist.savefig('distancesM67.png')
 
-def distanceQuantile(color, absMagKinda, color_err, absMagKinda_err, tgasCutMatched, xdgmm, distanceFile='distance.npz', quantile=0.05):
+def distanceQuantile(color, absMagKinda, color_err, absMagKinda_err, tgasCutMatched, xdgmm, distanceFile='distance.npz', quantile=0.05, nDistanceSamples=512):
     try:
         data = np.load(distanceFile)
         distanceQuantile = data['distanceQuantile']
@@ -751,7 +751,7 @@ if __name__ == '__main__':
         xdgmmFilename = 'xdgmm.'             + str(ngauss) + 'gauss.' + iter + '.' + survey + '.' + dataFilename + '.fit'
         distanceFile  = 'distanceQuantiles.' + str(ngauss) + 'gauss.' + iter + '.' + survey + '.' + dataFilename
         dustFile      = 'dustCorrection.'    + str(ngauss) + 'gauss.' + iter + '.' + survey + '.' + dataFilename
-        priorFile     = 'prior.'             + str(ngauss) + 'gauss. '+ iter + '.' + survey + '.' + dataFilename + '.png'
+        priorFile     = 'prior.'             + str(ngauss) + 'gauss.' + iter + '.' + survey + '.' + dataFilename + '.png'
 
         color = colorArray(mag1, mag2, dustEBV, bandDictionary)
         absMagKinda = absMagKindaArray(absmag, dustEBV, bandDictionary)
@@ -803,7 +803,7 @@ if __name__ == '__main__':
         os.rename('plot_sample.png', priorFile)
 
         #using prior calculate distances
-        distance, distanceMedian = distanceQuantile(color, absMagKinda, color_err, absMagKinda_err, tgasCutMatched, xdgmm, distanceFile=distanceFile, quantile=0.05)
+        distance, distanceMedian = distanceQuantile(color, absMagKinda, color_err, absMagKinda_err, tgasCutMatched, xdgmm, distanceFile=distanceFile, quantile=0.05, nDistanceSamples=128)
 
         #using distance, calculate dust
         try:

### 2017-03-07 testXD.py
distance check

@@ -430,7 +430,7 @@ def distanceTest(tgasCutMatched, nPosteriorPoints, data1, data2, err1, err2, xli
 
 def distanceQuantile(color, absMagKinda, color_err, absMagKinda_err, tgasCutMatched, xdgmm, distanceFile='distance.npz', quantile=0.05, nDistanceSamples=512):
     try:
-        data = np.load(distanceFile)
+        data = np.load(distanceFile+'.txt')
         distanceQuantile = data['distanceQuantile']
         distanceMedian = data['distanceMedian']
         print 'distance file is: ', distanceFile
@@ -469,6 +469,7 @@ def distanceQuantile(color, absMagKinda, color_err, absMagKinda_err, tgasCutMatc
             #find the distance at the 5% quantile
             distanceQuantile[index] = np.percentile(sampleDistance, quantile*100.)
             distanceMedian[index] = np.percentile(sampleDistance, 0.5*100.)
+            pdb.set_trace()
         np.savez(distanceFile, distanceQuantile=distanceQuantile, distanceMedian=distanceMedian)
     return distanceQuantile, distanceMedian
 

### 2017-03-07 testXD.py
next iterations

@@ -428,9 +428,9 @@ def distanceTest(tgasCutMatched, nPosteriorPoints, data1, data2, err1, err2, xli
     plt.tight_layout()
     figDist.savefig('distancesM67.png')
 
-def distanceQuantile(color, absMagKinda, color_err, absMagKinda_err, tgasCutMatched, xdgmm, distanceFile='distance.npz', quantile=0.05, nDistanceSamples=512):
+def distanceQuantile(color, absMagKinda, color_err, absMagKinda_err, tgasCutMatched, xdgmm, distanceFile='distance.npz', quantile=0.05, nDistanceSamples=512, nPosteriorPoints=1000):
     try:
-        data = np.load(distanceFile+'.txt')
+        data = np.load(distanceFile)
         distanceQuantile = data['distanceQuantile']
         distanceMedian = data['distanceMedian']
         print 'distance file is: ', distanceFile
@@ -463,13 +463,12 @@ def distanceQuantile(color, absMagKinda, color_err, absMagKinda_err, tgasCutMatc
             #normalize prior pdf
             posteriorDistance = summedPosteriorAbsmagKinda[positive]*xparallaxMAS[positive]**2.*10.**(0.2*apparentMagnitude)
             distance = 1./xparallaxMAS[positive]
-            #sample the PDF nDistanceSamples
-            sampleDistance = samples(distance[::-1], posteriorDistance[::-1], nDistanceSamples, plot=False)
 
-            #find the distance at the 5% quantile
-            distanceQuantile[index] = np.percentile(sampleDistance, quantile*100.)
-            distanceMedian[index] = np.percentile(sampleDistance, 0.5*100.)
-            pdb.set_trace()
+            cdf = scipy.integrate.cumtrapz(posteriorDistance[::-1], x=distance[::-1])
+            cdfInv = scipy.interpolate.interp1d(cdf, distance[::-1][:-1])
+            distanceQuantile[index] = cdfInv(quantile)
+            distanceMedian[index] = cdfInv(0.5)
+
         np.savez(distanceFile, distanceQuantile=distanceQuantile, distanceMedian=distanceMedian)
     return distanceQuantile, distanceMedian
 
@@ -744,15 +743,23 @@ if __name__ == '__main__':
[…]

### 2017-03-08 testXD.py
dust stuff

@@ -443,6 +443,8 @@ def distanceQuantile(color, absMagKinda, color_err, absMagKinda_err, tgasCutMatc
         distanceQuantile = np.zeros(nstars)
         distanceMedian = np.zeros(nstars)
         start = time.time()
+
+        fig, ax = plt.subplots()
         for index in range(nstars):
             if np.mod(index, 10000) == 0.0:
                 end = time.time()
@@ -466,9 +468,18 @@ def distanceQuantile(color, absMagKinda, color_err, absMagKinda_err, tgasCutMatc
 
             cdf = scipy.integrate.cumtrapz(posteriorDistance[::-1], x=distance[::-1])
             cdfInv = scipy.interpolate.interp1d(cdf, distance[::-1][:-1])
-            distanceQuantile[index] = cdfInv(quantile)
-            distanceMedian[index] = cdfInv(0.5)
 
+            try:
+                distanceQuantile[index] = cdfInv(quantile)
+                distanceMedian[index] = cdfInv(0.5)
+            except ValueError: 
+                print np.max(cdf)
+                plt.cla()
+                ax.plot(distance[1:], cdf)
+                ax.set_xlabel('distance')
+                ax.set_ylabel('cdf')
+                fig.savefig('cdfplots/cdf.' + str(index) + '.png')
+                #distanceMedian[index] = np.nan
         np.savez(distanceFile, distanceQuantile=distanceQuantile, distanceMedian=distanceMedian)
     return distanceQuantile, distanceMedian
 

### 2017-03-08 paperDistance/paper.tex

@@
 
-We build a prior for the distance measurement of Gaia stars using the Gaia measured parallaxes plus APASS photometry to build a color-magnitude-like diagram that serves as the prior for distance measurements of Gaia stars.
-Instead of making assumptions about the physical distributions of stars in the galaxy, or using stellar models for the modeled HR diagram of stars, we build a prior from the data themselves.
-
 LA: Insert itemized or enumerated list of assumptions here. DWH will
@@
 
-magSN = parallaxSN = 16 (excluding systematic parallax error in tgas)
-train gaussian means, mus, Vs on progressively lower SN dataset
+\begin{itemize}
+\item The photometric uncertainties in the J band are small compared with the uncertainties in the parallax
+\item Our empirical bayes approach is ok as long as we only infer information about the data contained within the prior
+\item the uncertainties are gaussian
+\item most stars lie where we expect them to on the CMD
+\item number of gaussians that's efficient? or do x-validation
+\item can't think
+\end{itemize}
+
+We would like to infer the distance to stars using the measured (noisy) parallax from Gaia.
+Gaia gives us the likelihood of the parallax of a star, given some unknown distance and uncertainties (assumed gaussian)
+\begin{equation}
+\label{eq:likelihood}
+P(\varpi |d, \sigma^2_{\varpi}, \textbf{I})
+\end{equation}
+
+where $d$ is distance to the star, $\varpi$ is the observed parallax, and $\sigma^2_{\varpi}$ is the uncertainty in that measurement, and I are other various prior assumptions (mostly about the noise properties?).
+This is the probability of the data given the model, or distance, but we want to flip it around and infer the distance to the star given the measured parallax and it's uncertainty.
+We do this using probability theory/bayes rule, and turn the likelihood of the data into the posterior probablity of the distance
 
@@
 \label{eq:bayes}
-P(\theta|\textbf{E}, \textbf{I}) = P(\theta ) \frac{P(\textbf{E} |\theta,\textbf{I})}{P(\textbf{E})},
+P(d|\varpi, \sigma^2_{\varpi}, \textbf{I}) = \frac{P(\varpi |d, \sigma^2_{\varpi}, \textbf{I}) P(d)}{\textbf{Z}},
 \end{equation}
 
-where $\theta$ is the model parameters that generate the evidence $\textbf{E}$, and P($\theta$) is the prior knowledge of the model, and $P(\textbf{E} |\theta,\textbf{I})$ is the likelihood of the evidence, given the model and some other prior assumptions.
-Our evidence is the parallaxes and G band magnitudes from Gaia, and the B and V band magnitudes from APASS.
-We build the prior from this evidence using extreme deconvolution
+where $P(d)$ is our prior knowledge of the distance to the star, and $\textbf{Z}$ is a normalization constant to make the posterior sum to 1.
+The strength in our inference method lies in how we generate our prior knowledge of the distances to stars.
+Here, instead of assuming the spatial distribution of stars in the milky way, or relying on stellar models, as previous work as shown, we build the color-magnitude diagram of the observed stars from their observations as our prior.
+The absolute magnitude of the CMD carries information about the distances to stars and where they can physically lie in this space.
+Toy model?
 
-The data are a transformation of the dust corrected absolute magnitude, and the dust corrected colors.
+To generate our CMD prior we use the J and K band from 2MASS, as well as the parallax from Gaia. The data are a transformation of the dust corrected absolute magnitude, and the dust corrected colors.
 
@@
 \begin{aligned}
-\mathrm{data vector} \;\;\;\; &\vec{y} = [10^{0.2M_{J,n}}, J_n - K_n] \;  n \in \mathrm{stars} \\
+\mathrm{data vector} \;\;\;\; &\vec{y} = [10^{0.2M_{J,n}}, (J- K)_n] \;  n \in \mathrm{stars} \\
 \mathrm{with \; absolute \; magnitude} \;\;\;\; &M_{J,n} = J_n - DM_n - Q_JE(B-V)_n \\
 & DM_n = -5\mathrm{log_{10}}\frac{1}{\varpi} \\
-\mathrm{and \; color} \;\;\;\; &J_n - K_n = J_n - K_n - Q_{JK}E(B-V)_n
+\mathrm{and \; color} \;\;\;\; &(J - K)_n = J_n - K_n - Q_{JK}E(B-V)_n
 
@@
 
-We fit the data using XD which maximizes the likelihood
+where $E(B-V)$ is the output of the Schlafly dust model, and $Q$ is the correction factor for that band assuming some reddening law.
+To generate our prior, we fit the data in the CMD space using XD (Bovy et al) which deconvolves the data with its noise, which are heterogeneous measurement uncertainties.
+It generates the denoised, underlying distribution from which the noisey data were drawn from.
+It represents the distribution as a mixture of gaussians which maximizes the likelihood of the noisy data.
 
@@
 
-Crossvalidation
+Crossvalidation?
 
@@
 
-We use stars cross matched in the TGAS and APASS set of observations.
-We require there are no NaNs in any APASS photometric band, as well as every band having positive errors.
-We check that the match between APASS and TGAS is good using extreme APASS-WISE colors.
-We correct for dust using 3D Bayestar (Schlafly+Finkbeiner) with an estimate of the distance from sampling of posterior Gaia distances from Adrain.
-There is minimal dust to the matched stars.
-To build our model here we use the Gaia G band magnitude, parallax and APASS B-V color.
+We use stars crossmatched in TGAS, 2MASS and within the observing footprint of Panstarrs to access the Schlafly 3D dust model.
+We require the signal to noise of both the photometry and parallax be $> 0.001$, that the photometry have nonzero errors, and remove a small selection of 2MASS stars that have zero color and zero J band magnitude.
+We correct for dust using 3D Bayestar (Schlafly+Finkbeiner) with an estimate of the distance from iteratively sampling our posterior distances, more below.
 
@@
 For each distance we take the median of the samples of the dust model.
-We apply each dust correction to our 2MASS photometry given by Equation \ref{eq:dust} where $K_{\lambda}$ = [3.626, 2.742, 3.303, 2.285, 1.698] for bands [B, V, g, r, i] respectively, taken from table (REF TABLE SFD), and $E(B-V)$ is the output of the Green et al dust models.
+We apply each dust correction to our 2MASS photometry given by Equation \ref{eq:dust} where $K_{\lambda}$ = [3.626, 2.742, 3.303, 2.285, 1.698, 0.709, 0.449, 0.302] for bands [B, V, g, r, i, J, H, K] respectively, taken from table (REF TABLE SFD), and $E(B-V)$ is the output of the Green et al dust models.
 We then regenerate our prior using the dust corrected photometry.
-We do not update the errors, but keep them the same as before the dust correction, which is wrong and under representing our errors but this leads again to a lighter deconvolution.
+We do not update the errors, but keep them the same as before the dust correction, which is wrong and under representing our errors but this leads again to a lighter deconvolution (Do we want to change this and add in the covariance for the dust?).
 Should I quote some statistics about the dust corrections?
@@
 \label{eq:dust}
-m_{\mathrm{band, corrected}} = m_{\mathrm{band}} - K_{\lambda} \times E(B-V)
+m_{\mathrm{band, corrected}} = m_{\mathrm{band}} - Q_{\lambda} \times E(B-V)
 \end{equation}

### 2017-03-09 testXD.py
fix distance grid

@@ -428,11 +428,10 @@ def distanceTest(tgasCutMatched, nPosteriorPoints, data1, data2, err1, err2, xli
     plt.tight_layout()
     figDist.savefig('distancesM67.png')
 
-def distanceQuantile(color, absMagKinda, color_err, absMagKinda_err, tgasCutMatched, xdgmm, distanceFile='distance.npz', quantile=0.05, nDistanceSamples=512, nPosteriorPoints=1000):
+def distanceQuantile(color, absMagKinda, color_err, absMagKinda_err, tgasCutMatched, xdgmm, distanceFile='distance.npy', quantile=0.05, nDistanceSamples=512, nPosteriorPoints=1000, iter='1st'):
     try:
         data = np.load(distanceFile)
-        distanceQuantile = data['distanceQuantile']
-        distanceMedian = data['distanceMedian']
+        distance = data['distance']
         print 'distance file is: ', distanceFile
     except IOError:
         print 'distance file does not exist: ', distanceFile
@@ -441,8 +440,10 @@ def distanceQuantile(color, absMagKinda, color_err, absMagKinda_err, tgasCutMatc
         #dustEBV = np.zeros(nstars)
         #dustEBV50 = np.zeros(nstars)
         distanceQuantile = np.zeros(nstars)
-        distanceMedian = np.zeros(nstars)
         start = time.time()
+        logDistance = np.linspace(-2, 1, nPosteriorPoints)
+        xparallaxMAS = 1./10.**logDistance
+        positive = xparallaxMAS > 0.
 
         fig, ax = plt.subplots()
         for index in range(nstars):
@@ -455,33 +456,92 @@ def distanceQuantile(color, absMagKinda, color_err, absMagKinda_err, tgasCutMatc
             meanData, covData = matrixize(color[index], absMagKinda[index], color_err[index], absMagKinda_err[index])
             meanData = meanData[0]
             covData = covData[0]
-            windowFactor = 5. #the number of sigma to sample in mas for plotting
-            minParallaxMAS = tgasCutMatched['parallax'][index] - windowFactor*tgasCutMatched['parallax_error'][index]
-            maxParallaxMAS = tgasCutMatched['parallax'][index] + windowFactor*tgasCutMatched['parallax_error'][index]
+            #windowFactor = 5. #the number of sigma to sample in mas for plotting
[…]

### 2017-03-09 testXD.py
distance stuff

@@ -428,7 +428,7 @@ def distanceTest(tgasCutMatched, nPosteriorPoints, data1, data2, err1, err2, xli
     plt.tight_layout()
     figDist.savefig('distancesM67.png')
 
-def distanceQuantile(color, absMagKinda, color_err, absMagKinda_err, tgasCutMatched, xdgmm, distanceFile='distance.npz', quantile=0.05, nDistanceSamples=512, nPosteriorPoints=1000):
+def distanceQuantile(color, absMagKinda, color_err, absMagKinda_err, tgasCutMatched, xdgmm, distanceFile='distance.npz', quantile=0.05, nDistanceSamples=512, nPosteriorPoints=1000, windowFactor=10.):
     try:
         data = np.load(distanceFile)
         distanceQuantile = data['distanceQuantile']
@@ -455,7 +455,6 @@ def distanceQuantile(color, absMagKinda, color_err, absMagKinda_err, tgasCutMatc
             meanData, covData = matrixize(color[index], absMagKinda[index], color_err[index], absMagKinda_err[index])
             meanData = meanData[0]
             covData = covData[0]
-            windowFactor = 5. #the number of sigma to sample in mas for plotting
             minParallaxMAS = tgasCutMatched['parallax'][index] - windowFactor*tgasCutMatched['parallax_error'][index]
             maxParallaxMAS = tgasCutMatched['parallax'][index] + windowFactor*tgasCutMatched['parallax_error'][index]
             apparentMagnitude = bandDictionary[absmag]['array'][bandDictionary[absmag]['key']][index]
@@ -822,13 +821,13 @@ if __name__ == '__main__':
         os.rename('plot_sample.png', priorFile)
 
         #using prior calculate distances
-        distance, distanceMedian = distanceQuantile(color, absMagKinda, color_err, absMagKinda_err, tgasCutMatched, xdgmm, distanceFile=distanceFile, quantile=0.05, nDistanceSamples=128, nPosteriorPoints=nPosteriorPoints)
+        distance, distanceMedian = distanceQuantile(color, absMagKinda, color_err, absMagKinda_err, tgasCutMatched, xdgmm, distanceFile=distanceFile, quantile=0.05, nDistanceSamples=128, nPosteriorPoints=nPosteriorPoints, windowFactor=15.)
 
         #using distance, calculate dust
         try:
             data = np.load(dustFile)
             dustEBVnew = data['ebv']
-            dustEBVMedianNew = data['ebv50']
[…]

### 2017-03-09 testXD.py
distance grid

@@ -9,7 +9,7 @@ import numpy as np
 import pdb
 import time
 from sklearn.model_selection import train_test_split
-
+from astropy.io import fits
 from xdgmm import XDGMM
 from sklearn.learning_curve import validation_curve
 from sklearn.cross_validation import ShuffleSplit
@@ -311,11 +311,11 @@ def plotPosteriorEachStar(meanPost, covPost, ampPost, summedPosterior, meanLike,
     fig.savefig('example.png')
 
 
-def distanceTest(tgasCutMatched, nPosteriorPoints, data1, data2, err1, err2, xlim, ylim, plot2DPost=False):
+def distanceTest(tgas, nPosteriorPoints, data1, data2, err1, err2, xlim, ylim, plot2DPost=False):
     """
     test posterior accuracy using distances to cluster M67
     """
-    indicesM67 = m67indices(tgasCutMatched, plot=False, db=0.5, dl=0.5)
+    indicesM67 = m67indices(tgas, plot=False, db=0.5, dl=0.5)
 
     #all the plot stuff
     figDist, axDist = plt.subplots(2, 3, figsize=(20, 15))
@@ -346,8 +346,8 @@ def distanceTest(tgasCutMatched, nPosteriorPoints, data1, data2, err1, err2, xli
 
         #plotting for each star in the cluster
         windowFactor = 5. #the number of sigma to sample in mas for plotting
-        minParallaxMAS = tgasCutMatched['parallax'][index] - windowFactor*tgasCutMatched['parallax_error'][index]
-        maxParallaxMAS = tgasCutMatched['parallax'][index] + windowFactor*tgasCutMatched['parallax_error'][index]
+        minParallaxMAS = tgas['parallax'][index] - windowFactor*tgas['parallax_error'][index]
+        maxParallaxMAS = tgas['parallax'][index] + windowFactor*tgas['parallax_error'][index]
         apparentMagnitude = bandDictionary[absmag]['array'][bandDictionary[absmag]['key']][index]
         xparallaxMAS, xabsMagKinda = plotXarrays(minParallaxMAS, maxParallaxMAS, apparentMagnitude, nPosteriorPoints=nPosteriorPoints)
 
@@ -428,7 +428,7 @@ def distanceTest(tgasCutMatched, nPosteriorPoints, data1, data2, err1, err2, xli
     plt.tight_layout()
     figDist.savefig('distancesM67.png')
 
[…]

### 2017-03-09 testXD.py
distance stuff

@@ -529,7 +529,7 @@ def distanceQuantile(color, absMagKinda, color_err, absMagKinda_err, tgasCutMatc
                 ax.set_ylabel('cdf')
                 ax.legend()
                 ax.set_title('$J-K$ ' +  '{0:.2f}'.format(float(color[index])) + ' $M_J$ ' + '{0:.2f}'.format(float(absMagKinda2absMag(absMagKinda[index]))))
-                fig.savefig('cdfplots/cdf.Mid.' + str(index) + '.' + iter '.png')
+                fig.savefig('cdfplots/cdf.Mid.' + str(index) + '.' + iter + '.png')
             """
             try:
                 distanceQuantile[index] = cdfInv(quantile)

### 2017-03-09 testXD.py
taking out some plotting

@@ -429,7 +429,7 @@ def distanceTest(tgas, nPosteriorPoints, data1, data2, err1, err2, xlim, ylim, p
     figDist.savefig('distancesM67.png')
 
 
-def distanceQuantile(color, absMagKinda, color_err, absMagKinda_err, tgas, xdgmm, distanceFile='distance.npy', quantile=0.05, nDistanceSamples=512, nPosteriorPoints=1000, iter='1st'):
+def distanceQuantile(color, absMagKinda, color_err, absMagKinda_err, tgas, xdgmm, distanceFile='distance.npy', quantile=0.05, nDistanceSamples=512, nPosteriorPoints=1000, iter='1st', plotPost=False):
     try:
         data = np.load(distanceFile)
         distance = data['distance']
@@ -446,7 +446,14 @@ def distanceQuantile(color, absMagKinda, color_err, absMagKinda_err, tgas, xdgmm
         xparallaxMAS = 1./10.**logDistance
         positive = xparallaxMAS > 0.
 
-        fig, ax = plt.subplots()
+        nMidMin = 0
+        nMidPost = 0
+        nMidFlat = 0
+        nSmallMin = 0
+        nSmallMax = 0
+        nSmallFlat = 0
+
+        if plotPost: fig, ax = plt.subplots()
         for index in range(nstars):
             if np.mod(index, 10000) == 0.0:
                 end = time.time()
@@ -489,50 +496,63 @@ def distanceQuantile(color, absMagKinda, color_err, absMagKinda_err, tgas, xdgmm
                 distanceQuantile[index] = 10.**cdfInv(quantile)
                 if np.mod(index, 10000) == 0.0:
                     label = 'posterior is good, log distance is ' + '{0:.2f}'.format(float(cdfInv(quantile)))
+                    if plotPost:
+                        plt.cla()
+                        ax.plot(logDistance[:-1], cdf, label=label, lw=2)
+                        ax.set_xlabel('log distance [kpc]')
+                        ax.set_ylabel('cdf')
+                        ax.legend()
+                        ax.set_title('$J-K$ ' +  '{0:.2f}'.format(float(color[index])) + ' $M_J$ ' + '{0:.2f}'.format(float(absMag)))
+                        fig.savefig('cdfplots/cdf.good.' + str(index) + '.' + iter + '.png')
+            #if the posterior lies way outside the distance window [0.01-10] kpc then set distance to which ever side of window has higher probability
+            elif np.max(cdf) < quantile:
[…]

### 2017-03-09 testXD.py
minor bugs

@@ -491,6 +491,7 @@ def distanceQuantile(color, absMagKinda, color_err, absMagKinda_err, tgas, xdgmm
             assert np.sum(np.isnan(posteriorLogDistance)) == 0., 'there are Nans in my posterior ' + str(index)
             absMag = absMagKinda2absMag(absMagKinda[index])
             if np.isnan(absMag): absMag = absMagKinda[index]
+
             #if the posterior lies well within the distance window then do the right thing
             if np.max(cdf) > 0.95:
                 distanceQuantile[index] = 10.**cdfInv(quantile)
@@ -504,19 +505,20 @@ def distanceQuantile(color, absMagKinda, color_err, absMagKinda_err, tgas, xdgmm
                         ax.legend()
                         ax.set_title('$J-K$ ' +  '{0:.2f}'.format(float(color[index])) + ' $M_J$ ' + '{0:.2f}'.format(float(absMag)))
                         fig.savefig('cdfplots/cdf.good.' + str(index) + '.' + iter + '.png')
+
             #if the posterior lies way outside the distance window [0.01-10] kpc then set distance to which ever side of window has higher probability
             elif np.max(cdf) < quantile:
                 #print 'The CDF did not reach above ', str(quantile), ' for ', str(index)
-                if P_minDist > P_maxDist:
+                if P_minDist > P_maxDist: #pdf lies at small distances
                     distanceQuantile[index] = 10.**minDist
                     label = 'posterior small and outside range, set to min Dist for ', str(index)
                     print label
                     nSmallMin += 1
-                if P_minDist < P_maxDist:
+                if P_minDist < P_maxDist: #pdf lies at large distances
                     distanceQuantile[index] = 10.**maxDist
                     label = 'posterior outside range, set to max Dist'
                     nSmallMax += 1
-                if P_minDist == P_maxDist:
+                if P_minDist == P_maxDist: #pdf is flat
                     label = 'The posterior is flat with value '+ str(P_minDist)+', Dist = 0.0 for ', str(index)
                     print label
                     nSmallFlat += 1
@@ -532,16 +534,16 @@ def distanceQuantile(color, absMagKinda, color_err, absMagKinda_err, tgas, xdgmm
             #if the posterior lies just on the edge, then if the pdf appears to be rising with distance set to quantile, else set to min distance
             else:
                 #print 'The max of the CDF is between ', str(quantile), ' and 0.95 for ', str(index)
[…]

### 2017-03-09 testXD.py
all the right logic

@@ -440,7 +440,7 @@ def distanceQuantile(color, absMagKinda, color_err, absMagKinda_err, tgas, xdgmm
         sourceID = np.zeros(nstars, dtype='>i8')
         #dustEBV = np.zeros(nstars)
         #dustEBV50 = np.zeros(nstars)
-        distanceQuantile = np.zeros(nstars)
+        distance = np.zeros(nstars)
         start = time.time()
         logDistance = np.linspace(-2, 1, nPosteriorPoints)
         xparallaxMAS = 1./10.**logDistance
@@ -448,11 +448,14 @@ def distanceQuantile(color, absMagKinda, color_err, absMagKinda_err, tgas, xdgmm
 
         nMidMin = 0
         nMidPost = 0
-        nMidFlat = 0
+        nMidFlatMin = 0
+        nMidFlatMax = 0
         nSmallMin = 0
         nSmallMax = 0
-        nSmallFlat = 0
-
+        nSmallFlatMin = 0
+        nSmallFlatMax = 0
+        debug=False
+        plotPost = True
         if plotPost: fig, ax = plt.subplots()
         for index in range(nstars):
             if np.mod(index, 10000) == 0.0:
@@ -465,15 +468,22 @@ def distanceQuantile(color, absMagKinda, color_err, absMagKinda_err, tgas, xdgmm
             meanData, covData = matrixize(color[index], absMagKinda[index], color_err[index], absMagKinda_err[index])
             meanData = meanData[0]
             covData = covData[0]
-
-            #windowFactor = 5. #the number of sigma to sample in mas for plotting
-            #minParallaxMAS = tgas['parallax'][index] - windowFactor*tgas['parallax_error'][index]
-            #maxParallaxMAS = tgas['parallax'][index] + windowFactor*tgas['parallax_error'][index]
-
             apparentMagnitude = bandDictionary[absmag]['array'][bandDictionary[absmag]['key']][index]
-            #xparallaxMAS, xabsMagKinda = plotXarrays(minParallaxMAS, maxParallaxMAS, apparentMagnitude, nPosteriorPoints=nPosteriorPoints)
-            #positive = xparallaxMAS > 0.
             xabsMagKinda = parallax2absMagKinda(xparallaxMAS, apparentMagnitude)
+
+            if debug:
+
+                windowFactor = 15. #the number of sigma to sample in mas for plotting
[…]

### 2017-03-14 paperDistance/paper.tex
bugfix

@@
 
+%@archiver{PUT FIGURE NAME HERE}
+
 % # To-do list
-% - We need a standard name for the \tgas--\tmass\ Intersection.
+% - We need a standard name for the \tgas--\tmass--\PanSTARRS\ Intersection.
 % - zeroth draft the introduction
@@
 a related project (CITE BORIS), in which we use a
-somewhat less flexible model for the \cmd, but in which we take a much more principled
+slightly less appropriate model for the \cmd, but in which we take a much more principled
 approach to the inference.
@@
 \mathrm{with \; absolute \; magnitude} \;\;\;\; &M_{J,n} = J_n - DM_n - Q_JE(B-V)_n \\
-& DM_n = -5\mathrm{log_{10}}\frac{1}{\varpi} \\
+& DM_n = -5\log_{10}\frac{1}{\varpi} \\
 \mathrm{and \; color} \;\;\;\; &(J - K)_n = J_n - K_n - Q_{JK}E(B-V)_n
-
 \end{aligned}
@@
 
+Talk about structures in CMD; red bump - acknowledge Ana ! s
+What's That Feature ?!!
+
+Fraction of stars in various parts of the diagram?
+
+Compare with literature stellar models
+
 How this compares to other Bayesian attempts at inferring distances,

### 2017-03-14 paperDistance/paper.tex
placeholders for assumptions and pedagogy

@@
 
+\section{Why does this work?}
+
+LA+DWH: Do we we want to put a pedagogical section in right here?
+
+Hogg would propose a straight-line fit (possibly with intrinsic
+width); showing that it can be used to de-noise noisy y-data.
+
 \section{Assumptions and methods}
@@
 \begin{itemize}
-\item The photometric uncertainties in the J band are small compared with the uncertainties in the parallax
-\item Our empirical bayes approach is ok as long as we only infer information about the data contained within the prior
-\item the uncertainties are gaussian
-\item most stars lie where we expect them to on the CMD
-\item number of gaussians that's efficient? or do x-validation
-\item can't think
+\item Some kind of stationarity assumption.
+\item Our empirical bayes approach is ok---that is, our numbers are large enough.
+\item The photometric uncertainties in the J band are small compared with the uncertainties in the parallax; the uncertainties are gaussian
+\item number of gaussians
+\item no use of physical models, baby!
 \end{itemize}

### 2017-03-14 paperDistance/paper.tex
move to modern

@@
 
-\documentclass[12pt, preprint]{aastex6}
-\hypersetup{colorlinks = false}
-
-%% Begin DWH mods
-\setlength{\voffset}{0in}
-\setlength{\hoffset}{0in}
-\setlength{\textwidth}{6in}
-\setlength{\textheight}{9in}
-\setlength{\headheight}{0ex}
-\setlength{\footnotesep}{0in}
-\setlength{\topmargin}{-\headsep}
-\setlength{\oddsidemargin}{0.25in}
-\setlength{\evensidemargin}{0.25in}
-\linespread{1.09}
-\setlength{\parindent}{1.2\baselineskip}
-\makeatletter % you know you are living your life wrong when you need to do this
-\long\def\frontmatter@title@above{
-\vspace*{-\headsep}\vspace*{\headheight}
-\noindent\footnotesize
-{\noindent\footnotesize\textsc{\@journalinfo}}\par
-{\noindent\scriptsize Preprint typeset using \LaTeX\ style AASTeX6
-with modifications by DWH.
-}\par\vspace*{-\baselineskip}\vspace*{0.625in}
-}%
-\long\def\frontmatter@abstractheading{%
-\makeaffils
-  \vspace*{-\baselineskip}\vspace*{1.5pt}
-  \vspace*{0.13189in}
- \begingroup
-  \centering
-  \abstractname
-  \vskip 1mm
-  \par
- \endgroup
- \everypar{\rightskip=0.0in\leftskip=\rightskip}\par
-}%
-\def\frontmatter@keys@format{\vspace*{0.5mm}%
-  \settowidth{\keys@width}{\normalsize\@keys@name}%
-  \rightskip=0.0in\leftskip=\rightskip\parindent=0pt%
-    \hangindent=\keys@width\hangafter=1\normalsize\raggedright}%
-\def\twodigits#1{\ifnum#1<10 0\fi\the#1}
-\def\mydate{\leavevmode\hbox{\the\year-\twodigits\month-\twodigits\day}}
-\makeatother
-%% End DWH mods
+\documentclass[modern]{aastex61}
+\usepackage{amsmath}
+\setlength{\parindent}{\baselineskip}
 
@@
 
-\usepackage{amsmath}
-
-\newcommand{\myemail}{landerson@simonsfoundation.org}
-\slugcomment{Draft manuscript not ready for distribution. \textbf{Copyright 2017 the authors.}}
-\shorttitle{Improving Gaia parallaxes}
-\shortauthors{Anderson, Hogg, Leistedt, Price-Whelan}
-\begin{document}\sloppy\sloppypar\raggedbottom\frenchspacing
+\shorttitle{improving gaia parallaxes}
+\shortauthors{anderson et al.}
+\begin{document}\sloppy\sloppypar\raggedbottom\frenchspacing % trust me
 
@@
        with a data-driven model of the color--magnitude diagram}
-\author{Lauren Anderson\altaffilmark{1},
-        David W. Hogg\altaffilmark{1,2,3,4},
-        Boris Leistedt\altaffilmark{2},
-        Adrian Price-Whelan\altaffilmark{5}}
-\altaffiltext{1}{Center for Computational Astrophysics, Flatiron Institute, Simons Foundation, New York City}
-\altaffiltext{2}{Center for Cosmology and Particle Physics, Department of Physics, New York University}
-\altaffiltext{3}{Center for Data Science, New York University}
-\altaffiltext{4}{Max-Planck-Institut f\"ur Astronomie, Heidelberg}
-\altaffiltext{5}{Princeton University Observatory, Princeton University}
 
-\begin{abstract}
+\author{Lauren Anderson}
+\affil{Center for Computational Astrophysics, Flatiron Institute, 162 Fifth Ave, New York, NY 10010, USA}
+
+\author{David W. Hogg}
+\affil{Center for Computational Astrophysics, Flatiron Institute, 162 Fifth Ave, New York, NY 10010, USA}
+\affil{Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, Newy York, NY 10003, USA}
+\affil{Center for Data Science, New York University, 60 Fifth Ave, New York, NY 10011, USA}
+\affil{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg}
+
+\author{Jo Bovy}
+\affil{Center for Computational Astrophysics, Flatiron Institute, 162 Fifth Ave, New York, NY 10010, USA}
+\affil{Department of Astronomy and Astrophysics, University of Toronto, 50 St. George Street, Toronto, ON M5S 3H4, Canada}
+
+\author{Boris Leistedt}
+\affil{Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, Newy York, NY 10003, USA}
+
+\author{Adrian Price-Whelan}
+\affil{Princeton University Observatory, Princeton University, 4 Ivy Lane, Princeton, NJ 08544, USA}
+
+\begin{abstract}\noindent % trust me
 The \gaia\ \tgas\ \project{Catalog} contains more than 2 million parallax measurements.
@@
 
-LA: Insert itemized or enumerated list of assumptions here. DWH will
-typeset it to look okay! Be exhaustive and then we will edit down.
-
+We make the following assumptions.
+The method we present is correct and justifiable under these assumptions, each of which is questionable in its own right.
+We return to criticize these assumptions---and the method that flows from them---in the Discussion below.
 \begin{itemize}
@@
 \item The photometric uncertainties in the J band are small compared with the uncertainties in the parallax; the uncertainties are gaussian
-\item number of gaussians
+\item Mixture-of-gaussian model with fixed number of gaussians
 \item no use of physical models, baby!

### 2017-03-14 paperDistance/paper.tex
assumptions draft:

@@
 
-We make the following assumptions.
-The method we present is correct and justifiable under these assumptions, each of which is questionable in its own right.
-We return to criticize these assumptions---and the method that flows from them---in the Discussion below.
+We make the following assumptions. The method we present is correct
+and justifiable under these assumptions, each of which is questionable
+in its own right. We return to criticize these assumptions---and the
+method that flows from them---in the Discussion below.
 \begin{itemize}
-\item Some kind of stationarity assumption.
-\item Our empirical bayes approach is ok---that is, our numbers are large enough.
-\item The photometric uncertainties in the J band are small compared with the uncertainties in the parallax; the uncertainties are gaussian
-\item Mixture-of-gaussian model with fixed number of gaussians
-\item no use of physical models, baby!
+\item We make an assumption of \emph{stationarity}; that is, that the
+  stars measured at lower signal-to-noise have analogs measured at
+  higher signal-to-noise.  In detail, this assumption is wrong, of
+  course (for example, distant halo stars are different than local
+  disk stars), but the stationarity assumption is fairly weak; as long
+  as there is support in the high signal-to-noise stars for the types
+  of stars seen at low signal-to-noise, the results will not be
+  strongly biased, but this assumption is deep and fundamental to this
+  work.
+\item We assume that we have large numbers of stars, large enough that
+  an empirical-Bayes or maximum-marginalized likelihood is a safe
+  approximation to full Bayesian inference. This assumption will be
+  least true in the least-populated parts of the \cmd. In the extreme
+  case of a stellar type with only one (statistically distinct)
+  exemplar in the data set, this assumption will be maximally wrong.
+\item We assume that the \tgas\ parallax uncertainties dominate the
+  noise in any estimates of absolute magnitude. We further assume that
+  the parallax and color uncertainties are Gaussian in form with
+  correctly known variances. In detail, we assume that the [INSERT
+    CORRECT TERMINOLOGY FROM GAIA HERE] parallax uncertainty estimates
+  are correct. Here we are making use of the fact that, for
+  de-noising, under-estimation of measurement noise is conservative
+  (it leads to under-deconvolution).
+\item We assume that the \cmd\ can be represented with a mixture of
+  Gaussians in a particular transformed space. This mixture we fix
+  (with only heuristic analysis) at 128, 512, and 2048 Gaussians in
+  different experiments.
+\item We make no use of stellar models; our only assumptions about
+  stellar physics are generic and implicit (for example, that the
+  \cmd\ is somehow smooth and compact).
 \end{itemize}

### 2017-03-14 paperDistance/paper.tex
bibliography

--- 2017-03-14 paperDistance/paper.tex
@@
 
-%@archiver{PUT FIGURE NAME HERE}
-
 % # To-do list
-% - We need a standard name for the \tgas--\tmass--\PanSTARRS\ Intersection.
+% - We need a standard name for the \tgas--\tmass\ Intersection.
 % - zeroth draft the introduction
@@
 
+\newcommand{\bibtex}{\textsc{Bib}\!\TeX} % bibtex. Not quite the correct typesetting, but close enough
 \newcommand{\foreign}[1]{\textsl{#1}}
@@
 
-The novel use of \xd\ here is to treat its output---the deconvolved \cmd---as
+The novel use of \xd\ \citep{bovy11} here is to treat its output---the deconvolved \cmd---as
 a prior for use in inference of individual-star distances.
@@
 a related project (CITE BORIS), in which we use a
-slightly less appropriate model for the \cmd, but in which we take a much more principled
+somewhat less flexible model for the \cmd, but in which we take a much more principled
 approach to the inference.
@@
 
-\section{Why does this work?}
-
-LA+DWH: Do we we want to put a pedagogical section in right here?
-
-Hogg would propose a straight-line fit (possibly with intrinsic
-width); showing that it can be used to de-noise noisy y-data.
-
 \section{Assumptions and methods}
@@
 \begin{itemize}
-\item Some kind of stationarity assumption.
-\item Our empirical bayes approach is ok---that is, our numbers are large enough.
-\item The photometric uncertainties in the J band are small compared with the uncertainties in the parallax; the uncertainties are gaussian
-\item number of gaussians
-\item no use of physical models, baby!
+\item The photometric uncertainties in the J band are small compared with the uncertainties in the parallax
+\item Our empirical bayes approach is ok as long as we only infer information about the data contained within the prior
+\item the uncertainties are gaussian
+\item most stars lie where we expect them to on the CMD
+\item number of gaussians that's efficient? or do x-validation
+\item can't think
 \end{itemize}
@@
 \mathrm{with \; absolute \; magnitude} \;\;\;\; &M_{J,n} = J_n - DM_n - Q_JE(B-V)_n \\
-& DM_n = -5\log_{10}\frac{1}{\varpi} \\
+& DM_n = -5\mathrm{log_{10}}\frac{1}{\varpi} \\
 \mathrm{and \; color} \;\;\;\; &(J - K)_n = J_n - K_n - Q_{JK}E(B-V)_n
+
 \end{aligned}
@@
 where $E(B-V)$ is the output of the Schlafly dust model, and $Q$ is the correction factor for that band assuming some reddening law.
-To generate our prior, we fit the data in the CMD space using XD (Bovy et al) which deconvolves the data with its noise, which are heterogeneous measurement uncertainties.
+To generate our prior, we fit the data in the CMD space using XD which deconvolves the data with its noise, which are heterogeneous measurement uncertainties.
 It generates the denoised, underlying distribution from which the noisey data were drawn from.
@@
 
+\bibliography{lmanders}
 %\begin{thebibliography}{}

### 2017-03-14 demo_plots.py

@@ -63,34 +63,39 @@ def plot_sample(x_true, y_true, x, y, samplex, sampley, xdgmm, xlabel='x', ylabe
     setup_text_plots(fontsize=16, usetex=True)
     plt.clf()
     alpha = 0.01
-    fig = plt.figure(figsize=(12, 9))
-    fig.subplots_adjust(left=0.1, right=0.95,
-                        bottom=0.1, top=0.95,
-                        wspace=0.02, hspace=0.02)
+    alpha_points = 0.01
+    figData = plt.figure(figsize=(12, 4.5))
+    figPrior = plt.figure(figsize=(12, 4.5))
+    for fig in [figData, figPrior]:
+        fig.subplots_adjust(left=0.1, right=0.95,
+                            bottom=0.1, top=0.95,
+                            wspace=0.02, hspace=0.02)
 
-    ax1 = fig.add_subplot(221)
+    ax1 = figData.add_subplot(121)
     ax1.scatter(x_true, y_true, s=1, lw=0, c='k', alpha=alpha)
 
-    ax2 = fig.add_subplot(222)
+    ax2 = figData.add_subplot(122)
 
-    ax2.scatter(x, y, s=1, lw=0, c='k', alpha=alpha)
-    ax2.errorbar(x, y, xerr=xerr, yerr=yerr, fmt="none", ecolor='black', zorder=0, lw=0.5, mew=0, alpha=0.05)
+    ax2.scatter(x, y, s=1, lw=0, c='k', alpha=alpha_points)
+    ax2.errorbar(x, y, xerr=xerr, yerr=yerr, fmt="none", zorder=0, lw=0.05, mew=0, alpha=1.0, cmap='Greys')
 
-    ax3 = fig.add_subplot(223)
-    ax3.scatter(samplex, sampley, s=4, lw=0, c='k', alpha=alpha)
-    xlim = ax3.get_xlim()
-    ylim = [10, -5] #ax3.get_ylim()
-
-    ax4 = fig.add_subplot(224)
+    ax3 = figPrior.add_subplot(121)
     for i in range(xdgmm.n_components):
         points = drawEllipse.plotvector(xdgmm.mu[i], xdgmm.V[i])
-        ax4.plot(points[0, :], absMagKinda2absMag(points[1,:]), 'k-', alpha=xdgmm.weights[i]/np.max(xdgmm.weights))
[…]

### 2017-03-14 testXD.py

@@ -455,10 +455,19 @@ def distanceQuantile(color, absMagKinda, color_err, absMagKinda_err, tgas, xdgmm
         nSmallMax = 0
         nSmallFlatMin = 0
         nSmallFlatMax = 0
-        debug=False
-        plotPost = False
+        debug = False
+        plotPost = True
+
+        distSmalldata = np.load('distanceQuantiles.128gauss.dQ0.05.6th.2MASS.All.npz.save')
+        distSmall = distSmalldata['distance']
+
+        distLargedata = np.load('distanceQuantiles.128gauss.dQ0.5.1st.2MASS.All.npz.save')
+        distLarge = distLargedata['distance']
+
+        delta = distLarge - distSmall
+
         if plotPost: fig, ax = plt.subplots()
-        for index in range(nstars):
+        for index in np.where(delta < 0)[0]: #range(nstars):
             if np.mod(index, 10000) == 0.0:
                 end = time.time()
                 print index, ' took ', str(end - start), 'seconds, projecting will be ', str((end-start)*((nstars-index)/10000.))
@@ -509,16 +518,16 @@ def distanceQuantile(color, absMagKinda, color_err, absMagKinda_err, tgas, xdgmm
             #if the posterior lies well within the distance window then do the right thing
             if np.max(cdf) > 0.95:
                 distance[index] = 10.**cdfInv(quantile)
-                if np.mod(index, 10000) == 0.0:
-                    label = 'posterior is good, log distance is ' + '{0:.2f}'.format(float(cdfInv(quantile)))
-                    if plotPost:
-                        plt.cla()
-                        ax.plot(logDistance[:-1], cdf, label=label, lw=2)
-                        ax.set_xlabel('log distance [kpc]')
-                        ax.set_ylabel('cdf')
-                        ax.legend()
-                        ax.set_title('$J-K$ ' +  '{0:.2f}'.format(float(color[index])) + ' $M_J$ ' + '{0:.2f}'.format(float(absMag)))
-                        fig.savefig('cdfplots/cdf.good.' + str(index) + '.' + iter + '.png')
+                #if np.mod(index, 10000) == 0.0:
[…]

### 2017-03-14 paperDistance/paper.tex
typos, bibtex
repo url; grants

@@
 \affil{Center for Computational Astrophysics, Flatiron Institute, 162 Fifth Ave, New York, NY 10010, USA}
-\affil{Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, Newy York, NY 10003, USA}
+\affil{Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, New York, NY 10003, USA}
 \affil{Center for Data Science, New York University, 60 Fifth Ave, New York, NY 10011, USA}
@@
 \author{Boris Leistedt}
-\affil{Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, Newy York, NY 10003, USA}
+\affil{Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, New York, NY 10003, USA}
 
@@
 This work has made use of data from the European Space Agency (\acronym{ESA})
-mission Gaia (http://www.cosmos.esa.int/gaia), processed by the Gaia
-Data Processing and Analysis Consortium (\acronym{DPAC},
-http://www.cosmos.esa.int/web/gaia/dpac/consortium). Funding for the
+mission Gaia\footnote{\url{http://www.cosmos.esa.int/gaia}}, processed by the Gaia
+Data Processing and Analysis Consortium (\acronym{DPAC}\footnote{\url{http://www.cosmos.esa.int/web/gaia/dpac/consortium}}). Funding for the
 \acronym{DPAC} has been provided by national institutions, in particular the
@@
 
-This project was partially supported by [DWH GIVE GRANT NUMBERS]. It
-made use of the \acronym{NASA} Astrophysics Data System.
+This project was partially supported by This research was partially supported by
+  the \acronym{NSF} (\acronym{AST-1517237}),
+  \acronym{NASA} (grant \acronym{NNX12AI50G}),
+  and the Moore-Sloan Data Science Environment at \acronym{NYU}.
+It made use of the \acronym{NASA} Astrophysics Data System.
+All the code used in this project is available online\footnote{\url{https://github.com/andersdot/stellarTwins}} under an open-source license.
 
-\appendix
-
-\section{Appendix material}
+%% \appendix
+%% \section{Appendix material}
 

### 2017-03-14 paperDistance/paper.tex
2mass, refs, etc

@@
 \newcommand{\gaia}{\project{Gaia}}
+\newcommand{\panstarrs}{\project{Pan\acronym{STARRS}}}
 \newcommand{\xd}{\acronym{XD}}
@@
 
-Here we build and use just such a data-driven model.
-In particular, we use all of the \gaia\ \tgas\ data, matched to the \tmass\
-photometry, to make a model of the noise-free (or low-noise)
-color--magnitude diagram (\cmd) of stars.
-We use
-an Empirical-Bayes approach known as Extreme Deconvolution (DWH CITE).
-This method deconvolves heteroskedastic data to derive an estimate
-of the noise-free or low-noise distribution that \emph{would have}
-been observed with far better data.
+Here we build and use just such a data-driven model.  In particular,
+we use all of the \gaia\ \tgas\ data, matched to the
+\tmass\ photometry (\citealt{skrutskie06}), and in the footprint of
+the \panstarrs\ Dust Map (CITE GREEN), to make a model of the
+noise-free (or low-noise) color--magnitude diagram (\cmd) of stars.
+We use an Empirical-Bayes approach known as Extreme Deconvolution (XD;
+\citealt{bovy11}).
+This method deconvolves heteroskedastic data to derive an
+estimate of the noise-free or low-noise distribution that \emph{would
+  have} been observed with far better data.
 This method has been used in astrophysics
@@
 
-The novel use of \xd\ \citep{bovy11} here is to treat its output---the deconvolved \cmd---as
+The novel use of \xd\ here is to treat its output---the deconvolved \cmd---as
 a prior for use in inference of individual-star distances.
@@
 is among our long-term goals and motivations.
-Along those lines, this paper can be seen as a companion to
-a related project (CITE BORIS), in which we use a
-slightly less appropriate model for the \cmd, but in which we take a much more principled
+Along those lines, this paper can be seen as a companion to a related
+project (CITE BORIS), in which we use a slightly less appropriate
+model for the \cmd, but in which we take a much more principled
 approach to the inference.
@@
 \acknowledgments It is a pleasure to thank
-  Jo Bovy (Toronto),
+  Ana Bonaca (Harvard),
   Andy Casey (Monash),
@@
 Center for Computational Astrophysics for comments and input.
-
 This project was developed in part at the 2016 \acronym{NYC} Gaia Sprint, hosted
@@
 
-[IS THERE A \tmass\ ACKNOWLEDGEMENT?]
+This publication makes use of data products from the Two Micron All
+Sky Survey, which is a joint project of the
+University of Massachusetts and the Infrared Processing and Analysis
+Center/California Institute of Technology, funded by the National
+Aeronautics and Space Administration and the National Science
+Foundation.
 

### 2017-03-14 paperDistance/paper.tex
posterior outputs warnings

@@
 
+DWH: Placeholder for the output format for our data products. Should
+be the last paragraph or two of this section.
+
 \section{Discussion}
@@
 
-Note the dangers of using the stars from our output in hierarchical inferences!
+The outputs of this project include posterior pdfs and posterior
+samplings for stellar parallaxes or distances.
+It is tempting to treat these outputs as equivalent to some kind of catalog of
+measurements, as we treated the original parallax measurements from \tgas.
+That is, it is tempting to treat these as simply ``better'' measurements.
+Taken one star at a time, this is permitted and true, and the basis for our
+claim that we have de-noised the \tgas\ Catalog.
+However, there are important differences, as there are with all
+probabilistic catalogs (HOGG CITE LANG, FINKBEINER), between the
+measurements and the de-noised measurements.
+The most important difference, for users of our output, between our
+output and the \tgas\ data input, is that the \tgas\ data is the representation
+of a likelihood function, whereas our ouptut is a
+representation of a posterior pdf, one per star.
+If used naively, posterior information can lead to serious statistical
+errors, because each data point has had a prior multiplied in.
+Effectively multiplicative uses of the data will effectively take the prior
+pdf to a large power.
+We have given examples of correct uses of posterior samplings or pdfs
+for subsequent inferences in previous work (HOGG CITE ECC and DFM); we
+encourage power-users to consult those methodological
+contributions before engaging with these outputs.
 

### 2017-03-15 tightPosterior.py
plots for paper

@@ -1,0 +1,109 @@
+import numpy as np
+from astropy.io import fits
+import matplotlib.pyplot as plt
+import matplotlib as mpl
+from dustmaps.bayestar import BayestarQuery
+from astropy.coordinates import SkyCoord
+import astropy.units as units
+import comparePrior
+
+def distanceFilename(ngauss, quantile, iter, survey, dataFilename):
+    return 'distanceQuantiles.' + str(ngauss) + 'gauss.dQ' + str(quantile) + '.' + iter + '.' + survey + '.' + dataFilename
+
+def dustFilename(ngauss, quantile, iter, survey, dataFilename):
+    return 'dustCorrection.'    + str(ngauss) + 'gauss.dQ' + str(quantile) + '.' + iter + '.' + survey + '.' + dataFilename
+
+def dataArrays(survey='2MASS'):
+
+    tgas = fits.getdata("stacked_tgas.fits", 1)
+    Apass = fits.getdata('tgas-matched-apass-dr9.fits')
+    twoMass = fits.getdata('tgas-matched-2mass.fits')
+
+    if survey == 'APASS':
+        mag1 = 'B'
+        mag2 = 'V'
+        absmag = 'G'
+        xlabel='B-V'
+        ylabel = r'M$_\mathrm{G}$'
+        xlim = [-0.2, 2]
+        ylim = [9, -2]
+
+    if survey == '2MASS':
+        mag1 = 'J'
+        mag2 = 'K'
+        absmag = 'J'
+        xlabel = 'J-K$_s$'
+        ylabel = r'M$_\mathrm{J}$'
+        xlim = [-0.25, 1.25]
+        ylim = [6, -4]
+
+    bandDictionary = {'B':{'key':'bmag', 'err_key':'e_bmag', 'array':twoMass},
+                     'V':{'key':'vmag', 'err_key':'e_vmag', 'array':Apass},
+                     'J':{'key':'j_mag', 'err_key':'j_cmsig', 'array':twoMass},
+                     'K':{'key':'k_mag', 'err_key':'k_cmsig', 'array':twoMass},
+                     'G':{'key':'phot_g_mean_mag', 'array':tgas}}
+
+    nonzeroError = (bandDictionary[mag1]['array'][bandDictionary[mag1]['err_key']] != 0.0) & \
+                   (bandDictionary[mag2]['array'][bandDictionary[mag2]['err_key']] != 0.0)
+
+    bayes = BayestarQuery(max_samples=1)
[…]

### 2017-03-15 paperDistance/paper.tex
adjust outline

@@
 % - We need a standard name for the \tgas--\tmass--\PanSTARRS\ Intersection.
-% - zeroth draft the introduction
-% - zeroth draft the method section
-% - zeroth draft the assumptions list
-% - reformat single-column
+% - zeroth draft / re-write the method section
+% - make a ``why does this work'' section
 
@@
 \usepackage{amsmath}
-\setlength{\parindent}{\baselineskip}
+\setlength{\parindent}{\baselineskip} % trust me!
+\renewcommand{\twocolumngrid}{} % trust me!
 
@@
 \title{Improving \textsl{Gaia} parallaxes
-       with a data-driven model of the color--magnitude diagram}
+       with a data-driven model \\
+       of the color--magnitude diagram}
 
@@
 Talk about structures in CMD; red bump - acknowledge Ana ! s
-What's That Feature ?!!
-
-Fraction of stars in various parts of the diagram?
+What's That Feature? Similarly; Fraction of stars in various parts of the diagram?
 
@@
 \bibliography{lmanders}
-%\begin{thebibliography}{}
-%\end{thebibliography}
 

### 2017-03-15 paperDistance/paper.tex
dust assumption

@@
   (it leads to under-deconvolution).
+\item We treat the \panstarrs-based three-dimensional dust maps (CITE
+  SCHLAFLY) as correct in their median dust estimates, and their
+  effects on color and magnitude. We iteratively update each star's
+  dust correction as its inferred distance updates as we optimize the
+  empirical-Bayes model. Our assumption is that these corrections are
+  good enough, and uncertainties in these are not dominant, for either
+  color or absolute magnitude.
 \item We assume that the \cmd\ can be represented with a mixture of

### 2017-03-16 paperDistance/paper.tex
argmax
rename bibtex; cite hogg and dfm

@@
 
-\newcommand{\bibtex}{\textsc{Bib}\!\TeX} % bibtex. Not quite the correct typesetting, but close enough
+% text macros
 \newcommand{\foreign}[1]{\textsl{#1}}
@@
 
+% math macros
+\DeclareMathOperator*{\argmax}{arg\,max}
+
+% start and title
 \shorttitle{improving gaia parallaxes}
@@
 \begin{document}\sloppy\sloppypar\raggedbottom\frenchspacing % trust me
-
 \title{Improving \textsl{Gaia} parallaxes
@@
 pdf to a large power.
-We have given examples of correct uses of posterior samplings or pdfs
-for subsequent inferences in previous work (HOGG CITE ECC and DFM); we
+We have (in previous work)
+given examples of correct uses of posterior samplings or pdfs
+for subsequent inferences (\citealt{hogg08}; \citealt{dfm14}); we
 encourage power-users to consult those methodological
@@
 
-\bibliography{lmanders}
+\bibliography{gaia}
 

### 2017-03-16 paperDistance/paper.tex
methods

--- 2017-03-15 paperDistance/paper.tex
@@
 
-\author{Lauren Anderson}
+\author[0000-0001-5725-9329]{Lauren Anderson}
 \affil{Center for Computational Astrophysics, Flatiron Institute, 162 Fifth Ave, New York, NY 10010, USA}
@@
 \begin{equation}
-\label{eq:likelihood}
+\label{eq:gaialike}
 P(\varpi |d, \sigma^2_{\varpi}, \textbf{I})
@@
 where $d$ is distance to the star, $\varpi$ is the observed parallax, and $\sigma^2_{\varpi}$ is the uncertainty in that measurement, and I are other various prior assumptions (mostly about the noise properties?).
-This is the probability of the data given the model, or distance, but we want to flip it around and infer the distance to the star given the measured parallax and it's uncertainty.
-We do this using probability theory/bayes rule, and turn the likelihood of the data into the posterior probablity of the distance
+This is the probability of the data given the model distance, but we want to flip it around and infer the distance to the star given the measured parallax and it's uncertainty.
+We do this using probability theory/bayes rule, and turn the likelihood of the data into the posterior probability of the distance
 
@@
 \label{eq:bayes}
-P(d|\varpi, \sigma^2_{\varpi}, \textbf{I}) = \frac{P(\varpi |d, \sigma^2_{\varpi}, \textbf{I}) P(d)}{\textbf{Z}},
+P(d_{true}|\varpi, \sigma^2_{\varpi}, \textbf{I}) = \frac{P(\varpi |d_{true} \sigma^2_{\varpi}, \textbf{I}) P(d)}{\textbf{Z}},
 \end{equation}
@@
 
+True values:
+  C - true color,
+  M - true absolute magnitude
+  Y = [10.^0.2(M+2), C]
+data:
+    (J-K)_c
+    \varpi
+    J_c
+    y = [\varpi10^0.2J_c, (J-K)_c]
+
+  prior assumptions:
+    \sigma_J, sigma_K, sigma_\varpi, Greene map
+
+Our true, latent values are C, the true J-K color, and M the true absolute magnitude, which we encapsulate in the true vector $\vec{Y}$.
+The observed data is the dust corrected J-K color $(J-K)_n$, the Gaia parallax $\varpi_n$, and the dust corrected J band magnitude $J_c$, which we encapsulate in the data vector.
+Our prior assumptions $I$ are the uncertainties in each of these measurements, as well as the Greene dust map
+
 \begin{equation}
@@
 \begin{aligned}
-\mathrm{data vector} \;\;\;\; &\vec{y} = [10^{0.2M_{J,n}}, (J- K)_n] \;  n \in \mathrm{stars} \\
+\mathrm{true vector} \;\;\;\; &\vec{Y} = [10^{0.2\,M + 2}, C] \\
+\mathrm{data vector} \;\;\;\; &\vec{y} = [\varpi\,10^{0.2J_{c,n}}, (J- K)_{c,n}] \;  n \in \mathrm{stars} \\
 \mathrm{with \; absolute \; magnitude} \;\;\;\; &M_{J,n} = J_n - DM_n - Q_JE(B-V)_n \\
 & DM_n = -5\log_{10}\frac{1}{\varpi} \\
-\mathrm{and \; color} \;\;\;\; &(J - K)_n = J_n - K_n - Q_{JK}E(B-V)_n
+\mathrm{and \; color} \;\;\;\; &(J - K)_n = J_n - K_n - Q_{JK}E(B-V)_n \\
+\mathrm{assumptions} \;\;\;\; &I = \left\{\left\{\sigma^2_n\right\}_{n=1}^N, \mathrm{assumptions}\right\}
 \end{aligned}
@@
 where $E(B-V)$ is the output of the Schlafly dust model, and $Q$ is the correction factor for that band assuming some reddening law.
+
+Likelihood function:
+  P(y|Y) = N(y|Y,I)
+Working with these variables, we then have a likelihood for the data, both the Gaia parallaxes and 2mass photometry of
+
+\begin{equation}
+\label{eq:fulllike}
+P(\vec{y}|\vec{Y}) = N(\vec{y}|\vec{Y}, I)
+\end{equation}
+
 To generate our prior, we fit the data in the CMD space using XD which deconvolves the data with its noise, which are heterogeneous measurement uncertainties.
 It generates the denoised, underlying distribution from which the noisey data were drawn from.
-It represents the distribution as a mixture of gaussians which maximizes the likelihood of the noisy data.
+It represents the distribution as a mixture of gaussians which maximizes the marginalized likelihood of the noisy data, marginalized over the true, latent variables.
+
+XD generates the gaussian parameters
 
 \begin{equation}
-\label{eq:xdlikelihood}
-P(\vec{y}|\alpha, I)
+\label{eq:xdmml}
+\begin{aligned}
+
+\alpha =& \left\{\mathrm{amp}_k, \mu_k, V_k\right\}_{k=1}^K \; k \in \mathrm{Gaussians} \\
+       =& \argmax_{\alpha} \Pi_n P(\vec{y}_n|\alpha) \\
+       =& \argmax_{\alpha} \Pi_n \int P(\vec{y}_n|\vec{Y}_n,I)P(\vec{Y}|\alpha)d\vec{Y}
+\end{aligned}
 \end{equation}
 
-with the output mixture of gaussians
+\begin{equation}
+\label{eq:posterior}
+P(Y|y, \alpha, I) = 1/Z p(y| Y,I)p(Y|\alpha, I)
+\end{equation}
 
 \begin{equation}
-\label{eq:xd}
-\begin{aligned}
-\alpha = &\left\{\mathrm{amp}_k, \mu_k, V_k\right\}_{k=1}^K \; k \in \mathrm{Gaussians} \\
-I = &\left\{\left\{\sigma^2_n\right\}_{n=1}^N, \mathrm{assumptions}\right\}
+\label{eq:parallaxPost}
+P(\vec{Y}_1|\vec{y}, \alpha, I) => p(\varpi_{true}|y, \alpha, I) \, f(\varpi_{true})
+\[
+f(\varpi_{true}) = \begin{cases}
+              1: -2 < log_{10} \varpi < 2\\
+              0: otherwise
+              \end{cases}
+\]
+\end{equation}
+
 \end{aligned}

### 2017-03-16 paperDistance/paper.tex

@@
 % - zeroth draft / re-write the method section
-% - make a ``why does this work'' section
+% - add ORCIDs like Hogg's
 
@@
 
-\author{David W. Hogg}
+\author[0000-0003-2866-9403]{David W. Hogg}
 \affil{Center for Computational Astrophysics, Flatiron Institute, 162 Fifth Ave, New York, NY 10010, USA}

### 2017-03-16 paperDistance/paper.tex
notation

@@
 \label{eq:gaialike}
-P(\varpi |d, \sigma^2_{\varpi}, \textbf{I})
+P(\varpi |\varpi_{true}, \sigma^2_{\varpi}, \textbf{I})
 \end{equation}
 
-where $d$ is distance to the star, $\varpi$ is the observed parallax, and $\sigma^2_{\varpi}$ is the uncertainty in that measurement, and I are other various prior assumptions (mostly about the noise properties?).
+where $d$ is distance to the star, $\varpi$ is the observed parallax, and $\sigma^2_{\varpi}$ is the uncertainty in that measurement, and I are other various prior assumptions (mostly about the noise properties which we will drop in our notation but is implicitly there).
 This is the probability of the data given the model distance, but we want to flip it around and infer the distance to the star given the measured parallax and it's uncertainty.
@@
 \label{eq:bayes}
-P(d_{true}|\varpi, \sigma^2_{\varpi}, \textbf{I}) = \frac{P(\varpi |d_{true} \sigma^2_{\varpi}, \textbf{I}) P(d)}{\textbf{Z}},
+P(\varpi_{true}|\varpi, \sigma^2_{\varpi}) = \frac{P(\varpi |\varpi_{true}, \sigma^2_{\varpi}) P(\varpi_{true})}{\textbf{Z}},
 \end{equation}
 
-where $P(d)$ is our prior knowledge of the distance to the star, and $\textbf{Z}$ is a normalization constant to make the posterior sum to 1.
+where $P(\varpi_{true})$ is our prior knowledge of the distance to the star, and $\textbf{Z}$ is a normalization constant to make the posterior sum to 1.
 The strength in our inference method lies in how we generate our prior knowledge of the distances to stars.
@@
 To generate our CMD prior we use the J and K band from 2MASS, as well as the parallax from Gaia. The data are a transformation of the dust corrected absolute magnitude, and the dust corrected colors.
-
-True values:
-  C - true color,
-  M - true absolute magnitude
-  Y = [10.^0.2(M+2), C]
-data:
-    (J-K)_c
-    \varpi
-    J_c
-    y = [\varpi10^0.2J_c, (J-K)_c]
-
-  prior assumptions:
-    \sigma_J, sigma_K, sigma_\varpi, Greene map
 
@@
 \begin{aligned}
-\mathrm{true vector} \;\;\;\; &\vec{Y} = [10^{0.2\,M + 2}, C] \\
-\mathrm{data vector} \;\;\;\; &\vec{y} = [\varpi\,10^{0.2J_{c,n}}, (J- K)_{c,n}] \;  n \in \mathrm{stars} \\
-\mathrm{with \; absolute \; magnitude} \;\;\;\; &M_{J,n} = J_n - DM_n - Q_JE(B-V)_n \\
-& DM_n = -5\log_{10}\frac{1}{\varpi} \\
-\mathrm{and \; color} \;\;\;\; &(J - K)_n = J_n - K_n - Q_{JK}E(B-V)_n \\
-\mathrm{assumptions} \;\;\;\; &I = \left\{\left\{\sigma^2_n\right\}_{n=1}^N, \mathrm{assumptions}\right\}
+\mathrm{true \, vector} \;\;\;\; &\vec{Y} = [10^{0.2\,M_{J,n} + 2}, C_n] \\
+\mathrm{data \, vector} \;\;\;\; &\vec{y} = [\varpi\,10^{0.2J^c_n}, (J- K)^c_n] \;  n \in \mathrm{stars} \\
+\mathrm{with \; absolute \; magnitude} \;\;\;\; &M^c_{J,n} = J_n + 5\log_{10}\varpi_n - Q_JE(B-V)_n \\
+\mathrm{and \; color} \;\;\;\; &(J - K)^c_n = J_n - K_n - Q_{JK}E(B-V)_n \\
+\mathrm{assumptions} \;\;\;\; &I = \left\{\left\{\sigma^2_{\varpi, n}\right\}, \left\{\sigma^2_{J,n}\right\}, \left\{\sigma^2_{K,n}\right\}, \mathrm{Dust \, Map}\right\}
 \end{aligned}
@@
 
-Likelihood function:
-  P(y|Y) = N(y|Y,I)
 Working with these variables, we then have a likelihood for the data, both the Gaia parallaxes and 2mass photometry of
@@
 \label{eq:fulllike}
-P(\vec{y}|\vec{Y}) = N(\vec{y}|\vec{Y}, I)
+P(\vec{y}|\vec{Y}) = \mathcal{N}(\vec{y}|\vec{Y}, \sigma_{\vec{y}})
 \end{equation}
@@
 \label{eq:xdmml}
-\begin{aligned}
-
+\begin{split}
 \alpha =& \left\{\mathrm{amp}_k, \mu_k, V_k\right\}_{k=1}^K \; k \in \mathrm{Gaussians} \\
-       =& \argmax_{\alpha} \Pi_n P(\vec{y}_n|\alpha) \\
-       =& \argmax_{\alpha} \Pi_n \int P(\vec{y}_n|\vec{Y}_n,I)P(\vec{Y}|\alpha)d\vec{Y}
-\end{aligned}
+       =& \argmax_{\alpha} \, \Pi_n \, P(\vec{y}_n|\alpha) \\
+       =& \argmax_{\alpha} \, \Pi_n \, \int P(\vec{y}_n|\vec{Y}_n) \, P(\vec{Y}|\alpha)d\vec{Y} \\
+\end{split}
 \end{equation}
@@
 \label{eq:posterior}
-P(Y|y, \alpha, I) = 1/Z p(y| Y,I)p(Y|\alpha, I)
+P(\vec{Y}|\vec{y}, \alpha) = \frac{1}{Z} \, P(\vec{y}|\vec{Y}) \, P(\vec{Y}|\alpha)
 \end{equation}
@@
 \label{eq:parallaxPost}
-P(\vec{Y}_1|\vec{y}, \alpha, I) => p(\varpi_{true}|y, \alpha, I) \, f(\varpi_{true})
+\begin{split}
+P(\vec{Y}_0|\vec{y}, \alpha) => P(\varpi_{true}|y, \alpha) \, f(\varpi_{true}) \\
 \[
 f(\varpi_{true}) = \begin{cases}
-              1: -2 < log_{10} \varpi < 2\\
-              0: otherwise
+              1: -2 < \mathrm{log}_{10} \varpi_{true} < 2\\
+              0: \mathrm{otherwise}
               \end{cases}
 \]
+\end{split}
 \end{equation}

### 2017-03-16 paperDistance/paper.tex
tweaks

--- 2017-03-16 paperDistance/paper.tex
@@
 
-\author[0000-0001-5725-9329]{Lauren Anderson}
+\author{Lauren Anderson}
 \affil{Center for Computational Astrophysics, Flatiron Institute, 162 Fifth Ave, New York, NY 10010, USA}
@@
 \begin{equation}
-\label{eq:gaialike}
+\label{eq:likelihood}
 P(\varpi |d, \sigma^2_{\varpi}, \textbf{I})
@@
 where $d$ is distance to the star, $\varpi$ is the observed parallax, and $\sigma^2_{\varpi}$ is the uncertainty in that measurement, and I are other various prior assumptions (mostly about the noise properties?).
-This is the probability of the data given the model distance, but we want to flip it around and infer the distance to the star given the measured parallax and it's uncertainty.
-We do this using probability theory/bayes rule, and turn the likelihood of the data into the posterior probability of the distance
+This is the probability of the data given the model, or distance, but we want to flip it around and infer the distance to the star given the measured parallax and it's uncertainty.
+We do this using probability theory/bayes rule, and turn the likelihood of the data into the posterior probablity of the distance
 
@@
 \label{eq:bayes}
-P(d_{true}|\varpi, \sigma^2_{\varpi}, \textbf{I}) = \frac{P(\varpi |d_{true} \sigma^2_{\varpi}, \textbf{I}) P(d)}{\textbf{Z}},
+P(d|\varpi, \sigma^2_{\varpi}, \textbf{I}) = \frac{P(\varpi |d, \sigma^2_{\varpi}, \textbf{I}) P(d)}{\textbf{Z}},
 \end{equation}
@@
 
-True values:
-  C - true color,
-  M - true absolute magnitude
-  Y = [10.^0.2(M+2), C]
-data:
-    (J-K)_c
-    \varpi
-    J_c
-    y = [\varpi10^0.2J_c, (J-K)_c]
-
-  prior assumptions:
-    \sigma_J, sigma_K, sigma_\varpi, Greene map
-
-Our true, latent values are C, the true J-K color, and M the true absolute magnitude, which we encapsulate in the true vector $\vec{Y}$.
-The observed data is the dust corrected J-K color $(J-K)_n$, the Gaia parallax $\varpi_n$, and the dust corrected J band magnitude $J_c$, which we encapsulate in the data vector.
-Our prior assumptions $I$ are the uncertainties in each of these measurements, as well as the Greene dust map
-
 \begin{equation}
@@
 \begin{aligned}
-\mathrm{true vector} \;\;\;\; &\vec{Y} = [10^{0.2\,M + 2}, C] \\
-\mathrm{data vector} \;\;\;\; &\vec{y} = [\varpi\,10^{0.2J_{c,n}}, (J- K)_{c,n}] \;  n \in \mathrm{stars} \\
+\mathrm{data vector} \;\;\;\; &\vec{y} = [10^{0.2M_{J,n}}, (J- K)_n] \;  n \in \mathrm{stars} \\
 \mathrm{with \; absolute \; magnitude} \;\;\;\; &M_{J,n} = J_n - DM_n - Q_JE(B-V)_n \\
 & DM_n = -5\log_{10}\frac{1}{\varpi} \\
-\mathrm{and \; color} \;\;\;\; &(J - K)_n = J_n - K_n - Q_{JK}E(B-V)_n \\
-\mathrm{assumptions} \;\;\;\; &I = \left\{\left\{\sigma^2_n\right\}_{n=1}^N, \mathrm{assumptions}\right\}
+\mathrm{and \; color} \;\;\;\; &(J - K)_n = J_n - K_n - Q_{JK}E(B-V)_n
 \end{aligned}
@@
 where $E(B-V)$ is the output of the Schlafly dust model, and $Q$ is the correction factor for that band assuming some reddening law.
-
-Likelihood function:
-  P(y|Y) = N(y|Y,I)
-Working with these variables, we then have a likelihood for the data, both the Gaia parallaxes and 2mass photometry of
+To generate our prior, we fit the data in the CMD space using XD which deconvolves the data with its noise, which are heterogeneous measurement uncertainties.
+It generates the denoised, underlying distribution from which the noisey data were drawn from.
+It represents the distribution as a mixture of gaussians which maximizes the likelihood of the noisy data.
 
 \begin{equation}
-\label{eq:fulllike}
-P(\vec{y}|\vec{Y}) = N(\vec{y}|\vec{Y}, I)
+\label{eq:xdlikelihood}
+P(\vec{y}|\alpha, I)
 \end{equation}
 
-To generate our prior, we fit the data in the CMD space using XD which deconvolves the data with its noise, which are heterogeneous measurement uncertainties.
-It generates the denoised, underlying distribution from which the noisey data were drawn from.
-It represents the distribution as a mixture of gaussians which maximizes the marginalized likelihood of the noisy data, marginalized over the true, latent variables.
-
-XD generates the gaussian parameters
+with the output mixture of gaussians
 
 \begin{equation}
-\label{eq:xdmml}
+\label{eq:xd}
 \begin{aligned}
-
-\alpha =& \left\{\mathrm{amp}_k, \mu_k, V_k\right\}_{k=1}^K \; k \in \mathrm{Gaussians} \\
-       =& \argmax_{\alpha} \Pi_n P(\vec{y}_n|\alpha) \\
-       =& \argmax_{\alpha} \Pi_n \int P(\vec{y}_n|\vec{Y}_n,I)P(\vec{Y}|\alpha)d\vec{Y}
-\end{aligned}
-\end{equation}
-
-\begin{equation}
-\label{eq:posterior}
-P(Y|y, \alpha, I) = 1/Z p(y| Y,I)p(Y|\alpha, I)
-\end{equation}
-
-\begin{equation}
-\label{eq:parallaxPost}
-P(\vec{Y}_1|\vec{y}, \alpha, I) => p(\varpi_{true}|y, \alpha, I) \, f(\varpi_{true})
-\[
-f(\varpi_{true}) = \begin{cases}
-              1: -2 < log_{10} \varpi < 2\\
-              0: otherwise
-              \end{cases}
-\]
-\end{equation}
-
+\alpha = &\left\{\mathrm{amp}_k, \mu_k, V_k\right\}_{k=1}^K \; k \in \mathrm{Gaussians} \\
+I = &\left\{\left\{\sigma^2_n\right\}_{n=1}^N, \mathrm{assumptions}\right\}
 \end{aligned}
@@
 This work has made use of data from the European Space Agency (\acronym{ESA})
-mission Gaia\footnote{\url{http://www.cosmos.esa.int/gaia}}, processed by the Gaia
-Data Processing and Analysis Consortium (\acronym{DPAC}\footnote{\url{http://www.cosmos.esa.int/web/gaia/dpac/consortium}}). Funding for the
+mission \gaia\footnote{\url{http://www.cosmos.esa.int/gaia}}, processed by the Gaia
+Data Processing and Analysis Consortium\footnote{\url{http://www.cosmos.esa.int/web/gaia/dpac/consortium}} (\acronym{DPAC}). Funding for the
 \acronym{DPAC} has been provided by national institutions, in particular the

### 2017-03-16 paperDistance/paper.tex
laurens first go at methods

--- 2017-03-16 paperDistance/paper.tex
@@
 \newcommand{\cmd}{\acronym{CMD}}
-
+\newcommand{\given}{\,|\,}
 % math macros
@@
 
-We would like to infer the distance to stars using the measured (noisy) parallax from Gaia.
-Gaia gives us the likelihood of the parallax of a star, given some unknown distance and uncertainties (assumed gaussian)
+We would like to infer the true parallax $\varpi_{true}$ to a star using the measured noisy parallax $\varpi$ and $\sigma_{\varpi}$ from \gaia.
+\gaia\ gives us the likelihood of the parallax of a star, given some unknown true parallax and uncertainties, which we assume to be a Gaussian
 \begin{equation}
+p(\varpi \given \varpi_{true}, \sigma_{\varpi},I) = \mathcal{N}(\varpi \given \varpi_{true}, \sigma_{\varpi})
 \label{eq:gaialike}
-P(\varpi |\varpi_{true}, \sigma^2_{\varpi}, \textbf{I})
 \end{equation}
 
-where $d$ is distance to the star, $\varpi$ is the observed parallax, and $\sigma^2_{\varpi}$ is the uncertainty in that measurement, and I are other various prior assumptions (mostly about the noise properties which we will drop in our notation but is implicitly there).
-This is the probability of the data given the model distance, but we want to flip it around and infer the distance to the star given the measured parallax and it's uncertainty.
-We do this using probability theory/bayes rule, and turn the likelihood of the data into the posterior probability of the distance
+where $\varpi_{true}$ represents the true parallax, or distance, to the star, $\varpi$ is the observed parallax, and $\sigma_{\varpi}$ is the uncertainty in that measurement.
+There are other various prior assumptions (mostly about the noise properties which we will drop in our notation but is implicitly there and labeled $I$ below).
+$\mathcal{N}$ is the Gaussian function $\mathcal{N}(x \given \mu, \sigma)$ with mean $\mu$ and variance $\sigma$.
+This is the probability of the data given the true parallax, but we want to flip it around and infer the true parallax, or distance, to the star given the measured parallax and it's uncertainty.
+We do this using probability theory/bayes rule, and turn the likelihood of the data into the posterior probability of the true parallax
 
 \begin{equation}
+p(\varpi_{true} \given \varpi, \sigma^2_{\varpi}) = \frac{p(\varpi \given \varpi_{true}, \sigma^2_{\varpi}) \, p(\varpi_{true})}{Z},
 \label{eq:bayes}
-P(\varpi_{true}|\varpi, \sigma^2_{\varpi}) = \frac{P(\varpi |\varpi_{true}, \sigma^2_{\varpi}) P(\varpi_{true})}{\textbf{Z}},
 \end{equation}
 
-where $P(\varpi_{true})$ is our prior knowledge of the distance to the star, and $\textbf{Z}$ is a normalization constant to make the posterior sum to 1.
-The strength in our inference method lies in how we generate our prior knowledge of the distances to stars.
-Here, instead of assuming the spatial distribution of stars in the milky way, or relying on stellar models, as previous work as shown, we build the color-magnitude diagram of the observed stars from their observations as our prior.
-The absolute magnitude of the CMD carries information about the distances to stars and where they can physically lie in this space.
+where $p(\varpi_{true})$ is our prior knowledge of the parallax to the star, and $Z$ is a normalization constant to properly normalize the posterior.
+The strength in our inference method lies in how we generate our prior knowledge of the parallax to a star, $p(\varpi_{true})$.
+For our prior, instead of assuming the spatial distribution of stars in the Milky Way, or relying on stellar models, as previous work as shown, we build the \cmd\ from the observed colors and absolute magnitudes of our full set of stars.
+The absolute magnitude of the \cmd\ carries information about the distances, or parallaxes, to stars and the tight correlation of observed stars within this space dictates where each indivdually observed star can physically lie in this space.
 Toy model?
 
-To generate our CMD prior we use the J and K band from 2MASS, as well as the parallax from Gaia. The data are a transformation of the dust corrected absolute magnitude, and the dust corrected colors.
+To generate our \cmd\ prior we use the $J$ and $K$ band photometry from \tmass, as well as the parallax from \gaia.
+The data are a transformation of the dust corrected absolute magnitude, and the dust corrected colors.
 
-Our true, latent values are C, the true J-K color, and M the true absolute magnitude, which we encapsulate in the true vector $\vec{Y}$.
-The observed data is the dust corrected J-K color $(J-K)_n$, the Gaia parallax $\varpi_n$, and the dust corrected J band magnitude $J_c$, which we encapsulate in the data vector.
+\begin{equation}
+10^{0.2\,M^c_{J,n} + 2} = \varpi\,10^{0.2J^c_n}
+\label{eq:transform}
+\end{equation}
+
+The transformation of the dust corrected absolute magnitude, Equation \ref{eq:transform}, is required to to keep the uncertainties gaussian, which, under some assumptions along with this transformation, we maintain.
+Gaussian uncertainties are required to use \xd\, more about this and our dust corrections below.
+
+Our true, latent values are the true $J-K$ color $C_n$ and the true $J$ band absolute magnitude $M_{J,n}$ for each star, which we encapsulate in the true vector $\vec{Y}$.
+The observed data is the dust corrected $J-K$ color $(J-K)^c_n$, the \gaia\ parallax $\varpi_n$ in units of $[mas]$, and the dust corrected $J$ band magnitude $J^c_n$, which we encapsulate in the data vector $\vec{y}$.
 Our prior assumptions $I$ are the uncertainties in each of these measurements, as well as the Greene dust map
@@
 \begin{equation}
-\label{eq:data}
 \begin{aligned}
@@
 \mathrm{data \, vector} \;\;\;\; &\vec{y} = [\varpi\,10^{0.2J^c_n}, (J- K)^c_n] \;  n \in \mathrm{stars} \\
-\mathrm{with \; absolute \; magnitude} \;\;\;\; &M^c_{J,n} = J_n + 5\log_{10}\varpi_n - Q_JE(B-V)_n \\
-\mathrm{and \; color} \;\;\;\; &(J - K)^c_n = J_n - K_n - Q_{JK}E(B-V)_n \\
+\mathrm{with \; absolute \; magnitude} \;\;\;\; &M^c_{J,n} = J_n + 5\log_{10}\varpi_n - Q_J\,E(B-V)_n \\
+\mathrm{and \; color} \;\;\;\; &(J - K)^c_n = J_n - K_n - Q_{JK}\,E(B-V)_n \\
 \mathrm{assumptions} \;\;\;\; &I = \left\{\left\{\sigma^2_{\varpi, n}\right\}, \left\{\sigma^2_{J,n}\right\}, \left\{\sigma^2_{K,n}\right\}, \mathrm{Dust \, Map}\right\}
 \end{aligned}
+\label{eq:data}
 \end{equation}
 
-where $E(B-V)$ is the output of the Schlafly dust model, and $Q$ is the correction factor for that band assuming some reddening law.
+where $E(B-V)_n$ is the output of the Green et al dust model, and $Q_{\lambda}$ is the correction factor for a photometric band assuming some reddening law.
 
-Working with these variables, we then have a likelihood for the data, both the Gaia parallaxes and 2mass photometry of
+Working with these variables, we then have a likelihood for the data, encapsulating both the \gaia\ parallaxes and \tmass\ photometry of
 
 \begin{equation}
+p(\vec{y} \given \vec{Y}) = \mathcal{N}(\vec{y} \given \vec{Y}, \sigma_{\vec{y}})
 \label{eq:fulllike}
-P(\vec{y}|\vec{Y}) = \mathcal{N}(\vec{y}|\vec{Y}, \sigma_{\vec{y}})
 \end{equation}
 
-To generate our prior, we fit the data in the CMD space using XD which deconvolves the data with its noise, which are heterogeneous measurement uncertainties.
-It generates the denoised, underlying distribution from which the noisey data were drawn from.
-It represents the distribution as a mixture of gaussians which maximizes the marginalized likelihood of the noisy data, marginalized over the true, latent variables.
+To generate our prior, we fit the data in the slightly transformed \cmd\ space, Equation \ref{eq:transform}, using \xd\ which deconvolves the data with its uncertainties, which are heterogeneous measurement uncertainties.
+It generates the denoised, underlying distribution from which the uncertain data were drawn from.
+It represents the distribution as a mixture of gaussians which maximizes the marginalized likelihood of the data, marginalized over the true data.
 
-XD generates the gaussian parameters
+\xd\ generates the gaussian parameters
+
+\begin{eqnarray}
+\alpha &=& \left\{\mathrm{amp}_k, \mu_k, V_k\right\}_{k=1}^K \; k \in \mathrm{Gaussians} \nonumber\\
+       &=& \argmax_{\alpha} \, \Pi_n \, p(\vec{y}_n \given \alpha) \nonumber\\
+       &=& \argmax_{\alpha} \, \Pi_n \, \int p(\vec{y}_n \given \vec{Y}_n) \, p(\vec{Y} \given \alpha)\mathrm{d\vec{Y}}
+\label{eq:xdmml}
+\end{eqnarray}
+
+To infer $\varpi_{true}$ from the likelihood of the \gaia\ data, Equation \ref{eq:gaialike}, we first do the inference in the 2D space of the \cmd
 
 \begin{equation}
-\label{eq:xdmml}
-\begin{split}
-\alpha =& \left\{\mathrm{amp}_k, \mu_k, V_k\right\}_{k=1}^K \; k \in \mathrm{Gaussians} \\
-       =& \argmax_{\alpha} \, \Pi_n \, P(\vec{y}_n|\alpha) \\
-       =& \argmax_{\alpha} \, \Pi_n \, \int P(\vec{y}_n|\vec{Y}_n) \, P(\vec{Y}|\alpha)d\vec{Y} \\
-\end{split}
+p(\vec{Y} \given \vec{y}, \alpha) = \frac{1}{Z} \, p(\vec{y} \given \vec{Y}) \, p(\vec{Y} \given \alpha)
+\label{eq:posterior}
 \end{equation}
 
-\begin{equation}
-\label{eq:posterior}
-P(\vec{Y}|\vec{y}, \alpha) = \frac{1}{Z} \, P(\vec{y}|\vec{Y}) \, P(\vec{Y}|\alpha)
-\end{equation}
+We then project the 2D posterior into the 1D absolute magnitude/parallax space
 
-\begin{equation}
-\label{eq:parallaxPost}
-\begin{split}
-P(\vec{Y}_0|\vec{y}, \alpha) => P(\varpi_{true}|y, \alpha) \, f(\varpi_{true}) \\
-\[
-f(\varpi_{true}) = \begin{cases}
+\begin{eqnarray}
+p(\vec{Y}_0 \given \vec{y}, \alpha) &\rightarrow& p(\varpi_{true} \given y, \alpha) \, f(\varpi_{true}) \\
+f(\varpi_{true}) &=& \begin{cases}
               1: -2 < \mathrm{log}_{10} \varpi_{true} < 2\\
@@
               \end{cases}
-\]
-\end{split}
-\end{equation}
+\label{eq:parallaxPost}
+\end{eqnarray}
 
-\end{aligned}
-\end{equation}
+And with a jacobian this becomes a posterior on distance to the star if you wish.
 
@@
 
-We use stars crossmatched in TGAS, 2MASS and within the observing footprint of Panstarrs to access the Schlafly 3D dust model.
+We use stars crossmatched in TGAS, 2MASS and within the observing footprint of \panstarrs to access the Schlafly 3D dust model.
 We require the signal to noise of both the photometry and parallax be $> 0.001$, that the photometry have nonzero errors, and remove a small selection of 2MASS stars that have zero color and zero J band magnitude.
@@
 Stars with poor signal to noise have a likelihood that is consistent with infinite distance and can therefore have severe dust corrections.
-This is most obvious in the giants, which tend to have the lowest signal to noise, and can therefore have dust corrections $\sim 1-2$ magnitudes.
-To apply a dust correction to the 2MASS photometry, we first generate our prior using \xd\ on the observed, attenuated photometry.
+This is most obvious in the giant stars, which tend to have the lowest signal to noise, and can therefore have dust corrections $> 2$ magnitudes.
+To apply a dust correction to the \tmass photometry, we first generate our prior using \xd\ on the observed, attenuated photometry.
 Using this raw prior we infer the distances to all the stars in our prior more precisely, increasing our signal to noise for each star and therefore minimize severe distances and their associated severe dust corrections.
@@
 \begin{equation}
+m_{\mathrm{band, corrected}} = m_{\mathrm{band}} - Q_{\lambda} \times E(B-V)
 \label{eq:dust}
-m_{\mathrm{band, corrected}} = m_{\mathrm{band}} - Q_{\lambda} \times E(B-V)
 \end{equation}

### 2017-03-16 paperDistance/paper.tex
slight change to notation

@@
 
-We would like to infer the true parallax $\varpi_{true}$ to a star using the measured noisy parallax $\varpi$ and $\sigma_{\varpi}$ from \gaia.
+Given these assumptions, we would like to infer the true parallax $\varpi_{true}$ to a star using the measured noisy parallax $\varpi$ and $\sigma_{\varpi}$ from \gaia.
 \gaia\ gives us the likelihood of the parallax of a star, given some unknown true parallax and uncertainties, which we assume to be a Gaussian
@@
 \mathrm{true \, vector} \;\;\;\; &\vec{Y} = [10^{0.2\,M_{J,n} + 2}, C_n] \\
-\mathrm{data \, vector} \;\;\;\; &\vec{y} = [\varpi\,10^{0.2J^c_n}, (J- K)^c_n] \;  n \in \mathrm{stars} \\
-\mathrm{with \; absolute \; magnitude} \;\;\;\; &M^c_{J,n} = J_n + 5\log_{10}\varpi_n - Q_J\,E(B-V)_n \\
+\mathrm{data \, vector} \;\;\;\; &\vec{y} = [\varpi_n\,10^{0.2J^c_n}, (J- K)^c_n] \;  n \in \mathrm{stars} \\
+\mathrm{with \; apparent \; magnitude} \;\;\;\; &J^c_n = J_n - Q_J\,E(B-V)_n \\
 \mathrm{and \; color} \;\;\;\; &(J - K)^c_n = J_n - K_n - Q_{JK}\,E(B-V)_n \\

### 2017-03-16 paperDistance/paper.tex
m67 section

--- 2017-03-16 paperDistance/paper.tex
@@
 
-We use stars crossmatched in TGAS, 2MASS and within the observing footprint of \panstarrs to access the Schlafly 3D dust model.
-We require the signal to noise of both the photometry and parallax be $> 0.001$, that the photometry have nonzero errors, and remove a small selection of 2MASS stars that have zero color and zero J band magnitude.
-We correct for dust using 3D Bayestar (Schlafly+Finkbeiner) with an estimate of the distance from iteratively sampling our posterior distances, more below.
+We use stars crossmatched in TGAS, 2MASS and within the observing footprint of \panstarrs\ to access the Greene 3D dust model.
+We require the photometry have real values and nonzero, real errors, and remove a small selection of \tmass\ stars that have zero color and zero J band magnitude.
 
@@
 
+We correct for dust using 3D Bayestar (Schlafly+Finkbeiner) with an estimate of the distance from iteratively sampling our posterior distances.
 Dust corrections are non trivial for stars with poor signal to noise.
@@
 This is most obvious in the giant stars, which tend to have the lowest signal to noise, and can therefore have dust corrections $> 2$ magnitudes.
-To apply a dust correction to the \tmass photometry, we first generate our prior using \xd\ on the observed, attenuated photometry.
-Using this raw prior we infer the distances to all the stars in our prior more precisely, increasing our signal to noise for each star and therefore minimize severe distances and their associated severe dust corrections.
-With more precise distance posteriors, we take the $5\%$ quantile of each distance posterior, so the closest bit of the posterior, and sample the Green et al 3D dust map at that distance.
+To apply a dust correction to the \tmass\ photometry, we first generate our prior using the observed, attenuated photometry.
+Using this raw prior we infer more precise distances to all the stars which we can use to get a measurement of dust.
+This approach minimizes severe distances from the likelihoods alone and their associated severe dust corrections.
+With more these more precise distance posteriors, we take the $5\%$ quantile of each distance posterior, so the closest bit of the posterior, and sample the Green et al 3D dust map at that distance.
 For each distance we take the median of the samples of the dust model.
-We apply each dust correction to our 2MASS photometry given by Equation \ref{eq:dust} where $K_{\lambda}$ = [3.626, 2.742, 3.303, 2.285, 1.698, 0.709, 0.449, 0.302] for bands [B, V, g, r, i, J, H, K] respectively, taken from table (REF TABLE SFD), and $E(B-V)$ is the output of the Green et al dust models.
+We apply each dust correction to our \tmass\ photometry given by Equation \ref{eq:data} where $Q_{\lambda}$ = [0.709, 0.302] for bands [J, K] respectively, taken from table (REF TABLE SFD), and $E(B-V)$ is the output of the Green et al dust models.
 We then regenerate our prior using the dust corrected photometry.
 We do not update the errors, but keep them the same as before the dust correction, which is wrong and under representing our errors but this leads again to a lighter deconvolution (Do we want to change this and add in the covariance for the dust?).
+We iterate this process 10 times, however the dust values seem to converge after about 2-3 iterations.
 Should I quote some statistics about the dust corrections?
@@
 
-\begin{equation}
-m_{\mathrm{band, corrected}} = m_{\mathrm{band}} - Q_{\lambda} \times E(B-V)
-\label{eq:dust}
-\end{equation}
+LMA: what are our results?
+\subsection{Distances to M67}
+As proof of concept we inferred the distances to stars in the open cluster M67.
+It is an open cluster in the constellation Cancer and is estimated to be 3-5 billion years old and 800-900 pc away.
+Previous work has estimated that bad things happen if you try to infer a distance as the inverse of the observed (noisy) parallax when the signal to noise of the parallax measurement drops below 5.
+With \gaia\ parallaxes having a minimum uncertainty of about 0.3 mas, for stars with the most certain parallaxes, the SN reaches 5 at about 700 pc, right where this cluster lies.
+So it is a great test case for comparing parallax inferences with or without a prior.
+We made a window of 1 degree squared on the sky with the cluster M67 lying in the middle.
+The cluster is visibly obvious as a clustering of stars in a visualization of the \gaia\ data in 2D on the sky.
+With our converged dust values, and it's associated prior, we calculate the posterior of the true parallaxes to the stars within this window on the sky around M67.
+Figure XXX shows the comparison of the likelihoods and posteriors.
+The vertical lines bracket the parallaxes of, or distances to, the stars associated with the cluster.
+The likelihoods of the observed parallaxes are far broader and the cluster is not obvious at all.
+In the posteriors there is a sharp peak in the probability at the previously determined distance to the cluster, the vertical lines, for far more stars than the likelihoods would imply.
+Our prior is not only making parallaxes more precise but also (possibly) increasing the accuracy.
+Too strong?
 

### 2017-03-17 paperDistance/paper.tex
m67 section
started messing with method section

@@
 
-Given these assumptions, we would like to infer the true parallax $\varpi_{true}$ to a star using the measured noisy parallax $\varpi$ and $\sigma_{\varpi}$ from \gaia.
-\gaia\ gives us the likelihood of the parallax of a star, given some unknown true parallax and uncertainties, which we assume to be a Gaussian
+We would like to infer the true parallax $\varpi_{\true}$ for
+each star given its parallax measurement $\varpi$ and associated
+uncertainty $\sigma_{\varpi}$ from \tgas.
+We assume (as \gaia\ suggests) that the likelihood is Gaussian
 \begin{equation}
-p(\varpi \given \varpi_{true}, \sigma_{\varpi},I) = \mathcal{N}(\varpi \given \varpi_{true}, \sigma_{\varpi})
+p(\varpi \given \varpi_{\true}, I) = \mathcal{N}(\varpi \given \varpi_{\true}, \sigma^2_{\varpi}) \quad ,
 \label{eq:gaialike}
 \end{equation}
-
-where $\varpi_{true}$ represents the true parallax, or distance, to the star, $\varpi$ is the observed parallax, and $\sigma_{\varpi}$ is the uncertainty in that measurement.
-There are other various prior assumptions (mostly about the noise properties which we will drop in our notation but is implicitly there and labeled $I$ below).
-$\mathcal{N}$ is the Gaussian function $\mathcal{N}(x \given \mu, \sigma)$ with mean $\mu$ and variance $\sigma$.
-This is the probability of the data given the true parallax, but we want to flip it around and infer the true parallax, or distance, to the star given the measured parallax and it's uncertainty.
-We do this using probability theory/bayes rule, and turn the likelihood of the data into the posterior probability of the true parallax
-
+where $I$ is a placeholder containing our assumptions, including all
+uncertainty estimates, and $\mathcal{N}(x \given \mu, \sigma^2)$ is a
+Gaussian function for $x$ with mean $\mu$ and variance $\sigma^2$.
+This is the probability of the data given the true parallax, but we
+want to flip it around and infer the true parallax, or distance, to
+the star given the measured parallax and it's uncertainty. We do this
+using probability theory
 \begin{equation}
-p(\varpi_{true} \given \varpi, \sigma^2_{\varpi}) = \frac{p(\varpi \given \varpi_{true}, \sigma^2_{\varpi}) \, p(\varpi_{true})}{Z},
+p(\varpi_{true} \given \varpi, \sigma^2_{\varpi}) = \frac{1}{Z}\,p(\varpi \given \varpi_{true}, \sigma^2_{\varpi}) \, p(\varpi_{true}) \quad ,
 \label{eq:bayes}
 \end{equation}
+where $p(\varpi_{true})$ is a prior probability density (pdf) for the
+parallax to the star, and $Z$ is a normalization constant.
+The strength of our inference method here lies in how we generate our prior
+pdf for the parallax $p(\varpi_{true})$.
+Instead of assuming a functional form for the spatial distribution of
+stars in the Milky Way (CITE TRI--CORYN), or relying on stellar
+models, we build a noise-deconvolved estimate of the color--magnitude
+diagram (\cmd) from the observed colors and absolute magnitudes of our
+full set of stars.
+This \cmd\ model becomes, for each star, a prior pdf for the distance
+to that star.
 
-where $p(\varpi_{true})$ is our prior knowledge of the parallax to the star, and $Z$ is a normalization constant to properly normalize the posterior.
-The strength in our inference method lies in how we generate our prior knowledge of the parallax to a star, $p(\varpi_{true})$.
-For our prior, instead of assuming the spatial distribution of stars in the Milky Way, or relying on stellar models, as previous work as shown, we build the \cmd\ from the observed colors and absolute magnitudes of our full set of stars.
-The absolute magnitude of the \cmd\ carries information about the distances, or parallaxes, to stars and the tight correlation of observed stars within this space dictates where each indivdually observed star can physically lie in this space.
-Toy model?
+[HOGG: NOW CONNECT THIS TO WHAT WE ACTUALLY DO WITH THE 2-d SPACE]
 

### 2017-03-17 paperDistance/paper.tex
math tweaks

@@
 \newcommand{\cmd}{\acronym{CMD}}
-\newcommand{\given}{\,|\,}
+
 % math macros
 \DeclareMathOperator*{\argmax}{arg\,max}
+\newcommand{\given}{\,|\,}
+\newcommand{\true}{\mathrm{true}}
 
@@
 \begin{equation}
-p(\varpi_{true} \given \varpi, \sigma^2_{\varpi}) = \frac{1}{Z}\,p(\varpi \given \varpi_{true}, \sigma^2_{\varpi}) \, p(\varpi_{true}) \quad ,
+p(\varpi_{\true} \given \varpi, \sigma^2_{\varpi}) = \frac{1}{Z}\,p(\varpi \given \varpi_{\true}, \sigma^2_{\varpi}) \, p(\varpi_{\true}) \quad ,
 \label{eq:bayes}
 \end{equation}
-where $p(\varpi_{true})$ is a prior probability density (pdf) for the
+where $p(\varpi_{\true})$ is a prior probability density (pdf) for the
 parallax to the star, and $Z$ is a normalization constant.
 The strength of our inference method here lies in how we generate our prior
-pdf for the parallax $p(\varpi_{true})$.
+pdf for the parallax $p(\varpi_{\true})$.
 Instead of assuming a functional form for the spatial distribution of
@@
 \begin{aligned}
-\mathrm{true \, vector} \;\;\;\; &\vec{Y} = [10^{0.2\,M_{J,n} + 2}, C_n] \\
-\mathrm{data \, vector} \;\;\;\; &\vec{y} = [\varpi_n\,10^{0.2J^c_n}, (J- K)^c_n] \;  n \in \mathrm{stars} \\
-\mathrm{with \; apparent \; magnitude} \;\;\;\; &J^c_n = J_n - Q_J\,E(B-V)_n \\
-\mathrm{and \; color} \;\;\;\; &(J - K)^c_n = J_n - K_n - Q_{JK}\,E(B-V)_n \\
-\mathrm{assumptions} \;\;\;\; &I = \left\{\left\{\sigma^2_{\varpi, n}\right\}, \left\{\sigma^2_{J,n}\right\}, \left\{\sigma^2_{K,n}\right\}, \mathrm{Dust \, Map}\right\}
+\mathrm{true \, vector} \quad &\vec{Y} = [10^{0.2\,M_{J,n} + 2}, C_n] \\
+\mathrm{data \, vector} \quad &\vec{y} = [\varpi_n\,10^{0.2J^c_n}, (J- K)^c_n] \;  n \in \mathrm{stars} \\
+\mathrm{with \; apparent \; magnitude} \quad &J^c_n = J_n - Q_J\,E(B-V)_n \\
+\mathrm{and \; color} \quad &(J - K)^c_n = J_n - K_n - Q_{JK}\,E(B-V)_n \\
+\mathrm{assumptions} \quad &I = \left\{\left\{\sigma^2_{\varpi, n}\right\}, \left\{\sigma^2_{J,n}\right\}, \left\{\sigma^2_{K,n}\right\}, \mathrm{Dust \, Map}\right\}
 \end{aligned}
@@
 
-To infer $\varpi_{true}$ from the likelihood of the \gaia\ data, Equation \ref{eq:gaialike}, we first do the inference in the 2D space of the \cmd
+To infer $\varpi_{\true}$ from the likelihood of the \gaia\ data, Equation \ref{eq:gaialike}, we first do the inference in the 2D space of the \cmd
 
@@
 \begin{eqnarray}
-p(\vec{Y}_0 \given \vec{y}, \alpha) &\rightarrow& p(\varpi_{true} \given y, \alpha) \, f(\varpi_{true}) \\
-f(\varpi_{true}) &=& \begin{cases}
-              1: -2 < \mathrm{log}_{10} \varpi_{true} < 2\\
-              0: \mathrm{otherwise}
+p(\vec{Y}_0 \given \vec{y}, \alpha) &\rightarrow& p(\varpi_{\true} \given y, \alpha) \, f(\varpi_{\true}) \\
+f(\varpi_{\true}) &=& \begin{cases}
+              1 \quad \mathrm{if} \quad -2 < \mathrm{log}_{10} \varpi_{true} < 2\\
+              0 \quad \mathrm{otherwise}
               \end{cases}

### 2017-03-17 paperDistance/paper.tex
tiny bit

@@
 full set of stars.
-This \cmd\ model becomes, for each star, a prior pdf for the distance
+This \cmd\ model becomes, for each star---given its observed color and
+magnitude (and uncertainties on those)---a prior pdf for the distance
 to that star.
 
-[HOGG: NOW CONNECT THIS TO WHAT WE ACTUALLY DO WITH THE 2-d SPACE]
+That said, this is \emph{not} exactly how we do the inference here.
+Here we perform the inference in a two-dimensional space that is a
+transform of the \cmd....[HOGG: Syntactical marker]
 

### 2017-03-17 paperDistance/paper.tex
figures

--- 2017-03-17 paperDistance/paper.tex
@@
 \section{Why does this work?}
+\begin{figure}
+\centering
+  \includegraphics[width=\textwidth]{lineExample.png}
+\caption{Why does this work: Fitting a line}
+\label{fig:line}
+\end{figure}
 
@@
   case of a stellar type with only one (statistically distinct)
-  exemplar in the data set, this assumption will be maximally wrong.
+  exemplar in the data set (e.g. a white dwarf), this assumption will be maximally wrong.
 \item We assume that the \tgas\ parallax uncertainties dominate the
@@
   SCHLAFLY) as correct in their median dust estimates, and their
-  effects on color and magnitude. We iteratively update each star's
-  dust correction as its inferred distance updates as we optimize the
-  empirical-Bayes model. Our assumption is that these corrections are
+  effects on color and magnitude. As we optimize the empirical-Bayes model, we iteratively update each star's
+  dust correction as its updated, inferred distance. Our assumption is that these corrections are
   good enough, and uncertainties in these are not dominant, for either
@@
   stellar physics are generic and implicit (for example, that the
-  \cmd\ is somehow smooth and compact).
+  \cmd\ is somehow smooth and compact). Dust model dependent on stellar models?
 \end{itemize}
 
-We would like to infer the true parallax $\varpi_{\true}$ for
-each star given its parallax measurement $\varpi$ and associated
-uncertainty $\sigma_{\varpi}$ from \tgas.
-We assume (as \gaia\ suggests) that the likelihood is Gaussian
+Given these assumptions, we would like to infer the true parallax $\varpi_{true}$ to a star using the measured noisy parallax $\varpi$ and $\sigma_{\varpi}$ from \gaia.
+\gaia\ gives us the likelihood of the parallax of a star, given some unknown true parallax and uncertainties, which we assume to be a Gaussian
 \begin{equation}
-p(\varpi \given \varpi_{\true}, I) = \mathcal{N}(\varpi \given \varpi_{\true}, \sigma^2_{\varpi}) \quad ,
+p(\varpi \given \varpi_{true}, \sigma_{\varpi},I) = \mathcal{N}(\varpi \given \varpi_{true}, \sigma_{\varpi})
 \label{eq:gaialike}
 \end{equation}
-where $I$ is a placeholder containing our assumptions, including all
-uncertainty estimates, and $\mathcal{N}(x \given \mu, \sigma^2)$ is a
-Gaussian function for $x$ with mean $\mu$ and variance $\sigma^2$.
-This is the probability of the data given the true parallax, but we
-want to flip it around and infer the true parallax, or distance, to
-the star given the measured parallax and it's uncertainty. We do this
-using probability theory
+
+where $\varpi_{true}$ represents the true parallax, or distance, to the star, $\varpi$ is the observed parallax, and $\sigma_{\varpi}$ is the uncertainty in that measurement.
+There are other various prior assumptions (mostly about the noise properties which we will drop in our notation but is implicitly there and labeled $I$ below).
+$\mathcal{N}$ is the Gaussian function $\mathcal{N}(x \given \mu, \sigma)$ with mean $\mu$ and variance $\sigma$.
+This is the probability of the data given the true parallax, but we want to flip it around and infer the true parallax, or distance, to the star given the measured parallax and it's uncertainty.
+We do this using probability theory/bayes rule, and turn the likelihood of the data into the posterior probability of the true parallax
+
 \begin{equation}
-p(\varpi_{true} \given \varpi, \sigma^2_{\varpi}) = \frac{1}{Z}\,p(\varpi \given \varpi_{true}, \sigma^2_{\varpi}) \, p(\varpi_{true}) \quad ,
+p(\varpi_{true} \given \varpi, \sigma^2_{\varpi}) = \frac{p(\varpi \given \varpi_{true}, \sigma^2_{\varpi}) \, p(\varpi_{true})}{Z},
 \label{eq:bayes}
 \end{equation}
-where $p(\varpi_{true})$ is a prior probability density (pdf) for the
-parallax to the star, and $Z$ is a normalization constant.
-The strength of our inference method here lies in how we generate our prior
-pdf for the parallax $p(\varpi_{true})$.
-Instead of assuming a functional form for the spatial distribution of
-stars in the Milky Way (CITE TRI--CORYN), or relying on stellar
-models, we build a noise-deconvolved estimate of the color--magnitude
-diagram (\cmd) from the observed colors and absolute magnitudes of our
-full set of stars.
-This \cmd\ model becomes, for each star, a prior pdf for the distance
-to that star.
 
-[HOGG: NOW CONNECT THIS TO WHAT WE ACTUALLY DO WITH THE 2-d SPACE]
+where $p(\varpi_{true})$ is our prior knowledge of the parallax to the star, and $Z$ is a normalization constant to properly normalize the posterior.
+The strength in our inference method lies in how we generate our prior knowledge of the parallax to a star, $p(\varpi_{true})$.
+For our prior, instead of assuming the spatial distribution of stars in the Milky Way, or relying on stellar models, as previous work as shown, we build the \cmd\ from the observed colors and absolute magnitudes of our full set of stars.
+The absolute magnitude of the \cmd\ carries information about the distances, or parallaxes, to stars and the tight correlation of observed stars within this space dictates where each indivdually observed star can physically lie in this space.
+Toy model?
 
@@
 \section{Data and results}
+\begin{figure}
+\centering
+  \includegraphics[width=\textwidth]{data.png}
+\caption{The observed \cmd: point estimates on the left and including error bars on the right.}
+\label{fig:data}
+\end{figure}
 
@@
 \subsection{Dust}
+
+\begin{figure}
+\centering
+  \includegraphics[width=\textwidth]{dust.png}
+\caption{The converged dust values at the 5\% distance projected on the l,b galactic coordinates.
+This shows both the footprint of our analysis as well as the dust values we used.
+Each point represents a single star in our cross matched catalog.
+The large missing areas are due to the lack of dust values from \panstarrs\ since it only looked above -30 degrees in dec }
+\label{fig:dust}
+\end{figure}
 
@@
 
+\subsection{Prior}
+\begin{figure}
+\centering
+  \includegraphics[width=\textwidth]{priorNgaussComparison.png}
+\caption{Prior: the XD generated mixture of gaussians we use as our prior}
+\label{fig:prior}
+\end{figure}
+
 LMA: what are our results?
 \subsection{Distances to M67}
+\begin{figure}
+\centering
+  \includegraphics[width=\textwidth]{distancesM67.png}
+\caption{Proof of concept: parallax inference to the open cluster M67}
+\label{fig:m67}
+\end{figure}
+
 As proof of concept we inferred the distances to stars in the open cluster M67.
@@
 Too strong?
+
+\begin{figure}
+\centering
+  \includegraphics[width=\textwidth]{posterior.png}
+\caption{5 example posteriors showing the priors and likelihoods as well}
+\label{fig:posterior}
+\end{figure}
 

### 2017-03-17 tightPosterior.py

@@ -1,11 +1,11 @@
 import numpy as np
-from astropy.io import fits
 import matplotlib.pyplot as plt
 import matplotlib as mpl
-from dustmaps.bayestar import BayestarQuery
-from astropy.coordinates import SkyCoord
-import astropy.units as units
 import comparePrior
+import testXD
+from xdgmm import XDGMM
+import testXD
+import stellarTwins as st
 
 def distanceFilename(ngauss, quantile, iter, survey, dataFilename):
     return 'distanceQuantiles.' + str(ngauss) + 'gauss.dQ' + str(quantile) + '.' + iter + '.' + survey + '.' + dataFilename
@@ -13,97 +13,106 @@ def distanceFilename(ngauss, quantile, iter, survey, dataFilename):
 def dustFilename(ngauss, quantile, iter, survey, dataFilename):
     return 'dustCorrection.'    + str(ngauss) + 'gauss.dQ' + str(quantile) + '.' + iter + '.' + survey + '.' + dataFilename
 
-def dataArrays(survey='2MASS'):
-
-    tgas = fits.getdata("stacked_tgas.fits", 1)
-    Apass = fits.getdata('tgas-matched-apass-dr9.fits')
-    twoMass = fits.getdata('tgas-matched-2mass.fits')
-
-    if survey == 'APASS':
-        mag1 = 'B'
-        mag2 = 'V'
-        absmag = 'G'
-        xlabel='B-V'
-        ylabel = r'M$_\mathrm{G}$'
-        xlim = [-0.2, 2]
-        ylim = [9, -2]
-
-    if survey == '2MASS':
-        mag1 = 'J'
-        mag2 = 'K'
-        absmag = 'J'
-        xlabel = 'J-K$_s$'
-        ylabel = r'M$_\mathrm{J}$'
-        xlim = [-0.25, 1.25]
-        ylim = [6, -4]
-
-    bandDictionary = {'B':{'key':'bmag', 'err_key':'e_bmag', 'array':twoMass},
-                     'V':{'key':'vmag', 'err_key':'e_vmag', 'array':Apass},
-                     'J':{'key':'j_mag', 'err_key':'j_cmsig', 'array':twoMass},
-                     'K':{'key':'k_mag', 'err_key':'k_cmsig', 'array':twoMass},
-                     'G':{'key':'phot_g_mean_mag', 'array':tgas}}
-
[…]

### 2017-03-18 paperDistance/paper.tex

--- 2017-03-17 paperDistance/paper.tex
@@
   stellar physics are generic and implicit (for example, that the
-  \cmd\ is somehow smooth and compact).
+  \cmd\ is somehow smooth and compact). Dust model dependent on stellar models?
 \end{itemize}

### 2017-03-18 paperDistance/paper.tex
subsection heading
references

@@
 
-The \gaia\ Mission (DWH CITE) will deliver more than a billion stellar distances.
+The \gaia\ Mission (\citealt{gaia16}) will deliver more than a billion stellar distances.
 Only a small fraction (but large number) of these distance
@@
 low-resolution $B_p\,R_p$ spectrophotometry, are used to provide
-distance estimates to more distant stars (DWH CITE).
+distance estimates to more distant stars.
 The full stack required for these distance inferences is complex.
 It involves modeling not just stars, but also the dust in the Milky Way,
-and the response of the telescope itself (DWH CITE).
+and the response of the telescope itself.
 
-With projects like \project{The Cannon} (DWH CITE) and \project{Avast}
-(DWH CITE), we are exploring the extent to which our predictive models of
+With projects like \project{The Cannon} (\citealt{ness15}; \citealt{casey17}; \citealt{ho17})
+and \project{Avast} (Bedell et al.\ in preparation),
+we are exploring the extent to which our predictive models of
 stars could be purely data-driven or statistical.
@@
 Here we build and use just such a data-driven model.  In particular,
-we use all of the \gaia\ \tgas\ data, matched to the
+we use all of the \gaia\ \tgas\ data (\citealt{tgas}), matched to the
 \tmass\ photometry (\citealt{skrutskie06}), and in the footprint of
@@
 previously to model the velocity distribution of stars in the Solar
-Neighborhood (DWH CITE BOTH) and to perform quasar target selection
-in \project{Sloan Digital Sky Survey} data (DWH CITE BOTH).
+Neighborhood (\citealt{hogg05}; \citealt{bovy09}) and to
+perform quasar target selection
+in \project{Sloan Digital Sky Survey} data (\citealt{xdqso}; \citealt{xdqsoz}).
 Its advantage over other deconvolution methods is that it takes as input,
@@
 Along those lines, this paper can be seen as a companion to a related
-project (CITE BORIS), in which we use a slightly less appropriate
+project (Leistedt et al. in preparation), in which we use a slightly less appropriate
 model for the \cmd, but in which we take a much more principled
@@
 
-DWH: Placeholder for the output format for our data products. Should
-be the last paragraph or two of this section.
+\subsection{Output data}
+DWH: Placeholder for the output format for our data products.
 
@@
 safe, and it is a goal to explore full hierarchical modeling in the coming
-years (and see, for example, CITE BORIS).
+years (and see, for example, Leistedt et al. in preparation).
 
@@
 However, there are important differences, as there are with all
-probabilistic catalogs (HOGG CITE LANG, FINKBEINER), between the
+probabilistic catalogs (\citealt{hogg11}; \citealt{portillo17}), between the
 measurements and the de-noised measurements.

### 2017-03-18 paperDistance/paper.tex
references

@@
 Here we build and use just such a data-driven model. In particular,
-we use all of the \gaia\ \tgas\ data (\citealt{tgas}), matched to the
+we use all of the \gaia\ \tgas\ data (\citealt{tgas};
+these data were generated by the methodology of \citealt{michalik15}), matched to the
 \tmass\ photometry (\citealt{skrutskie06}), and in the footprint of
-the \panstarrs\ Dust Map (CITE GREEN), to make a model of the
+the \panstarrs\ Dust Map (\citealt{green16}), to make a model of the
 noise-free (or low-noise) color--magnitude diagram (\cmd) of stars.
@@
 However, since the data set is large, the \xd\ approximation is not bad; it is
-sometimes known as the ``empirical Bayes'' method, and is well studied (DWH REFS).
+sometimes known as the ``empirical Bayes'' method, and is well studied\footnote{For a start, try \url{https://en.wikipedia.org/wiki/Empirical_Bayes_method}.}.
 
@@
 On a different tack, there has been some discussion in the literature
-about how to use a parallax responsibly to infer a distance (CITE TRI+CBJ).
+about how to use a parallax responsibly to infer a distance (\citealt{astraatmadja16a}; \citealt{astraatmadja16b}).
 These papers use the \gaia\ measurement to construct a parallax likelihood,
@@
   (it leads to under-deconvolution).
-\item We treat the \panstarrs-based three-dimensional dust maps (CITE
-  SCHLAFLY) as correct in their median dust estimates, and their
+\item We treat the \panstarrs-based three-dimensional dust maps (\citealt{green15})
+  as correct in their median dust estimates, and their
   effects on color and magnitude. As we optimize the empirical-Bayes model, we iteratively update each star's
@@
 Instead of assuming a functional form for the spatial distribution of
-stars in the Milky Way (CITE TRI--CORYN), or relying on stellar
-models, we build a noise-deconvolved estimate of the color--magnitude
+stars in the Milky Way (\citealt{astraatmadja16b}), or relying on stellar
+models (\citealt{gaia16}), we build a noise-deconvolved estimate of the color--magnitude
 diagram (\cmd) from the observed colors and absolute magnitudes of our

### 2017-03-20 demo_plots.py
plot aesthetics

@@ -59,13 +59,13 @@ def absMagKinda2absMag(absMagKinda):
     return 5.*np.log10(10.*absMagKinda_in_arcseconds)
 
 
-def plot_sample(x_true, y_true, x, y, samplex, sampley, xdgmm, xlabel='x', ylabel='y', xerr=None, yerr=None):
+def plot_sample(x_true, y_true, x, y, samplex, sampley, xdgmm, xlabel='x', ylabel='y', xerr=None, yerr=None, ylim=(6, -6), xlim=(0.5, 1.5)):
     setup_text_plots(fontsize=16, usetex=True)
     plt.clf()
-    alpha = 0.01
+    alpha = 0.1
     alpha_points = 0.01
-    figData = plt.figure(figsize=(12, 4.5))
-    figPrior = plt.figure(figsize=(12, 4.5))
+    figData = plt.figure(figsize=(12, 5.5))
+    figPrior = plt.figure(figsize=(12, 5.5))
     for fig in [figData, figPrior]:
         fig.subplots_adjust(left=0.1, right=0.95,
                             bottom=0.1, top=0.95,
@@ -89,8 +89,8 @@ def plot_sample(x_true, y_true, x, y, samplex, sampley, xdgmm, xlabel='x', ylabe
 
     ax4 = figPrior.add_subplot(122)
     ax4.scatter(samplex, sampley, s=4, lw=0, c='k', alpha=alpha)
-    xlim = ax4.get_xlim()
-    ylim = [10, -5] #ax3.get_ylim()
+    #xlim = ax4.get_xlim()
+    #ylim = ylim #ax3.get_ylim()
 
 
     titles = ["Observed Distribution", "Obs+Noise Distribution",
@@ -100,7 +100,7 @@ def plot_sample(x_true, y_true, x, y, samplex, sampley, xdgmm, xlabel='x', ylabe
     ax = [ax1, ax2, ax3, ax4]
 
     for i in range(4):
-        ax[i].set_xlim(-0.5, 1.5)
+        ax[i].set_xlim(xlim)
         ax[i].set_ylim(ylim[0], ylim[1]*1.1)
 
         #ax[i].xaxis.set_major_locator(plt.MultipleLocator([-1, 0, 1]))

### 2017-03-20 testXD.py

@@ -934,7 +934,7 @@ if __name__ == '__main__':
         xlabel = 'J-K$_s$'
         ylabel = r'M$_\mathrm{J}$'
         xlim = [-0.25, 1.25]
-        ylim = [6, -4]
+        ylim = [6, -6]
 
     iteration = ['1st', '2nd', '3rd', '4th', '5th', '6th', '7th', '8th', '9th', '10th']
     previteration = ['0th', '1st', '2nd', '3rd', '4th', '5th', '6th', '7th', '8th', '9th']
@@ -962,7 +962,7 @@ if __name__ == '__main__':
             assert np.sum(dustEBV) != 0.0, 'dust for iteration ' + str(iter) +  ' not read in properly'
 
         color = colorArray(mag1, mag2, dustEBV, bandDictionary)
-        absMagKinda, apparentMagnitude = absMagKindaArray(absmag, dustEBV, bandDictionary)
+        absMagKinda, apparentMagnitude = absMagKindaArray(absmag, dustEBV, bandDictionary, tgas['parallax'])
 
         color_err = np.sqrt(bandDictionary[mag1]['array'][bandDictionary[mag1]['err_key']]**2. + bandDictionary[mag2]['array'][bandDictionary[mag2]['err_key']]**2.)
         absMagKinda_err = tgas['parallax_error']*10.**(0.2*bandDictionary[absmag]['array'][bandDictionary[absmag]['key']])
@@ -1006,8 +1006,11 @@ if __name__ == '__main__':
 
         #plot 2x2 visual of prior w/ samples
         sample = xdgmm.sample(Nsamples)
+        yerr_minus = absMagKinda2absMag(absMagKinda+absMagKinda_err) - absMagKinda2absMag(absMagKinda)
+        yerr_plus = absMagKinda2absMag(absMagKinda) - absMagKinda2absMag(absMagKinda-absMagKinda_err)
+
         dp.plot_sample(color, absMagKinda2absMag(absMagKinda), color, absMagKinda2absMag(absMagKinda),
-                    sample[:,0],absMagKinda2absMag(sample[:,1]),xdgmm, xerr=color_err, yerr=absMagKinda2absMag(absMagKinda_err), xlabel=xlabel, ylabel=ylabel)
[…]

### 2017-03-20 tightPosterior.py

@@ -41,7 +41,7 @@ if survey == '2MASS':
     ylim = [6, -4]
 
 
-fig, ax = plt.subplots(2, 3, figsize=(12, 7))
+fig, ax = plt.subplots(2, 3, figsize=(15, 7))
 ax = ax.flatten()
 
 tgas, twoMass, Apass, bandDictionary, indices = testXD.dataArrays()
@@ -79,8 +79,11 @@ xparallaxMAS = np.linspace(0, 10, nPosteriorPoints)
 for i in range(np.max(digit)):
     currentInd = np.where(digit == i)[0]
     index = currentInd[np.random.randint(0, high=np.sum(digit==i))]
+    yerr_minus = testXD.absMagKinda2absMag(absMagKinda[index]+absMagKinda_err[index]) - testXD.absMagKinda2absMag(absMagKinda[index])
+    yerr_plus = testXD.absMagKinda2absMag(absMagKinda[index]) - testXD.absMagKinda2absMag(absMagKinda[index]-absMagKinda_err[index])
+    print yerr_minus, yerr_plus
     ax[0].scatter(color[index], testXD.absMagKinda2absMag(absMagKinda[index]), c='black')
-    ax[0].errorbar(color[index], testXD.absMagKinda2absMag(absMagKinda[index]), xerr=color_err[index], yerr=testXD.absMagKinda2absMag(absMagKinda_err[index]), fmt="none", zorder=0, lw=0.5, mew=0, alpha=1.0, color='black', ecolor='black')
+    ax[0].errorbar(color[index], testXD.absMagKinda2absMag(absMagKinda[index]), xerr=[[color_err[index], color_err[index]]], yerr=[[yerr_minus, yerr_plus]], fmt="none", zorder=0, lw=0.5, mew=0, alpha=1.0, color='black', ecolor='black')
     ax[0].annotate(str(i+1), (color[index]+0.02, testXD.absMagKinda2absMag(absMagKinda[index])+0.02))
     meanData, covData = testXD.matrixize(color[index], absMagKinda[index], color_err[index], absMagKinda_err[index])
     meanPrior, covPrior = testXD.matrixize(color[index], absMagKinda[index], color_err[index], 1e5)
@@ -108,11 +111,17 @@ for i in range(np.max(digit)):
     priorParallax = summedPriorAbsMagKinda*10.**(0.2*apparentMagnitude[index])
[…]

### 2017-03-20 paperDistance/paper.tex

@@
 \tmass\ photometry (\citealt{skrutskie06}), and in the footprint of
-the \panstarrs\ Dust Map (\citealt{green16}), to make a model of the
+the \panstarrs\ Dust Map (\citealt{green15}), to make a model of the
 noise-free (or low-noise) color--magnitude diagram (\cmd) of stars.

### 2017-03-21 paperDistance/paper.tex
why does this work section

@@
 \centering
-  \includegraphics[width=\textwidth]{lineExample.png}
+  \includegraphics[width=\textwidth]{toy.png}
 \caption{Why does this work: Fitting a line}
@@
 
-LA+DWH: Do we we want to put a pedagogical section in right here?
+In this paper we use the denoised, observed \cmd\ as a prior to infer more precise parallax pdfs from \gaia\ measurements.
+For a toy example to show how this method works, instead of using the \cmd\ as our prior to get more precise measurements of parallax, we use a straight line $y = mx + b$ with some gaussian thickness $\Delta$ as our prior to infer some posterior belief over $y_{true,n}$ from noisy $y_n$ data.
+Here the thickness of the true distribution mimics the true thickness of the \cmd\ due to age, mass and metallicity.
+To set up the problem we have the true underlying distribution $y_{true, n} = m_{true}\,x_n + b_{true} + \Delta_n$ where $\Delta_n$ is drawn from $\mathcal{N}(\Delta_n \given 0, t^2)$, where $\mathcal{N}$ represents a Gaussian function, here with mean $0$ and variance $t^2$.
+We also have some noisy data $y_n$ with measurement uncertainty $\sigma_n$.
+To generate a measurement uncertainty $\sigma_n$, we draw from a uniform distribution a value $z_n$, and then stretch the value to get some more significant outliers: $\sigma_n = (2z_n)^3$.
+From this we generate our noisy data $y_n = y_{true,n} + \mathcal{N}(\mathrm{noise} \given 0, \sigma_n^2)$.
+So we have a thick true distribution, and some noisy draws from this distribution.
 
-Hogg would propose a straight-line fit (possibly with intrinsic
-width); showing that it can be used to de-noise noisy y-data.
+First we determine the underlying, true distribution that the noisy data $y_n$ are drawn from.
+This denoised distribution is our prior.
+We then use this pior to infer a posterior belief about the true values, $p(y_{true,n} \given y_n, \tilde{\sigma}_n^2)$.
+When the true values are inferred in this way, the associated variance $\tilde{\sigma}_n^2$ is the convolution of the true distribution variance $t^2$ and the measurement variance $\sigma_n^2$.
+Figure \ref{fig:line}, left panel, shows 1024 realizations of the noisy values $y_n$ and their associated measurement uncertainties $\sigma_n$ as the grey points.
+Here we highlight 5 random points in black to highlight the method, but the same method is being applied to all the data.
+We denoise these points using marginalized maximum likelihood estimation (MML), which marginalizes over the true data values (we never need to know them) and maximizes the likelihood of the data.
+Here our likelihood for each individual datum is $p(y_n \given m, b, t^2, \sigma_n^2) = \mathcal{N}(y_n \given m\,x_n + b, t^2 + \sigma_n^2)$ and the likelihood of the full data set is $\prod_n\, p(y_n \given m, b, t^2, \sigma_n^2)$.
+Maximizing this likelihood is the same as minimizing the $\chi^2 = \sum_n \frac{[y_n - [m\,x_n + b]]^2}{[\sigma_n^2 + t_n^2]} + \sum_n\,ln(2\pi\,[\sigma_n^2 + t^2])$.
+This MML gives us the best fit parameters $\hat{m}$, $\hat{b}$, and $\hat{t}$.
+In Figure \ref{fig:line}, center panel, the true functions $y_{true} = m_{true}\,x + b_{true} \pm t_{true}$ are shown as the red lines, and the best fit MML functions $\hat{y}_{true} = \hat{m}\, x + \hat{b} \pm \hat{t}$ are shown as the blue lines.
+Now that we have an estimate of the true underlying distribution $\hat{y}_{true}$, we use this as a prior to infer a posterior belief of the true values for each datum $p(y_{true,n} \given y_n, \tilde{\sigma}_n^2)$
+
+So
+\begin{eqnarray}
+p(y_{true,n} \given y_n, \tilde{\sigma}_n) &=& p(y_n \given y_{true,n}, \sigma_n)\,p(y_{true,n}) \\
+p(y_n \given y_{true,n}, \sigma_n) &=& \mathcal{N}(y_n \given y_{true,n}, \sigma_n) \\
+p(y_{true,n}) &=& \mathcal{N}(y_{true,n} \given \hat{m}\, x_n + \hat{b}, \hat{t})
+\label{eq:toyBayes}
+\end{eqnarray}
+
+Since everything is Gaussian, this can be done analytically by multiplying Gaussians.
+The blue points in Figure \ref{fig:line}, center panel, represent the posterior belief of the true value $p(y_{true,n} \given y_n, \tilde{\sigma}_n)$.
+The black points represent the original, noisy data $y_n \pm \sigma_n$, the same as those highlighted in panel A, and the red points are the true values $y_{true, n}$.
+Finally, Figure \ref{fig:line}, right panel, shows the full set of posterior beliefs for each $y_n$.
+The points are colored by their measurement variance $\sigma_n^2$ to show that noisier measurements are more influenced by the prior and get pulled closer to the center of the MML distribution, our prior, while more certain measurements remain closer to their measured values $y_n$.
 

### 2017-03-21 toy/toy.py

@@ -1,13 +1,27 @@
 import numpy as np
 import scipy.optimize as op
+import sys
 
-def make_fake_data(N=512):
+
+def make_fake_data(parsTrue, N=512):
     np.random.seed(42)
-    mtrue, btrue, ttrue = -1.37, 0.2, 0.4
+    mtrue, btrue, ttrue = parsTrue
     xns = np.random.uniform(size=N) * 2. - 1.
-    yntrues = mtrue * xns + btrue + np.random.normal(size=N) * ttrue
-    sigmans = (np.random.uniform(size=N) * 1.5) ** 3
-    yns = yntrues + np.random.normal(size=N) * sigmans
+    tns = np.random.normal(scale=ttrue, size=N)
+    #tns = np.random.normal(size=N) * ttrue
+    yntrues = mtrue * xns + btrue + tns
+    sigmans = (np.random.uniform(size=N) * 2.) ** 3.
+    #yns = yntrues + np.random.normal(size=N) * sigmans
+    yns = yntrues + np.random.normal(scale=sigmans)
+    fig, ax = plt.subplots(1,3)
+    ax[0].scatter(xns, tns, s=2)
+    ax[0].set_ylabel('sampled noise from ttrue gaussian')
+    ax[1].scatter(xns, sigmans, s=2)
+    ax[1].set_ylabel('sigmans')
+    ax[2].scatter(xns, np.random.normal(scale=sigmans), s=2)
+    ax[2].set_ylabel('sampled noise from sigmans gaussian')
+    plt.tight_layout()
+    fig.savefig('laurenUnderstanding.png')
     return xns, yns, sigmans, yntrues
 
 def objective(pars, xns, yns, sigmas):
@@ -28,37 +42,67 @@ def denoise_one_datum(xn, yn, sigman, m, b, t):
         np.sqrt(1. / (s2inv + t2inv))
 
 if __name__ == "__main__":
-    import pylab as plt
+    import matplotlib.pyplot as plt
+    import matplotlib as mpl
+    mpl.rcParams['xtick.labelsize'] = 18
+    mpl.rcParams['ytick.labelsize'] = 18
+    nexamples =  np.int(sys.argv[1])
+    xlim = (-1, 1)
+    ylim = (-5, 5)
[…]

### 2017-03-21 paperDistance/paper.tex
toy example caption

@@
   \includegraphics[width=\textwidth]{toy.png}
-\caption{Why does this work: Fitting a line}
+\caption{ {\bf Why does this work}: An example of generating posterior beliefs of noisy data by using the denoised distribution of data as a prior.
+The left panel is 1024 realizations of noisy data drawn from the gaussian distribution represented by the $1\sigma$ red lines shown in the center panel.
+The black points are 5 randomly chosen data to highlight the method.
+The blue lines in the center panel represent the MML best estimate for the underlying distribution, which denoises the data.
+The blue points in the center panel represent the posterior belief of the true values of the noisy black points, and the red points represent the actual true values.
+The right panel shows the posterior belief of the true value for all the points.
+They are colored by the measurement variance showing that data with larger measured variance are more influenced by the prior and lie towards the center of the MML distribution, or prior.
+Where as more certain measurements tend to lie closer to their measured value.
+The blue points are better than the black for doing stuff.}
 \label{fig:line}
@@
 
-So
+So using Bayes theorem
 \begin{eqnarray}

### 2017-03-22 paperDistance/paper.tex
small tweaks, more tomorrow

@@
 In this paper we use the denoised, observed \cmd\ as a prior to infer more precise parallax pdfs from \gaia\ measurements.
-For a toy example to show how this method works, instead of using the \cmd\ as our prior to get more precise measurements of parallax, we use a straight line $y = mx + b$ with some gaussian thickness $\Delta$ as our prior to infer some posterior belief over $y_{true,n}$ from noisy $y_n$ data.
+Here we will demonstrate this method with a simple toy mode.
+Instead of using the \cmd\ as our prior to get more precise measurements of parallax, we use a straight line $y = mx + b$ with some gaussian thickness $t$ as our prior.
+We use this prior to infer some posterior belief over $y_{\true,n}$ from noisy $y_n$ data.
 Here the thickness of the true distribution mimics the true thickness of the \cmd\ due to age, mass and metallicity.
-To set up the problem we have the true underlying distribution $y_{true, n} = m_{true}\,x_n + b_{true} + \Delta_n$ where $\Delta_n$ is drawn from $\mathcal{N}(\Delta_n \given 0, t^2)$, where $\mathcal{N}$ represents a Gaussian function, here with mean $0$ and variance $t^2$.
+To set up the problem we have the true underlying distribution $y_{\true, n} = m_{\true}\,x_n + b_{\true} + \Delta_n$ where $\Delta_n$ is drawn from $\mathcal{N}(\Delta_n \given 0, t^2)$, where $\mathcal{N}$ represents a Gaussian function, here with mean $0$ and variance $t^2$.
 We also have some noisy data $y_n$ with measurement uncertainty $\sigma_n$.
 To generate a measurement uncertainty $\sigma_n$, we draw from a uniform distribution a value $z_n$, and then stretch the value to get some more significant outliers: $\sigma_n = (2z_n)^3$.
-From this we generate our noisy data $y_n = y_{true,n} + \mathcal{N}(\mathrm{noise} \given 0, \sigma_n^2)$.
+From this we generate our noisy data $y_n = y_{\true,n} + \epsilon$ where $\epsilon$ is drawn from $\mathcal{N}(\epsilon \given 0, \sigma_n^2)$.
 So we have a thick true distribution, and some noisy draws from this distribution.
@@
 This denoised distribution is our prior.
-We then use this pior to infer a posterior belief about the true values, $p(y_{true,n} \given y_n, \tilde{\sigma}_n^2)$.
+We then use this pior to infer a posterior belief about the true values, $p(y_{\true,n} \given y_n, \tilde{\sigma}_n^2)$.
 When the true values are inferred in this way, the associated variance $\tilde{\sigma}_n^2$ is the convolution of the true distribution variance $t^2$ and the measurement variance $\sigma_n^2$.
@@
 This MML gives us the best fit parameters $\hat{m}$, $\hat{b}$, and $\hat{t}$.
-In Figure \ref{fig:line}, center panel, the true functions $y_{true} = m_{true}\,x + b_{true} \pm t_{true}$ are shown as the red lines, and the best fit MML functions $\hat{y}_{true} = \hat{m}\, x + \hat{b} \pm \hat{t}$ are shown as the blue lines.
-Now that we have an estimate of the true underlying distribution $\hat{y}_{true}$, we use this as a prior to infer a posterior belief of the true values for each datum $p(y_{true,n} \given y_n, \tilde{\sigma}_n^2)$
+In Figure \ref{fig:line}, center panel, the true functions $y_{\true} = m_{\true}\,x + b_{\true} \pm t_{\true}$ are shown as the red lines, and the best fit MML functions $\hat{y}_{\true} = \hat{m}\, x + \hat{b} \pm \hat{t}$ are shown as the blue lines.
+Now that we have an estimate of the true underlying distribution $\hat{y}_{\true}$, we use this as a prior to infer a posterior belief of the true values for each datum $p(y_{\true,n} \given y_n, \tilde{\sigma}_n^2)$.
+Here is a good point to step back and acknowledge that our method is odd/strange/different/new (which maybe you're also currently thinking but can't quite express why yet).
+We are using the set of data to determine the denoised distribution, and then using that denoised distribution to make inferences on the data.
+We are technically using the data twice!
+If we were going to be fully, technically correct, we should denoise the distribution, but leave one datum out; the one datum we want to infer the true y value for\footnote{DWH: Remember what this brain dump is for}.
+Leaving one out, and then estimating the prior and then inferring the posterior for each 1024 points, let alone over a million points is way too expensive.
+Luckily with large data sets, even at 1024, 1 datum is $\sim0.1\%$ of your dataset and therefore leaving that point out will not change the MML parameters of your denoised distribution, and with the \gaia\ dataset of over a million stars, we can safely step into the Empirical Bayes regime.
 
@@
 \begin{eqnarray}
-p(y_{true,n} \given y_n, \tilde{\sigma}_n) &=& p(y_n \given y_{true,n}, \sigma_n)\,p(y_{true,n}) \\
-p(y_n \given y_{true,n}, \sigma_n) &=& \mathcal{N}(y_n \given y_{true,n}, \sigma_n) \\
-p(y_{true,n}) &=& \mathcal{N}(y_{true,n} \given \hat{m}\, x_n + \hat{b}, \hat{t})
+p(y_{\true,n} \given y_n, \tilde{\sigma}_n) &=& p(y_n \given y_{\true,n}, \sigma_n)\,p(y_{\true,n}) \\
+p(y_n \given y_{\true,n}, \sigma_n) &=& \mathcal{N}(y_n \given y_{\true,n}, \sigma_n) \\
+p(y_{\true,n}) &=& \mathcal{N}(y_{\true,n} \given \hat{m}\, x_n + \hat{b}, \hat{t})
 \label{eq:toyBayes}
@@
 Since everything is Gaussian, this can be done analytically by multiplying Gaussians.
-The blue points in Figure \ref{fig:line}, center panel, represent the posterior belief of the true value $p(y_{true,n} \given y_n, \tilde{\sigma}_n)$.
-The black points represent the original, noisy data $y_n \pm \sigma_n$, the same as those highlighted in panel A, and the red points are the true values $y_{true, n}$.
+The blue points in Figure \ref{fig:line}, center panel, represent the posterior belief of the true value $p(y_{\true,n} \given y_n, \tilde{\sigma}_n)$.
+The black points represent the original, noisy data $y_n \pm \sigma_n$, the same as those highlighted in panel A, and the red points are the true values $y_{\true, n}$.
 Finally, Figure \ref{fig:line}, right panel, shows the full set of posterior beliefs for each $y_n$.
 The points are colored by their measurement variance $\sigma_n^2$ to show that noisier measurements are more influenced by the prior and get pulled closer to the center of the MML distribution, our prior, while more certain measurements remain closer to their measured values $y_n$.
+
+LA: Conclusions and stuff about why this stuff is good
 

### 2017-03-22 toy/toy.py

@@ -41,38 +41,26 @@ def denoise_one_datum(xn, yn, sigman, m, b, t):
     return (yn * s2inv + (m * xn + b) * t2inv) / (s2inv + t2inv), \
         np.sqrt(1. / (s2inv + t2inv))
 
-if __name__ == "__main__":
-    import matplotlib.pyplot as plt
-    import matplotlib as mpl
-    mpl.rcParams['xtick.labelsize'] = 18
-    mpl.rcParams['ytick.labelsize'] = 18
-    nexamples =  np.int(sys.argv[1])
-    xlim = (-1, 1)
-    ylim = (-5, 5)
-    fig, axes = plt.subplots(1, 3, figsize=(16, 5))
-    axes = axes.flatten()
-    mtrue, btrue, ttrue = -1.37, 0.2, 0.8
-    parsTrue = [mtrue, btrue, ttrue] #m, b, t
+def plot(fig, axes, figNoise, axesNoise, xns, yns, sigmans, yntrues, m, b, t, mtrue, btrue, ttrue, ydns, sigmadns, nexamples=5):
 
     alpha_all = 0.05
     alpha_chosen = 1.0
+    xlim = (-1, 1)
+    ylim = (-5, 5)
 
     dataMap = mpl.cm.get_cmap('Blues')
     dataColor = dataMap(0.75)
     trueMap = mpl.cm.get_cmap('Reds')
     trueColor = trueMap(0.75)
-    xns, yns, sigmans, yntrues = make_fake_data(parsTrue)
+
     for ax in axes[:-1]:
         ax.errorbar(xns, yns, yerr=sigmans, fmt="o", color="k", alpha=alpha_all ,mew=0)
         ax.errorbar(xns[0:nexamples], yns[0:nexamples], yerr=sigmans[0:nexamples], fmt="o", color="k", zorder=37, alpha=alpha_chosen, mew=0)
 
-
-    m, b, t = get_best_fit_pars(xns, yns, sigmans)
-    print m, b, t
     xp = np.array(xlim)
     axes[1].plot(xp, m * xp + b + t, color=dataColor)
     axes[1].plot(xp, m * xp + b - t, color=dataColor)
-    axes[1].scatter(xns[0:nexamples], yntrues[0:nexamples], c=trueColor, s=20., lw=2, zorder=36, alpha=alpha_chosen, facecolors='None')
[…]

### 2017-04-10 paperDistance/paper.tex
various tweaks to the paper

@@
   \includegraphics[width=\textwidth]{toy.png}
-\caption{ {\bf Why does this work}: An example of generating posterior beliefs of noisy data by using the denoised distribution of data as a prior.
-The left panel is 1024 realizations of noisy data drawn from the gaussian distribution represented by the $1\sigma$ red lines shown in the center panel.
+\caption{ {\bf Why does this work}: An example of generating posterior beliefs of true values from noisy data by using the noise-deconvolved distribution of data as a prior.
+The upper left panel visualizes the simplified toy distribution.
+The red lines are the $1\sigma$ contours of the distribution, and the red points are 1024 samples from the distribution.
+These same $1\sigma$ contours are drawn in the bottom left panel.
+In the upper right and lower left panel, the grey points are these same 1024 samples of the distribution but with some measurement uncertainty added to them.
+The error bars show the $1\sigma$ measurement uncertainties.
 The black points are 5 randomly chosen data to highlight the method.
-The blue lines in the center panel represent the MML best estimate for the underlying distribution, which denoises the data.
-The blue points in the center panel represent the posterior belief of the true values of the noisy black points, and the red points represent the actual true values.
-The right panel shows the posterior belief of the true value for all the points.
-They are colored by the measurement variance showing that data with larger measured variance are more influenced by the prior and lie towards the center of the MML distribution, or prior.
-Where as more certain measurements tend to lie closer to their measured value.
+In the lower left panel, the blue lines represent the noise deconvolved, best estimate for the underlying distribution.
+The blue points represent the posterior belief of the true values of the noisy black points using the noise deconcolved distribution as the prior.
+The red points represent the actual true values.
+The lower right panel shows the posterior belief of the true value for all the points.
+They are colored by the measurement uncertainty, showing that data with larger measured variance are more influenced by the prior and lie towards the center of the noise deconvolved distribution, or prior.
+In contrast, more certain measurements tend to lie closer to their measured value.
 The blue points are better than the black for doing stuff.}
-\label{fig:line}
+\label{fig:toy}
 \end{figure}
 
-In this paper we use the denoised, observed \cmd\ as a prior to infer more precise parallax pdfs from \gaia\ measurements.
-Here we will demonstrate this method with a simple toy mode.
-Instead of using the \cmd\ as our prior to get more precise measurements of parallax, we use a straight line $y = mx + b$ with some gaussian thickness $t$ as our prior.
-We use this prior to infer some posterior belief over $y_{\true,n}$ from noisy $y_n$ data.
-Here the thickness of the true distribution mimics the true thickness of the \cmd\ due to age, mass and metallicity.
-To set up the problem we have the true underlying distribution $y_{\true, n} = m_{\true}\,x_n + b_{\true} + \Delta_n$ where $\Delta_n$ is drawn from $\mathcal{N}(\Delta_n \given 0, t^2)$, where $\mathcal{N}$ represents a Gaussian function, here with mean $0$ and variance $t^2$.
-We also have some noisy data $y_n$ with measurement uncertainty $\sigma_n$.
-To generate a measurement uncertainty $\sigma_n$, we draw from a uniform distribution a value $z_n$, and then stretch the value to get some more significant outliers: $\sigma_n = (2z_n)^3$.
-From this we generate our noisy data $y_n = y_{\true,n} + \epsilon$ where $\epsilon$ is drawn from $\mathcal{N}(\epsilon \given 0, \sigma_n^2)$.
-So we have a thick true distribution, and some noisy draws from this distribution.
+In this paper we use the noise deconvolved observed \cmd\ as a prior to infer more precise parallax pdfs from \gaia\ measurements.
+Here we will demonstrate this method with a simpler toy model.
+Instead of using the 2D distribution of the \cmd\ as our prior, for this simpler model we use a 2D distribution that is a 1D Gaussian in the y-direction, with a running mean $\mu = mx + b$ and some thickness $t$, as our prior.
+The $1\sigma$ contours as well as (noise free) samples from the distribution are shown in Figure \ref{fig:toy}, upper left panel.
+Noisy samples from this distribution are shown in the upper right panel.
+Using a representation of this distribution as our prior we can infer some posterior belief of $y_{\true,n}$, a true sample from the distribution, from $y_n$, a noisy sample of the distribution.
+Here the thickness of the distribution mimics the true thickness of the \cmd\ due to age, mass and metallicity.
+To set up the problem, we have the true underlying distribution of points and the noisy data
 
-First we determine the underlying, true distribution that the noisy data $y_n$ are drawn from.
-This denoised distribution is our prior.
-We then use this pior to infer a posterior belief about the true values, $p(y_{\true,n} \given y_n, \tilde{\sigma}_n^2)$.
+\begin{eqnarray}
+y_{\true, n} \, &=& \, m_{\true}\,x_n + b_{\true} + \Delta_n \\
+y_n \, &=& \, y_{true,n} + \epsilon_n
+\label{eq:ytrue}
+\end{eqnarray}
+
+where $\Delta_n$ is drawn from the Gaussian function $\mathcal{N}(\Delta_n \given 0, t^2)$, and $\epsilon_n$ is drawn from the Gaussian function $\mathcal{N}(\epsilon_n \given 0, \sigma_n^2)$, so we quote $\sigma_n$ as the measurement uncertainty.
+These represent true and noisy draws from a distribution that is a straight line with some thickness $t$, and are visualized the upper panels of Figure \ref{fig:toy}
+
+First we determine the underlying distribution that the noisy data $y_n$ are drawn from, shown as the blue lines in the lower left panel of Figure \ref{fig:toy}
+This denoised distribution becomes our prior.
+We then use this pior to infer a posterior belief about the true values, $p(y_{\true,n} \given y_n, \tilde{\sigma}_n^2)$, shown as the blue points in the lower panels of Figure \ref{fig:toy}
 When the true values are inferred in this way, the associated variance $\tilde{\sigma}_n^2$ is the convolution of the true distribution variance $t^2$ and the measurement variance $\sigma_n^2$.
-Figure \ref{fig:line}, left panel, shows 1024 realizations of the noisy values $y_n$ and their associated measurement uncertainties $\sigma_n$ as the grey points.
+Figure \ref{fig:toy}, upper right, shows 1024 realizations of the noisy values $y_n$ and their associated measurement uncertainties $\sigma_n$ as the grey points.
 Here we highlight 5 random points in black to highlight the method, but the same method is being applied to all the data.
 We denoise these points using marginalized maximum likelihood estimation (MML), which marginalizes over the true data values (we never need to know them) and maximizes the likelihood of the data.
-Here our likelihood for each individual datum is $p(y_n \given m, b, t^2, \sigma_n^2) = \mathcal{N}(y_n \given m\,x_n + b, t^2 + \sigma_n^2)$ and the likelihood of the full data set is $\prod_n\, p(y_n \given m, b, t^2, \sigma_n^2)$.
-Maximizing this likelihood is the same as minimizing the $\chi^2 = \sum_n \frac{[y_n - [m\,x_n + b]]^2}{[\sigma_n^2 + t_n^2]} + \sum_n\,ln(2\pi\,[\sigma_n^2 + t^2])$.
+Here our likelihood for each individual datum is
+
+\begin{equation}
+p(y_n \given m, b, t^2, \sigma_n^2) = \mathcal{N}(y_n \given m\,x_n + b, t^2 + \sigma_n^2)
+\label{eq:toyLike}
+\end{equation}
+
+and the likelihood of the full data set is
+
+\begin{equation}
+\prod_n\, p(y_n \given m, b, t^2, \sigma_n^2).
+\label{eq:toyLikeFull}
+\end{equation}
+
+Maximizing this likelihood is the same as minimizing
+
+\begin{equation}
+\chi^2 = \sum_n \frac{[y_n - [m\,x_n + b]]^2}{[\sigma_n^2 + t_n^2]} + \sum_n\,ln(2\pi\,[\sigma_n^2 + t^2])
+\label{eq:chisq}
+\end{equation}
+
 This MML gives us the best fit parameters $\hat{m}$, $\hat{b}$, and $\hat{t}$.
-In Figure \ref{fig:line}, center panel, the true functions $y_{\true} = m_{\true}\,x + b_{\true} \pm t_{\true}$ are shown as the red lines, and the best fit MML functions $\hat{y}_{\true} = \hat{m}\, x + \hat{b} \pm \hat{t}$ are shown as the blue lines.
+In Figure \ref{fig:toy}, center panel, the true functions $y_{\true} = m_{\true}\,x + b_{\true} \pm t_{\true}$ are shown as the red lines, and the best fit MML functions $\hat{y}_{\true} = \hat{m}\, x + \hat{b} \pm \hat{t}$ are shown as the blue lines.
 Now that we have an estimate of the true underlying distribution $\hat{y}_{\true}$, we use this as a prior to infer a posterior belief of the true values for each datum $p(y_{\true,n} \given y_n, \tilde{\sigma}_n^2)$.
+
 Here is a good point to step back and acknowledge that our method is odd/strange/different/new (which maybe you're also currently thinking but can't quite express why yet).
@@
 If we were going to be fully, technically correct, we should denoise the distribution, but leave one datum out; the one datum we want to infer the true y value for\footnote{DWH: Remember what this brain dump is for}.
-Leaving one out, and then estimating the prior and then inferring the posterior for each 1024 points, let alone over a million points is way too expensive.
+Leaving one out, and then estimating the prior and then inferring the posterior for each data point is way too expensive for the 1024 points in our toy model, and also for the 1.2 million stars in our actual data set.
 Luckily with large data sets, even at 1024, 1 datum is $\sim0.1\%$ of your dataset and therefore leaving that point out will not change the MML parameters of your denoised distribution, and with the \gaia\ dataset of over a million stars, we can safely step into the Empirical Bayes regime.
@@
 
-Since everything is Gaussian, this can be done analytically by multiplying Gaussians.
-The blue points in Figure \ref{fig:line}, center panel, represent the posterior belief of the true value $p(y_{\true,n} \given y_n, \tilde{\sigma}_n)$.
-The black points represent the original, noisy data $y_n \pm \sigma_n$, the same as those highlighted in panel A, and the red points are the true values $y_{\true, n}$.
-Finally, Figure \ref{fig:line}, right panel, shows the full set of posterior beliefs for each $y_n$.
+Since everything is Gaussian, this can be done analytically by multiplying Gaussians together.
+The blue points in Figure \ref{fig:toy}, lower right panel, represent the posterior belief of the true value $p(y_{\true,n} \given y_n, \tilde{\sigma}_n)$ for the highlighted black points.
+Finally, Figure \ref{fig:toy}, right panel, shows the full set of posterior beliefs for each $y_n$.
 The points are colored by their measurement variance $\sigma_n^2$ to show that noisier measurements are more influenced by the prior and get pulled closer to the center of the MML distribution, our prior, while more certain measurements remain closer to their measured values $y_n$.
+By doing this we have better estimates of the true y value for each point.
+Bringing the toy model back to our original problem, using the noise-deconvolved CMD as our prior allows us to get better estimates of the parallax for each star.
 
@@
   \includegraphics[width=\textwidth]{data.png}
-\caption{The observed \cmd: point estimates on the left and including error bars on the right.}
+\caption{The observed \cmd: To visualize the full dataset, in the left panel we use the point estimates of the \tmass\ $J-K$ color, the $J$ band apparent magnitude, and the \gaia\ parallax.
+The blue lines represent the $1\sigma$ and $2\sigma$ contours for the distribution.
+To give a sense of the uncertainties, in the right panel we subsample the dataset and include the $1\sigma$ uncertainties.
+The uncertainties are dominated by the parallax noise, with many stars diverging to infinitely far away and therefore very intrinsically bright.}
 \label{fig:data}
@@
 \subsection{Prior}
+
+\begin{figure}
+\centering
+  \includegraphics[width=\textwidth]{prior_ngauss128.png}
+\caption{Prior: the XD generated mixture of gaussians we use as our prior.
+The left panel shows a sampling of the prior, which is a mixture of gaussians shown in the right hand panel.
+The blue lines are the $1\sigma$ and $2\sigma$ contours.
+The gaussians have a funny shape due to the transformed space in which we fit the data.}
+\label{fig:prior}
+\end{figure}
+
 \begin{figure}
@@
   \includegraphics[width=\textwidth]{priorNgaussComparison.png}
-\caption{Prior: the XD generated mixture of gaussians we use as our prior}
+\caption{Prior: the XD generated mixture of gaussians we use as our prior.
+To determine the number of gaussians we used cross validation, using 70\% of our sample as a training set and 30\% as the test set.
+Increasing the number of gaussians we used continued to increase the likelihood of the test set of data, showing that we are no over fitting.
+Using more gaussians as our prior did not substantially change our posterior results, but it did start to highlight the features of the CMD.}
 \label{fig:prior}
@@
 
-LMA: what are our results?
-\subsection{Distances to M67}
+\begin{figure}
+\centering
+  \includegraphics[width=\textwidth]{posterior.png}
+\caption{5 example posteriors showing the priors and likelihoods as well}
+\label{fig:posterior}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=\textwidth]{posteriorCMD.png}
+\caption{the expectation values and subsampled error bars}
+\label{fig:posteriorCMD}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{delta.png}
+\caption{Fractional Change in Variance: The left panel shows the natural log of the fractional change in the variance of the posterior relative to the likelihood as a function of the stellar color.
+The coloring of the points in both panels is by the fractional change in variance to guide the coloring of the right panel.
+The blue lines are the $1\sigma$ and $2\sigma$ contours.
+The median fractional change is XXX.
+Some posteriors have significantly smaller variances shown in black, but some actually increase in variance, shown as the yellow bump at $(J-K)^C$ about XXX.
+The right panel shows the posterior point estimates on the CMD colored by fraction change in variance.
+There is significant changes for the red giant branch stars, as well as some main sequence stars.}
+\label{fig:delta}
+\end{figure}
+
+\subsection{Results: Distances to M67}
 \begin{figure}
@@
 
-\begin{figure}
-\centering
-  \includegraphics[width=\textwidth]{posterior.png}
-\caption{5 example posteriors showing the priors and likelihoods as well}
-\label{fig:posterior}
-\end{figure}
-
 \subsection{Output data}
 DWH: Placeholder for the output format for our data products.
+RA, DEC, l, b, tgas source ID, tycho-2 source ID, parallax samples, parallax posterior, flag (some fraction of posterior not in window, negative gaia parallax, ...)
 
@@
 
-Some discussion of possible failure modes. Including very rare kinds of stars.
+Some discussion of possible failure modes. Including very rare kinds of stars. Should we add gaussians to prior where we know stars exist that aren't well sampled in Gaia (ie WD)
 
@@
 like, eg, the Tri--Coryn papers. Favorably! And with \emph{fewer}
-assumptions, in fact!
+assumptions, in fact! What form should this comparison be in?
 

### 2017-04-10 comparePrior.py

@@ -24,6 +24,60 @@ def dustFilename(ngauss, quantile, iter, survey, dataFilename):
     return 'dustCorrection.'    + str(ngauss) + 'gauss.dQ' + str(quantile) + '.' + iter + '.' + survey + '.' + dataFilename
 
 
+def priorSample(ngauss=128, quantile=0.5, iter='8th', survey='2MASS', dataFilename='All.npz', Nsamples=1.2e6, xdgmmFilename='xdgmm.fit', xlabel='X', ylabel='Y', contourColor='k'):
+
+    setup_text_plots(fontsize=16, usetex=True)
+
+    xdgmm = XDGMM(filename=xdgmmFilename)
+    figPrior = plt.figure(figsize=(12, 5.5))
+    figPrior.subplots_adjust(left=0.1, right=0.95,
+                            bottom=0.15, top=0.95,
+                            wspace=0.1, hspace=0.1)
+    sample = xdgmm.sample(Nsamples)
+    negParallax = sample[:,1] < 0
+    nNegP = np.sum(negParallax)
+    while nNegP > 0:
+        sampleNew = xdgmm.sample(nNegP)
+        sample[negParallax] = sampleNew
+        negParallax = sample[:,1] < 0
+        nNegP = np.sum(negParallax)
+
+    samplex = sample[:,0]
+    sampley = testXD.absMagKinda2absMag(sample[:,1])
+    ax3 = figPrior.add_subplot(121)
+    alpha = 0.1
+    xlim = [-0.25, 1.25]
+    ylim = [6, -6]
+
+    levels = 1.0 - np.exp(-0.5 * np.arange(1.0, 2.1, 1.0) ** 2)
+    corner.hist2d(samplex, sampley, ax=ax3, levels=levels, bins=200, plot_datapoints=False, no_fill_contours=True, plot_density=False, color=contourColor)
+    ax3.scatter(samplex, sampley, s=1, lw=0, c='k', alpha=alpha)
+
+    ax4 = figPrior.add_subplot(122)
+    for i in range(xdgmm.n_components):
+        points = drawEllipse.plotvector(xdgmm.mu[i], xdgmm.V[i])
+        ax4.plot(points[0, :], testXD.absMagKinda2absMag(points[1,:]), 'k-', alpha=xdgmm.weights[i]/np.max(xdgmm.weights))
+
+    titles = ["Extreme Deconvolution\n  resampling",
+              "Extreme Deconvolution\n  cluster locations"]
[…]

### 2017-04-10 demo_plots.py

@@ -103,7 +103,7 @@ def kdeDensity(ax, x, y, threshold=0.01, bins=100, s=1, lw=0, alpha=1):
     #fig.colorbar(im)
     return ax
 
-def plot_sample(x, y, samplex, sampley, xdgmm, xlabel='x', ylabel='y', xerr=None, yerr=None, ylim=(6, -6), xlim=(0.5, 1.5), errSubsample=1.2e6, thresholdScatter=0.1, binsScatter=200, c='black',  norm=None, cmap=None, contourColor='k', posterior=False):
+def plot_sample(x, y, samplex, sampley, xdgmm, xlabel='x', ylabel='y', xerr=None, yerr=None, ylim=(6, -6), xlim=(0.5, 1.5), errSubsample=1.2e6, thresholdScatter=0.1, binsScatter=200, c='black',  norm=None, cmap=None, contourColor='k', posterior=False, ind=None):
     prng = np.random.RandomState(1234567890)
     setup_text_plots(fontsize=16, usetex=True)
     plt.clf()
@@ -124,7 +124,7 @@ def plot_sample(x, y, samplex, sampley, xdgmm, xlabel='x', ylabel='y', xerr=None
     #im = ax1.scatter(x, y, s=1, lw=0, c=c, alpha=alpha, norm=norm, cmap=cmap)
 
     ax2 = figData.add_subplot(122)
-    ind = prng.randint(0, len(x), size=errSubsample)
+    if ind is None: ind = prng.randint(0, len(x), size=errSubsample)
     #ax2.scatter(x[ind], y[ind], s=1, lw=0, c='black', alpha=alpha_points)
     ax2.errorbar(x[ind], y[ind], xerr=xerr[ind], yerr=[yerr[0][ind], yerr[1][ind]], fmt="none", zorder=0, mew=0, ecolor='black', alpha=0.5, elinewidth=0.5)
 

### 2017-04-10 tightPosterior.py

@@ -66,13 +66,24 @@ projectedDimension = 1  #which dimension to project the prior onto
 ndim = 2
 xparallaxMAS = np.linspace(0, 10, nPosteriorPoints)
 
+
+y = absMagKinda
+yplus  = y + absMagKinda_err
+yminus = y - absMagKinda_err
+parallaxErrGoesNegative = yminus < 0
+absMagYMinus = testXD.absMagKinda2absMag(yminus)
+absMagYMinus[parallaxErrGoesNegative] = -50.
+yerr_minus = testXD.absMagKinda2absMag(y) - absMagYMinus
+yerr_plus = testXD.absMagKinda2absMag(yplus) - testXD.absMagKinda2absMag(y)
+
+
 #plot likelihood and posterior in each axes
 for iteration in np.arange(20, 40):
-    fig, ax = plt.subplots(2, 3, figsize=(15, 7))
+    fig, ax = plt.subplots(2, 3, figsize=(15, 9))
     ax = ax.flatten()
     fig.subplots_adjust(left=0.1, right=0.9,
                                 bottom=0.1, top=0.9,
-                                wspace=0.4, hspace=0.4)
+                                wspace=0.4, hspace=0.5)
 
     #plot prior in upper left
     xdgmmFilename = 'xdgmm.' + str(ngauss) + 'gauss.dQ' + str(quantile) + '.' + iter + '.2MASS.All.npz.fit'
@@ -80,26 +91,17 @@ for iteration in np.arange(20, 40):
     testXD.plotPrior(xdgmm, ax[0], c='k', lw=1)
     ax[0].set_xlim(xlim)
     ax[0].set_ylim(ylim)
-    ax[0].set_xlabel('$(J-K)^C$')
-    ax[0].set_ylabel('$M_J^C$')
+    ax[0].set_xlabel('$(J-K)^C$', fontsize=18)
+    ax[0].set_ylabel('$M_J^C$', fontsize=18)
 
     for i in range(np.max(digit)):
-        currentInd = np.where((digit == i) & (tgas['parallax_error'] > 0.75))[0]
+        currentInd = np.where((digit == i))[0]
         index = currentInd[np.random.randint(0, high=len(currentInd))]
-        y = absMagKinda
-        yplus  = y + absMagKinda_err
-        yminus = y - absMagKinda_err
-        absMagYMinus = testXD.absMagKinda2absMag(yminus)
-        absMagYMinus[yminus <0] = -50.
[…]

### 2017-04-10 toy/toy.py

@@ -10,7 +10,7 @@ def make_fake_data(parsTrue, N=512):
     tns = np.random.normal(scale=ttrue, size=N)
     #tns = np.random.normal(size=N) * ttrue
     yntrues = mtrue * xns + btrue + tns
-    sigmans = (np.random.uniform(size=N) * 2.) ** 3.
+    sigmans = (np.random.uniform(size=N)*1.5) ** 3.
     #yns = yntrues + np.random.normal(size=N) * sigmans
     yns = yntrues + np.random.normal(scale=sigmans)
     fig, ax = plt.subplots(1,3)
@@ -53,34 +53,39 @@ def plot(fig, axes, figNoise, axesNoise, xns, yns, sigmans, yntrues, m, b, t, mt
     trueMap = mpl.cm.get_cmap('Reds')
     trueColor = trueMap(0.75)
 
-    for ax in axes[:-1]:
+    for ax in axes[1:-1]:
         ax.errorbar(xns, yns, yerr=sigmans, fmt="o", color="k", alpha=alpha_all ,mew=0)
         ax.errorbar(xns[0:nexamples], yns[0:nexamples], yerr=sigmans[0:nexamples], fmt="o", color="k", zorder=37, alpha=alpha_chosen, mew=0)
 
     xp = np.array(xlim)
-    axes[1].plot(xp, m * xp + b + t, color=dataColor)
-    axes[1].plot(xp, m * xp + b - t, color=dataColor)
-    axes[1].scatter(xns[0:nexamples], yntrues[0:nexamples], c=trueColor, lw=2, zorder=36, alpha=alpha_chosen, facecolors='None')
-    axes[1].plot(xp, mtrue*xp + btrue + ttrue, color=trueColor, zorder=35)
-    axes[1].plot(xp, mtrue*xp + btrue - ttrue, color=trueColor, zorder=34)
-    r1 = axes[1].add_patch(mpl.patches.Rectangle((-10,-10), 0.1, 0.1, color='black', alpha=alpha_chosen))
-    r2 = axes[1].add_patch(mpl.patches.Rectangle((-10,-10), 0.1, 0.1, color=trueColor, alpha=alpha_chosen))
-    r3 = axes[1].add_patch(mpl.patches.Rectangle((-10,-10), 0.1, 0.1, color=dataColor, alpha=alpha_chosen))
[…]

### 2017-04-17 paperDistance/paper.tex
paper additions

@@
 Noisy samples from this distribution are shown in the upper right panel.
-Using a representation of this distribution as our prior we can infer some posterior belief of $y_{\true,n}$, a true sample from the distribution, from $y_n$, a noisy sample of the distribution.
+Using a representation of this distribution as our prior we can infer some posterior belief of $y_{\true,n}$, a true sample of the distribution, from $y_n$, a noisy sample of the distribution.
 Here the thickness of the distribution mimics the true thickness of the \cmd\ due to age, mass and metallicity.
@@
 
-where $\Delta_n$ is drawn from the Gaussian function $\mathcal{N}(\Delta_n \given 0, t^2)$, and $\epsilon_n$ is drawn from the Gaussian function $\mathcal{N}(\epsilon_n \given 0, \sigma_n^2)$, so we quote $\sigma_n$ as the measurement uncertainty.
-These represent true and noisy draws from a distribution that is a straight line with some thickness $t$, and are visualized the upper panels of Figure \ref{fig:toy}
+where $\Delta_n$ is drawn from the Gaussian function $\mathcal{N}(\Delta_n \given 0, t^2)$, and $\epsilon_n$ is drawn from the Gaussian function $\mathcal{N}(\epsilon_n \given 0, \sigma_n^2)$, so $\sigma_n$ is the measurement uncertainty.
+These represent true and noisy draws from a distribution that is a straight line with some thickness $t$, and are visualized in the upper panels of Figure \ref{fig:toy}
 
@@
 This denoised distribution becomes our prior.
-We then use this pior to infer a posterior belief about the true values, $p(y_{\true,n} \given y_n, \tilde{\sigma}_n^2)$, shown as the blue points in the lower panels of Figure \ref{fig:toy}
+We then use this pior to infer a posterior belief about the true values, $p(y_{\true,n} \given y_n, \tilde{\sigma}_n^2)$, whose expectation value is shown as the blue points in the lower panels of Figure \ref{fig:toy}
 When the true values are inferred in this way, the associated variance $\tilde{\sigma}_n^2$ is the convolution of the true distribution variance $t^2$ and the measurement variance $\sigma_n^2$.
@@
 Here we highlight 5 random points in black to highlight the method, but the same method is being applied to all the data.
-We denoise these points using marginalized maximum likelihood estimation (MML), which marginalizes over the true data values (we never need to know them) and maximizes the likelihood of the data.
+We learn the noise-deconvolved distribution using marginalized maximum likelihood estimation (MML), which marginalizes over the true data values (we never need to know them) and maximizes the likelihood of the data.
 Here our likelihood for each individual datum is
@@
 This MML gives us the best fit parameters $\hat{m}$, $\hat{b}$, and $\hat{t}$.
-In Figure \ref{fig:toy}, center panel, the true functions $y_{\true} = m_{\true}\,x + b_{\true} \pm t_{\true}$ are shown as the red lines, and the best fit MML functions $\hat{y}_{\true} = \hat{m}\, x + \hat{b} \pm \hat{t}$ are shown as the blue lines.
+In Figure \ref{fig:toy}, lower left panel, the true functions $y_{\true} = m_{\true}\,x + b_{\true} \pm t_{\true}$ are shown as the red lines, and the best fit MML functions $\hat{y}_{\true} = \hat{m}\, x + \hat{b} \pm \hat{t}$ are shown as the blue lines.
 Now that we have an estimate of the true underlying distribution $\hat{y}_{\true}$, we use this as a prior to infer a posterior belief of the true values for each datum $p(y_{\true,n} \given y_n, \tilde{\sigma}_n^2)$.
@@
 Here is a good point to step back and acknowledge that our method is odd/strange/different/new (which maybe you're also currently thinking but can't quite express why yet).
-We are using the set of data to determine the denoised distribution, and then using that denoised distribution to make inferences on the data.
+We are using the set of data to determine the noise-deconvolved distribution, and then using that noise-deconvolved distribution to make inferences on the data.
 We are technically using the data twice!
-If we were going to be fully, technically correct, we should denoise the distribution, but leave one datum out; the one datum we want to infer the true y value for\footnote{DWH: Remember what this brain dump is for}.
+If we were going to be fully, technically correct, we should learn the distribution, but leave one datum out; the one datum we want to infer the true y value for\footnote{DWH: Remember what this brain dump is for}.
 Leaving one out, and then estimating the prior and then inferring the posterior for each data point is way too expensive for the 1024 points in our toy model, and also for the 1.2 million stars in our actual data set.
-Luckily with large data sets, even at 1024, 1 datum is $\sim0.1\%$ of your dataset and therefore leaving that point out will not change the MML parameters of your denoised distribution, and with the \gaia\ dataset of over a million stars, we can safely step into the Empirical Bayes regime.
+Luckily with large data sets, even at 1024, 1 datum is $\sim0.1\%$ of your dataset and therefore leaving that point out will not change the MML parameters of your noise-deconvolved distribution, and with the \gaia\ dataset of over a million stars, we can safely step into the Empirical Bayes regime.
 
@@
 The blue points in Figure \ref{fig:toy}, lower right panel, represent the posterior belief of the true value $p(y_{\true,n} \given y_n, \tilde{\sigma}_n)$ for the highlighted black points.
-Finally, Figure \ref{fig:toy}, right panel, shows the full set of posterior beliefs for each $y_n$.
+Finally, Figure \ref{fig:toy}, lower right panel, shows the full set of posterior beliefs for all $y_n$.
 The points are colored by their measurement variance $\sigma_n^2$ to show that noisier measurements are more influenced by the prior and get pulled closer to the center of the MML distribution, our prior, while more certain measurements remain closer to their measured values $y_n$.
@@
 Bringing the toy model back to our original problem, using the noise-deconvolved CMD as our prior allows us to get better estimates of the parallax for each star.
-
-LA: Conclusions and stuff about why this stuff is good
 
@@
 To apply a dust correction to the \tmass\ photometry, we first generate our prior using the observed, attenuated photometry.
-Using this raw prior we infer more precise distances to all the stars which we can use to get a measurement of dust.
+Using this raw prior we infer more precise distances to all the stars.
+We then use this more precise distance posterior to get a measurement of dust in the 3D dust map.
 This approach minimizes severe distances from the likelihoods alone and their associated severe dust corrections.
-With more these more precise distance posteriors, we take the $5\%$ quantile of each distance posterior, so the closest bit of the posterior, and sample the Green et al 3D dust map at that distance.
-For each distance we take the median of the samples of the dust model.
+In detail, we take the $5\%$ quantile of each distance posterior, so the closest bit of the posterior, and sample the Green et al 3D dust map at that distance.
+We take the $5\%$ quantile, as apposed to the median for example, to do a light dust correction because over correcting for dust can shoot you into a different space of the CMD and then bias your distance inference.
+For each distance we sample the median of the probabilistic dust model.
 We apply each dust correction to our \tmass\ photometry given by Equation \ref{eq:data} where $Q_{\lambda}$ = [0.709, 0.302] for bands [J, K] respectively, taken from table (REF TABLE SFD), and $E(B-V)$ is the output of the Green et al dust models.
-We then regenerate our prior using the dust corrected photometry.
+We then regenerate our prior using the lightly dust corrected photometry.
 We do not update the errors, but keep them the same as before the dust correction, which is wrong and under representing our errors but this leads again to a lighter deconvolution (Do we want to change this and add in the covariance for the dust?).
 We iterate this process 10 times, however the dust values seem to converge after about 2-3 iterations.
+The iterative process allows for possibly slightly high dust values to be corrected to slightly too low dust values until a distance and corresponding dust value is more consistent with the CMD.
 Should I quote some statistics about the dust corrections?
@@
 
+\subsection{What's That Feature?}
+We used cross validation to show that the likelihood of the data continued to increase from 16 Gaussians to 2048 Gaussians.
+We chose to do our inference using 128 Gaussians, but increasing the number of Gaussians was also a good check to see what features in the CMD appeared to be real, and which ones smoothed out with increasing numbers of Gaussians.
+The most obvious structures that persisted to our maximum number of Gaussians are:
+\begin{itemize}
+\item Main Sequence
+\item Binary Sequence
+\item Possible metallicity variation in MS lifetime
+\item Red clump
+\item Red bump
+\item Red Giant Branch
+\item Asymptotic Giant Branch
+\item planetary nebula
+\end{itemize}
+Fraction of stars in various parts of the diagram?
+
 Come back to the assumptions. What happens if we change or violate them?
@@
 Number of gaussians in \xd, and the w parameter. That we set them heuristically.
-
-Talk about structures in CMD; red bump - acknowledge Ana ! s
-What's That Feature? Similarly; Fraction of stars in various parts of the diagram?
 

### 2017-04-17 paperDistance/paper.tex
paper stuff

@@
 
-We would like to infer the true parallax $\varpi_{\true}$ for
-each star given its parallax measurement $\varpi$ and associated
-uncertainty $\sigma_{\varpi}$ from \tgas.
-We assume (as \gaia\ suggests) that the likelihood is Gaussian
+\subsection{Setting up the Inference}
+Think about presentation, prioritizing/highlighting prior but logically building it from an full inference perspective and needing a prior, then dive into prior stuff.
+It's confusing right now jumping between inference and prior building !!
+\subsection{Generating the Prior}
+
+Under these assumptions, we would like to infer the true parallax $\varpi_{\true}$ for each star given its measured parallax $\varpi$ and associated uncertainty $\sigma_{\varpi}$ from \tgas.
+The strength of our inference method here lies in how we generate our prior information on the true parallax.
+Supplementary to this trigonometric measure of distance, the photometry of stars can improve the distance inference.
+The intrinsic temperatures and luminosities of stars tightly correlate leading to the observed colors and absolute magnitudes of stars also tightly correlating, usually visualized as the color--magnitude diagram (\cmd).
+The tight correlation of the \cmd\ benefits our inference because the absolute magnitude of a star contains distance information.
+We include this distance information in the inference of the true parallax by using the \cmd\ as our prior.
+Specifically, we use the \cmd\ generated from the data themselves as our prior.
+Instead of assuming a functional form for the spatial distribution of stars in the Milky Way (\citealt{astraatmadja16b}), or relying on stellar models (\citealt{gaia16}), we build a noise-deconvolved estimate of the \cmd\ from the observed colors and absolute magnitudes of our full set of stars.
+This \cmd\ model becomes, for each star---given its observed color and magnitude (and uncertainties on those)---a prior pdf for the distance
+to that star.
+
+Because our prior in the \cmd\ space, our inference will include not only the parallax from \gaia, but also the photometry of the stars.
+For inferring the true parallax, including generating our \cmd\ prior, we use the $J$ and $K$ band photometry from \tmass, as well as the parallax from \gaia.
+Our true, latent values are the $(J-K)^{true}_n$ color and the $J$ band absolute magnitude $M^{true}_{J,n}$ for each star, which we encapsulate in the true vector $\mathbf{Y}$.
+The observed data is the dust corrected color $(J-K)^c_n$, the \gaia\ parallax $\varpi_n$ in units of $[mas]$, and the dust corrected $J$ band magnitude $J^c_n$, which we encapsulate in the data vector $\mathbf{y}$.
+Instead of transforming the parallax and apparent magnitude into an absolute magnitude, the data vector $\mathbf{y}$ contains a transformation of the dust corrected absolute magnitude
+
 \begin{equation}
-p(\varpi \given \varpi_{\true}, I) = \mathcal{N}(\varpi \given \varpi_{\true}, \sigma^2_{\varpi}) \quad ,
-\label{eq:gaialike}
+10^{0.2\,M^c_{J,n} + 2} = \varpi\,10^{0.2J^c_n}
+\label{eq:transform}
 \end{equation}
-where $I$ is a placeholder containing our assumptions, including all
-uncertainty estimates, and $\mathcal{N}(x \given \mu, \sigma^2)$ is a
-Gaussian function for $x$ with mean $\mu$ and variance $\sigma^2$.
-This is the probability of the data given the true parallax, but we
-want to flip it around and infer the true parallax, or distance, to
-the star given the measured parallax and it's uncertainty. We do this
-using probability theory
+
+The transformation of the dust corrected absolute magnitude, Equation \ref{eq:transform}, is required to to keep the uncertainties Gaussian, which, under some assumptions along with this transformation, we maintain.
+Gaussian uncertainties are required to generate our prior using \xd\, but more about this and our dust corrections below.
+Our prior assumptions $I$ are the uncertainties in each of these measurements, as well as the Greene dust map
+
+\begin{equation}
+\begin{aligned}
+\mathrm{true \, vector} \quad &\mathbf{Y}_n = [10^{0.2\,M^{true}_{J,n} + 2}, (J-K)^{true}_n] \\
+\mathrm{data \, vector} \quad &\mathbf{y}_n = [\varpi_n\,10^{0.2J^c_n}, (J- K)^c_n] \;  n \in \mathrm{stars} \\
+\mathrm{with \; apparent \; magnitude} \quad &J^c_n = J_n - Q_J\,E(B-V)_n \\
+\mathrm{and \; color} \quad &(J - K)^c_n = J_n - K_n - Q_{JK}\,E(B-V)_n \\
+\mathrm{assumptions} \quad &I_n = \left\{\left\{\sigma^2_{\varpi, n}\right\}, \left\{\sigma^2_{J,n}\right\}, \left\{\sigma^2_{K,n}\right\}, \mathrm{Dust \, Map}\right\}
+\end{aligned}
+\label{eq:data}
+\end{equation}
+
+where $E(B-V)_n$ is the output of the Green et al dust model, and $Q_{\lambda}$ is the correction factor for a photometric band assuming some reddening law.
+
+If life were simple, our inference would be a 1D problem using Bayes theorem looking something like:
 \begin{equation}
@@
 parallax to the star, and $Z$ is a normalization constant.
-The strength of our inference method here lies in how we generate our prior
-pdf for the parallax $p(\varpi_{\true})$.
-Instead of assuming a functional form for the spatial distribution of
-stars in the Milky Way (\citealt{astraatmadja16b}), or relying on stellar
-models (\citealt{gaia16}), we build a noise-deconvolved estimate of the color--magnitude
-diagram (\cmd) from the observed colors and absolute magnitudes of our
-full set of stars.
-This \cmd\ model becomes, for each star---given its observed color and
-magnitude (and uncertainties on those)---a prior pdf for the distance
-to that star.
 
-That said, this is \emph{not} exactly how we do the inference here.
-Here we perform the inference in a two-dimensional space that is a
-transform of the \cmd....[HOGG: Syntactical marker]
+But life isn't simple, and our prior is a 2D function in color and transformed absolute magnitude space, so our full inference in done in 2D space as well.
+We assume (as \gaia\ suggests) that the parallax likelihood is Gaussian, as well as the $(J-K)^c_n$ color likelihood. So our full likelihood is
+\begin{equation}
+p(\mathbf{y}_n \given \mathbf{Y}_n, I_n) = \mathcal{N}(\mathbf{y}_n \given \mathbf{Y}_n, V_n)
+\end{equation}
 
-To generate our \cmd\ prior we use the $J$ and $K$ band photometry from \tmass, as well as the parallax from \gaia.
-The data are a transformation of the dust corrected absolute magnitude, and the dust corrected colors.
+with \\
+$V_n = $
+\[
+\begin{bmatrix}
+\sigma_{J,n}^2 + \sigma_{K,n}^2 & 0 \\
+0 & (\sigma_{\varpi}\,10^{0.2\,J_n^C})^2
+\end{bmatrix}
+\]
+
+where $\mathcal{N}$ is a 2D Gaussian function with mean $\mathbf{Y}$ and variance $V_n$.
+Here we have assumed that the parallas uncertainty dominates over the $J$ band uncertainty, so we take $J_n^C$ as a point estimate and only propagate the parallax uncertainty.
+This is to keep our uncertainties Gaussian so we can use \xd\ generate our prior below.
+This is the probability of the data vector given the true vector, but we
+want to flip it around and infer the true vector and therefore the true parallax, or distance. We do this
+analogous to Equation \ref{eq:bayes}.
 
 \begin{equation}
-10^{0.2\,M^c_{J,n} + 2} = \varpi\,10^{0.2J^c_n}
-\label{eq:transform}
+p(\mathbf{Y}_n \given \mathbf{y}_n, I_n) = p(\mathbf{y}_n \given \mathbf{Y}_n, I_n) \, p(\mathbf{Y}_n)
 \end{equation}
 
-The transformation of the dust corrected absolute magnitude, Equation \ref{eq:transform}, is required to to keep the uncertainties gaussian, which, under some assumptions along with this transformation, we maintain.
-Gaussian uncertainties are required to use \xd\, more about this and our dust corrections below.
-
-Our true, latent values are the true $J-K$ color $C_n$ and the true $J$ band absolute magnitude $M_{J,n}$ for each star, which we encapsulate in the true vector $\vec{Y}$.
-The observed data is the dust corrected $J-K$ color $(J-K)^c_n$, the \gaia\ parallax $\varpi_n$ in units of $[mas]$, and the dust corrected $J$ band magnitude $J^c_n$, which we encapsulate in the data vector $\vec{y}$.
-Our prior assumptions $I$ are the uncertainties in each of these measurements, as well as the Greene dust map
-
-\begin{equation}
-\begin{aligned}
-\mathrm{true \, vector} \quad &\vec{Y} = [10^{0.2\,M_{J,n} + 2}, C_n] \\
-\mathrm{data \, vector} \quad &\vec{y} = [\varpi_n\,10^{0.2J^c_n}, (J- K)^c_n] \;  n \in \mathrm{stars} \\
-\mathrm{with \; apparent \; magnitude} \quad &J^c_n = J_n - Q_J\,E(B-V)_n \\
-\mathrm{and \; color} \quad &(J - K)^c_n = J_n - K_n - Q_{JK}\,E(B-V)_n \\
-\mathrm{assumptions} \quad &I = \left\{\left\{\sigma^2_{\varpi, n}\right\}, \left\{\sigma^2_{J,n}\right\}, \left\{\sigma^2_{K,n}\right\}, \mathrm{Dust \, Map}\right\}
-\end{aligned}
-\label{eq:data}
-\end{equation}
-
-where $E(B-V)_n$ is the output of the Green et al dust model, and $Q_{\lambda}$ is the correction factor for a photometric band assuming some reddening law.
-
-Working with these variables, we then have a likelihood for the data, encapsulating both the \gaia\ parallaxes and \tmass\ photometry of
-
-\begin{equation}
-p(\vec{y} \given \vec{Y}) = \mathcal{N}(\vec{y} \given \vec{Y}, \sigma_{\vec{y}})
-\label{eq:fulllike}
-\end{equation}
+where $p(\mathbf{Y}_n)$ is our prior.
 
 To generate our prior, we fit the data in the slightly transformed \cmd\ space, Equation \ref{eq:transform}, using \xd\ which deconvolves the data with its uncertainties, which are heterogeneous measurement uncertainties.
-It generates the denoised, underlying distribution from which the uncertain data were drawn from.
+It generates the underlying distribution from which the uncertain data were drawn from.
 It represents the distribution as a mixture of gaussians which maximizes the marginalized likelihood of the data, marginalized over the true data.
@@
 \alpha &=& \left\{\mathrm{amp}_k, \mu_k, V_k\right\}_{k=1}^K \; k \in \mathrm{Gaussians} \nonumber\\
-       &=& \argmax_{\alpha} \, \Pi_n \, p(\vec{y}_n \given \alpha) \nonumber\\
-       &=& \argmax_{\alpha} \, \Pi_n \, \int p(\vec{y}_n \given \vec{Y}_n) \, p(\vec{Y} \given \alpha)\mathrm{d\vec{Y}}
+       &=& \argmax_{\alpha} \, \Pi_n \, p(\mathbf{y}_n \given \alpha) \nonumber\\
+       &=& \argmax_{\alpha} \, \Pi_n \, \int p(\mathbf{y}_n \given \mathbf{Y}_n) \, p(\mathbf{Y}_n \given \alpha)\mathrm{d\mathbf{Y}_n}
 \label{eq:xdmml}
@@
 
-To infer $\varpi_{\true}$ from the likelihood of the \gaia\ data, Equation \ref{eq:gaialike}, we first do the inference in the 2D space of the \cmd
+To infer $\varpi_{\true}$ we first do the inference in the 2D space of the \cmd
 
 \begin{equation}
-p(\vec{Y} \given \vec{y}, \alpha) = \frac{1}{Z} \, p(\vec{y} \given \vec{Y}) \, p(\vec{Y} \given \alpha)
+p(\mathbf{Y} \given \mathbf{y}, \alpha) = \frac{1}{Z} \, p(\mathbf{y} \given \mathbf{Y}) \, p(\mathbf{Y} \given \alpha)
 \label{eq:posterior}
@@
 
-We then project the 2D posterior into the 1D absolute magnitude/parallax space
+We then project the 2D posterior into the 1D transformed absolute magnitude space, and do a little algebra to get
 
 \begin{eqnarray}
-p(\vec{Y}_0 \given \vec{y}, \alpha) &\rightarrow& p(\varpi_{\true} \given y, \alpha) \, f(\varpi_{\true}) \\
+p(\mathbf{Y}_0 \given \mathbf{y}, \alpha) &\rightarrow& p(\varpi_{\true} \given y, \alpha) \, f(\varpi_{\true}) \\
 f(\varpi_{\true}) &=& \begin{cases}
@@
 \end{eqnarray}
-
-And with a jacobian this becomes a posterior on distance to the star if you wish.
+with $f(\varpi_{true})$ being a window function to insure the true parallax is positive and put it on a similar grid for all the posteriors.
+And with a jacobian this becomes a posterior on distance.
 
@@
 Using more gaussians as our prior did not substantially change our posterior results, but it did start to highlight the features of the CMD.}
-\label{fig:prior}
+\label{fig:priorCompare}
 \end{figure}
@@
 
-Long discussion of dust.
+Some discussion of the inconsistency of how we treat dust within the probabilistic framework.
 

### 2017-04-19 paperDistance/paper.tex
paper tweaks

@@
 \subsection{Setting up the Inference}
-Think about presentation, prioritizing/highlighting prior but logically building it from an full inference perspective and needing a prior, then dive into prior stuff.
-It's confusing right now jumping between inference and prior building !!
-\subsection{Generating the Prior}
+ \emph{Think about presentation, prioritizing/highlighting prior but logically building it from an full inference perspective and needing a prior, then dive into prior stuff.
+It's confusing right now jumping between inference and prior building !!}
 
@@
 The strength of our inference method here lies in how we generate our prior information on the true parallax.
-Supplementary to this trigonometric measure of distance, the photometry of stars can improve the distance inference.
+Supplementary to this trigonometric measure of distance, the photometry of stars is also informative of their distances.
 The intrinsic temperatures and luminosities of stars tightly correlate leading to the observed colors and absolute magnitudes of stars also tightly correlating, usually visualized as the color--magnitude diagram (\cmd).
-The tight correlation of the \cmd\ benefits our inference because the absolute magnitude of a star contains distance information.
+The tight correlation of the \cmd\ benefits our inference because the luminosity or absolute magnitude of a star, when compared with its brightness or apparent magnitude, contains distance information.
 We include this distance information in the inference of the true parallax by using the \cmd\ as our prior.
@@
 Instead of assuming a functional form for the spatial distribution of stars in the Milky Way (\citealt{astraatmadja16b}), or relying on stellar models (\citealt{gaia16}), we build a noise-deconvolved estimate of the \cmd\ from the observed colors and absolute magnitudes of our full set of stars.
-This \cmd\ model becomes, for each star---given its observed color and magnitude (and uncertainties on those)---a prior pdf for the distance
-to that star.
+This \cmd\ model becomes, for each star---given its observed color and magnitude (and uncertainties on those)---a prior pdf for the distance to that star.
 
-Because our prior in the \cmd\ space, our inference will include not only the parallax from \gaia, but also the photometry of the stars.
+Because our prior is the \cmd, our inference will include not only the parallax from \gaia, but also the photometry of the stars.
 For inferring the true parallax, including generating our \cmd\ prior, we use the $J$ and $K$ band photometry from \tmass, as well as the parallax from \gaia.
@@
 Gaussian uncertainties are required to generate our prior using \xd\, but more about this and our dust corrections below.
-Our prior assumptions $I$ are the uncertainties in each of these measurements, as well as the Greene dust map
+Our prior assumptions $I$ are the uncertainties in each of these measurements, as well as the Greene dust map. So our full terminology is
 
@@
 
-If life were simple, our inference would be a 1D problem using Bayes theorem looking something like:
+If life were simple, our inference would be a 1D problem using Bayes theorem and look something like:
 \begin{equation}
@@
 where $\mathcal{N}$ is a 2D Gaussian function with mean $\mathbf{Y}$ and variance $V_n$.
-Here we have assumed that the parallas uncertainty dominates over the $J$ band uncertainty, so we take $J_n^C$ as a point estimate and only propagate the parallax uncertainty.
+Here we have assumed that the parallax uncertainty dominates over the $J$ band uncertainty, so we take $J_n^C$ as a point estimate and only propagate the parallax uncertainty.
 This is to keep our uncertainties Gaussian so we can use \xd\ generate our prior below.
-This is the probability of the data vector given the true vector, but we
+
+This likelihood is the probability of the data vector given the true vector, but we
 want to flip it around and infer the true vector and therefore the true parallax, or distance. We do this
@@
 
+\subsection{Generating the Prior}
+
 To generate our prior, we fit the data in the slightly transformed \cmd\ space, Equation \ref{eq:transform}, using \xd\ which deconvolves the data with its uncertainties, which are heterogeneous measurement uncertainties.
 It generates the underlying distribution from which the uncertain data were drawn from.
-It represents the distribution as a mixture of gaussians which maximizes the marginalized likelihood of the data, marginalized over the true data.
+It represents the distribution as a mixture of Gaussians which maximizes the marginalized likelihood of the data, marginalized over the true data.
 
@@
 \end{eqnarray}
+
+where $\alpha$ encapsulates the parameters of each Gaussian in the mixture of Gaussians that defines our \cmd\ prior.
+
+\subsection{Back to the Inference}
 
@@
 
+\subsection{Comparison with other priors}
+\begin{figure}
+\centering
+  \includegraphics[width=\textwidth]{deltaSimple.png}
+\caption{Other example prior: The same visualization as Figure \ref{fig:delta} of the change in the posterior, but here, if instead of prior information containing the \cmd\, the prior information is an exponentially declining spatial density of stars.
+}
+\label{fig:otherPrior}
+\end{figure}
+
+There are other prior information options when inferring distances to stars.
+A commonly used prior for inferring distances (cite: Tri, CBJ) is the exponentially declining spatial density of stars.
+It is parameterized by a scale length, and has the nice property that it is smooth out to very large distances so the posteriors are also very smooth.
+It is a fairly weak prior and basically sets a smooth scale at which it's challenging for stars to lie at, and also allows for negative parallaxes.
+It is a fairly broad prior so the median \gaia\ star has an increased posterior variance.
+It has the largest affect on the very noisy and neegative parallaxes and has the nice property of bringing these noisy or negative measurements to reasonable distances.
+Given this prior, what does the observed \cmd\ look like?
+The \cmd\ is broad and the red clump highly stretched.
+
 \subsection{What's That Feature?}
@@
 Some discussion of possible failure modes. Including very rare kinds of stars. Should we add gaussians to prior where we know stars exist that aren't well sampled in Gaia (ie WD)
+Inference with our prior is bad for underrepresented populations in the \cmd. For example, it appears there are only $2$ white dwarfs in \gaia\ DR1. With t
 
@@
   Andy Casey (Monash),
+  Dustin Lang (Toronto) for cross matched 2MASS-TGAS catalogue
 and the attendees at the Stars Group Meeting at the Flatiron Institute

### 2017-04-19 paperDistance/paper.tex
tiny tweaks

@@
   case of a stellar type with only one (statistically distinct)
-  exemplar in the data set (e.g. a white dwarf), this assumption will be maximally wrong.
+  exemplar in the data set (for example, a white dwarf), this assumption will be maximally wrong.
 \item We assume that the \tgas\ parallax uncertainties dominate the
@@
   stellar physics are generic and implicit (for example, that the
-  \cmd\ is somehow smooth and compact). Dust model dependent on stellar models?
+  \cmd\ is somehow smooth and compact).
 \end{itemize}
@@
 
-Come back to the assumptions. What happens if we change or violate them?
+Come back to the assumptions. What happens if we change or violate them? Long discussion of the noise model. We took serious short-cuts. Call out to Boris.
 
-Long discussion of the noise model. We took serious short-cuts. Call out to Boris.
-
-Some discussion of the inconsistency of how we treat dust within the probabilistic framework.
+LMA: Some discussion of the inconsistency of how we treat dust within the probabilistic framework.
 

### 2017-04-19 paperDistance/paper.tex
paper additions dust

--- 2017-04-19 paperDistance/paper.tex
@@
 In detail, we take the $5\%$ quantile of each distance posterior, so the closest bit of the posterior, and sample the Green et al 3D dust map at that distance.
-We take the $5\%$ quantile, as apposed to the median for example, to do a light dust correction because over correcting for dust can shoot you into a different space of the CMD and then bias your distance inference.
+We take the $5\%$ quantile, as apposed to the median for example, to do a light dust correction because over correcting for dust can shoot the likelihood into an unphysical, different space of the CMD and then bias the distance inference.
 For each distance we sample the median of the probabilistic dust model.
-We apply each dust correction to our \tmass\ photometry given by Equation \ref{eq:data} where $Q_{\lambda}$ = [0.709, 0.302] for bands [J, K] respectively, taken from table (REF TABLE SFD), and $E(B-V)$ is the output of the Green et al dust models.
+We apply each light dust correction to our \tmass\ photometry given by Equation \ref{eq:data} where $Q_{\lambda}$ = [0.709, 0.302] for bands [J, K] respectively, taken from table (REF TABLE SFD), and $E(B-V)$ is the output of the Green et al dust models.
 We then regenerate our prior using the lightly dust corrected photometry.
-We do not update the errors, but keep them the same as before the dust correction, which is wrong and under representing our errors but this leads again to a lighter deconvolution (Do we want to change this and add in the covariance for the dust?).
+We do not update the errors, but keep them the same as before the dust correction, which is wrong and under representing our errors but this leads again to a lighter deconvolution.
+We could add the covariance of the dust to our likelihoods, but this is beyond the scope of this demonstration of concept.
 We iterate this process 10 times, however the dust values seem to converge after about 2-3 iterations.
-The iterative process allows for possibly slightly high dust values to be corrected to slightly too low dust values until a distance and corresponding dust value is more consistent with the CMD.
-Should I quote some statistics about the dust corrections?
-mostly small?
-maybe difference before after correcting for distance?
+The iterative process allows for slight variations in distance and corresponding dust to become more consistent with the \cmd.
+Within our probabilistic framework this handling of dust is not proper.
+We are taking a point estimate of the probabilistic dust model at a quantile distance of our posterior.
+A full, proper account would sample the dust model at distances sampled from the posterior and propagate those samples back into \xd during the iterative proccess.
+The issues is that this dust sample distribution would not be a Gaussian distribution which \xd\ requires.
+We could calculate this more complex dust distribution and estimate it as a Gaussian distribution to feed to \xd, however, this is beyond the scope of this paper.
+
+Should I quote some statistics about the dust corrections? mostly small? maybe difference before/after applying prior to distance?
 
@@
 \subsection{What's That Feature?}
-We used cross validation to show that the likelihood of the data continued to increase from 16 Gaussians to 2048 Gaussians.
-We chose to do our inference using 128 Gaussians, but increasing the number of Gaussians was also a good check to see what features in the CMD appeared to be real, and which ones smoothed out with increasing numbers of Gaussians.
-The most obvious structures that persisted to our maximum number of Gaussians are:
+The denoised expectation values of the parallax projected into the \cmd\ show interest features.
 \begin{itemize}
-\item Main Sequence
-\item Binary Sequence
-\item Possible metallicity variation in MS lifetime
-\item Red clump
-\item Red bump
-\item Red Giant Branch
-\item Asymptotic Giant Branch
-\item planetary nebula
+\item (1) Narrow Lower Main Sequence
+\item (2) Wide Upper Main Sequence
+\item (3) Plausible Binary Sequence
+\item (4) Prominent Red clump
+\item (5) Narrow Red Giant Branch
+\item (6) Noticeable turn off
 \end{itemize}
-Fraction of stars in various parts of the diagram?
+The lower main sequence is very narrow while the upper main sequence is much wider.
+This may be an age variation on the main sequence mixed with a malquist bias of only seeing a very small volume of lower main sequence stars but a much larger volume of upper main sequence stars.
+There is a plausible binary sequence of stars with about twice the brightness as the main sequence.
+The helium burning red clump is (surprisingly) prominent for the volume DR1 probed, and the red giant branch is straight/narrow.
+The main sequence turn off is noticeable and also surprisingly narrow, possibly a reflection of the star formation history of the local volume also mixed with a malmquist bias.
 

### 2017-04-22 paperDistance/paper.tex
list

--- 2017-04-19 paperDistance/paper.tex
@@
 method that flows from them---in the Discussion below.
-\begin{itemize}
-\item We make an assumption of \emph{stationarity}; that is, that the
+\begin{description}
+\item[stationarity] We make an assumption of \emph{stationarity}; that is, that the
   stars measured at lower signal-to-noise have analogs measured at
@@
   work.
-\item We assume that we have large numbers of stars, large enough that
+\item[big data] We assume that we have large numbers of stars, large enough that
   an empirical-Bayes or maximum-marginalized likelihood is a safe
@@
   exemplar in the data set (for example, a white dwarf), this assumption will be maximally wrong.
-\item We assume that the \tgas\ parallax uncertainties dominate the
+\item[noise model] We assume that the \tgas\ parallax uncertainties dominate the
   noise in any estimates of absolute magnitude. We further assume that
@@
   (it leads to under-deconvolution).
-\item We treat the \panstarrs-based three-dimensional dust maps (\citealt{green15})
+\item[dust] We treat the \panstarrs-based three-dimensional dust maps (\citealt{green15})
   as correct in their median dust estimates, and their
@@
   color or absolute magnitude.
-\item We assume that the \cmd\ can be represented with a mixture of
+\item[mixture of Gaussians] We assume that the \cmd\ can be represented with a mixture of
   Gaussians in a particular transformed space. This mixture we fix
@@
   different experiments.
-\item We make no use of stellar models; our only assumptions about
+\item[no physics] We make no use of stellar models; our only assumptions about
   stellar physics are generic and implicit (for example, that the
   \cmd\ is somehow smooth and compact).
-\end{itemize}
+\end{description}
 

### 2017-04-22 paperDistance/paper.tex

--- 2017-04-19 paperDistance/paper.tex
@@
   case of a stellar type with only one (statistically distinct)
-  exemplar in the data set (e.g. a white dwarf), this assumption will be maximally wrong.
-\item We assume that the \tgas\ parallax uncertainties dominate the
+  exemplar in the data set (for example, a white dwarf), this assumption will be maximally wrong.
+\item[noise model] We assume that the \tgas\ parallax uncertainties dominate the
   noise in any estimates of absolute magnitude. We further assume that
@@
   different experiments.
-\item We make no use of stellar models; our only assumptions about
+\item[no physics] We make no use of stellar models; our only assumptions about
   stellar physics are generic and implicit (for example, that the
-  \cmd\ is somehow smooth and compact). Dust model dependent on stellar models?
-\end{itemize}
+  \cmd\ is somehow smooth and compact).
+\end{description}
 

### 2017-04-24 paperDistance/paper.tex
paper stuff

--- 2017-04-19 paperDistance/paper.tex
@@
 The iterative process allows for slight variations in distance and corresponding dust to become more consistent with the \cmd.
-Within our probabilistic framework this handling of dust is not proper.
-We are taking a point estimate of the probabilistic dust model at a quantile distance of our posterior.
-A full, proper account would sample the dust model at distances sampled from the posterior and propagate those samples back into \xd during the iterative proccess.
-The issues is that this dust sample distribution would not be a Gaussian distribution which \xd\ requires.
-We could calculate this more complex dust distribution and estimate it as a Gaussian distribution to feed to \xd, however, this is beyond the scope of this paper.
 
@@
   \includegraphics[width=\textwidth]{posteriorCMD.png}
-\caption{the expectation values and subsampled error bars}
+\caption{the expectation values and subsampled error bars.
+Notice that some parts of this \cmd\ are tighter than the prior, or noise-deconvolved \cmd.
+The visual of the prior is a sampling of the Gaussians, but this is the expectation value so this lies much closer to the center of the distribution.
+Similar to the toy model, where the pdf maximum of the noisier points lies in the center of the distribution.}
 \label{fig:posteriorCMD}
@@
 
-Some discussion of the inconsistency of how we treat dust within the probabilistic framework.
+\subsection{Proper Dust Corrections}
+Within our probabilistic framework, the way we handle dust is not proper.
+We are taking a point estimate of the probabilistic dust model at a quantile distance of our posterior, correcting our photometry using this point estimate and rederiving the prior in \xd.
+A full, proper account would sample the dust model at distances sampled from the posterior and propagate those samples back into \xd during the iterative proccess.
+The issues is that this dust sample distribution would not be a Gaussian distribution which \xd\ requires.
+We could calculate this more complex dust distribution and estimate it as a Gaussian distribution to feed to \xd, however, this is beyond the scope of this paper.
 
-Some discussion of possible failure modes. Including very rare kinds of stars. Should we add gaussians to prior where we know stars exist that aren't well sampled in Gaia (ie WD)
-Inference with our prior is bad for underrepresented populations in the \cmd. For example, it appears there are only $2$ white dwarfs in \gaia\ DR1. With t
+\subsection{Stationarity Assumption}
+Both the low and high signal to noise objects are being drawn from the same distribution, independent of noise.
+This isn't precisely true because most Gaia objects have similar noise but their signal depends on their distance.
+Nearby stars have higher signal (parallax) because they are closer and the nearby stars are mostly main sequence stars (IMF, timing, malmquist).
+We can see bright giants to much larger distances and therefore lower signal or parallax and lower signal to noise.
+So the red giant branch is a population of relatively lower signal to noise objects than the main sequence.
+But I think this is ok, that the low and high signal to noise objects have to be drawn from the same distribution within some feature or patch of the space.
+This is all assuming the Gaia noise model is correct.
+If the uncertainties are over estimated then we could be over deconvolving (generating a tighter distribution than the truth) And in some extreme over estimate of uncertainties the lower signal to noise giants could be pulled to the main sequence.
+But with the quoted uncertainties this seems unlikely.
+As you move up the giant branch and probe farther giants the signal to noise will continue to diminish but I think this is okay since the color uncertainties are not large.
+Possibly a metallicity trend with distance, placing more nearby (higher S/N) objects at slightly different positions on the CMD pulling lower S/N objects to that distribution for our prior.
+But if something has small error bars is good data is good data both for inference and generating the prior
 
-Number of gaussians in \xd, and the w parameter. That we set them heuristically.
+For very rare kinds of stars.
+Inference with our prior is bad for underrepresented populations in the \cmd.
+For example, it appears there are only $2$ white dwarfs in \gaia\ DR1.
+With them being very underrepresented in the prior, the amplitudes of the gaussians representing them is very small, so the WDs are more likely to be pulled to mimic a much farther away MS star.
+
+\subsection{Cross Validation}
+For generating our prior, we use a mixture of 128 Gaussians.
+This number, 128, was chosen heuristically.
+It makes the method computationally tractable, and is sufficient for this demonstration of concept, but more Gaussians might be desired for a more final data product.
+We would ideally use the optimal number of Gaussians determined by cross validation, but preliminary tests using cross validation suggest the optimal number of Gaussians is very large.
 

### 2017-04-27 paperDistance/paper.tex
put in Hoggs corrections to abstract and intro

@@
 \begin{abstract}\noindent % trust me
-The \gaia\ \tgas\ \project{Catalog} contains more than 2 million parallax measurements.
+
 Conversion of a noisy parallax measurement into a posterior belief over distances requires inference with a prior.
 Usually this prior represents some kind of belief about the Milky Way.
-However, there is multi-band photometry for the \tgas\ stars from imaging surveys;
+However, there is multi-band photometry for a large fractino of the \gaia\ \tgas\ catalog from imaging surveys;
 this imaging is incredibly informative about stellar distances.
-Here we use color information on \tgas\ stars from \tmass\ to build a noise-deconvolved empirical prior distribution of stars in color--magnitude space.
+Here we use color information on $1.2$ million \tgas\ stars from \tmass\ to build a noise-deconvolved empirical prior distribution for stars in color--magnitude space.
 This data-driven model contains no knowledge of the physics of stellar interiors or photospheres, nor of the Milky Way, but rather derives its precision from its generative model of noisy parallax measurements and an assumption of stationarity.
@@
 The \xd-optimized prior is used to perform parallax and distance inferences for every star, yielding a precise stellar distance estimate and uncertainty (and full posterior) for every star.
-Our posterior parallax estimates are more precise than the Gaia catalog outputs by a factor of [XXX] for the median \tgas\ star; and more precise than previous examples of Bayesian distance estimates by a factor of [YYY].
+Our posterior parallax estimates are more precise than the \tgas\ catalog entries by a median factor of [XXX]; and more precise than previous examples of Bayesian distance estimates by a factor of [YYY].
 The precision can be attributed to the statistics concept of shrinkage.
-We independently validate our distances by looking at members of Milky Way star clusters; for example, M67 is not visible at all in the \tgas\ parallax estimates, but appears clearly in our posterior parallax estimates.
+We independently validate our distances by looking at members of Milky Way star clusters; for example, M67 is not visible as a cluster in the \tgas\ parallax estimates, but appears clearly in our posterior parallax estimates.
 All our results, including a posterior parallax and distance sampling for [ZZZ] \tgas\ stars, are available in companion electronic tables.
@@
 Only a small fraction (but large number) of these distance
-measurements will be purely astrometric:
+measurements will be both precise and purely astrometric:
 \gaia\ uses astrometric parallax to determine the distances of the closer
-stars, and calibrate spectrophotometric models.
+stars, and calibrate to spectrophotometric models.
 These spectrophotometric models, in turn, along with \gaia's on-board
@@
 Because there are small issues with each of these components, there are small
-color and luminosity differences with the data---the models are wrong in detail---and
+color and luminosity differences with the data---the models are inaccurate in detail---and
 the physical models therefore produce distance estimates that are biased.
@@
 However, the use of physical models for distance estimation is not necessary.
-Because there are stars at all luminosities with good parallax information,
+Because there are stars at all luminosities with useful parallax information,
 it is possible to build a stellar photometric model from the data themselves.
@@
 we use all of the \gaia\ \tgas\ data (\citealt{tgas};
-these data were generated by the methodology of \citealt{michalik15}), matched to the
-\tmass\ photometry (\citealt{skrutskie06}), and in the footprint of
+these data were generated by the methodology of \citealt{michalik15}), that match to the
+\tmass\ photometry (\citealt{skrutskie06}) and lie in the footprint of
 the \panstarrs\ Dust Map (\citealt{green15}), to make a model of the
@@
 of the deconvolved distribution,
-its use as a prior is only approximately valid; it constitutes using the
-data twice.
+its use as a prior is only approximately valid.
 However, since the data set is large, the \xd\ approximation is not bad; it is
@@
 similar stars.
-That is a much weaker assumption than has been made in other Bayesian
+That is a much weaker assumption than has been made in most other Bayesian
 distance estimations.
@@
 This project is fundamentally a demonstration of concept:
-We are only using the ``small'' (relative to the expected final \gaia\ data set)
+We are only using a subset of the ``small'' (relative to the expected final \gaia\ data set)
 \tgas\ Catalog.
@@
 However, as we show below, we get very good results.
-It is promising that we will be able to obtain similar or even better results
+It is promising that we will be able to obtain even better results
 in later \gaia\ data releases, and eventually put distances on $>10^9$
@@
 
-First we determine the underlying distribution that the noisy data $y_n$ are drawn from, shown as the blue lines in the lower left panel of Figure \ref{fig:toy}
-This denoised distribution becomes our prior.
-We then use this pior to infer a posterior belief about the true values, $p(y_{\true,n} \given y_n, \tilde{\sigma}_n^2)$, whose expectation value is shown as the blue points in the lower panels of Figure \ref{fig:toy}
+First we determine the underlying distribution that the noisy data $y_n$ are drawn from.
+The $1\sigma$ contours of this noise deconvolved distribution are shown as the blue lines in the lower left panel of Figure \ref{fig:toy}.
+This noise deconvolved distribution becomes our prior.
+We then use this pior to infer a posterior belief about the true values, $p(y_{\true,n} \given y_n, \tilde{\sigma}_n^2)$, whose expectation value and $1\sigma$ uncertainty is shown as the blue points in the lower panels of Figure \ref{fig:toy}.
 When the true values are inferred in this way, the associated variance $\tilde{\sigma}_n^2$ is the convolution of the true distribution variance $t^2$ and the measurement variance $\sigma_n^2$.
@@
 
-and the likelihood of the full data set is
+It is a Gaussian because the noise model is a Gaussian. The likelihood of the full data set is then the product of the individual likelihoods
 
@@
 
-Maximizing this likelihood is the same as minimizing
+Maximizing this joint likelihood of all the data is the same as minimizing
 

### 2017-05-01 comparePrior.py
figure tweaks

@@ -374,9 +374,16 @@ def compareSimpleGaia(ngauss=128, quantile=0.05, iter='10th', survey='2MASS', da
     yim = (-1, 5)
     for file in ['posteriorSimple.npz', postFile]:
         data = np.load(file)
+        posterior = data['posterior']
+        samples = np.zeros(np.shape(posterior)[0])
+        xparallaxMAS = np.logspace(-2, 2, np.shape(posterior)[1])
+        for i, p in enumerate(posterior):
+            samples[i] = testXD.samples(xparallaxMAS, posterior[p], 1, plot=False)
         mean = data['mean']
         var = data['var']
         absMag = testXD.absMagKinda2absMag(mean*10.**(0.2*apparentMagnitude))
+        absMagSample = testXD.absMagKinda2absMag(samples*10.**(0.2*apparentMagnitude))
+
 
         neg = tgas['parallax'] < 0
         fig, ax = plt.subplots(1, 2)
@@ -438,7 +445,7 @@ def compareSimpleGaia(ngauss=128, quantile=0.05, iter='10th', survey='2MASS', da
         ax2 = figVarDiff.add_subplot(122)
 
         ax1.scatter(x, absMag[notnans], s=1, lw=0, c=y, alpha=0.05, norm=norm, cmap=cmap)
-        ax2.scatter(x, absMag[notnans], s=1, lw=0, c=tgas['parallax_error'][notnans]**2., alpha=0.05, norm=cNorm, cmap=cmap)
+        ax2.scatter(x, absMag[notnans], s=1, lw=0, c=tgas['parallax_error'][notnans]**2., alpha=0.05, cmap=cmap)
 
         titles = ["Colored by change in variance", "Colored by observed variance"]
 
@@ -459,6 +466,33 @@ def compareSimpleGaia(ngauss=128, quantile=0.05, iter='10th', survey='2MASS', da
             ax[i].set_ylabel(ylabel, fontsize = 18)
 
         figVarDiff.savefig('denoisedVariance_' + file.split('.')[0] + '.png')
+        figVarDiff.clf()
+
+        ax1 = figVarDiff.add_subplot(121)
+        ax2 = figVarDiff.add_subplot(122)
+
+        ax1.scatter(x, absMag[notnans], s=1, lw=0, c=y, alpha=0.05, norm=norm, cmap=cmap)
[…]

### 2017-05-01 comparePrior.py
figure tweaks

@@ -378,7 +378,7 @@ def compareSimpleGaia(ngauss=128, quantile=0.05, iter='10th', survey='2MASS', da
         samples = np.zeros(np.shape(posterior)[0])
         xparallaxMAS = np.logspace(-2, 2, np.shape(posterior)[1])
         for i, p in enumerate(posterior):
-            samples[i] = testXD.samples(xparallaxMAS, posterior[p], 1, plot=False)
+            samples[i] = testXD.samples(xparallaxMAS, p, 1, plot=False)
         mean = data['mean']
         var = data['var']
         absMag = testXD.absMagKinda2absMag(mean*10.**(0.2*apparentMagnitude))

### 2017-05-01 comparePrior.py
figure tweaks

@@ -378,7 +378,7 @@ def compareSimpleGaia(ngauss=128, quantile=0.05, iter='10th', survey='2MASS', da
         samples = np.zeros(np.shape(posterior)[0])
         xparallaxMAS = np.logspace(-2, 2, np.shape(posterior)[1])
         for i, p in enumerate(posterior):
-            samples[i] = testXD.samples(xparallaxMAS, p, 1, plot=False)
+            samples[i] = testXD.samples(xparallaxMAS, p, 1, plot=False)[0]
         mean = data['mean']
         var = data['var']
         absMag = testXD.absMagKinda2absMag(mean*10.**(0.2*apparentMagnitude))

### 2017-05-01 comparePrior.py
figure tweaks

@@ -377,7 +377,8 @@ def compareSimpleGaia(ngauss=128, quantile=0.05, iter='10th', survey='2MASS', da
         posterior = data['posterior']
         samples = np.zeros(np.shape(posterior)[0])
         xparallaxMAS = np.logspace(-2, 2, np.shape(posterior)[1])
-        for i, p in enumerate(posterior):
+        for i, p in enumerate(posterior[0:10,:]):
+            print p
             samples[i] = testXD.samples(xparallaxMAS, p, 1, plot=False)[0]
         mean = data['mean']
         var = data['var']

### 2017-05-01 comparePrior.py
figure tweaks

@@ -377,9 +377,9 @@ def compareSimpleGaia(ngauss=128, quantile=0.05, iter='10th', survey='2MASS', da
         posterior = data['posterior']
         samples = np.zeros(np.shape(posterior)[0])
         xparallaxMAS = np.logspace(-2, 2, np.shape(posterior)[1])
-        for i, p in enumerate(posterior[0:10,:]):
-            print p
-            samples[i] = testXD.samples(xparallaxMAS, p, 1, plot=False)[0]
+        for i, p in enumerate(posterior):
+            try: samples[i] = testXD.samples(xparallaxMAS, p, 1, plot=False)[0]
+            except IndexError: import pdb; pdb.set_trace()
         mean = data['mean']
         var = data['var']
         absMag = testXD.absMagKinda2absMag(mean*10.**(0.2*apparentMagnitude))

### 2017-05-01 comparePrior.py
figure tweaks

@@ -379,7 +379,7 @@ def compareSimpleGaia(ngauss=128, quantile=0.05, iter='10th', survey='2MASS', da
         xparallaxMAS = np.logspace(-2, 2, np.shape(posterior)[1])
         for i, p in enumerate(posterior):
             try: samples[i] = testXD.samples(xparallaxMAS, p, 1, plot=False)[0]
-            except IndexError: import pdb; pdb.set_trace()
+            except IndexError: samples[i] = -999
         mean = data['mean']
         var = data['var']
         absMag = testXD.absMagKinda2absMag(mean*10.**(0.2*apparentMagnitude))

### 2017-05-02 paperDistance/paper.tex
paper tweaks ;

@@
 In this paper we use the noise deconvolved observed \cmd\ as a prior to infer more precise parallax pdfs from \gaia\ measurements.
-Here we will demonstrate this method with a simpler toy model.
-Instead of using the 2D distribution of the \cmd\ as our prior, for this simpler model we use a 2D distribution that is a 1D Gaussian in the y-direction, with a running mean $\mu = mx + b$ and some thickness $t$, as our prior.
-The $1\sigma$ contours as well as (noise free) samples from the distribution are shown in Figure \ref{fig:toy}, upper left panel.
-Noisy samples from this distribution are shown in the upper right panel.
-Using a representation of this distribution as our prior we can infer some posterior belief of $y_{\true,n}$, a true sample of the distribution, from $y_n$, a noisy sample of the distribution.
+Understanding this method is challenging because we are using the data both for the prior and the likelihood in our inference.
+To ease understanding, here we will demonstrate this method with a simpler toy model.
+Instead of inferring more precise parallax pdfs using the 2D distribution of the \cmd\ as our prior, for this simpler model we will infer more precise pdfs of the value $Y$.
+We will do this simpler inference using a 2D distribution that is a 1D Gaussian in the y-direction, with a running mean $\mu = mx + b$ and some thickness $t$, as our prior.
 Here the thickness of the distribution mimics the true thickness of the \cmd\ due to age, mass and metallicity.
-To set up the problem, we have the true underlying distribution of points and the noisy data
+The $1\sigma$ contours as well as (noise free) samples from the toy model distribution are shown in Figure \ref{fig:toy}, upper left panel.
+Similarly, noisy samples from this distribution are shown in the upper right panel.
+Using a representation of the true distribution (which we learn from the noisy data) as our prior, we can infer some posterior belief of $y_{\true,n}$, a true sample of the distribution, from $y_n$, a noisy sample of the distribution.
+
+Setting up the problem in more detail, we have the true underlying distribution of points and the noisy data
 
@@
 y_{\true, n} \, &=& \, m_{\true}\,x_n + b_{\true} + \Delta_n \\
-y_n \, &=& \, y_{true,n} + \epsilon_n
+y_n \, &=& \, y_{\true,n} + \epsilon_n
 \label{eq:ytrue}
@@
 These represent true and noisy draws from a distribution that is a straight line with some thickness $t$, and are visualized in the upper panels of Figure \ref{fig:toy}
+The upper left shows 1024 realizations of the true values, and the upper right, as well as lower left, show 1024 realizations of the noisy values $y_n$ and their associated measurement uncertainties $\sigma_n$ as the grey points.
+Here we highlight 5 random points in black to highlight the method, but the same method is being applied to all the data.
 
-First we determine the underlying distribution that the noisy data $y_n$ are drawn from.
-The $1\sigma$ contours of this noise deconvolved distribution are shown as the blue lines in the lower left panel of Figure \ref{fig:toy}.
-This noise deconvolved distribution becomes our prior.
-We then use this pior to infer a posterior belief about the true values, $p(y_{\true,n} \given y_n, \tilde{\sigma}_n^2)$, whose expectation value and $1\sigma$ uncertainty is shown as the blue points in the lower panels of Figure \ref{fig:toy}.
-When the true values are inferred in this way, the associated variance $\tilde{\sigma}_n^2$ is the convolution of the true distribution variance $t^2$ and the measurement variance $\sigma_n^2$.
-Figure \ref{fig:toy}, upper right, shows 1024 realizations of the noisy values $y_n$ and their associated measurement uncertainties $\sigma_n$ as the grey points.
-Here we highlight 5 random points in black to highlight the method, but the same method is being applied to all the data.
-We learn the noise-deconvolved distribution using marginalized maximum likelihood estimation (MML), which marginalizes over the true data values (we never need to know them) and maximizes the likelihood of the data.
+To infer a pdf of the true y value for each measured y value we need a prior.
+For our prior we will use an estimate of the underlying true distribution that the noisy data $y_n$ are drawn from, which we build from the noisy y data themselves.
+We learn the noise-deconvolved function using marginalized maximum likelihood estimation (MML), which marginalizes over the true data values (we never need to know them) and maximizes the likelihood of the data.
 Here our likelihood for each individual datum is
@@
 If we were going to be fully, technically correct, we should learn the distribution, but leave one datum out; the one datum we want to infer the true y value for\footnote{DWH: Remember what this brain dump is for}.
-Leaving one out, and then estimating the prior and then inferring the posterior for each data point is way too expensive for the 1024 points in our toy model, and also for the 1.2 million stars in our actual data set.
+Leaving one datum out to estimate the prior and then inferring the posterior for that datum is way too expensive for the 1024 points in our toy model, and also for the 1.2 million stars in our actual data set.
 Luckily with large data sets, even at 1024, 1 datum is $\sim0.1\%$ of your dataset and therefore leaving that point out will not change the MML parameters of your noise-deconvolved distribution, and with the \gaia\ dataset of over a million stars, we can safely step into the Empirical Bayes regime.
+(Talk more about emperical bayes here?)
 
@@
 Since everything is Gaussian, this can be done analytically by multiplying Gaussians together.
-The blue points in Figure \ref{fig:toy}, lower right panel, represent the posterior belief of the true value $p(y_{\true,n} \given y_n, \tilde{\sigma}_n)$ for the highlighted black points.
-Finally, Figure \ref{fig:toy}, lower right panel, shows the full set of posterior beliefs for all $y_n$.
-The points are colored by their measurement variance $\sigma_n^2$ to show that noisier measurements are more influenced by the prior and get pulled closer to the center of the MML distribution, our prior, while more certain measurements remain closer to their measured values $y_n$.
-By doing this we have better estimates of the true y value for each point.
+In Figure \ref{fig:toy} lower left panel, the blue points represent the posterior belief of the true value $p(y_{\true,n} \given y_n, \tilde{\sigma}_n)$ for the highlighted black points.
+The red points are the true y values.
+The blue points in Figure \ref{fig:toy}, lower right panel, show the expectation value and $1\sigma$ uncertainty of the posterior belief for all 1024 $y_n$.
+When the true values are inferred in this way, the associated variance $\tilde{\sigma}_n^2$ is the convolution of the true distribution variance $t^2$ and the measurement variance $\sigma_n^2$.
+The points are colored by their measurement variance $\sigma_n^2$ which shows that noisier measurements are more influenced by the prior and get pulled closer to the center of the MML distribution, our prior.
+In contrast, more certain measurements remain closer to their measured values $y_n$.
+By inferring the true y value from the measured y value in addition to the prior information of the distribution the y value was drawn from, this we have better estimates of the true y value for each point.
 Bringing the toy model back to our original problem, using the noise-deconvolved CMD as our prior allows us to get better estimates of the parallax for each star.
@@
   (it leads to under-deconvolution).
-\item We treat the \panstarrs-based three-dimensional dust maps (\citealt{green15})
+\item[dust] We treat the \panstarrs-based three-dimensional dust maps (\citealt{green15})
   as correct in their median dust estimates, and their
   effects on color and magnitude. As we optimize the empirical-Bayes model, we iteratively update each star's
-  dust correction as its updated, inferred distance. Our assumption is that these corrections are
+  dust correction at its updated, inferred distance. Our assumption is that these corrections are
   good enough, and uncertainties in these are not dominant, for either
   color or absolute magnitude.
-\item We assume that the \cmd\ can be represented with a mixture of
+\item[mixture of Gaussians] We assume that the \cmd\ can be represented with a mixture of
   Gaussians in a particular transformed space. This mixture we fix
@@
 The strength of our inference method here lies in how we generate our prior information on the true parallax.
-Supplementary to this trigonometric measure of distance, the photometry of stars is also informative of their distances.
+Supplementary to the trigonometric measure of distance, the photometry of stars is also informative of their distances.
 The intrinsic temperatures and luminosities of stars tightly correlate leading to the observed colors and absolute magnitudes of stars also tightly correlating, usually visualized as the color--magnitude diagram (\cmd).
@@
 \begin{equation}
-10^{0.2\,M^c_{J,n} + 2} = \varpi\,10^{0.2J^c_n}
+y_0 = \varpi\,10^{0.2J^c_n} = 10^{0.2\,M^c_{J,n} + 2}
 \label{eq:transform}
@@
 \end{equation}
-where $p(\varpi_{\true})$ is a prior probability density (pdf) for the
-parallax to the star, and $Z$ is a normalization constant.
+where $p(\varpi \given \varpi_{\true}, \sigma^2_{\varpi})$ is the likelihood of the observed parallax, $p(\varpi_{\true})$ is a prior probability density (pdf) for the
+parallax, and $Z$ is a normalization constant.
 
 But life isn't simple, and our prior is a 2D function in color and transformed absolute magnitude space, so our full inference in done in 2D space as well.
-We assume (as \gaia\ suggests) that the parallax likelihood is Gaussian, as well as the $(J-K)^c_n$ color likelihood. So our full likelihood is
+We assume (as \gaia\ suggests) that the parallax likelihood is Gaussian, as well as the $(J-K)^c_n$ color likelihood. So our full likelihood of the data is
 \begin{equation}
@@
 where $\mathcal{N}$ is a 2D Gaussian function with mean $\mathbf{Y}$ and variance $V_n$.
-Here we have assumed that the parallax uncertainty dominates over the $J$ band uncertainty, so we take $J_n^C$ as a point estimate and only propagate the parallax uncertainty.
+Here we have assumed that the parallax uncertainty is significantly larger than the $J$ band uncertainty, and therefore the dominant uncertainty in the transformed absolute magnitude, so we take $J_n^C$ as a point estimate and only propagate the parallax uncertainty.
 This is to keep our uncertainties Gaussian so we can use \xd\ generate our prior below.
@@
 
-To generate our prior, we fit the data in the slightly transformed \cmd\ space, Equation \ref{eq:transform}, using \xd\ which deconvolves the data with its uncertainties, which are heterogeneous measurement uncertainties.
+To generate our prior, we fit the data in the slightly transformed \cmd\ space, using Equation \ref{eq:transform} instead of absolute magnitude.
+We fit this transfromed \cmd\ using \xd\ which deconvolves the data with its uncertainties, which are heterogeneous measurement uncertainties.
 It generates the underlying distribution from which the uncertain data were drawn from.
-It represents the distribution as a mixture of Gaussians which maximizes the marginalized likelihood of the data, marginalized over the true data.
+\xd\ generates a represention of the distribution as a mixture of Gaussians which maximizes the marginalized likelihood of the data, marginalizing over the true data.
 
@@
 
-We use stars crossmatched in TGAS, 2MASS and within the observing footprint of \panstarrs\ to access the Greene 3D dust model.
-We require the photometry have real values and nonzero, real errors, and remove a small selection of \tmass\ stars that have zero color and zero J band magnitude.
+We use stars crossmatched in \tgas\ and \tmass\, and lie within the observing footprint of \panstarrs\ to access the Greene 3D dust model.
+We require the photometry have real values, and nonzero, real errors, and remove a small selection of \tmass\ stars that have zero color and zero J band magnitude.
 
@@
 To apply a dust correction to the \tmass\ photometry, we first generate our prior using the observed, attenuated photometry.
-Using this raw prior we infer more precise distances to all the stars.
+Using this raw prior we infer more precise distances to all the stars in our sample.
 We then use this more precise distance posterior to get a measurement of dust in the 3D dust map.
@@
 It is an open cluster in the constellation Cancer and is estimated to be 3-5 billion years old and 800-900 pc away.
-Previous work has estimated that bad things happen if you try to infer a distance as the inverse of the observed (noisy) parallax when the signal to noise of the parallax measurement drops below 5.
-With \gaia\ parallaxes having a minimum uncertainty of about 0.3 mas, for stars with the most certain parallaxes, the SN reaches 5 at about 700 pc, right where this cluster lies.
+Previous work (cite CBJ) has estimated that bad things happen if you try to infer a distance as the inverse of the observed (noisy) parallax when the signal to noise of the parallax measurement drops below 5.
+With \gaia\ parallaxes having a minimum uncertainty of about 0.3 mas, the SN reaches 5 for the most certain parallax measurements at about 700 pc, right where this cluster lies.
 So it is a great test case for comparing parallax inferences with or without a prior.
-We made a window of 1 degree squared on the sky with the cluster M67 lying in the middle.
-The cluster is visibly obvious as a clustering of stars in a visualization of the \gaia\ data in 2D on the sky.
+We made a window of 1 square degree on the sky with the cluster M67 lying in the middle.
+The cluster is visibly obvious as a clustering of stars in a visualization of the \gaia\ data on the 2D sky.
 With our converged dust values, and it's associated prior, we calculate the posterior of the true parallaxes to the stars within this window on the sky around M67.
-Figure XXX shows the comparison of the likelihoods and posteriors.
-The vertical lines bracket the parallaxes of, or distances to, the stars associated with the cluster.
-The likelihoods of the observed parallaxes are far broader and the cluster is not obvious at all.
-In the posteriors there is a sharp peak in the probability at the previously determined distance to the cluster, the vertical lines, for far more stars than the likelihoods would imply.
+Figure \ref{fig:m67} shows the comparison of the likelihoods and posteriors.
+The vertical lines bracket the parallaxes of, or distances to, the cluster.
+The likelihoods of the observed parallaxes, shown in the top row, are broad and the cluster is not obvious at all.
+The posteriors, shown in the bottom row, show a sharp peak in the probability at the previously determined distance to the cluster, the vertical lines, for far more stars than the likelihoods would imply.
 Our prior is not only making parallaxes more precise but also (possibly) increasing the accuracy.
@@
 
+Come back to the assumptions. What happens if we change or violate them? Long discussion of the noise model. We took serious short-cuts. Call out to Boris.
+
+LMA: Some discussion of the inconsistency of how we treat dust within the probabilistic framework.
+=======
 Come back to the assumptions. What happens if we change or violate them?

### 2017-05-03 paperDistance/paper.tex
paper tweaks

@@
 
-\author{Jo Bovy}
-\affil{Center for Computational Astrophysics, Flatiron Institute, 162 Fifth Ave, New York, NY 10010, USA}
-\affil{Department of Astronomy and Astrophysics, University of Toronto, 50 St. George Street, Toronto, ON M5S 3H4, Canada}
+%\author{Jo Bovy}
+%\affil{Center for Computational Astrophysics, Flatiron Institute, 162 Fifth Ave, New York, NY 10010, USA}
+%\affil{Department of Astronomy and Astrophysics, University of Toronto, 50 St. George Street, Toronto, ON M5S 3H4, Canada}
 
-\author{Boris Leistedt}
-\affil{Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, New York, NY 10003, USA}
+%\author{Boris Leistedt}
+%\affil{Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, New York, NY 10003, USA}
 
-\author{Adrian Price-Whelan}
-\affil{Princeton University Observatory, Princeton University, 4 Ivy Lane, Princeton, NJ 08544, USA}
+%\author{Adrian Price-Whelan}
+%\affil{Princeton University Observatory, Princeton University, 4 Ivy Lane, Princeton, NJ 08544, USA}
 
@@
 That is, the model for the \cmd\ we have built is representative for
-the \tgas\ Catalog.
+(a subset of) the \tgas\ Catalog.
 Inasmuch as that is true, we expect any (sensible) distance estimates
-we generate from our distance posterior pdfs to be weakly biased.
+we generate from our distance posterior pdfs to be only weakly biased.
 
@@
 In particular, there appear to be essentially no white dwarf stars,
-and other classes (like hot main-sequence stars) have few to no exemplars
+and other classes (like bright red giant branch stars) have few to no exemplars
 in the data set at high signal-to-noise (in parallax).
@@
 errors, because each data point has had a prior multiplied in.
-Effectively multiplicative uses of the data will effectively take the prior
+Multiplicative uses of the data will effectively take the prior
 pdf to a large power.
@@
 There are other prior information options when inferring distances to stars.
-A commonly used prior for inferring distances (cite: Tri, CBJ) is the exponentially declining spatial density of stars.
+One option is the exponentially declining spatial density of stars (cite: Tri, CBJ).
 It is parameterized by a scale length, and has the nice property that it is smooth out to very large distances so the posteriors are also very smooth.
@@
 It is a fairly broad prior so the median \gaia\ star has an increased posterior variance.
-It has the largest affect on the very noisy and neegative parallaxes and has the nice property of bringing these noisy or negative measurements to reasonable distances.
+It has the largest affect on the most noisy and negative parallaxes and has the nice property of bringing these noisy or negative measurements to reasonable parallaxes.
 Given this prior, what does the observed \cmd\ look like?
@@
 
-Come back to the assumptions. What happens if we change or violate them? Long discussion of the noise model. We took serious short-cuts. Call out to Boris.
-
-LMA: Some discussion of the inconsistency of how we treat dust within the probabilistic framework.
-=======
 Come back to the assumptions. What happens if we change or violate them?
@@
 \subsection{Proper Dust Corrections}
-Within our probabilistic framework, the way we handle dust is not proper.
+Within our probabilistic framework, we improperly handle dust.
 We are taking a point estimate of the probabilistic dust model at a quantile distance of our posterior, correcting our photometry using this point estimate and rederiving the prior in \xd.
@@
   Dustin Lang (Toronto) for cross matched 2MASS-TGAS catalogue
+  Jo Bovy (Toronto)
+  Boris Leistedt (NYU)
+  Adrian Price-Whelan (Princeton)
 and the attendees at the Stars Group Meeting at the Flatiron Institute

### 2017-05-03 paperDistance/paper.tex
paper tweaks

@@
 Within our probabilistic framework, we improperly handle dust.
-We are taking a point estimate of the probabilistic dust model at a quantile distance of our posterior, correcting our photometry using this point estimate and rederiving the prior in \xd.
-A full, proper account would sample the dust model at distances sampled from the posterior and propagate those samples back into \xd during the iterative proccess.
-The issues is that this dust sample distribution would not be a Gaussian distribution which \xd\ requires.
+We are taking a point estimate of the probabilistic dust model at a quantile distance of our posterior, correcting our photometry using this point estimate and rederiving the prior.
+A full, proper account would sample the dust model at distances sampled from the posterior and propagate those samples back into the prior during the iterative proccess.
+The issue, the reason we don't do a full probabilistic treatment, is that this proper dust distribution would not be a Gaussian distribution which \xd\ requires.
 We could calculate this more complex dust distribution and estimate it as a Gaussian distribution to feed to \xd, however, this is beyond the scope of this paper.
@@
 \subsection{Stationarity Assumption}
-Both the low and high signal to noise objects are being drawn from the same distribution, independent of noise.
-This isn't precisely true because most Gaia objects have similar noise but their signal depends on their distance.
+Stationarity assumes both the low and high signal to noise objects are being drawn from the same distribution, independent of noise.
+This isn't precisely true because most Gaia objects have similar noise $(\sim 3 mas)$, but their signal (parallax) depends on their distance.
 Nearby stars have higher signal (parallax) because they are closer and the nearby stars are mostly main sequence stars (IMF, timing, malmquist).
-We can see bright giants to much larger distances and therefore lower signal or parallax and lower signal to noise.
-So the red giant branch is a population of relatively lower signal to noise objects than the main sequence.
+However, we can see bright giants to much larger distances and therefore lower signal or parallax, leading to lower signal to noise.
+So the red giant branch is a population of relatively lower signal to noise objects compared with the main sequence.
 But I think this is ok, that the low and high signal to noise objects have to be drawn from the same distribution within some feature or patch of the space.

### 2017-05-04 comparePrior.py
paper tweaks

@@ -229,7 +229,7 @@ def dataViz(survey='2MASS', ngauss=128, quantile=0.05, dataFilename='All.npz', i
                     xdgmm, xerr=color_err[notnans], yerr=[yerr_minus, yerr_plus], xlabel=xlabel, ylabel=ylabel, xlim=xlim,
                     ylim=ylim, errSubsample=1.2e3, thresholdScatter=2., binsScatter=200, c=c, norm=cNorm, cmap='Blues', contourColor=contourColor, posterior=True)
 
-        dataFile = 'inferredDistances_data_' + file.split('.')[0] + '.png'
+        dataFile = 'inferredDistances_data_' + file.split('.')[0] + '.pdf'
         priorFile = 'prior_' + str(ngauss) +'gauss.png'
         os.rename('plot_sample.data.png', dataFile)
         os.rename('plot_sample.prior.png', priorFile)

### 2017-05-05 comparePrior.py
save data

@@ -13,6 +13,7 @@ import stellarTwins as st
 import scipy.integrate
 import corner
 from astroML.plotting import setup_text_plots
+from astropy.io import ascii
 
 def prior(xdgmm, ax):
     for gg in range(xdgmm.n_components):
@@ -225,6 +226,8 @@ def dataViz(survey='2MASS', ngauss=128, quantile=0.05, dataFilename='All.npz', i
         yerr_minus = absMag - absMagYMinus
         yerr_plus = testXD.absMagKinda2absMag(yplus) - absMag
         #notnan = ~np.isnan(color[notnans]) & ~np.isnan(absMag)
+
+        ascii.write([color[notnans], absMag, color_err[notnans], yerr_minus, yerr_plus, c], 'cmdExpectation.txt', names=['color', 'absMag', 'color_err', 'absMag_errMinus', 'absMag_errPlus', 'logDeltaVar'])
         dp.plot_sample(color[notnans], absMag, sample[:,0], testXD.absMagKinda2absMag(sample[:,1]),
                     xdgmm, xerr=color_err[notnans], yerr=[yerr_minus, yerr_plus], xlabel=xlabel, ylabel=ylabel, xlim=xlim,
                     ylim=ylim, errSubsample=1.2e3, thresholdScatter=2., binsScatter=200, c=c, norm=cNorm, cmap='Blues', contourColor=contourColor, posterior=True)

### 2017-05-05 comparePrior.py
plot tweaks

@@ -127,7 +127,7 @@ def dustViz(ngauss=128, quantile=0.5, iter='8th', survey='2MASS', dataFilename='
     cb.set_label(r'E($B-V$ )')
     fig.savefig('dustViz.dQ' + str(quantile) + '.png')
 
-def dataViz(survey='2MASS', ngauss=128, quantile=0.05, dataFilename='All.npz', iter='10th', Nsamples=3e5, contourColor='k', dustFile='dust.npz'):
+def dataViz(survey='2MASS', ngauss=128, quantile=0.05, dataFilename='All.npz', iter='10th', Nsamples=3e5, contourColor='k', dustFile='dust.npz', sdss5=False):
 
     if survey == 'APASS':
         mag1 = 'B'
@@ -192,10 +192,10 @@ def dataViz(survey='2MASS', ngauss=128, quantile=0.05, dataFilename='All.npz', i
     """
     dp.plot_sample(color[positive], testXD.absMagKinda2absMag(y), sample[:,0], testXD.absMagKinda2absMag(sample[:,1]),
                 xdgmm, xerr=color_err[positive], yerr=[yerr_minus, yerr_plus], xlabel=xlabel, ylabel=ylabel, xlim=xlim, ylim=ylim, errSubsample=2.4e3, thresholdScatter=2., binsScatter=200, contourColor=contourColor)
-    dataFile = 'data_noDust.png'
-    priorFile = 'prior_' + str(ngauss) +'gauss.png'
-    os.rename('plot_sample.data.png', dataFile)
-    os.rename('plot_sample.prior.png', priorFile)
+    dataFile = 'data_noDust.pdf'
+    priorFile = 'prior_' + str(ngauss) +'gauss.pdf'
+    os.rename('plot_sample.data.pdf', dataFile)
+    os.rename('plot_sample.prior.pdf', priorFile)
     #import pdb; pdb.set_trace()
     data = np.load(dustFile)
     dustEBV = data['ebv']
@@ -227,12 +227,12 @@ def dataViz(survey='2MASS', ngauss=128, quantile=0.05, dataFilename='All.npz', i
         #notnan = ~np.isnan(color[notnans]) & ~np.isnan(absMag)
         dp.plot_sample(color[notnans], absMag, sample[:,0], testXD.absMagKinda2absMag(sample[:,1]),
                     xdgmm, xerr=color_err[notnans], yerr=[yerr_minus, yerr_plus], xlabel=xlabel, ylabel=ylabel, xlim=xlim,
[…]

### 2017-05-05 demo_plots.py

@@ -1,6 +1,7 @@
 import numpy as np
 import matplotlib
 matplotlib.use('pdf')
+#from matplotlib.backends.backend_pdf import PdfPages
 from matplotlib import pyplot as plt
 from astroML.plotting.tools import draw_ellipse
 from astroML.plotting import setup_text_plots
@@ -103,7 +104,7 @@ def kdeDensity(ax, x, y, threshold=0.01, bins=100, s=1, lw=0, alpha=1):
     #fig.colorbar(im)
     return ax
 
-def plot_sample(x, y, samplex, sampley, xdgmm, xlabel='x', ylabel='y', xerr=None, yerr=None, ylim=(6, -6), xlim=(0.5, 1.5), errSubsample=1.2e6, thresholdScatter=0.1, binsScatter=200, c='black',  norm=None, cmap=None, contourColor='k', posterior=False, ind=None):
+def plot_sample(x, y, samplex, sampley, xdgmm, xlabel='x', ylabel='y', xerr=None, yerr=None, ylim=(6, -6), xlim=(0.5, 1.5), errSubsample=1.2e6, thresholdScatter=0.1, binsScatter=200, c='black',  norm=None, cmap=None, contourColor='k', posterior=False, ind=None, plot_contours=True), sdss5=False:
     prng = np.random.RandomState(1234567890)
     setup_text_plots(fontsize=16, usetex=True)
     plt.clf()
@@ -120,7 +121,8 @@ def plot_sample(x, y, samplex, sampley, xdgmm, xlabel='x', ylabel='y', xerr=None
     levels = 1.0 - np.exp(-0.5 * np.arange(1.0, 2.1, 1.0) ** 2)
     cNorm  = plt.matplotlib.colors.LogNorm(vmin=3, vmax=1e5)
     #ax1.hist2d(x, y, bins=100, norm=cNorm, cmap='Greys')
-    im = corner.hist2d(x, y, ax=ax1, levels=levels, bins=200, no_fill_contours=True, plot_density=False, color=contourColor)
+    if sdss5: plot_contours=False
+    im = corner.hist2d(x, y, ax=ax1, levels=levels, bins=200, no_fill_contours=True, plot_density=False, color=contourColor, plot_contours=plot_contours)
     #im = ax1.scatter(x, y, s=1, lw=0, c=c, alpha=alpha, norm=norm, cmap=cmap)
 
     ax2 = figData.add_subplot(122)
[…]

### 2017-05-05 stellarTwins.py

@@ -1,5 +1,5 @@
 from astropy.io import fits
-import h5py
+#import h5py
 import numpy as np
 import matplotlib.pyplot as plt
 from scipy.spatial import ckdtree as kdtree

### 2017-05-05 paperDistance/paper.tex

@@
 Usually this prior represents some kind of belief about the Milky Way.
-However, there is multi-band photometry for a large fractino of the \gaia\ \tgas\ catalog from imaging surveys;
+However, there is multi-band photometry for a large fraction of the \gaia\ \tgas\ catalog from imaging surveys;
 this imaging is incredibly informative about stellar distances.
-Here we use color information on $1.2$ million \tgas\ stars from \tmass\ to build a noise-deconvolved empirical prior distribution for stars in color--magnitude space.
+Here we use color information of $1.2$ million \tgas\ stars from \tmass\ to build a noise-deconvolved empirical prior distribution for stars in color--magnitude space.
 This data-driven model contains no knowledge of the physics of stellar interiors or photospheres, nor of the Milky Way, but rather derives its precision from its generative model of noisy parallax measurements and an assumption of stationarity.
@@
 The algorithm is run not in absolute-magnitude space but in a transformed space where the measurement uncertainty is closer to Gaussian.
-The \xd-optimized prior is used to perform parallax and distance inferences for every star, yielding a precise stellar distance estimate and uncertainty (and full posterior) for every star.
-Our posterior parallax estimates are more precise than the \tgas\ catalog entries by a median factor of [XXX]; and more precise than previous examples of Bayesian distance estimates by a factor of [YYY].
-The precision can be attributed to the statistics concept of shrinkage.
-We independently validate our distances by looking at members of Milky Way star clusters; for example, M67 is not visible as a cluster in the \tgas\ parallax estimates, but appears clearly in our posterior parallax estimates.
+The \xd-optimized prior is used to perform parallax inferences for every star, yielding a precise photometric parallax estimate and uncertainty (and full posterior) for every star.
+Our posterior photometric parallax estimates are more precise than the \tgas\ catalog entries by a median factor of [XXX]; and more precise than previous examples of Bayesian distance estimates by a factor of [YYY].
+%The precision can be attributed to the statistics concept of shrinkage.
+We independently validate our distances by looking at members of a Milky Way star cluster M67, which is not visible as a cluster in the \tgas\ parallax estimates, but appears clearly as a cluster in our posterior parallax estimates.
 All our results, including a posterior parallax and distance sampling for [ZZZ] \tgas\ stars, are available in companion electronic tables.
@@
 
-The \gaia\ Mission (\citealt{gaia16}) will deliver more than a billion stellar distances.
-Only a small fraction (but large number) of these distance
-measurements will be both precise and purely astrometric:
-\gaia\ uses astrometric parallax to determine the distances of the closer
-stars, and calibrate to spectrophotometric models.
+The \gaia\ Mission (\citealt{gaia16}) will deliver more than a billion stellar parallax measurements.
+Only a small fraction (but large number) of these
+measurements will be both precise and purely astrometric.
+\gaia\ will use astrometric parallax to determine the distances of the more precise
+stars, and then calibrate spectrophotometric models.
 These spectrophotometric models, in turn, along with \gaia's on-board
 low-resolution $B_p\,R_p$ spectrophotometry, are used to provide
-distance estimates to more distant stars.
-The full stack required for these distance inferences is complex.
-It involves modeling not just stars, but also the dust in the Milky Way,
+parallax estimates to less precise stars.
+The full stack required for these parallax inferences is complex.
+It involves modeling the stars, as well as the dust in the Milky Way,
 and the response of the telescope itself.
+It might be possible to eliminate some of this complexity using data driven models which don't rely on physical models.
 
-With projects like \project{The Cannon} (\citealt{ness15}; \citealt{casey17}; \citealt{ho17})
+Projects like \project{The Cannon} (\citealt{ness15}; \citealt{casey17}; \citealt{ho17})
 and \project{Avast} (Bedell et al.\ in preparation),
-we are exploring the extent to which our predictive models of
+explore the extent to which our predictive models of
 stars could be purely data-driven or statistical.
@@
 physics-driven models, but they are also usually less interpretable.
-However, the \gaia\ use case is ideal for thinking about purely statistical
+However, the \gaia\ data set is ideal for thinking about purely statistical
 models of stars:
-The distance to a star is a simple latent variable, and the nearby
+The distance to a star is a latent variable, and the nearby
 stars constitute a high signal-to-noise ``training set''.
@@
 
-Here we deliver a demonstration-of-concept project that shows that a
+Here we deliver a demonstration-of-concept that shows that a
 data-driven model of the \gaia\ data could in principle deliver very
-precise distances, without any input of the physics or numerical models of stellar
+precise parallaxes, without any input of the physics or numerical models of stellar
 evolution, interiors, or photospheres.
+HOGG:What is a photometric parallax? <Enter here>
 Most stellar models used to generate photometric distances or photometric
@@
 radiative transfer, nuclear reactions, and atomic and molecular transitions.
-Because there are small issues with each of these components, there are small
-color and luminosity differences with the data---the models are inaccurate in detail---and
+Because there are small issues with each of these components, the models are inaccurate in detail, and
 the physical models therefore produce distance estimates that are biased.
-Besides, they build a long list of physical assumptions (about
+In addition, they build a long list of physical assumptions (about
 nuclear reactions, convection, and thermal timescales, for example)
-into the distance estimates.
-That is, it is impossible, when using physical models,
-to deliver photometric distance estimates that involve minimal assumptions.
+into the parallax estimates.
+When using physical models, it is impossible to deliver photometric parallax estimates that involve minimal assumptions.
 
 However, the use of physical models for distance estimation is not necessary.
-Because there are stars at all luminosities with useful parallax information,
-it is possible to build a stellar photometric model from the data themselves.
-Different stars are observed at different levels of parallax precision,
-so it requires relatively high technology to build this data-driven model
+It is possible to build a stellar photometric model from the data themselves,
+because there are stars at all luminosities with useful parallax information,.
+One challenge is that different stars are observed at different levels of parallax precision,
+so it requires relatively sophisticated technology to build this data-driven model
 using all the data available, fairly and responsibly.
-We have this technology!
+Luckily, we have this technology!
 
-Here we build and use just such a data-driven model. In particular,
-we use all of the \gaia\ \tgas\ data (\citealt{tgas};
+Here we build and use just such a data-driven model to infer more precise parallaxes for each star.
+In particular, we use all of the \gaia\ \tgas\ data (\citealt{tgas};
 these data were generated by the methodology of \citealt{michalik15}), that match to the
@@
 noise-free (or low-noise) color--magnitude diagram (\cmd) of stars.
-We use an Empirical-Bayes approach known as Extreme Deconvolution (XD;
+We build the \cmd\ using an Empirical-Bayes approach known as Extreme Deconvolution (XD;
 \citealt{bovy11}).
@@
 and handles in a principled way, heteroskedastic data.
-Its principle disadvantage for the present purposes is that it requires
+Its disadvantage, for the present purposes, is that it requires
 or implies a strictly Gaussian noise model.
-In the below, this requires us to transform the \cmd\ to a space in which
+As we show below, this requires us to transform the \cmd\ to a space in which
 the noise is approximately Gaussian.
 
-The novel use of \xd\ here is to treat its output---the deconvolved \cmd---as
-a prior for use in inference of individual-star distances.
+We then use the novel \xd\ output---the deconvolved \cmd---as
+a prior for use in inference of individual-star parallaxes.
 These inferences, one per star, provide much more precise parallax, distance,
@@
 of the deconvolved distribution,
-its use as a prior is only approximately valid.
+its use as a prior is technically using the data twice and is therefore only approximately valid.
 However, since the data set is large, the \xd\ approximation is not bad; it is
@@
 
-On a different tack, there has been some discussion in the literature
+Although not a fully Bayesian hierarchical inference, the methodology presented in this paper makes exceedingly weak assumptions
+about the Milky Way, and similarly weak assumptions about the properties of
+stars.
+To date, there has been some discussion in the literature
 about how to use a parallax responsibly to infer a distance (\citealt{astraatmadja16a}; \citealt{astraatmadja16b}).
@@
 Way.
-The methodology presented in this paper makes exceedingly weak assumptions
-about the Milky Way, and similarly weak assumptions about the properties of
-stars.
 The strongest new assumption made in this work is that for every star
@@
 space-based photometry that the \gaia\ Mission will deliver.
-However, as we show below, we get very good results.
+However, as we show below, we get very good results in terms of parallax precision.
 It is promising that we will be able to obtain even better results
 in later \gaia\ data releases, and eventually put distances on $>10^9$
-\gaia\ stars using photometric distance methodologies but without any
+\gaia\ stars using photometric parallax methodologies but without any
 commitment to---or even use of---physical models of stars or the Milky Way.
@@
 Instead of inferring more precise parallax pdfs using the 2D distribution of the \cmd\ as our prior, for this simpler model we will infer more precise pdfs of the value $Y$.
-We will do this simpler inference using a 2D distribution that is a 1D Gaussian in the y-direction, with a running mean $\mu = mx + b$ and some thickness $t$, as our prior.
+We will do this simpler inference using a prior that is also a 2D distribution: a 1D Gaussian in the y-direction, with a running mean $\mu = mx + b$ and some thickness $t$.
 Here the thickness of the distribution mimics the true thickness of the \cmd\ due to age, mass and metallicity.

### 2017-05-08 paperDistance/paper.tex
edited toy model section

@@
 We will do this simpler inference using a prior that is also a 2D distribution: a 1D Gaussian in the y-direction, with a running mean $\mu = mx + b$ and some thickness $t$.
-Here the thickness of the distribution mimics the true thickness of the \cmd\ due to age, mass and metallicity.
+Here the thickness of the toy distribution mimics the true thickness of the \cmd\ due to age, mass and metallicity.
 The $1\sigma$ contours as well as (noise free) samples from the toy model distribution are shown in Figure \ref{fig:toy}, upper left panel.
 Similarly, noisy samples from this distribution are shown in the upper right panel.
-Using a representation of the true distribution (which we learn from the noisy data) as our prior, we can infer some posterior belief of $y_{\true,n}$, a true sample of the distribution, from $y_n$, a noisy sample of the distribution.
+Using a representation of the true distribution (which we learn from the noisy data) as our prior, we can infer some posterior belief over $y_{\true,n}$, a true sample of the distribution, from $y_n$, a noisy sample of the distribution.
 
-Setting up the problem in more detail, we have the true underlying distribution of points and the noisy data
+Setting up the problem in more detail, we have the true samples from the underlying distribution and the noisy samples
 
@@
 
-Here is a good point to step back and acknowledge that our method is odd/strange/different/new (which maybe you're also currently thinking but can't quite express why yet).
-We are using the set of data to determine the noise-deconvolved distribution, and then using that noise-deconvolved distribution to make inferences on the data.
-We are technically using the data twice!
-If we were going to be fully, technically correct, we should learn the distribution, but leave one datum out; the one datum we want to infer the true y value for\footnote{DWH: Remember what this brain dump is for}.
+Here is a good point to step back and acknowledge that our method is a bit strange.
+We are using the set of data to determine the noise-deconvolved distribution, which is an estimate of the true distribution, and then using that noise-deconvolved distribution to make inferences on the data.
+We are technically using the data twice!To be fully, technically correct, we should learn the true distribution while leaving one datum out, the one datum we want to infer the true y value for\footnote{DWH: Remember what this brain dump is for}.
 Leaving one datum out to estimate the prior and then inferring the posterior for that datum is way too expensive for the 1024 points in our toy model, and also for the 1.2 million stars in our actual data set.
-Luckily with large data sets, even at 1024, 1 datum is $\sim0.1\%$ of your dataset and therefore leaving that point out will not change the MML parameters of your noise-deconvolved distribution, and with the \gaia\ dataset of over a million stars, we can safely step into the Empirical Bayes regime.
+Luckily with large data sets, even at 1024, 1 datum is $\sim0.1\%$ of your dataset and therefore leaving that point out will not change the MML parameters of your noise-deconvolved distribution.
+With the \gaia\ dataset of over a million stars, we can safely step into the Empirical Bayes regime.
 (Talk more about emperical bayes here?)
 
-So using Bayes theorem
+So, returning the inference, we use Bayes theorem to turn the likelihood of the noisy data into a posterior believe over the true y value.
 \begin{eqnarray}
 p(y_{\true,n} \given y_n, \tilde{\sigma}_n) &=& p(y_n \given y_{\true,n}, \sigma_n)\,p(y_{\true,n}) \\
-p(y_n \given y_{\true,n}, \sigma_n) &=& \mathcal{N}(y_n \given y_{\true,n}, \sigma_n) \\
-p(y_{\true,n}) &=& \mathcal{N}(y_{\true,n} \given \hat{m}\, x_n + \hat{b}, \hat{t})
+\mathrm{where}\;p(y_n \given y_{\true,n}, \sigma_n) &=& \mathcal{N}(y_n \given y_{\true,n}, \sigma_n) \\
+\mathrm{and}\;p(y_{\true,n}) &=& \mathcal{N}(y_{\true,n} \given \hat{m}\, x_n + \hat{b}, \hat{t})
 \label{eq:toyBayes}
 \end{eqnarray}
-
 Since everything is Gaussian, this can be done analytically by multiplying Gaussians together.
 In Figure \ref{fig:toy} lower left panel, the blue points represent the posterior belief of the true value $p(y_{\true,n} \given y_n, \tilde{\sigma}_n)$ for the highlighted black points.
-The red points are the true y values.
-The blue points in Figure \ref{fig:toy}, lower right panel, show the expectation value and $1\sigma$ uncertainty of the posterior belief for all 1024 $y_n$.
-When the true values are inferred in this way, the associated variance $\tilde{\sigma}_n^2$ is the convolution of the true distribution variance $t^2$ and the measurement variance $\sigma_n^2$.
-The points are colored by their measurement variance $\sigma_n^2$ which shows that noisier measurements are more influenced by the prior and get pulled closer to the center of the MML distribution, our prior.
-In contrast, more certain measurements remain closer to their measured values $y_n$.
-By inferring the true y value from the measured y value in addition to the prior information of the distribution the y value was drawn from, this we have better estimates of the true y value for each point.
-Bringing the toy model back to our original problem, using the noise-deconvolved CMD as our prior allows us to get better estimates of the parallax for each star.
+The red points are the associated true y values.
+In Figure \ref{fig:toy}, lower right panel, the blue points show the expectation value and $1\sigma$ uncertainty of the posterior belief for all 1024 $y_n$.
+When the true values are inferred in this way, the associated variance $\tilde{\sigma}_n^2$ is the convolution of the true distribution variance, $t^2$, and the measurement variance, $\sigma_n^2$.
+The points are colored by their measurement variance $\sigma_n^2$, which shows that noisier measurements are more influenced by the prior and get pulled closer to the center of the MML distribution, our prior.
+In contrast, more certain measurements remain closer to their measured values, $y_n$.
+We now have better estimates of each true y value by inferring each true y value from the measured y value and using an estimate of the true distribution the measured y value was drawn from as the prior.
+Bringing the toy model back to our original problem, we can get better estimates of the true parallax for each star by inferring the true parallax from the measured parallax and using the noise-deconvolved \cmd, an estimate of the true \cmd, as our prior.
 
@@
 
-We make the following assumptions. The method we present is correct
+To infer the true parallax to a star using the noise-deconvolved \cmd\ as a prior, we make the following assumptions. The method we present is correct
 and justifiable under these assumptions, each of which is questionable
 in its own right. We return to criticize these assumptions---and the
-method that flows from them---in the Discussion below.
+method that flows from them---in the discussion section below.
 \begin{description}

### 2017-05-08 comparePrior.py
tweaks

@@ -193,10 +193,10 @@ def dataViz(survey='2MASS', ngauss=128, quantile=0.05, dataFilename='All.npz', i
     """
     dp.plot_sample(color[positive], testXD.absMagKinda2absMag(y), sample[:,0], testXD.absMagKinda2absMag(sample[:,1]),
                 xdgmm, xerr=color_err[positive], yerr=[yerr_minus, yerr_plus], xlabel=xlabel, ylabel=ylabel, xlim=xlim, ylim=ylim, errSubsample=2.4e3, thresholdScatter=2., binsScatter=200, contourColor=contourColor)
-    dataFile = 'data_noDust.png'
-    priorFile = 'prior_' + str(ngauss) +'gauss.png'
-    os.rename('plot_sample.data.png', dataFile)
-    os.rename('plot_sample.prior.png', priorFile)
+    dataFile = 'data_noDust.pdf'
+    priorFile = 'prior_' + str(ngauss) +'gauss.pdf'
+    os.rename('plot_sample.data.pdf', dataFile)
+    os.rename('plot_sample.prior.pdf', priorFile)
     #import pdb; pdb.set_trace()
     data = np.load(dustFile)
     dustEBV = data['ebv']
@@ -226,16 +226,16 @@ def dataViz(survey='2MASS', ngauss=128, quantile=0.05, dataFilename='All.npz', i
         yerr_minus = absMag - absMagYMinus
         yerr_plus = testXD.absMagKinda2absMag(yplus) - absMag
         #notnan = ~np.isnan(color[notnans]) & ~np.isnan(absMag)
-
-        ascii.write([color[notnans], absMag, color_err[notnans], yerr_minus, yerr_plus, c], 'cmdExpectation.txt', names=['color', 'absMag', 'color_err', 'absMag_errMinus', 'absMag_errPlus', 'logDeltaVar'])
+        contourColor = 'k'
+        ascii.write([color[notnans], absMag, color_err[notnans], yerr_minus, yerr_plus, c[notnans]], 'cmdExpectation.txt', names=['color', 'absMag', 'color_err', 'absMag_errMinus', 'absMag_errPlus', 'logDeltaVar'])
         dp.plot_sample(color[notnans], absMag, sample[:,0], testXD.absMagKinda2absMag(sample[:,1]),
                     xdgmm, xerr=color_err[notnans], yerr=[yerr_minus, yerr_plus], xlabel=xlabel, ylabel=ylabel, xlim=xlim,
[…]

### 2017-05-08 demo_plots.py

@@ -103,7 +103,7 @@ def kdeDensity(ax, x, y, threshold=0.01, bins=100, s=1, lw=0, alpha=1):
     #fig.colorbar(im)
     return ax
 
-def plot_sample(x, y, samplex, sampley, xdgmm, xlabel='x', ylabel='y', xerr=None, yerr=None, ylim=(6, -6), xlim=(0.5, 1.5), errSubsample=1.2e6, thresholdScatter=0.1, binsScatter=200, c='black',  norm=None, cmap=None, contourColor='k', posterior=False, ind=None):
+def plot_sample(x, y, samplex, sampley, xdgmm, xlabel='x', ylabel='y', xerr=None, yerr=None, ylim=(6, -6), xlim=(0.5, 1.5), errSubsample=1.2e6, thresholdScatter=0.1, binsScatter=200, c='black',  norm=None, cmap=None, contourColor='k', posterior=False, ind=None, rasterized=True):
     prng = np.random.RandomState(1234567890)
     setup_text_plots(fontsize=16, usetex=True)
     plt.clf()
@@ -120,7 +120,7 @@ def plot_sample(x, y, samplex, sampley, xdgmm, xlabel='x', ylabel='y', xerr=None
     levels = 1.0 - np.exp(-0.5 * np.arange(1.0, 2.1, 1.0) ** 2)
     cNorm  = plt.matplotlib.colors.LogNorm(vmin=3, vmax=1e5)
     #ax1.hist2d(x, y, bins=100, norm=cNorm, cmap='Greys')
-    im = corner.hist2d(x, y, ax=ax1, levels=levels, bins=200, no_fill_contours=True, plot_density=False, color=contourColor)
+    im = corner.hist2d(x, y, ax=ax1, levels=levels, bins=200, no_fill_contours=True, plot_density=False, color=contourColor, rasterized=rasterized, plot_contours=False)
     #im = ax1.scatter(x, y, s=1, lw=0, c=c, alpha=alpha, norm=norm, cmap=cmap)
 
     ax2 = figData.add_subplot(122)
@@ -131,7 +131,7 @@ def plot_sample(x, y, samplex, sampley, xdgmm, xlabel='x', ylabel='y', xerr=None
     ax3 = figPrior.add_subplot(121)
     #ax3.hist2d(x, y, bins=100, norm=cNorm, cmap='Greys')
     #kdeDensity(ax3, samplex, sampley, threshold=thresholdScatter, bins=binsScatter, s=1, lw=0, alpha=alpha)
[…]

### 2017-05-09 paperDistance/paper.tex
done with methods, onto data and results

@@
   stars measured at lower signal-to-noise have analogs measured at
-  higher signal-to-noise.  In detail, this assumption is wrong, of
-  course (for example, distant halo stars are different than local
+  higher signal-to-noise.  In detail, this assumption is wrong, (for example, distant halo stars are different than local
   disk stars), but the stationarity assumption is fairly weak; as long
@@
   of stars seen at low signal-to-noise, the results will not be
-  strongly biased, but this assumption is deep and fundamental to this
+  strongly biased. Although a weak assumption, it is deep and fundamental to this
   work.
 \item[big data] We assume that we have large numbers of stars, large enough that
-  an empirical-Bayes or maximum-marginalized likelihood is a safe
+  empirical Bayes, or maximum-marginalized likelihood, is a safe
   approximation to full Bayesian inference. This assumption will be
   least true in the least-populated parts of the \cmd. In the extreme
-  case of a stellar type with only one (statistically distinct)
-  exemplar in the data set (for example, a white dwarf), this assumption will be maximally wrong.
+  case, if a portion of the \cmd\ has only one (statistically distinct)
+  exemplar, for example, a white dwarf, inference for these cases will be biased.
 \item[noise model] We assume that the \tgas\ parallax uncertainties dominate the
@@
   are correct. Here we are making use of the fact that, for
-  de-noising, under-estimation of measurement noise is conservative
-  (it leads to under-deconvolution).
+  de-noising, under-estimation of measurement noise is conservative, and leads to under-deconvolution.
 \item[dust] We treat the \panstarrs-based three-dimensional dust maps (\citealt{green15})
   as correct in their median dust estimates, and their
-  effects on color and magnitude. As we optimize the empirical-Bayes model, we iteratively update each star's
+  effects on color and magnitude. As we optimize the empirical Bayes model, we iteratively update each star's
   dust correction at its updated, inferred distance. Our assumption is that these corrections are
-  good enough, and uncertainties in these are not dominant, for either
-  color or absolute magnitude.
-\item[mixture of Gaussians] We assume that the \cmd\ can be represented with a mixture of
+  good enough, and uncertainties in these are not dominant, for neither
+  color nor absolute magnitude.
+\item[mixture of Gaussians] We assume that the \cmd\ can be represented by a mixture of
   Gaussians in a particular transformed space. This mixture we fix
-  (with only heuristic analysis) at 128, 512, and 2048 Gaussians in
-  different experiments.
+  (with only heuristic analysis) at 128 Gaussians.
 \item[no physics] We make no use of stellar models; our only assumptions about
@@
 
-Under these assumptions, we would like to infer the true parallax $\varpi_{\true}$ for each star given its measured parallax $\varpi$ and associated uncertainty $\sigma_{\varpi}$ from \tgas.
-The strength of our inference method here lies in how we generate our prior information on the true parallax.
+Under these assumptions, we would like to infer the true parallax, $\varpi_{\true}$, for each star given its measured parallax, $\varpi$, and associated uncertainty, $\sigma_{\varpi}$, from \tgas.
+The strength of our inference lies in how we generate our prior information of $\varpi_{\true}$.
 Supplementary to the trigonometric measure of distance, the photometry of stars is also informative of their distances.
-The intrinsic temperatures and luminosities of stars tightly correlate leading to the observed colors and absolute magnitudes of stars also tightly correlating, usually visualized as the color--magnitude diagram (\cmd).
+The intrinsic temperatures and luminosities of stars tightly correlate leading to the observed colors and absolute magnitudes of stars also tightly correlating, usually visualized as the \cmd.
 The tight correlation of the \cmd\ benefits our inference because the luminosity or absolute magnitude of a star, when compared with its brightness or apparent magnitude, contains distance information.
-We include this distance information in the inference of the true parallax by using the \cmd\ as our prior.
+We include this distance information in the inference of t$\varpi_{\true}$ by using the \cmd\ as our prior.
 Specifically, we use the \cmd\ generated from the data themselves as our prior.
 Instead of assuming a functional form for the spatial distribution of stars in the Milky Way (\citealt{astraatmadja16b}), or relying on stellar models (\citealt{gaia16}), we build a noise-deconvolved estimate of the \cmd\ from the observed colors and absolute magnitudes of our full set of stars.
-This \cmd\ model becomes, for each star---given its observed color and magnitude (and uncertainties on those)---a prior pdf for the distance to that star.
+This \cmd\ model becomes, for each star---given its observed color and magnitude (and uncertainties on those)---a prior pdf for the parallax of that star.
 
 Because our prior is the \cmd, our inference will include not only the parallax from \gaia, but also the photometry of the stars.
-For inferring the true parallax, including generating our \cmd\ prior, we use the $J$ and $K$ band photometry from \tmass, as well as the parallax from \gaia.
+For inferring $\varpi_{\true}$, we use the $J$ and $K$ band photometry from \tmass, as well as the parallax from \gaia.
 Our true, latent values are the $(J-K)^{true}_n$ color and the $J$ band absolute magnitude $M^{true}_{J,n}$ for each star, which we encapsulate in the true vector $\mathbf{Y}$.
@@
 \end{equation}
-
-The transformation of the dust corrected absolute magnitude, Equation \ref{eq:transform}, is required to to keep the uncertainties Gaussian, which, under some assumptions along with this transformation, we maintain.
-Gaussian uncertainties are required to generate our prior using \xd\, but more about this and our dust corrections below.
+The transformation of the dust corrected absolute magnitude is required to to keep the uncertainties Gaussian, which, under some assumptions along with this transformation, we maintain.
+Gaussian uncertainties are required to generate our prior using \xd, but more about this and our dust corrections below.
 Our prior assumptions $I$ are the uncertainties in each of these measurements, as well as the Greene dust map. So our full terminology is
@@
 \end{equation}
-
 where $E(B-V)_n$ is the output of the Green et al dust model, and $Q_{\lambda}$ is the correction factor for a photometric band assuming some reddening law.
@@
 \end{equation}
-where $p(\varpi \given \varpi_{\true}, \sigma^2_{\varpi})$ is the likelihood of the observed parallax, $p(\varpi_{\true})$ is a prior probability density (pdf) for the
+where $p(\varpi \given \varpi_{\true}, \sigma^2_{\varpi})$ is the likelihood of the observed parallax, $p(\varpi_{\true})$ is a prior pdf for the
 parallax, and $Z$ is a normalization constant.
-
 But life isn't simple, and our prior is a 2D function in color and transformed absolute magnitude space, so our full inference in done in 2D space as well.
@@
 \end{equation}
-
 with \\
@@
 \]
-
 where $\mathcal{N}$ is a 2D Gaussian function with mean $\mathbf{Y}$ and variance $V_n$.
-Here we have assumed that the parallax uncertainty is significantly larger than the $J$ band uncertainty, and therefore the dominant uncertainty in the transformed absolute magnitude, so we take $J_n^C$ as a point estimate and only propagate the parallax uncertainty.
-This is to keep our uncertainties Gaussian so we can use \xd\ generate our prior below.
+Here we have assumed that the parallax uncertainty is significantly larger than the $J$ band uncertainty, and therefore the dominant uncertainty in the transformed absolute magnitude.
+So we take $J_n^C$ as a point estimate and only propagate the parallax uncertainty.
+Again, this is to keep our uncertainties Gaussian so we can use \xd\ generate our prior below.
 
@@
 \end{equation}
-
-where $p(\mathbf{Y}_n)$ is our prior.
+where $p(\mathbf{y}_n \given \mathbf{Y}_n, I_n)$ is the likelihood of the data vector, and $p(\mathbf{Y}_n)$ is our prior.
 
@@
 We fit this transfromed \cmd\ using \xd\ which deconvolves the data with its uncertainties, which are heterogeneous measurement uncertainties.
-It generates the underlying distribution from which the uncertain data were drawn from.
+It generates an estimate of the underlying distribution from which the uncertain data were drawn from.
 \xd\ generates a represention of the distribution as a mixture of Gaussians which maximizes the marginalized likelihood of the data, marginalizing over the true data.
@@
 \end{eqnarray}
-
 where $\alpha$ encapsulates the parameters of each Gaussian in the mixture of Gaussians that defines our \cmd\ prior.
@@
 
-To infer $\varpi_{\true}$ we first do the inference in the 2D space of the \cmd
+To infer $\varpi_{\true}$, we first do the inference in the 2D space of the \cmd. Using Bayes theorem,
 
@@
 \end{equation}
-
-We then project the 2D posterior into the 1D transformed absolute magnitude space, and do a little algebra to get
+We then project the 2D posterior into the 1D transformed absolute magnitude space, and do a little algebra to solve for $\varpi_{\true}$
 
@@
 p(\mathbf{Y}_0 \given \mathbf{y}, \alpha) &\rightarrow& p(\varpi_{\true} \given y, \alpha) \, f(\varpi_{\true}) \\
-f(\varpi_{\true}) &=& \begin{cases}
+\mathrm{where} \; f(\varpi_{\true}) &=& \begin{cases}
               1 \quad \mathrm{if} \quad -2 < \mathrm{log}_{10} \varpi_{true} < 2\\
@@
 \end{eqnarray}
-with $f(\varpi_{true})$ being a window function to insure the true parallax is positive and put it on a similar grid for all the posteriors.
-And with a jacobian this becomes a posterior on distance.
+with $f(\varpi_{true})$ being a window function to insure $\varpi_{\true}$ is positive and to put $\varpi_{\true}$ on a similar grid for all the posteriors.
 
@@
 
-We use stars crossmatched in \tgas\ and \tmass\, and lie within the observing footprint of \panstarrs\ to access the Greene 3D dust model.
+We use stars crossmatched in \tgas\ and \tmass\, that also lie within the observing footprint of \panstarrs\ to access the Greene 3D dust model.
 We require the photometry have real values, and nonzero, real errors, and remove a small selection of \tmass\ stars that have zero color and zero J band magnitude.
@@
 
-We correct for dust using 3D Bayestar (Schlafly+Finkbeiner) with an estimate of the distance from iteratively sampling our posterior distances.
-Dust corrections are non trivial for stars with poor signal to noise.
+We need to correct the photometry used in our method, both for the generation of the prior and the likelihood, for dust.
+With the emergence of 3D dust maps, it is now possible to estimate dust corrections to stars within the Milky Way.
+The challenge is getting a proper estimate of the distance to the star before inferring $\varpi_{\true}$, since we need a dust correction prior to doing our inference.
+Dust corrections are particularly non trivial for stars with poor parallax signal to noise.
 Stars with poor signal to noise have a likelihood that is consistent with infinite distance and can therefore have severe dust corrections.
 This is most obvious in the giant stars, which tend to have the lowest signal to noise, and can therefore have dust corrections $> 2$ magnitudes.
+Although the dust model is probabilistic, as well as our distance inference, to move forward, we break from a fully Bayesian framework, in which we might infer the distance and the dust simultaneously (more about this in the discussion).
+Instead, we correct for dust using a point estimate from 3D dust map (Schlafly+Finkbeiner) sampled at a distance estimated from iteratively sampling our posterior pdf over $\varpi_{\true}$.
 To apply a dust correction to the \tmass\ photometry, we first generate our prior using the observed, attenuated photometry.
-Using this raw prior we infer more precise distances to all the stars in our sample.
-We then use this more precise distance posterior to get a measurement of dust in the 3D dust map.
-This approach minimizes severe distances from the likelihoods alone and their associated severe dust corrections.
+Using this raw prior, we infer more precise parallaxes to all the stars in our sample.
+We then use this more precise parallax posterior to get a measurement of dust in the 3D dust map.
+This approach minimizes severe distances, and their associated severe dust corrections, when using the likelihoods alone.
 In detail, we take the $5\%$ quantile of each distance posterior, so the closest bit of the posterior, and sample the Green et al 3D dust map at that distance.

### 2017-05-11 paperDistance/paper.tex
mostly through the discussion

@@
 Although the dust model is probabilistic, as well as our distance inference, to move forward, we break from a fully Bayesian framework, in which we might infer the distance and the dust simultaneously (more about this in the discussion).
-Instead, we correct for dust using a point estimate from 3D dust map (Schlafly+Finkbeiner) sampled at a distance estimated from iteratively sampling our posterior pdf over $\varpi_{\true}$.
-To apply a dust correction to the \tmass\ photometry, we first generate our prior using the observed, attenuated photometry.
+Instead, we correct for dust using a point estimate from the 3D dust map (Schlafly+Finkbeiner) sampled at a distance estimated from iteratively inferring and then sampling our posterior pdf over $\varpi_{\true}$.
+In more detail, to apply a dust correction to the \tmass\ photometry, we first generate our prior using the observed, attenuated photometry.
 Using this raw prior, we infer more precise parallaxes to all the stars in our sample.
-We then use this more precise parallax posterior to get a measurement of dust in the 3D dust map.
-This approach minimizes severe distances, and their associated severe dust corrections, when using the likelihoods alone.
-In detail, we take the $5\%$ quantile of each distance posterior, so the closest bit of the posterior, and sample the Green et al 3D dust map at that distance.
-We take the $5\%$ quantile, as apposed to the median for example, to do a light dust correction because over correcting for dust can shoot the likelihood into an unphysical, different space of the CMD and then bias the distance inference.
-For each distance we sample the median of the probabilistic dust model.
-We apply each light dust correction to our \tmass\ photometry given by Equation \ref{eq:data} where $Q_{\lambda}$ = [0.709, 0.302] for bands [J, K] respectively, taken from table (REF TABLE SFD), and $E(B-V)$ is the output of the Green et al dust models.
+We then use this more precise parallax posterior to get a measurement of dust in the 3D dust map for each star.
+And repeat.
+This approach minimizes severe distances, and their associated severe dust corrections, which would occur if using the likelihoods alone.
+In detail, we take the $5\%$ quantile of each distance posterior, so the closest bit of the posterior, and sample the 3D dust map at that distance.
+We take the $5\%$ quantile, as apposed to the median, for example, to do a light dust correction, because over correcting for dust can shoot the likelihood into an unphysical, different space of the \cmd and bias the parallax inference.
+For each distance, we sample the median of the probabilistic dust model.
+We apply each light dust correction to our \tmass\ photometry given by Equation \ref{eq:data} where $Q_{\lambda}$ = [0.709, 0.302] for bands [J, K] respectively, taken from table (REF TABLE SFD).
 We then regenerate our prior using the lightly dust corrected photometry.
-We do not update the errors, but keep them the same as before the dust correction, which is wrong and under representing our errors but this leads again to a lighter deconvolution.
+We do not update the errors, but keep them the same as before the dust correction, which is wrong and under representing our errors but this leads, again, to a lighter deconvolution.
 We could add the covariance of the dust to our likelihoods, but this is beyond the scope of this demonstration of concept.
@@
 
-\subsection{Prior}
+\subsection{Output Data}
+Using a data driven model of the observed \cmd\ we inferred a posterior belief of the true parallax of each star. Given our method, we call our posterior parallax the photometric parallax.
+DWH: Placeholder for the output format for our data products.
+RA, DEC, l, b, tgas source ID, tycho-2 source ID, parallax samples, parallax posterior, flag (some fraction of posterior not in window, negative gaia parallax, ...)
 
@@
 
-\begin{figure}
-\centering
-  \includegraphics[width=\textwidth]{priorNgaussComparison.png}
-\caption{Prior: the XD generated mixture of gaussians we use as our prior.
-To determine the number of gaussians we used cross validation, using 70\% of our sample as a training set and 30\% as the test set.
-Increasing the number of gaussians we used continued to increase the likelihood of the test set of data, showing that we are no over fitting.
-Using more gaussians as our prior did not substantially change our posterior results, but it did start to highlight the features of the CMD.}
-\label{fig:priorCompare}
-\end{figure}
+%\begin{figure}
+%\centering
+%  \includegraphics[width=\textwidth]{priorNgaussComparison.png}
+%\caption{Prior: the XD generated mixture of gaussians we use as our prior.
+To determine the number of gaussians we used cross validation, using 70\% of our %sample as a training set and 30\% as the test set.
+Increasing the number of gaussians we used continued to increase the likelihood of the test set of data, %showing that we are no over fitting.
+Using more gaussians as our prior did not substantially change our posterior results, but it did start to highlight the %features of the CMD.}
+%\label{fig:priorCompare}
+%\end{figure}
 
@@
 
-\subsection{Results: Distances to M67}
+\subsection{Distances to M67}
 \begin{figure}
@@
 
-As proof of concept we inferred the distances to stars in the open cluster M67.
+As proof of concept of our demonstration of concept, we inferred the distances to stars in the open cluster M67.
 It is an open cluster in the constellation Cancer and is estimated to be 3-5 billion years old and 800-900 pc away.
@@
 The cluster is visibly obvious as a clustering of stars in a visualization of the \gaia\ data on the 2D sky.
-With our converged dust values, and it's associated prior, we calculate the posterior of the true parallaxes to the stars within this window on the sky around M67.
+With our converged dust values, and its associated prior, we calculate the posterior over $\varpi_{\true}$ for each star within this window on the sky around M67.
 Figure \ref{fig:m67} shows the comparison of the likelihoods and posteriors.
-The vertical lines bracket the parallaxes of, or distances to, the cluster.
+The vertical lines bracket the previously measured parallaxes of, or distances to, the cluster.
 The likelihoods of the observed parallaxes, shown in the top row, are broad and the cluster is not obvious at all.
-The posteriors, shown in the bottom row, show a sharp peak in the probability at the previously determined distance to the cluster, the vertical lines, for far more stars than the likelihoods would imply.
+The posterior pdf, shown in the bottom row, are more sharply peaked at the previously determined distance to the cluster, the vertical lines, for far more stars than the likelihoods.
 Our prior is not only making parallaxes more precise but also (possibly) increasing the accuracy.
-Too strong?
-
-\subsection{Output data}
-DWH: Placeholder for the output format for our data products.
-RA, DEC, l, b, tgas source ID, tycho-2 source ID, parallax samples, parallax posterior, flag (some fraction of posterior not in window, negative gaia parallax, ...)
 
@@
 models, nor models of the Milky Way.
-That is, the geometric parallaxes can calibrate a set of photometric
-models that are purely statistical---a model of the data rather than
+We used the geometric parallaxes to calibrate a photometric
+model that is purely statistical, which is a model of the data rather than
 a model of stars \foreign{per se}.
@@
 
-We obtained the photometric parallaxes in this project
-by building a data-driven model of the
-color--magnitude diagram (\cmd) of the stars in the \tgas\ data set,
-and using it as a prior pdf for Bayesian interences.
+Transition clause/sentence. We obtained the photometric parallaxes in this project
+by building a data-driven model of \cmd\ of the stars in the \tgas\ data set,
+and using it as a prior pdf for Bayesian inference.
 The posterior pdfs for distance that we obtain are, in general, much
@@
 
-Although we have not performed principled hierarchical Bayesian inference in this
-project, we built our prior for each star's distance by
+Transition clause/sentence. Although we have not performed principled hierarchical Bayesian inference in this
+project, we built our prior for each star's parallax by
 performing a statistically responsible deconvolution
@@
 assumptions.
-This deconvolution makes use of a reasonable noise model and an
-assumption of stationarity to shrink the parallax uncertainties, and
-do so dramatically for the stars measured at lower signal-to-noise.
-Because the method we use (Extreme Deconvolution or \xd, which is
-an Empirical Bayesian maximum-marginalized-likelihood estimator) accounts for
+This deconvolution, using a reasonable noise model and an
+assumption of stationarity, shrinks the parallax uncertainties, most dramatically for the stars measured at lower signal-to-noise.
+Because the method we use, \xd, which is
+an Empirical Bayesian maximum-marginalized-likelihood estimato, accounts for
 heteroskedastic noise, we were able to build our photometric parallax
 model---which is a model of the \cmd---using
-all the data, not just the data with the highest signal-to-noise or
-largest parallaxes.
-That is, the model for the \cmd\ we have built is representative for
-(a subset of) the \tgas\ Catalog.
-Inasmuch as that is true, we expect any (sensible) distance estimates
-we generate from our distance posterior pdfs to be only weakly biased.
+all the data, not just the data with the highest signal-to-noise.
+So, the model for the \cmd\ we built is representative for
+our subset of the \tgas\ Catalog.
+Because our model is representitive of our full subset catalog, we expect any (sensible) parallax estimates
+we generate from our parallax posterior pdfs to be only weakly biased.
 
-The Empirical Bayesian methodology is a good approximation to full
-hierarchical inference in the limit of large numbers of data points.
-That is, when no individual star is carrying a lot of weight in the
+Transition clause/sentence. The empirical Bayesian methodology is a good approximation to full
+hierarchical inference in the limit of large numbers of data points, such that no individual star is carrying a lot of weight in the
 model of the \cmd.
-With its millions of stars, the \tgas\ Catalog may appear to safely
+With its millions of stars, the \tgas\ Catalog appears to safely
 live in this limit.
 However, because there are parts of the \cmd\ that are poorly populated,
-even in \tgas, the Empirical Bayes approximation may be bad for some
-color ranges.
+even in \tgas, the empirical Bayes approximation may be bad for some
+portions of the \cmd.
 In particular, there appear to be essentially no white dwarf stars,
@@
 
-In some sense, the biggest limitation of the data-driven \cmd\ model
-created in this project is that we made no use of any kind of \tgas\ selection
+Transition clause/sentence. The biggest limitation of our method for generating the \cmd\ from the data is that we made no use of any kind of \tgas\ selection
 function or completeness estimate or inverse-selection-volume corrections.
-That is, the model is a model of the \tgas\ Catalog
+This model is a model of the \tgas\ Catalog
 (or really the \tgas--\tmass\ intersection), not of the stars in
@@
 Importantly, because we used \emph{all} of the stars in the \tgas--\tmass\
-intersection, the model \emph{is} representive of the full catalog,
+intersection, the model \emph{is} representive of the full intersection,
 even the distant parts, where all the parallaxes are measured at low
@@
 the high signal-to-noise sources; such a model would be biased towards
-stellar types and compositions found locally, lower-luminosity stars,
+stellar types and compositions found locally, as well as lower-luminosity stars,
 and stars which happened to get good parallaxes.
@@
 to inferences within that same \tgas--\tmass\ intersection.
-For example, it would be biasing to use this same prior to infer
+It would be biasing to use this same prior to infer
 distances for the full billion-star catalog released in \gaia\ DR1
@@
 
-The outputs of this project include posterior pdfs and posterior
-samplings for stellar parallaxes or distances.
+Transition clause/statement. The outputs of this project include posterior pdfs for stellar parallaxes.
 It is tempting to treat these outputs as equivalent to some kind of catalog of
@@
 measurements and the de-noised measurements.
-The most important difference, for users of our output, between our
-output and the \tgas\ data input, is that the \tgas\ data is the representation
+For users of our output, the most important difference between our
+output and the \tgas\ data input is that the \tgas\ data is the representation
 of a likelihood function, whereas our ouptut is a
 representation of a posterior pdf, one per star.
-If used naively, posterior information can lead to serious statistical
+Posterior information used naively can lead to serious statistical
 errors, because each data point has had a prior multiplied in.
@@
 
-There are other prior information options when inferring distances to stars.
+There are other options for prior information when inferring distances to stars.
 One option is the exponentially declining spatial density of stars (cite: Tri, CBJ).
 It is parameterized by a scale length, and has the nice property that it is smooth out to very large distances so the posteriors are also very smooth.
-It is a fairly weak prior and basically sets a smooth scale at which it's challenging for stars to lie at, and also allows for negative parallaxes.
-It is a fairly broad prior so the median \gaia\ star has an increased posterior variance.
-It has the largest affect on the most noisy and negative parallaxes and has the nice property of bringing these noisy or negative measurements to reasonable parallaxes.
+It is a fairly weak prior.
+It sets a smooth scale distance beyond which the posterior struggles to maintain much probability, and also converts a negative parallax likelihood into a posterior of very small parallax.
+It is a fairly broad prior so the median \gaia\ star has an increased posterior variance of XXX.
+It has the largest effect on the most noisy and negative parallaxes, and has the nice property of bringing these noisy or negative measurements to reasonable parallaxes.
 Given this prior, what does the observed \cmd\ look like?
@@
 The lower main sequence is very narrow while the upper main sequence is much wider.
-This may be an age variation on the main sequence mixed with a malquist bias of only seeing a very small volume of lower main sequence stars but a much larger volume of upper main sequence stars.
+This may be an age variation on the main sequence mixed with a malquist bias of seeing a much larger volume of upper main sequence stars.
 There is a plausible binary sequence of stars with about twice the brightness as the main sequence.
-The helium burning red clump is (surprisingly) prominent for the volume DR1 probed, and the red giant branch is straight/narrow.
+The helium burning red clump is (surprisingly) prominent for the volume DR1 probed, and the red giant branch is fairly straight and narrow.
 The main sequence turn off is noticeable and also surprisingly narrow, possibly a reflection of the star formation history of the local volume also mixed with a malmquist bias.
@@
 Within our probabilistic framework, we improperly handle dust.
-We are taking a point estimate of the probabilistic dust model at a quantile distance of our posterior, correcting our photometry using this point estimate and rederiving the prior.
-A full, proper account would sample the dust model at distances sampled from the posterior and propagate those samples back into the prior during the iterative proccess.
-The issue, the reason we don't do a full probabilistic treatment, is that this proper dust distribution would not be a Gaussian distribution which \xd\ requires.
+We are taking a point estimate of the probabilistic dust model at a quantile of our distance posterior, correcting our photometry using this point estimate, and then re-deriving the prior.
+A full, proper account would sample the dust model at distances sampled from the posterior, and propagate those samples back into the prior during the iterative process.
+The reason we don't do a full probabilistic treatment, is that this proper dust distribution would not be a Gaussian distribution, which \xd\ requires.
 We could calculate this more complex dust distribution and estimate it as a Gaussian distribution to feed to \xd, however, this is beyond the scope of this paper.
@@
 Stationarity assumes both the low and high signal to noise objects are being drawn from the same distribution, independent of noise.
-This isn't precisely true because most Gaia objects have similar noise $(\sim 3 mas)$, but their signal (parallax) depends on their distance.
-Nearby stars have higher signal (parallax) because they are closer and the nearby stars are mostly main sequence stars (IMF, timing, malmquist).
-However, we can see bright giants to much larger distances and therefore lower signal or parallax, leading to lower signal to noise.
+This isn't precisely true because most \gaia\ objects have similar noise $(\sim 3 \, \mathrm{mas})$, but their signal (parallax) depends on their distance.
+Nearby stars have higher signal (parallax) because they are closer and the nearby stars are mostly main sequence stars due to the IMF, timing, and malmquist bias.
+However, we can see bright giants to much larger distances, and therefore lower signal or parallax, leading to lower signal to noise.
 So the red giant branch is a population of relatively lower signal to noise objects compared with the main sequence.
-But I think this is ok, that the low and high signal to noise objects have to be drawn from the same distribution within some feature or patch of the space.
+But this does not violate the stationarity assumption in detail, because the low and high signal to noise objects have to be drawn from the same distribution within some feature or patch of the space.
 This is all assuming the Gaia noise model is correct.
@@
 But with the quoted uncertainties this seems unlikely.
-As you move up the giant branch and probe farther giants the signal to noise will continue to diminish but I think this is okay since the color uncertainties are not large.
+As you move up the giant branch, and probe farther giants, the signal to noise will continue to diminish, but because the color uncertainties are not large, the S/N spread within a patch of the \cmd\ is well mixed.
 Possibly a metallicity trend with distance, placing more nearby (higher S/N) objects at slightly different positions on the CMD pulling lower S/N objects to that distribution for our prior.

### 2017-05-11 paperDistance/paper.tex
cmd feature plot

@@
 
+%How this compares to other Bayesian attempts at inferring distances,
+%like, eg, the Tri--Coryn papers. Favorably! And with \emph{fewer}
+%assumptions, in fact! What form should this comparison be in?
 There are other options for prior information when inferring distances to stars.
@@
 \subsection{What's That Feature?}
+\begin{figure}
+\centering
+  \includegraphics[width=\textwidth]{whatsThatFeature.pdf}
+\caption{Features in the \cmd: The expectation values from the posterior pdfs of the true parallax}
+\label{fig:wtf}
+\end{figure}
+
 The denoised expectation values of the parallax projected into the \cmd\ show interest features.
@@
 Come back to the assumptions. What happens if we change or violate them?
-
+Put in citations
 Long discussion of the noise model. We took serious short-cuts. Call out to Boris.
@@
 Compare with literature stellar models
-
-How this compares to other Bayesian attempts at inferring distances,
-like, eg, the Tri--Coryn papers. Favorably! And with \emph{fewer}
-assumptions, in fact! What form should this comparison be in?
 

### 2017-05-12 paperDistance/paper.tex
conclusion reminder
some touches

@@
 We now have better estimates of each true y value by inferring each true y value from the measured y value and using an estimate of the true distribution the measured y value was drawn from as the prior.
+HOGG: describe why the blue and red points have a different morphology
 Bringing the toy model back to our original problem, we can get better estimates of the true parallax for each star by inferring the true parallax from the measured parallax and using the noise-deconvolved \cmd, an estimate of the true \cmd, as our prior.
@@
 
-Come back to the assumptions. What happens if we change or violate them?
-Put in citations
-Long discussion of the noise model. We took serious short-cuts. Call out to Boris.
+HOGG: Come back to the assumptions. What happens if we change or violate them?
+Anderson: Put in citations
+HOGG: Long discussion of the noise model. We took serious short-cuts. Call out to Boris.
 
@@
 Compare with literature stellar models
+
+\section{Conclusion}
+Summarizing remarks:
+
+Concluding remarks:
 

### 2017-05-16 paperDistance/paper.tex
remove four HOGGs; add four HOGGs

--- 2017-05-11 paperDistance/paper.tex
@@
 % - zeroth draft / re-write the method section
-% - add ORCIDs like Hogg's
 
@@
 evolution, interiors, or photospheres.
-HOGG:What is a photometric parallax? <Enter here>
-Most stellar models used to generate photometric distances or photometric
-parallaxes have been physical models, based on gravity, fluid mechanics,
+That is, this is a data-driven or purely empirical \emph{photometric
+  parallax} method.
+A photometric parallax is a parallax (or distance) estimate based on
+the photometry (colors and apparent magnitude or magnitudes) of a
+star.
+Fundamentally, photometric parallaxes capitalize on the fact that the
+apparent magnitude of a star is a strong function of distance, and
+that the absolute magnitude is a strong function of color (or surface
+temperature) and evolutionary stage.
+
+Most stellar models used to generate photometric
+parallaxes have been \emph{physical} models, based on gravity, fluid mechanics,
 radiative transfer, nuclear reactions, and atomic and molecular transitions.
 Because there are small issues with each of these components, the models are inaccurate in detail, and
-the physical models therefore produce distance estimates that are biased.
+the physical models therefore produce parallax (and hence distance) estimates that are biased.
 In addition, they build a long list of physical assumptions (about
@@
 We now have better estimates of each true y value by inferring each true y value from the measured y value and using an estimate of the true distribution the measured y value was drawn from as the prior.
-Bringing the toy model back to our original problem, we can get better estimates of the true parallax for each star by inferring the true parallax from the measured parallax and using the noise-deconvolved \cmd, an estimate of the true \cmd, as our prior.
 
-\section{Assumptions and methods}
+In detail, the posterior points shown in the lower-right panel of
+Figure~\ref{fig:toy} look more concentrated (vertically) than the true
+values shown in the upper-left panel.
+This is because the points in the lower-right (posterior) panel show
+not typical values or typical draws from the posterior; they show the
+modes (and, with error bars, the variances) of the posterior pdfs.
+When an original data point has a large variance, the posterior ends
+up close to the prior, which has a mode at the center line; the
+posterior modes therefore concentrate more towards the center line
+than any representative sampling.
+A plot of posterior samplings (as opposed to a plot of posterior modes) would
+show a scatter like that in the upper-left (truth) panel.
+
+Bringing the toy model back to our original problem:
+We can get better estimates of the true parallax for each star by inferring the true parallax from the measured parallax and using the noise-deconvolved \cmd, an estimate of the true \cmd, as our prior.
+
+\section{Assumptions and methods}\label{sec:method}
 
@@
 least some aspects of \emph{accuracy} have been improved as well, by
-showing that the posterior distance estimates to stars in stellar
-clusters [HOGG: Is this too general?] are much more clustered than likelihood-based distance
+showing that the posterior distance estimates to stars in (at least one) stellar
+cluster are much more clustered than likelihood-based distance
 estimates.
@@
 
-Come back to the assumptions. What happens if we change or violate them?
-Put in citations
-Long discussion of the noise model. We took serious short-cuts. Call out to Boris.
+In Section~\ref{sec:method} we listed a set of assumptions upon which
+the photometric parallax method in this project is based.
+Here we return to these assumptions, and discuss them critically in more
+detail.
+\begin{description}
+\item[stationarity] Stationarity assumes both the low and high
+  signal to noise objects are being drawn from the same distribution,
+  independent of noise. This isn't precisely true because most
+  \gaia\ objects have similar noise $(\sim 3 \, \mathrm{mas})$, but
+  their signal (parallax) depends on their distance. Nearby stars have
+  higher signal (parallax) because they are closer and the nearby
+  stars are mostly main sequence stars due to the IMF, timing, and
+  malmquist bias. However, we can see bright giants to much larger
+  distances, and therefore lower signal or parallax, leading to lower
+  signal to noise. So the red giant branch is a population of
+  relatively lower signal to noise objects compared with the main
+  sequence. But this does not violate the stationarity assumption in
+  detail, because the low and high signal to noise objects have to be
+  drawn from the same distribution within some feature or patch of the
+  space. This is all assuming the Gaia noise model is correct. If the
+  uncertainties are over estimated then we could be over deconvolving
+  (generating a tighter distribution than the truth) And in some
+  extreme over estimate of uncertainties the lower signal to noise
+  giants could be pulled to the main sequence. But with the quoted
+  uncertainties this seems unlikely. As you move up the giant branch,
+  and probe farther giants, the signal to noise will continue to
+  diminish, but because the color uncertainties are not large, the S/N
+  spread within a patch of the \cmd\ is well mixed. Possibly a
+  metallicity trend with distance, placing more nearby (higher S/N)
+  objects at slightly different positions on the CMD pulling lower S/N
+  objects to that distribution for our prior. But if something has
+  small error bars is good data is good data both for inference and
+  generating the prior
 
-\subsection{Proper Dust Corrections}
-Within our probabilistic framework, we improperly handle dust.
-We are taking a point estimate of the probabilistic dust model at a quantile of our distance posterior, correcting our photometry using this point estimate, and then re-deriving the prior.
-A full, proper account would sample the dust model at distances sampled from the posterior, and propagate those samples back into the prior during the iterative process.
-The reason we don't do a full probabilistic treatment, is that this proper dust distribution would not be a Gaussian distribution, which \xd\ requires.
-We could calculate this more complex dust distribution and estimate it as a Gaussian distribution to feed to \xd, however, this is beyond the scope of this paper.
-
-\subsection{Stationarity Assumption}
-Stationarity assumes both the low and high signal to noise objects are being drawn from the same distribution, independent of noise.
-This isn't precisely true because most \gaia\ objects have similar noise $(\sim 3 \, \mathrm{mas})$, but their signal (parallax) depends on their distance.
-Nearby stars have higher signal (parallax) because they are closer and the nearby stars are mostly main sequence stars due to the IMF, timing, and malmquist bias.
-However, we can see bright giants to much larger distances, and therefore lower signal or parallax, leading to lower signal to noise.
-So the red giant branch is a population of relatively lower signal to noise objects compared with the main sequence.
-But this does not violate the stationarity assumption in detail, because the low and high signal to noise objects have to be drawn from the same distribution within some feature or patch of the space.
-This is all assuming the Gaia noise model is correct.
-If the uncertainties are over estimated then we could be over deconvolving (generating a tighter distribution than the truth) And in some extreme over estimate of uncertainties the lower signal to noise giants could be pulled to the main sequence.
-But with the quoted uncertainties this seems unlikely.
-As you move up the giant branch, and probe farther giants, the signal to noise will continue to diminish, but because the color uncertainties are not large, the S/N spread within a patch of the \cmd\ is well mixed.
-Possibly a metallicity trend with distance, placing more nearby (higher S/N) objects at slightly different positions on the CMD pulling lower S/N objects to that distribution for our prior.
-But if something has small error bars is good data is good data both for inference and generating the prior
-
-For very rare kinds of stars.
-Inference with our prior is bad for underrepresented populations in the \cmd.
-For example, it appears there are only $2$ white dwarfs in \gaia\ DR1.
-With them being very underrepresented in the prior, the amplitudes of the gaussians representing them is very small, so the WDs are more likely to be pulled to mimic a much farther away MS star.
+  For very rare kinds of stars. Inference with our prior is bad for
+  underrepresented populations in the \cmd. For example, it appears
+  there are only $2$ white dwarfs in \gaia\ DR1. With them being very
+  underrepresented in the prior, the amplitudes of the gaussians
+  representing them is very small, so the WDs are more likely to be
+  pulled to mimic a much farther away MS star.
+\item[big data] HOGG 
+\item[noise model] HOGG
+\item[dust] Within our probabilistic framework, we improperly handle
+  dust. We are taking a point estimate of the probabilistic dust model
+  at a quantile of our distance posterior, correcting our photometry
+  using this point estimate, and then re-deriving the prior. A full,
+  proper account would sample the dust model at distances sampled from
+  the posterior, and propagate those samples back into the prior
+  during the iterative process. The reason we don't do a full
+  probabilistic treatment, is that this proper dust distribution would
+  not be a Gaussian distribution, which \xd\ requires. We could
+  calculate this more complex dust distribution and estimate it as a
+  Gaussian distribution to feed to \xd, however, this is beyond the
+  scope of this paper.
+\item[mixture of Gaussians] HOGG
+\item[no physics] HOGG
+\end{description}
 
@@
 
-Compare with literature stellar models
+Anderson: Compare with literature stellar models?? This is an enhanced goal; strike this.
 

### 2017-05-16 paperDistance/paper.tex
assumption discussion needs work

@@
 
+\subsection{Critical discussion of assumptions}
+
 In Section~\ref{sec:method} we listed a set of assumptions upon which
@@
 \begin{description}
-\item[stationarity] Stationarity assumes both the low and high
-  signal to noise objects are being drawn from the same distribution,
-  independent of noise. This isn't precisely true because most
-  \gaia\ objects have similar noise $(\sim 3 \, \mathrm{mas})$, but
-  their signal (parallax) depends on their distance. Nearby stars have
-  higher signal (parallax) because they are closer and the nearby
-  stars are mostly main sequence stars due to the IMF, timing, and
-  malmquist bias. However, we can see bright giants to much larger
-  distances, and therefore lower signal or parallax, leading to lower
-  signal to noise. So the red giant branch is a population of
+\item[stationarity] (HOGG needs to clean this paragraph)
+  We have assumed that (locally) the morphology of
+  the \cmd\ is independent of signal-to-noise; that is, we have assumed that
+  both the low and high
+  signal-to-noise objects are being drawn from the same distribution.
+  This isn't precisely true because nearby stars in \tgas\ are
+  mostly main sequence stars, while we can see luminous giants to much larger
+  distances. So the red giant branch is a population of
   relatively lower signal to noise objects compared with the main
-  sequence. But this does not violate the stationarity assumption in
+  sequence. This does not violate the stationarity assumption in
   detail, because the low and high signal to noise objects have to be
@@
   objects to that distribution for our prior. But if something has
-  small error bars is good data is good data both for inference and
-  generating the prior
+  small error bars it is good data is good data both for inference and
+  generating the prior.
 
